### 2016-05-08 writeup/rocha_etal.tex
Updates tex/big

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% AGUtmpl.tex: this template file is for articles formatted with LaTeX2e,
% Modified July 2014
%
% This template includes commands and instructions
% given in the order necessary to produce a final output that will
% satisfy AGU requirements.
%
% PLEASE DO NOT USE YOUR OWN MACROS
% DO NOT USE \newcommand, \renewcommand, or \def.
%
% FOR FIGURES, DO NOT USE \psfrag or \subfigure.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% All questions should be e-mailed to latex@agu.org.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Step 1: Set the \documentclass
%
% There are two options for article format: two column (default)
% and draft.
%
% PLEASE USE THE DRAFT OPTION TO SUBMIT YOUR PAPERS.
% The draft option produces double spaced output.
%
% Choose the journal abbreviation for the journal you are
% submitting to:

% jgrga JOURNAL OF GEOPHYSICAL RESEARCH
% gbc   GLOBAL BIOCHEMICAL CYCLES
% grl   GEOPHYSICAL RESEARCH LETTERS
% pal   PALEOCEANOGRAPHY
% ras   RADIO SCIENCE
% rog   REVIEWS OF GEOPHYSICS
% tec   TECTONICS
% wrr   WATER RESOURCES RESEARCH
% gc    GEOCHEMISTRY, GEOPHYSICS, GEOSYSTEMS
% sw    SPACE WEATHER
% ms    JAMES
% ef    EARTH'S FUTURE
% ea    EARTH AND SPACE SCIENCE
%
%
%
% (If you are submitting to a journal other than jgrga,
% substitute the initials of the journal for "jgrga" below.)

%\documentclass[draft,grl]{agutex2015}
\documentclass[grl]{agutex2015}

% To create numbered lines:

% If you don't already have lineno.sty, you can download it from
% http://www.ctan.org/tex-archive/macros/latex/contrib/ednotes/
% (or search the internet for lineno.sty ctan), available at TeX Archive Network (CTAN).
% Take care that you always use the latest version.

% To activate the commands, uncomment \usepackage{lineno}
% and \linenumbers*[1]command, below:

% \usepackage{lineno}
% \linenumbers*[1]
%  To add line numbers to lines with equations:
%  \begin{linenomath*}
%  \begin{equation}
%  \end{equation}
%  \end{linenomath*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Figures and Tables
%
%
% DO NOT USE \psfrag or \subfigure commands.
%
%
%  Uncomment the following command to include .eps files
%  (comment out this line for draft format):
%\usepackage[dvipdf]{graphicx}
\usepackage{graphicx}

%
%  Uncomment the following command to allow illustrations to print
%   when using Draft:
%\setkeys{Gin}{draft=false}
%
% Substitute one of the following for [dvips] above
% if you are using a different driver program and want to
% proof your illustrations on your machine:
%
% [xdvi], [dvipdf], [dvipsone], [dviwindo], [emtex], [dviwin],
% [pctexps],  [pctexwin],  [pctexhp],  [pctex32], [truetex], [tcidvi],
% [oztex], [textures]
%
% See how to enter figures and tables at the end of the article, after
% references.
%
%% ------------------------------------------------------------------------ %%
%
%  ENTER PREAMBLE
%
%% ------------------------------------------------------------------------ %%

% Author names in capital letters:
\authorrunninghead{ROCHA ET AL.}

% Shorter version of title entered in capital letters:
\titlerunninghead{Seasonality in dominant upper-ocean dynamics at submesoscales}

%Corresponding author mailing address and e-mail address:
%\authoraddr{Corresponding author: A. B. Smith,
%Department of Hydrology and Water Resources, University of
%Arizona, Harshbarger Building 11, Tucson, AZ 85721, USA.
%(a.b.smith@hwr.arizona.edu)}

\begin{document}

%% ------------------------------------------------------------------------ %%
%
%  TITLE
%
%% ------------------------------------------------------------------------ %%

\title{Seasonality in dominant upper-ocean dynamics at submesoscales}
%
% e.g., \title{Terrestrial ring current:
% Origin, formation, and decay $\alpha\beta\Gamma\Delta$}
%

%% ------------------------------------------------------------------------ %%
%
%  AUTHORS AND AFFILIATIONS
%
%% ------------------------------------------------------------------------ %%

%Use \author{\altaffilmark{}} and \altaffiltext{}

% \altaffilmark will produce footnote;
% matching \altaffiltext will appear at bottom of page.

\authors{Cesar, Sarah, Teri, \& Bill \altaffilmark{1}}
% Eric Brown,\altaffilmark{1,2} Rick Williams,\altaffilmark{3}
% John B. McDougall\altaffilmark{4}, and S. Visconti\altaffilmark{5}}

\altaffiltext{1}{Scripps Institution of Oceanography, University of California San Diego, La Jolla, CA, USA}

%\altaffiltext{2}{Department of Geography, Ohio State University,
%Columbus, Ohio, USA.}

%\altaffiltext{3}{Department of Space Sciences, University of
%Michigan, Ann Arbor, Michigan, USA.}

%\altaffiltext{4}{Division of Hydrologic Sciences, Desert Research
%Institute, Reno, Nevada, USA.}

%\altaffiltext{5}{Dipartimento di Idraulica, Trasporti ed
%Infrastrutture Civili, Politecnico di Torino, Turin, Italy.}

%% ------------------------------------------------------------------------ %%
%
%  KEYPOINTS
%
%% ------------------------------------------------------------------------ %%

% Key points are 1 to 3 points that the author provides,
% that are 100 characters or less, that are ultimately published
% with the article.
%% for example:
% \keypoints{\item Here is the first keypoint. what happens if it is a
% long keypoint, like this one. We want to see this wrap please.
% \item This is the second.
% \item And here is the third keypoint
% }

\keypoints{\item Submesoscale (10-100 km) vertical vorticity and lateral strain peaks in late
                 Winter/early Spring;
                 divergence and stratification peaks in late Summer/early Fall.
           \item Quasi-balanced eddies dominate the
                  variability in late Winter/early Spring; inertia-gravity waves
                  dominate in late Summer/early Fall.
            \item These results have implications for the diagnosis of
                  surface velocity from high-resolution altimeters.
                  }

%% Keypoints will print underneath the abstract.

%% ------------------------------------------------------------------------ %%
%
%  ABSTRACT
%
%% ------------------------------------------------------------------------ %%

% >> Do NOT include any \begin...\end commands within
% >> the body of the abstract.

\begin{abstract}
(Type abstract here)
\end{abstract}

%% ------------------------------------------------------------------------ %%
%
%  BEGIN ARTICLE
%
%% ------------------------------------------------------------------------ %%

% The body of the article must start with a \begin{article} command
%
% \end{article} must follow the references section, before the figures
%  and tables.

\begin{article}

%% ------------------------------------------------------------------------ %%
%
%  TEXT
%
%% ------------------------------------------------------------------------ %%

\section{Introduction}

\subsection{Motivation}
Recent studies have suggested that submesoscale upper-ocean flows
undergo strong seasonal cycle \citep{sasaki_etal2014,qiu_etal2014,
callies_etal2015, thompson_etal2016,buckingham_etal2016}.

Observations at submesoscales are limited.

SWOT and COMPIRA will provide submesoscale-resolving SSH measurements
with global coverage.

\subsection{This paper}
Using the output of a $1/24^\circ$ state estimate, we show that (vertical)
vorticity and (horizontal) divergence undergo strong seasonal cycle that are
out of phase. This seasonal cycle is associated with changes in the upper-ocean
stratification. As previously argued, in late Winter/early Spring,
deep mixed-layers are prone to
shallow baroclinic instabilities that are roughly in geostrophic balance
and flux energy upscale \citep{sasaki_etal2014,callies_etal2016},
driving a mild seasonal modulation at mesoscales \citep{sasaki_etal2014,qiu_etal2014}.
The divergence part of the flow, however, peaks in late Summer/early Fall,
when the upper-ocean is strongly stratified. Hence these results suggest a
strong seasonal modulation of the dominant upper-ocean submesoscale dynamics
with the dominance quasi-balanced turbulence in late Winter/early Spring and
quasi-linear inertia-gravity waves in late Summer/early Fall.

\section{The LLC2160 model output}
We use the output of the latitude-longitude polar cap (LLC)
MITgcm  numerical simulation. The output analyzed here, the LLC2160,
is the 1/24$^\circ$ --- 90-vertical-levels simulation from a hierarchy of global
forward numerical solutions on a LLC grid \citep{forget_etal2015} that were
spun up from a ECCO2 adjoint-method state estimate. The LLC2160 was forced by
tides and high-resolution surface fluxes similarly to the LLC4320 simulation
used by \citet{rocha_etal2016} --- see their appendix D. We chose the second-higher
resolution  simulation (LLC2160) of the LLC hierarchy because it was run for longer:
the LLC2160 output spans two years form March 2011 to April 2013. The model time-step
was 45 s and the model variables were saved hourly. LLC hierarchy model setup
and control files are available online
(at \texttt{http://mitgcm.org/viewvc/MITgcm/MITgcm$\_$contrib/
llc$\_$hires}). The model output was provided by D. Menemenlis (JPL/NASA).

A key aspects: (1) tides. \citep{sasaki_etal2014}.

\section{Results}

\subsection{The seasonal cycle in bulk quantities}

\begin{figure*}[ht]
\begin{center}
\includegraphics[width=1.\textwidth]{figs/fig1.pdf}
 \caption{Snapshots of surface vorticity (a and b) and a transect
          of potential density at 165$^\circ$E (c and d). The snapshots were
          taken at 00:00 UTC.}
 \label{fig1}
 \end{center}
 \end{figure*}

 \begin{figure*}[ht]
   \begin{center}
     \includegraphics[width=.9\textwidth]{figs/fig2.pdf}
  \caption{Time series of horizontal velocity diagnostics at the surface
          (a through d) and stratification in the upper 50 m (e).}
  \label{fig2}
  \end{center}
\end{figure*}

\begin{figure*}[ht]
  \begin{center}
    \includegraphics[width=.9\textwidth]{figs/fig3.pdf}
 \caption{Surface (horizontal) KE and SSH variance wavenumber spectra. Solid lines
 are spectra based on hourly snapshots, dashed lines are spectra based on daily-averaged
 fields.}
 \label{fig3}
 \end{center}
\end{figure*}

\begin{figure}[ht]
  \begin{center}
    \includegraphics[width=.45\textwidth]{figs/fig4.pdf}
 \caption{Surface (horizontal) KE frequency spectra. Solid lines
 are spectra based on the 1/24$^\circ$ fields, dashed line are spectra
 based on 100-km-low-pass smoothed fields.}
 \label{fig4}
 \end{center}
\end{figure}

\subsection{The high-frequency signal}
Here we argue that those divergent motions are likely inertia-gravity waves.
We can show a couple of spectra and argue that they roughly satisfy polarization
relations.

We can also try and show some observations here. Any cruises with ADCP data.
While it may be hard to have enough data for the errorbars to be small, a
rough consistency may be better than nothing.

Perhaps a mooring data should show strong seasonal modulation at high-frequencies?

We must be able to present some observational evidence that the model is not
misleading at high-frequencies.

\subsection{Geostrophic velocities from SSH}
Here we argue that high-frequency (supra-daily) flows significantly project
on the sea-surface, and thus estimation of submesoscale (10-100 km) geostrophic
velocity from sea-surface height (SSH) is not warranted. Calculating spectral
KE fluxes from both velocity and geostrophic velocity (from SSH) would
contrast with the results in Sasaki et al.

\section{Summary and Discussion}
Our main finding is that
upper-ocean divergent flows, dominated by inertia-gravity waves, undergo
a strong seasonal cycle that is out-of-phase with the seasonal cycle of
the quasi-balanced flows. Therefore, the present results suggest significant
seasonal modulation of the dominant upper-ocean dynamics at submesoscales
(10-100 km): Quasi-balanced turbulence driven by shallow baroclinic instabilities
 in late Winter/early Spring
and inertia-gravity waves dominate in late Summer/early Fall.

That the surface velocity and SSH submesoscale (10-100 km) variability may be
dominated by ageostrophic flows in
Summer/Fall has consequences for the interpretation of data from
the future SWOT and COMPIRA altimeter missions,
which will deliver SSH measurements at submesoscales. To the extent that
high-frequency flows are dominated by incoherent internal tides and other
internal waves, it may be very difficult (if not impossible) to
separate submesoscale SSH variability associated with geostrophic motions
from high-frequency, ageostrophic flows. Thus, previous claims that
one will be able to easily obtain submesoscale surface
geostrophic velocities and monitor such seasonal cycle on global scales
\citep{sasaki_etal2014,qiu_etal2014} are overstated.

The present analysis focuses on a single patch of ocean in the
vicinities of the Kuroshio Extension, not
representative of other regions such as eastern boundary currents
and the middle of the subtropical gyre: care must be taken in overgeneralizing
these results with care.  The effects of smaller-scale/higher-frequency
``sub-submesoscale'' flows on the submesoscale surface velocity and SSH variability
are presently unknown.

%  this should go in a supplementary material
%\appendix
%\section*{How well does the model captures high-frequency modes?}
% \begin{figure*}[ht]
%   \begin{center}
%     \includegraphics[width=.9\textwidth]{figs/figA1.pdf}
%  \caption{A comp.}
%  \label{fig4}
%  \end{center}
%\end{figure*}

%%% End of body of article:

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional Appendix goes here
%
% \appendix resets counters and redefines section heads
% but doesn't print anything.
% After typing \appendix
%
%\section{Here Is Appendix Title}
% will show
% Appendix A: Here Is Appendix Title
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Optional Glossary or Notation section, goes here
%
%%%%%%%%%%%%%%
% Glossary is only allowed in Reviews of Geophysics
% \section*{Glossary}
% \paragraph{Term}
% Term Definition here
%
%%%%%%%%%%%%%%
% Notation -- End each entry with a period.
% \begin{notation}
% Term & definition.\\
% Second term & second definition.\\
% \end{notation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  ACKNOWLEDGMENTS

\begin{acknowledgments}
Thanks to D. Menemenlis (JPL/NASA) and the other MITgcm usual suspects.
This research was funded by NSF and NASA.
\end{acknowledgments}

%% ------------------------------------------------------------------------ %%
%%  REFERENCE LIST AND TEXT CITATIONS
%
% Either type in your references using
% \begin{thebibliography}{}
% \bibitem{}
% Text
% \end{thebibliography}
%
% Or,
%
% If you use BiBTeX for your references, please use the agufull08.bst file (available at % ftp://ftp.agu.org/journals/latex/journals/Manuscript-Preparation/) to produce your .bbl
% file and copy the contents into your paper here.
%
% Follow these steps:
% 1. Run LaTeX on your LaTeX file.
%
% 2. Make sure the bibliography style appears as \bibliographystyle{agufull08}. Run BiBTeX on your LaTeX
% file.
%
% 3. Open the new .bbl file containing the reference list and
%   copy all the contents into your LaTeX file here.
%
% 4. Comment out the old \bibliographystyle and \bibliography commands.
%
% 5. Run LaTeX on your new file before submitting.
%
% AGU does not want a .bib or a .bbl file. Please copy in the contents of your .bbl file here.

%\begin{thebibliography}{}

\bibliographystyle{agufull08}
\bibliography{rocha_etal}

%\providecommand{\natexlab}[1]{#1}
%\expandafter\ifx\csname urlstyle\endcsname\relax
%  \providecommand{\doi}[1]{doi:\discretionary{}{}{}#1}\else
%  \providecommand{\doi}{doi:\discretionary{}{}{}\begingroup
%  \urlstyle{rm}\Url}\fi
%
%\bibitem[{\textit{Atkinson and Sloan}(1991)}]{AtkinsonSloan}
%Atkinson, K., and I.~Sloan (1991), The numerical solution of first-kind
%  logarithmic-kernel integral equations on smooth open arcs, \textit{Math.
%  Comp.}, \textit{56}(193), 119--139.
%
%\bibitem[{\textit{Colton and Kress}(1983)}]{ColtonKress1}
%Colton, D., and R.~Kress (1983), \textit{Integral Equation Methods in
%  Scattering Theory}, John Wiley, New York.
%
%\bibitem[{\textit{Hsiao et~al.}(1991)\textit{Hsiao, Stephan, and
%  Wendland}}]{StephanHsiao}
%Hsiao, G.~C., E.~P. Stephan, and W.~L. Wendland (1991), On the {D}irichlet
%  problem in elasticity for a domain exterior to an arc, \textit{J. Comput.
%  Appl. Math.}, \textit{34}(1), 1--19.
%
%\bibitem[{\textit{Lu and Ando}(2012)}]{LuAndo}
%Lu, P., and M.~Ando (2012), Difference of scattering geometrical optics
%  components and line integrals of currents in modified edge representation,
%  \textit{Radio Sci.}, \textit{47},  RS3007, \doi{10.1029/2011RS004899}.

%\end{thebibliography}

%Reference citation examples:

%...as shown by \textit{Kilby} [2008].
%...as shown by {\textit  {Lewin}} [1976], {\textit  {Carson}} [1986], {\textit  {Bartholdy and Billi}} [2002], and {\textit  {Rinaldi}} [2003].
%...has been shown [\textit{Kilby et al.}, 2008].
%...has been shown [{\textit  {Lewin}}, 1976; {\textit  {Carson}}, 1986; {\textit  {Bartholdy and Billi}}, 2002; {\textit  {Rinaldi}}, 2003].
%...has been shown [e.g., {\textit  {Lewin}}, 1976; {\textit  {Carson}}, 1986; {\textit  {Bartholdy and Billi}}, 2002; {\textit  {Rinaldi}}, 2003].

%...as shown by \citet{jskilby}.
%...as shown by \citet{lewin76}, \citet{carson86}, \citet{bartoldy02}, and \citet{rinaldi03}.
%...has been shown \citep{jskilbye}.
%...has been shown \citep{lewin76,carson86,bartoldy02,rinaldi03}.
%...has been shown \citep [e.g.,][]{lewin76,carson86,bartoldy02,rinaldi03}.
%
% Please use ONLY \citet and \citep for reference citations.
% DO NOT use other cite commands (e.g., \cite, \citeyear, \nocite, \citealp, etc.).

%% ------------------------------------------------------------------------ %%
%
%  END ARTICLE
%
%% ------------------------------------------------------------------------ %%
\end{article}
%
%
%% Enter Figures and Tables here:
%
% DO NOT USE \psfrag or \subfigure commands.
%
% Figure captions go below the figure.
% Table titles go above tables; all other caption information
%  should be placed in footnotes below the table.
%
%----------------
% EXAMPLE FIGURE
%
 %\begin{figure}
 %\noindent\includegraphics[width=20pc]{samplefigure.eps}
 %\caption{Caption text here}
 %\label{figure_label}
 %\end{figure}
%
% ---------------
% EXAMPLE TABLE
%
%\begin{table}
%\caption{Time of the Transition Between Phase 1 and Phase 2\tablenotemark{a}}
%\centering
%\begin{tabular}{l c}
%\hline
% Run  & Time (min)  \\
%\hline
%  $l1$  & 260   \\
%  $l2$  & 300   \\
%  $l3$  & 340   \\
%  $h1$  & 270   \\
%  $h2$  & 250   \\
%  $h3$  & 380   \\
%  $r1$  & 370   \\
%  $r2$  & 390   \\
%\hline
%\end{tabular}
%\tablenotetext{a}{Footnote text here.}
%\end{table}

% See below for how to make sideways figures or tables.

\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

More Information and Advice:

%% ------------------------------------------------------------------------ %%
%
%  SECTION HEADS
%
%% ------------------------------------------------------------------------ %%

% Capitalize the first letter of each word (except for
% prepositions, conjunctions, and articles that are
% three or fewer letters).

% AGU follows standard outline style; therefore, there cannot be a section 1 without
% a section 2, or a section 2.3.1 without a section 2.3.2.
% Please make sure your section numbers are balanced.
% ---------------
% Level 1 head
%
% Use the \section{} command to identify level 1 heads;
% type the appropriate head wording between the curly
% brackets, as shown below.
%
%An example:
%\section{Level 1 Head: Introduction}
%
% ---------------
% Level 2 head
%
% Use the \subsection{} command to identify level 2 heads.
%An example:
%\subsection{Level 2 Head}
%
% ---------------
% Level 3 head
%
% Use the \subsubsection{} command to identify level 3 heads
%An example:
%\subsubsection{Level 3 Head}
%
%---------------
% Level 4 head
%
% Use the \subsubsubsection{} command to identify level 3 heads
% An example:
%\subsubsubsection{Level 4 Head} An example.
%
%% ------------------------------------------------------------------------ %%
%
%  IN-TEXT LISTS
%
%% ------------------------------------------------------------------------ %%
%
% Do not use bulleted lists; enumerated lists are okay.
% \begin{enumerate}
% \item
% \item
% \item
% \end{enumerate}
%
%% ------------------------------------------------------------------------ %%
%
%  EQUATIONS
%
%% ------------------------------------------------------------------------ %%

% Single-line equations are centered.
% Equation arrays will appear left-aligned.

Math coded inside display math mode \[ ...\]
 will not be numbered, e.g.,:
 \[ x^2=y^2 + z^2\]

 Math coded inside \begin{equation} and \end{equation} will
 be automatically numbered, e.g.,:
 \begin{equation}
 x^2=y^2 + z^2
 \end{equation}

% IF YOU HAVE MULTI-LINE EQUATIONS, PLEASE
% BREAK THE EQUATIONS INTO TWO OR MORE LINES
% OF SINGLE COLUMN WIDTH (20 pc, 8.3 cm)
% using double backslashes (\\).

% To create multiline equations, use the
% \begin{eqnarray} and \end{eqnarray} environment
% as demonstrated below.
\begin{eqnarray}
  x_{1} & = & (x - x_{0}) \cos \Theta \nonumber \\
        && + (y - y_{0}) \sin \Theta  \nonumber \\
  y_{1} & = & -(x - x_{0}) \sin \Theta \nonumber \\
        && + (y - y_{0}) \cos \Theta.
\end{eqnarray}

%If you don't want an equation number, use the star form:
%\begin{eqnarray*}...\end{eqnarray*}

% Break each line at a sign of operation
% (+, -, etc.) if possible, with the sign of operation
% on the new line.

% Indent second and subsequent lines to align with
% the first character following the equal sign on the
% first line.

% Use an \hspace{} command to insert horizontal space
% into your equation if necessary. Place an appropriate
% unit of measure between the curly braces, e.g.
% \hspace{1in}; you may have to experiment to achieve
% the correct amount of space.

%% ------------------------------------------------------------------------ %%
%
%  EQUATION NUMBERING: COUNTER
%
%% ------------------------------------------------------------------------ %%

% You may change equation numbering by resetting
% the equation counter or by explicitly numbering
% an equation.

% To explicitly number an equation, type \eqnum{}
% (with the desired number between the brackets)
% after the \begin{equation} or \begin{eqnarray}
% command.  The \eqnum{} command will affect only
% the equation it appears with; LaTeX will number
% any equations appearing later in the manuscript
% according to the equation counter.
%

% If you have a multiline equation that needs only
% one equation number, use a \nonumber command in
% front of the double backslashes (\\) as shown in
% the multiline equation above.

%% ------------------------------------------------------------------------ %%
%
%  SIDEWAYS FIGURE AND TABLE EXAMPLES
%
%% ------------------------------------------------------------------------ %%
%
% For tables and figures, add \usepackage{rotating} to the paper and add the rotating.sty file to the folder.
% AGU prefers the use of {sidewaystable} over {landscapetable} as it causes fewer problems.
%
% \begin{sidewaysfigure}
% \includegraphics[width=20pc]{samplefigure.eps}
% \caption{caption here}
% \label{label_here}
% \end{sidewaysfigure}
%
%
%
% \begin{sidewaystable}
% \caption{}
% \begin{tabular}
% Table layout here.
% \end{tabular}
% \end{sidewaystable}
%
%

### 2016-05-08 writeup/rocha_etal.tex
Updates section 2 and 3

@@
 the LLC2160 output spans two years form March 2011 to April 2013. The model time-step
-was 45 s and the model variables were saved hourly. LLC hierarchy model setup
+was 45s and the model variables were saved hourly. LLC hierarchy setup
 and control files are available online
 (at \texttt{http://mitgcm.org/viewvc/MITgcm/MITgcm$\_$contrib/
-llc$\_$hires}). The model output was provided by D. Menemenlis (JPL/NASA).
+llc$\_$hires}). For the LLC hierarchy, the sea-surface height is calculated
+using a linearized equation. The model output was provided by D. Menemenlis
+(JPL/NASA).
 
-A key aspects: (1) tides. \citep{sasaki_etal2014}.
+A key aspect of the simulation analyzed here is that it is forced by
+the 16 most significant tidal components.
+Barotropic tides interact with topography and generate internal
+tides that project on mesoscales to submesoscales
+\citep{rocha_etal2016}. Hence, tidal forcing fundamentally differentiates our analysis
+from other modeling studies that investigated upper dynamics
+and its seasonality \citep{sasaki_etal2014,qiu_etal2014}.
 
-\section{Results}
+The region of focus is the northwest Pacific, in the vicinities of the Kuroshio
+Extension, where previous studies have
+shown strong seasonality in the quasi-gesotrophic surface velocity field.
+We analyze a sub-domain of the LLC2160 simulation of about 2000 km$^2$
+spanning 155-175$^\circ$E; 25-40$^\circ$N. We focus on the surface velocity field
+and near-surface stratification.
 
-\subsection{The seasonal cycle in bulk quantities}
+\section{Seasonality in surface velocity and near-surface stratification}
+
+We calculate the vertical vorticity, horizontal divergence, and the magnitude
+of the lateral rate of strain:
+\begin{equation}\label{vort}
+\zeta \equiv v_x - u_y\, ,
+\end{equation}
+\begin{equation}\label{div}
+\delta \equiv u_x + v_y\, ,
+\end{equation}
+and
+\begin{equation}\label{strain}
+S \equiv [(v_x+u_y)^2 + (u_x-v_y)^2]^{1/2}\, ,
+\end{equation}
 

### 2016-05-09 writeup/rocha_etal.tex
Adds details of numerical method
Adds more content to secs 3 and 4

@@
 \citep{rocha_etal2016}. Hence, tidal forcing fundamentally differentiates our analysis
-from other modeling studies that investigated upper dynamics
+from recent modeling studies that investigated upper dynamics
 and its seasonality \citep{sasaki_etal2014,qiu_etal2014}.
@@
 We analyze a sub-domain of the LLC2160 simulation of about 2000 km$^2$
-spanning 155-175$^\circ$E; 25-40$^\circ$N. We focus on the surface velocity field
-and near-surface stratification.
+spanning 155-175$^\circ$E; 25-40$^\circ$N (figure \ref{fig1}a-b).
 
@@
 
-We calculate the vertical vorticity, horizontal divergence, and the magnitude
-of the lateral rate of strain:
-\begin{equation}\label{vort}
-\zeta \equiv v_x - u_y\, ,
-\end{equation}
-\begin{equation}\label{div}
-\delta \equiv u_x + v_y\, ,
-\end{equation}
-and
-\begin{equation}\label{strain}
-S \equiv [(v_x+u_y)^2 + (u_x-v_y)^2]^{1/2}\, ,
-\end{equation}
+Focusing on the lateral gradients of the horizontal velocity vector.
+The lateral gradients are calculated using a centered
+second-order finite differences scheme. We then calculate
+the vertical vorticity $\zeta \equiv v_x - u_y$,
+horizontal divergence $\delta \equiv u_x + v_y$, and the magnitude
+of the lateral rate of strain $S \equiv [(v_x+u_y)^2 + (u_x-v_y)^2]^{1/2}$.
+These quantities highlight the submesoscale structures in the flow
+and hint on their dynamics.
+
+Figures \ref{fig1}a-b show snapshots of vertical vorticity $\zeta$ in early Spring
+(April 15) and Fall (October 15). Clearly, a strong seasonality in vorticity
+exists: large values of fine-grained vertical vorticity are
+observed in early Spring with maximum values as high as $0.9f$, where $f$ is
+the local planetary vorticity, and root-mean-square (RMS) of about 0.25. In early
+Fall, the situation is the opposite: the vertical vorticity is coarse-grained,
+and its local maximum and RMS is no larger than $0.3f$ and $0.1f$, respectively.
+Indeed, the seasonal cycle in vorticity/strain is very strong (Figure~\ref{fig2}):
+the RMS vertical vorticity is about twice as large in late Winter/early Spring
+than it is in late Summer/early Fall. This picture is consistent with
+recent studies: shallow baroclinic instabilities energize the submesoscales
+in late Winter, drawing from the available potential energy stored in deep mixed
+layers \citep{sasaki_etal2014,callies_etal2015,callies_etal2016}.
 
@@
 
+The bulk of the vertical vorticity is associated with subinertial flows
+($T_f = 2\pi/f_0\approx 23.5$h, where $f_0$ is the inertial frequency at
+the mean latitude): daily-averaging the velocity fields reduces the vertical
+vorticity by less than 20$\%$ and the seasonal cycle remains strong (see
+green lines in figure \ref{fig2}). Most of this seasonal cycle is associated
+with submesoscale flows: low-pass filtering the velocity field
+with a cut-off scale of 100 km reduces the RMS vorticity by at least 40$\%$,
+yielding bulk statistics of velocity gradient roughly consistent with AVISO
+geostrophic velocity data (see lines blue and black in figure \ref{fig2}).
+
+The horizontal divergent part of the flow is significant: the RMS is $\delta/f > 0.1$).
+Most of these horizontal divergent flows project are supra-inertial and project
+on scales smaller than 100 km: daily-averaging the velocity fields or low-pass
+filtering (100 km cutoff) reduces the horizontal divergence by about 70$\%$.
+Furthermore, the seasonal cycle in horizontal divergence is out-of-phase with that
+of vertical vorticity: the horizontal divergence is about 30$\%$ larger in late
+Summer/early Fall than it is in late Winter/early Spring. The divergence associated
+with submesoscale subinertial eddies is small and its seasonal cycle is in phase
+with that of vorticity (see green lines).  The seasonal cycle of the horizontal
+divergence is roughly in phase with the seasonal cycle of the near-surface (0-50m)
+stratification.
+
  \begin{figure*}[ht]
@@
      \includegraphics[width=.9\textwidth]{figs/fig2.pdf}
-  \caption{Time series of horizontal velocity diagnostics at the surface
-          (a through d) and stratification in the upper 50 m (e).}
+  \caption{Time series of the root-mean-square of kinetic energy (a), vorticity
+  (b), lateral strain (c), horizontal divergence (d), and stratification in the upper 50 m (e).}
   \label{fig2}
@@
 \end{figure*}
+
+\section{Projection on horizontal scales}
+
+% Here we argue that those divergent motions are likely inertia-gravity waves.
+% We can show a couple of spectra and argue that they roughly satisfy polarization
+% relations.
+%
+% We can also try and show some observations here. Any cruises with ADCP data.
+% While it may be hard to have enough data for the errorbars to be small, a
+% rough consistency may be better than nothing.
+%
+% Perhaps a mooring data should show strong seasonal modulation at high-frequencies?
+%
+% We must be able to present some observational evidence that the model is not
+% misleading at high-frequencies.
+
+% Here we argue that high-frequency (supra-daily) flows significantly project
+% on the sea-surface, and thus estimation of submesoscale (10-100 km) geostrophic
+% velocity from sea-surface height (SSH) is not warranted. Calculating spectral
+% KE fluxes from both velocity and geostrophic velocity (from SSH) would
+% contrast with the results in Sasaki et al.
 
@@
 
+\section{Projection on time scales}
+To better understand the dynamics the projection of these flows on horizontal
+scales, we calculate wavenumber spectra of kinetic energy and sea-surface height
+(SSH) variance.
+
 \begin{figure}[ht]
@@
 \end{figure}
-
-\subsection{The high-frequency signal}
-Here we argue that those divergent motions are likely inertia-gravity waves.
-We can show a couple of spectra and argue that they roughly satisfy polarization
-relations.
-
-We can also try and show some observations here. Any cruises with ADCP data.
-While it may be hard to have enough data for the errorbars to be small, a
-rough consistency may be better than nothing.
-
-Perhaps a mooring data should show strong seasonal modulation at high-frequencies?
-
-We must be able to present some observational evidence that the model is not
-misleading at high-frequencies.
-
-\subsection{Geostrophic velocities from SSH}
-Here we argue that high-frequency (supra-daily) flows significantly project
-on the sea-surface, and thus estimation of submesoscale (10-100 km) geostrophic
-velocity from sea-surface height (SSH) is not warranted. Calculating spectral
-KE fluxes from both velocity and geostrophic velocity (from SSH) would
-contrast with the results in Sasaki et al.
 

### 2016-05-09 writeup/rocha_etal.tex
Adds significant text on sections 4 and 5

@@
 llc$\_$hires}). For the LLC hierarchy, the sea-surface height is calculated
-using a linearized equation. The model output was provided by D. Menemenlis
-(JPL/NASA).
+using a linearized equation. While the nominal horizontal resolution is 1/24$^\circ$,
+ horizontal wavenumber spectra suggests an effective resolution of about 20 km.
 
@@
 
-\begin{figure*}[ht]
-\begin{center}
-\includegraphics[width=1.\textwidth]{figs/fig1.pdf}
- \caption{Snapshots of surface vorticity (a and b) and a transect
-          of potential density at 165$^\circ$E (c and d). The snapshots were
-          taken at 00:00 UTC.}
- \label{fig1}
- \end{center}
- \end{figure*}
-
 The bulk of the vertical vorticity is associated with subinertial flows
@@
 stratification.
+\begin{figure*}[ht]
+\begin{center}
+\includegraphics[width=1.\textwidth]{figs/fig1.pdf}
+ \caption{Snapshots of surface vorticity (a and b) and a transect
+          of potential density at 165$^\circ$E (c and d). The snapshots were
+          taken at 00:00 UTC.}
+ \label{fig1}
+ \end{center}
+ \end{figure*}
 
@@
 \section{Projection on horizontal scales}
+
+To better understand the dynamics the projection of these flows on horizontal
+scales, we calculate wavenumber spectra of kinetic energy and sea-surface height
+(SSH) variance. For simplicity we calculate zonal wavenumber spectra at various
+latitudes. The two-year time averaged and any spatial linear trend were
+removed before calculating discrete Fourier transforms.Statistical significance
+is obtained by averaging over latitude and time. Using conservative temporal and
+spatial decorrelation estimate of 10 days and 250 km, we obtain roughly 50
+independent realizations of the spectra.
+
+Figure \ref{fig3}a depicts the horizontal wavenumber spectra of surface KE in
+April and October. At scales larger than 20km, the spectra based on hourly
+velocity snapshots are nearly indistinguishable within 95$\%$ confidence level
+(compare solid lines in figure \ref{fig3}a). Consistent with the results of
+\citet{rocha_etal2016} who analyzed the LLC4320 output in Drake Passage,
+there is significant high-frequency variabilty at submesoscales. Daily-averaging
+the velocity field suppresses spatial variability at scales small than about 250
+km both in April and October (compare solid lines against dashed lines in figure
+\ref{fig3}a). But this suppression is more dramatic in October, when the horizontally
+divergent flows peak. At scales
+smaller than 100km, 40$\%$ of the surface KE in April a is accounted for by suprainertial
+flows as opposed to 77$\%$ in October. The seasonal of subinertial submesoscale flows
+is significant, consistent with the results of \citet{sasaki_etal2014}, which are
+based on snapshots of a model without tidal forcing.
+
+These high-frequency submesoscale flows significantly project on the sea-surface.
+In October, when the horizontally divergent flows peak and the submesoscale
+balanced flows are minimum, there is a dramatic difference between the spectra
+based on hourly and daily-averaged SSH: at scales smaller than 100 km, the spectra
+of hourly SSH roughly follows a $-2$ power-law, whereas the spectra of daily-averaged
+SSH roughly follows a $-5$ power-law. At scales
+smaller than 100km, 44$\%$ of the surface KE in April is accounted for by suprainertial
+flows as opposed to 86$\%$ in October! The out-of-phase seasonal cycle of subinertial
+and suprainertial flows conspire to yield no significant seasonality in the spectra
+of hourly SSH.
 
@@
 \section{Projection on time scales}
-To better understand the dynamics the projection of these flows on horizontal
-scales, we calculate wavenumber spectra of kinetic energy and sea-surface height
-(SSH) variance.
+To better understand the which flows project onto high-wavenumbers, the projection
+of these flows on time scales, we calculate frequency spectra of surface KE
+(figure \ref{fig4}). To obtain statistically meaningful results, we average
+spectral estimates along a section at 165$^\circ$E.  There is also about 50
+independent realizations of spectral estimates.
+
+The surface KE spectra of subinertial flows very roughly follow a $-2$ power-law.
+The high-frequency flows are split near-inertial flows, tidal signals and their
+higher harmonics. Hence these high-frequency flows are dominates by inertia-gravity
+waves. (Note that the the inertial peak nearly merges with the
+diurnal tide peak). There is statistically significant seasonality. The seasonal
+cycle of the sub-inertial motions (at least for periods between 5-20 days)
+is out-of-phase with the seasonal cycle of supra-inertial flows: the former
+are more energetic in April and the latter in October.
+
+Most of these high-frequency inertia-gravity waves project onto scales smaller
+than 100 km. The surface KE spectra of 100-km-smoothed velocity field is significantly
+suppressed in supra-inertial variability --- there are no statistically significant
+changes at sub-inertial frequencies (see dashed lines in figure \ref{fig4}).
+
+% say something about relative energy in smoother vs. unsmoothed fields.
+% also say something about comparison with observations.
 

### 2016-05-09 writeup/rocha_etal.tex
Changes fig2

@@
 \usepackage{graphicx}
+
+% CR added this
+\usepackage{color}
 
@@
 second-order finite differences scheme. We then calculate
-the vertical vorticity $\zeta \equiv v_x - u_y$,
-horizontal divergence $\delta \equiv u_x + v_y$, and the magnitude
-of the lateral rate of strain $S \equiv [(v_x+u_y)^2 + (u_x-v_y)^2]^{1/2}$.
+the vertical vorticity $\zeta \equiv v_x - u_y$ and 
+horizontal divergence $\delta \equiv u_x + v_y$.
 These quantities highlight the submesoscale structures in the flow
@@
           taken at 00:00 UTC.}
+\vspace{-1.5cm}
  \label{fig1}
@@
      \includegraphics[width=.9\textwidth]{figs/fig2.pdf}
-  \caption{Time series of the root-mean-square of kinetic energy (a), vorticity
-  (b), lateral strain (c), horizontal divergence (d), and stratification in the upper 50 m (e).}
+  \caption{Time series of the root-mean-square of vorticity (a),
+  horizontal divergence (b), and stratification in the upper 50 m (c).
+  \textcolor{red}{May add a time series of baroclinic production $\overline{w'b'}$ in
+  the mixed layer. }.}
   \label{fig2}
@@
 \section{Projection on time scales}
-To better understand the which flows project onto high-wavenumbers, the projection
-of these flows on time scales, we calculate frequency spectra of surface KE
-(figure \ref{fig4}). To obtain statistically meaningful results, we average
-spectral estimates along a section at 165$^\circ$E.  There is also about 50
+To better understand the high-frequency/high-wavenumber flows  we calculate frequency
+spectra of surface KE (figure \ref{fig4}). To obtain statistically meaningful results,
+we average spectral estimates along a section at 165$^\circ$E.  There is also about 50
 independent realizations of spectral estimates.
@@
  are spectra based on the 1/24$^\circ$ fields, dashed line are spectra
- based on 100-km-low-pass smoothed fields.}
+ based on 100-km-smoothed fields.}
  \label{fig4}

### 2016-05-09 writeup/rocha_etal.tex
Improves introduction and adds appendix

@@
 
-\keypoints{\item Submesoscale (10-100 km) vertical vorticity and lateral strain peaks in late
-                 Winter/early Spring;
-                 divergence and stratification peaks in late Summer/early Fall.
-           \item Quasi-balanced eddies dominate the
+\keypoints{\item Submesoscale (10-100 km) vertical vorticity and horizontal divergence
+                 divergence undergo a strong seasonal cycle that are out-of-phase.
+           \item Quasi-balanced  turbulence dominates the
                   variability in late Winter/early Spring; inertia-gravity waves
@@
 
-\subsection{Motivation}
-Recent studies have suggested that submesoscale upper-ocean flows
-undergo strong seasonal cycle \citep{sasaki_etal2014,qiu_etal2014,
-callies_etal2015, thompson_etal2016,buckingham_etal2016}.
+Recent interest in upper-ocean dynamics has focused on the strong seasonal
+cycle of shallow baroclinic instabilities and their role in submesoscale (1-100 km)
+turbulence and mesoscale modulation \citep{sasaki_etal2014,qiu_etal2014,
+callies_etal2015, thompson_etal2016,buckingham_etal2016}. Contemporary studies
+have also suggested that inertia-gravity waves contribute significantly
+to the near-surface variability at scale submesoscales \citep{richman_etal2012,
+buhler_etal2014,rocha_etal2016}.
 
-Observations at submesoscales are limited.
-
-SWOT and COMPIRA will provide submesoscale-resolving SSH measurements
-with global coverage.
-
-\subsection{This paper}
-Using the output of a $1/24^\circ$ state estimate, we show that (vertical)
-vorticity and (horizontal) divergence undergo strong seasonal cycle that are
-out of phase. This seasonal cycle is associated with changes in the upper-ocean
-stratification. As previously argued, in late Winter/early Spring,
-deep mixed-layers are prone to
+Using the output of a $1/24^\circ$ global simulation, we show that the near-surface
+expression of inertia-gravity waves undergo strong seasonal cycle that is out-of-phase
+with the seasonal cycle of the balanced variability. This seasonal cycle is associated
+with changes in the upper-ocean
+stratification. Consistent with previous studies,
+deep late Winter mixed-layers are prone to
 shallow baroclinic instabilities that are roughly in geostrophic balance
 and flux energy upscale \citep{sasaki_etal2014,callies_etal2016},
-driving a mild seasonal modulation at mesoscales \citep{sasaki_etal2014,qiu_etal2014}.
-The divergence part of the flow, however, peaks in late Summer/early Fall,
+driving a mild seasonal modulation of the mesoscales \citep{sasaki_etal2014,qiu_etal2014}.
+The horizontally divergent part of the flow, however, peaks in late Summer/early Fall,
 when the upper-ocean is strongly stratified. Hence these results suggest a
-strong seasonal modulation of the dominant upper-ocean submesoscale dynamics
-with the dominance quasi-balanced turbulence in late Winter/early Spring and
+strong seasonal modulation of the dominant upper-ocean submesoscale dynamics:
+quasi-balanced turbulence in late Winter/early Spring and
 quasi-linear inertia-gravity waves in late Summer/early Fall.
@@
 %  this should go in a supplementary material
-%\appendix
-%\section*{How well does the model captures high-frequency modes?}
-% \begin{figure*}[ht]
-%   \begin{center}
-%     \includegraphics[width=.9\textwidth]{figs/figA1.pdf}
-%  \caption{A comp.}
-%  \label{fig4}
-%  \end{center}
-%\end{figure*}
+\appendix
+\section{How well does the model captures high-frequency modes?}
+ \begin{figure*}[ht]
+   \begin{center}
+     \includegraphics[width=.9\textwidth]{figs/figA1.pdf}
+  \caption{A comparison.}
+  \label{figA1}
+  \end{center}
+\end{figure*}
 

### 2016-05-09 writeup/rocha_etal.tex
Adds appendix

@@
 tides that project on mesoscales to submesoscales
-\citep{rocha_etal2016}. Hence, tidal forcing fundamentally differentiates our analysis
+\citep{rocha_etal2016}. Hence, tidal forcing fundamentally distinguishes our analysis
 from recent modeling studies that investigated upper dynamics
@@
 \section{How well does the model captures high-frequency modes?}
+To assess how well the LLC2160 represents the high-frequency variability,
+we compare the model near-surface velocity field against two available
+mooring data: KEO (the data is available online at \texttt{http://www.pmel.noaa.gov/ocs/KEO})
+and KESS mooring $\#$7 (the data is available online at \texttt{http://uskess.whoi.edu}).
+For proper comparison, we use model data closest to these two moorings (west of
+the region considered in the present study). Focusing on high-frequencies, we
+calculate frequency spectral estimates every month. There are about 48 independent
+spectral realizations.
+
+Figure \ref{figA1} shows KE frequency spectra for the model and mooring data.
+At the uppermost common depth (20 m), very low frequencies (periods > 10 days)
+and the inertial, diurnal, and semi-diurnal frequencies agree well --- the KESS
+ data are more energetic at intermedia frequencies. At 40 m, the model and observations
+ have consistent spectra at periods larger than about 6 h. The data from moorings
+are more energetic at higher frequencies and tidal harmonics are not observed,
+except for the first two harmonics that are apparent in the KESS data at 40 m.
+Three reasons for the lack of tidal harmonics in the mooring data are: 1) The high-
+frequency variability is dominated by high-frequency noise due to mooring vertical
+excursions; 2) The model does not have enough resolution to resolve the full spectrum
+of ultraviolet catastrophic wave-wave interactions; 3) The model is not forced at
+frequencies higher that 6h.
+
  \begin{figure*}[ht]
@@
      \includegraphics[width=.9\textwidth]{figs/figA1.pdf}
-  \caption{A comparison.}
+  \caption{A near-surface KE spectra comparison between model and two
+  available mooring observations.}
   \label{figA1}

### 2016-05-09 writeup/rocha_etal.tex
Fixes typos

@@
 % Shorter version of title entered in capital letters:
-\titlerunninghead{Seasonality in dominant upper-ocean dynamics at submesoscales}
+\titlerunninghead{Seasonality in governing upper-ocean dynamics at submesoscales}
 
@@
 
-\title{Seasonality in dominant upper-ocean dynamics at submesoscales}
+\title{Seasonality in governing upper-ocean dynamics at submesoscales}
 %
@@
 \section{The LLC2160 model output}
-We use the output of the latitude-longitude polar cap (LLC)
-MITgcm  numerical simulation. The output analyzed here, the LLC2160,
+We use the output of a latitude-longitude polar cap (LLC)
+MITgcm  numerical simulation  \citep{forget_etal2015}. The output analyzed here, the LLC2160,
 is the 1/24$^\circ$ --- 90-vertical-levels simulation from a hierarchy of global
-forward numerical solutions on a LLC grid \citep{forget_etal2015} that were
+forward numerical solutions on a LLC grid that were
 spun up from a ECCO2 adjoint-method state estimate. The LLC2160 was forced by
-tides and high-resolution surface fluxes similarly to the LLC4320 simulation
+tides and 6-hourly surface fluxes similarly to the LLC4320 simulation
 used by \citet{rocha_etal2016} --- see their appendix D. We chose the second-higher
@@
 llc$\_$hires}). For the LLC hierarchy, the sea-surface height is calculated
-using a linearized equation. While the nominal horizontal resolution is 1/24$^\circ$,
+using a linearized equation. While the nominal horizontal resolution  of the
+LLC2160 simulation is 1/24$^\circ$,
  horizontal wavenumber spectra suggests an effective resolution of about 20 km.
+ (see figure \ref{fig3}).
 
@@
 
-Focusing on the lateral gradients of the horizontal velocity vector.
+Focusing on the submesoscale variability, we present estimates of bulk quantities
+associated with the the horizontal velocity lateral gradient.
 The lateral gradients are calculated using a centered
@@
 green lines in figure \ref{fig2}). Most of this seasonal cycle is associated
-with submesoscale flows: low-pass filtering the velocity field
+with submesoscale flows: smoothing the velocity field
 with a cut-off scale of 100 km reduces the RMS vorticity by at least 40$\%$,
 yielding bulk statistics of velocity gradient roughly consistent with AVISO
-geostrophic velocity data (see lines blue and black in figure \ref{fig2}).
+geostrophic velocities (see lines blue and black in figure \ref{fig2}).
 
 The horizontal divergent part of the flow is significant: the RMS is $\delta/f > 0.1$).
-Most of these horizontal divergent flows project are supra-inertial and project
-on scales smaller than 100 km: daily-averaging the velocity fields or low-pass
-filtering (100 km cutoff) reduces the horizontal divergence by about 70$\%$.
-Furthermore, the seasonal cycle in horizontal divergence is out-of-phase with that
-of vertical vorticity: the horizontal divergence is about 30$\%$ larger in late
+Most of these horizontal divergent flows are supra-inertial and project
+on scales smaller than 100 km: daily-averaging or smoothing the velocity fields
+ (100 km cutoff) reduces the horizontal divergence by about 70$\%$.
+Furthermore, the seasonal cycle in horizontal divergence is out-of-phase with the
+seasonal cycle of vertical vorticity: the horizontal divergence is about 30$\%$ larger in late
 Summer/early Fall than it is in late Winter/early Spring. The divergence associated
@@
 
-To better understand the dynamics the projection of these flows on horizontal
+To better understand the projection of these flows onto different horizontal
 scales, we calculate wavenumber spectra of kinetic energy and sea-surface height
 (SSH) variance. For simplicity we calculate zonal wavenumber spectra at various
-latitudes. The two-year time averaged and any spatial linear trend were
+latitudes. The two-year time averaged and spatial linear trend were
 removed before calculating discrete Fourier transforms.Statistical significance
@@
 divergent flows peak. At scales
-smaller than 100km, 40$\%$ of the surface KE in April a is accounted for by suprainertial
-flows as opposed to 77$\%$ in October. The seasonal of subinertial submesoscale flows
+smaller than 100 km, 40$\%$ of the surface KE in April a is accounted for by suprainertial
+flows as opposed to 77$\%$ in October. The seasonality of subinertial submesoscale flows
 is significant, consistent with the results of \citet{sasaki_etal2014}, which are
@@
 SSH roughly follows a $-5$ power-law. At scales
-smaller than 100km, 44$\%$ of the surface KE in April is accounted for by suprainertial
+smaller than 100 km, 44$\%$ of the surface KE in April is accounted for by suprainertial
 flows as opposed to 86$\%$ in October! The out-of-phase seasonal cycle of subinertial
@@
 spectra of surface KE (figure \ref{fig4}). To obtain statistically meaningful results,
-we average spectral estimates along a section at 165$^\circ$E.  There is also about 50
-independent realizations of spectral estimates.
+we average spectral estimates along a section at 165$^\circ$E.  There is about 50
+independent spectral realizations.
 
@@
 The high-frequency flows are split near-inertial flows, tidal signals and their
-higher harmonics. Hence these high-frequency flows are dominates by inertia-gravity
-waves. (Note that the the inertial peak nearly merges with the
-diurnal tide peak). There is statistically significant seasonality. The seasonal
+higher harmonics. Hence these high-frequency flows are dominated by inertia-gravity
+waves. (Note that the inertial peak nearly merges with the
+diurnal tide peak.) There is statistically significant seasonality. The seasonal
 cycle of the sub-inertial motions (at least for periods between 5-20 days)
@@
 than 100 km. The surface KE spectra of 100-km-smoothed velocity field is significantly
-suppressed in supra-inertial variability --- there are no statistically significant
+suppressed in supra-inertial frequencies --- there are no statistically significant
 changes at sub-inertial frequencies (see dashed lines in figure \ref{fig4}).
@@
 seasonal modulation of the dominant upper-ocean dynamics at submesoscales
-(10-100 km): Quasi-balanced turbulence driven by shallow baroclinic instabilities
+(10-100 km): quasi-balanced turbulence driven by shallow baroclinic instabilities
  in late Winter/early Spring
@@
 which will deliver SSH measurements at submesoscales. To the extent that
-high-frequency flows are dominated by incoherent internal tides and other
+high-frequency flows are dominated by spatially incoherent internal tides and other
 internal waves, it may be very difficult (if not impossible) to
@@
 and the middle of the subtropical gyre: care must be taken in overgeneralizing
-these results with care.  The effects of smaller-scale/higher-frequency
+these results.  The effects of smaller-scale/higher-frequency
 ``sub-submesoscale'' flows on the submesoscale surface velocity and SSH variability

### 2016-05-09 writeup/rocha_etal.tex
Fixes typos

@@
 have also suggested that inertia-gravity waves contribute significantly
-to the near-surface variability at scale submesoscales \citep{richman_etal2012,
+to the near-surface variability at submesoscales \citep{richman_etal2012,
 buhler_etal2014,rocha_etal2016}.
@@
 Using the output of a $1/24^\circ$ global simulation, we show that the near-surface
-expression of inertia-gravity waves undergo strong seasonal cycle that is out-of-phase
+expression of inertia-gravity waves undergoes a strong seasonal cycle that is out-of-phase
 with the seasonal cycle of the balanced variability. This seasonal cycle is associated
@@
 %%% End of body of article:
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2016-05-11 writeup/rocha_etal.tex
Fixes typos

@@
 resolution  simulation (LLC2160) of the LLC hierarchy because it was run for longer:
-the LLC2160 output spans two years form March 2011 to April 2013. The model time-step
+the LLC2160 output spans two years from March 2011 to April 2013. The model time-step
 was 45s and the model variables were saved hourly. LLC hierarchy setup
@@
 \citep{rocha_etal2016}. Hence, tidal forcing fundamentally distinguishes our analysis
-from recent modeling studies that investigated upper dynamics
+from recent modeling studies that investigated upper-ocean dynamics
 and its seasonality \citep{sasaki_etal2014,qiu_etal2014}.
@@
 Extension, where previous studies have
-shown strong seasonality in the quasi-gesotrophic surface velocity field.
+shown strong seasonality in submesoscale turbulence.
 We analyze a sub-domain of the LLC2160 simulation of about 2000 km$^2$
@@
 Focusing on the submesoscale variability, we present estimates of bulk quantities
-associated with the the horizontal velocity lateral gradient.
+associated with the lateral gradient of the horizontal velocity.
 The lateral gradients are calculated using a centered
@@
 and its local maximum and RMS is no larger than $0.3f$ and $0.1f$, respectively.
-Indeed, the seasonal cycle in vorticity/strain is very strong (Figure~\ref{fig2}):
+Indeed, the seasonal cycle in vorticity is very strong (Figure~\ref{fig2}):
 the RMS vertical vorticity is about twice as large in late Winter/early Spring
@@
 with submesoscale flows: smoothing the velocity field
-with a cut-off scale of 100 km reduces the RMS vorticity by at least 40$\%$,
-yielding bulk statistics of velocity gradient roughly consistent with AVISO
-geostrophic velocities (see lines blue and black in figure \ref{fig2}).
+with a Hanning filter with cut-off scale of 100 km reduces the RMS vorticity by
+at least 40$\%$, yielding bulk statistics of velocity gradient roughly consistent
+with AVISO geostrophic velocities (see lines blue and black in figure \ref{fig2}).
 
@@
  (100 km cutoff) reduces the horizontal divergence by about 70$\%$.
-Furthermore, the seasonal cycle in horizontal divergence is out-of-phase with the
+Interestingly, the seasonal cycle in horizontal divergence is out-of-phase with the
 seasonal cycle of vertical vorticity: the horizontal divergence is about 30$\%$ larger in late
@@
 latitudes. The two-year time averaged and spatial linear trend were
-removed before calculating discrete Fourier transforms.Statistical significance
+removed, and the ensuing fields were Hanning spectral ``window''. Statistical significance
 is obtained by averaging over latitude and time. Using conservative temporal and
@@
 there is significant high-frequency variabilty at submesoscales. Daily-averaging
-the velocity field suppresses spatial variability at scales small than about 250
+the velocity field suppresses spatial variability at scales smaller than about 250
 km both in April and October (compare solid lines against dashed lines in figure
@@
 divergent flows peak. At scales
-smaller than 100 km, 40$\%$ of the surface KE in April a is accounted for by suprainertial
+smaller than 100 km, 40$\%$ of the surface KE in April is accounted for by supra-inertial
 flows as opposed to 77$\%$ in October. The seasonality of subinertial submesoscale flows
@@
 flows as opposed to 86$\%$ in October! The out-of-phase seasonal cycle of subinertial
-and suprainertial flows conspire to yield no significant seasonality in the spectra
+and suprainertial flows conspire to yield nearly insignificant seasonality in the spectra
 of hourly SSH.
@@
 cycle of the sub-inertial motions (at least for periods between 5-20 days)
-is out-of-phase with the seasonal cycle of supra-inertial flows: the former
+is out-of-phase with the seasonal cycle of supra-inertial flows: the former flows
 are more energetic in April and the latter in October.
@@
 
-That the surface velocity and SSH submesoscale (10-100 km) variability may be
+That the  surface velocity and SSH submesoscale variability may be
 dominated by ageostrophic flows in
@@
 internal waves, it may be very difficult (if not impossible) to
-separate submesoscale SSH variability associated with geostrophic motions
+separate SSH submesoscale variability associated with geostrophic motions
 from high-frequency, ageostrophic flows. Thus, previous claims that

### 2016-06-04 writeup/rocha_etal.tex
Updates figures
Adds new figure

@@
 
-\authors{Cesar, Sarah, Teri, \& Bill \altaffilmark{1}}
+\authors{Cesar et al. \altaffilmark{1}}
 % Eric Brown,\altaffilmark{1,2} Rick Williams,\altaffilmark{3}
@@
 \begin{abstract}
-(Type abstract here)
+    We show that internal waves are a loud noise.
 \end{abstract}
@@
 \begin{center}
-\includegraphics[width=1.\textwidth]{figs/fig1.pdf}
- \caption{Snapshots of surface vorticity (a and b) and a transect
-          of potential density at 165$^\circ$E (c and d). The snapshots were
+\hspace{-1.25cm}\includegraphics[width=.7\textwidth]{figs/fig1_1.pdf}\\
+\vspace{-.125cm}
+\includegraphics[width=.9\textwidth]{figs/fig1_2.pdf}
+ \caption{(a) The study region with the subregion where the LLC outputs are
+          analyzed. Snapshots of surface vorticity (b and c) and a transect
+          of potential density at 165$^\circ$E (d and e). The snapshots were
           taken at 00:00 UTC.}
@@
 
-\begin{figure}[ht]
-  \begin{center}
-    \includegraphics[width=.45\textwidth]{figs/fig4.pdf}
- \caption{Surface (horizontal) KE frequency spectra. Solid lines
- are spectra based on the 1/24$^\circ$ fields, dashed line are spectra
- based on 100-km-smoothed fields.}
- \label{fig4}
- \end{center}
-\end{figure}
+%\begin{figure}[ht]
+%  \begin{center}
+%    \includegraphics[width=.45\textwidth]{figs/fig4.pdf}
+% \caption{Surface (horizontal) KE frequency spectra. Solid lines
+% are spectra based on the 1/24$^\circ$ fields, dashed line are spectra
+% based on 100-km-smoothed fields.}
+% \label{fig4}
+% \end{center}
+%\end{figure}
 
@@
 %  this should go in a supplementary material
-\appendix
-\section{How well does the model captures high-frequency modes?}
+\clearpage
+
+\section{Supplemental material 1: Spectral errors}
+
+\section{Supplemental material 2: On the LLC simulations}
+
+\subsection{Seasonality in the upper-ocean stratification}
+
+\subsection{How well does the model captures high-frequency modes?}
 To assess how well the LLC2160 represents the high-frequency variability,
@@
 \end{figure*}
+
+\section{Supplemental material 3: Seasonality in pressure modes near-surface amplitude}
 
@@
 \begin{acknowledgments}
-Thanks to D. Menemenlis (JPL/NASA) and the other MITgcm usual suspects.
+Thanks to the MITgcm usual suspects.
 This research was funded by NSF and NASA.

### 2016-06-04 writeup/rocha_etal.tex
Fixes kenged
Update subsection heads

@@
 
-Using the output of a $1/24^\circ$ global simulation, we show that the near-surface
+Using the output of two new primitive equation $1/24^\circ$ and $1/24^\circ$ global
+ simulations with embedded tides, we show that the near-surface
 expression of inertia-gravity waves undergoes a strong seasonal cycle that is out-of-phase
@@
 
-\section{The LLC2160 model output}
+\section{The LLC model outputs}
 We use the output of a latitude-longitude polar cap (LLC)
@@
  \caption{(a) The study region with the subregion where the LLC outputs are
-          analyzed. Snapshots of surface vorticity (b and c) and a transect
+          analyzed. LLC 4320 (1/48$^\circ$) snapshots of surface vorticity (b and c) and a transect
           of potential density at 165$^\circ$E (d and e). The snapshots were
@@
      \includegraphics[width=.9\textwidth]{figs/fig2.pdf}
-  \caption{Time series of the root-mean-square of vorticity (a),
-  horizontal divergence (b), and stratification in the upper 50 m (c).
-  \textcolor{red}{May add a time series of baroclinic production $\overline{w'b'}$ in
-  the mixed layer. }.}
+  \caption{Time series of the root-mean-square (RMS) of vorticity (a) and
+  horizontal divergence (b) in the LLC outputs and gridded AVISO data.}
   \label{fig2}

### 2016-06-05 writeup/rocha_etal.tex
Edits figs
New supplemental figure
Edits captions of figures

@@
 
+\appendix
+
 \section{Supplemental material 1: Spectral errors}
@@
 \subsection{Seasonality in the upper-ocean stratification}
+
+\begin{figure*}[ht]
+   \begin{center}
+     \includegraphics[width=1.\textwidth]{figs/fig_s2_1.pdf}
+  \caption{Potential density: a comparison between Roemmich-Gilson Argo
+          climatology (a, b) and the LLC simulations 1/24$^\circ$ (c, d)
+          and  1/48$^\circ$ (e, f). \textcolor{red}{Need to re-plot  the 1/24$^\circ$
+          figures once extraction (currently down) becomes available}.}
+  \label{figS2_1}
+  \end{center}
+\end{figure*}
+
+\subsection{Spectra}
+
+\begin{figure*}[ht]
+   \begin{center}
+     \includegraphics[width=1.\textwidth]{figs/fig_s2_3.pdf}
+  \caption{A comparison between spectra of the two LLC simulations 1/24$^\circ$
+  (green) and 1/48$^\circ$ (purple). The statistical errorbars represent conservative
+  significance levels considering that there are only four independent relations
+  for the monthly spectra.}
+  \label{figS2_3}
+  \end{center}
+\end{figure*}
+
+\subsection{Seasonality in statistics: probability density functions}
+\begin{figure*}[ht]
+   \begin{center}
+     \includegraphics[width=1.\textwidth]{figs/fig_s2_2.pdf}
+  \caption{Time dependence of probability density functions of vertical vorticity (a, b)
+  and horizontal divergence (c, d). Monthly averages in April and October are
+  also shown (e, f).}
+  \label{figS2_2}
+  \end{center}
+\end{figure*}
 
@@
      \includegraphics[width=.9\textwidth]{figs/figA1.pdf}
-  \caption{A near-surface KE spectra comparison between model and two
-  available mooring observations.}
+  \caption{A comparison between frequency spectra of LLC simulations and two
+  available moored current meter records.}
   \label{figA1}
@@
 \section{Supplemental material 3: Seasonality in pressure modes near-surface amplitude}
+
+\begin{figure*}[ht]
+   \begin{center}
+     \includegraphics[width=1.\textwidth]{figs/fig_s3.pdf}
+  \caption{The seasonal variability of the WOA 2013 stratification averaged  over
+  the domain and the associated gravest three pressure modes. Only the upper 450 m is shown.}
+  \label{figS3}
+  \end{center}
+\end{figure*}
 
@@
 \begin{acknowledgments}
-Thanks to the MITgcm usual suspects.
-This research was funded by NSF and NASA.
+We thank the MITgcm community and our colleagues at the NASA Advanced
+Supercomputing (NAS) Division for their awesome support.
+This research was funded by NASA and NSF.
 \end{acknowledgments}

### 2016-06-06 writeup/rocha_etal.tex
Updates figures and bib
Fixes legends

@@
   \caption{Time series of the root-mean-square (RMS) of vorticity (a) and
-  horizontal divergence (b) in the LLC outputs and gridded AVISO data.}
+  lateral rate of strain (b), and horizontal divergence (c) in the LLC outputs and gridded AVISO data.}
   \label{fig2}
+  \end{center}
+\end{figure*}
+
+\begin{figure*}[ht]
+  \begin{center}
+    \includegraphics[width=.9\textwidth]{figs/fig3.pdf}
+ \caption{Seasonal variation of joint probability distributions: vorticity vs.
+          divergence (a through e) and vorticity vs. strain rate (f through j).
+          Dashed lines in (a) through (e) represent $\delta = \frac{\omega}{f_0}
+          \zeta$ where $\omega = 2\pi/$(12 h, 3 h) and $f_0$ is the inertial frequency
+          in the middle of the domain (32.5$^\circ$ N). Dashed lines in (f) through (j)
+          represent pure strain flow $\alpha = \zeta$.}
+ \label{fig3}
  \end{center}
@@
   \begin{center}
-    \includegraphics[width=.9\textwidth]{figs/fig3.pdf}
- \caption{Surface (horizontal) KE and SSH variance wavenumber spectra. Solid lines
+    \includegraphics[width=.9\textwidth]{figs/fig4.pdf}
+ \caption{Surface (horizontal) KE (a) and SSH variance wavenumber spectra (b)
+ in the1/48$^\circ$ simulation. Solid lines
  are spectra based on hourly snapshots, dashed lines are spectra based on daily-averaged

### 2016-06-07 writeup/rocha_etal.tex
Significant additions to the text

@@
 \usepackage{color}
+\usepackage{amsmath}
 
@@
 
-\keypoints{\item Submesoscale (10-100 km) vertical vorticity and horizontal divergence
-                 divergence undergo a strong seasonal cycle that are out-of-phase.
-           \item Quasi-balanced  turbulence dominates the
-                  variability in late Winter/early Spring; inertia-gravity waves
-                  dominate in late Summer/early Fall.
-            \item These results have implications for the diagnosis of
-                  surface velocity from high-resolution altimeters.
+\keypoints{\item Upper-ocean submesoscale (10-100 km) turbulence and innertia-gravity waves
+                 undergo strong seasonal cycles that are out-of-phase.
+           \item Submesoscale turbulence dominates the horizontal  velocity and
+                  sea-surface height variability in late Winter/early Spring.
+            \item Inertia-gravity waves are summertime enhanced near the surface and
+                  dominate the surface variability in late Summer/early Fall.
                   }
@@
 
-\section{The LLC model outputs}
-We use the output of a latitude-longitude polar cap (LLC)
-MITgcm  numerical simulation  \citep{forget_etal2015}. The output analyzed here, the LLC2160,
-is the 1/24$^\circ$ --- 90-vertical-levels simulation from a hierarchy of global
-forward numerical solutions on a LLC grid that were
-spun up from a ECCO2 adjoint-method state estimate. The LLC2160 was forced by
-tides and 6-hourly surface fluxes similarly to the LLC4320 simulation
-used by \citet{rocha_etal2016} --- see their appendix D. We chose the second-higher
-resolution  simulation (LLC2160) of the LLC hierarchy because it was run for longer:
-the LLC2160 output spans two years from March 2011 to April 2013. The model time-step
-was 45s and the model variables were saved hourly. LLC hierarchy setup
-and control files are available online
-(at \texttt{http://mitgcm.org/viewvc/MITgcm/MITgcm$\_$contrib/
-llc$\_$hires}). For the LLC hierarchy, the sea-surface height is calculated
-using a linearized equation. While the nominal horizontal resolution  of the
-LLC2160 simulation is 1/24$^\circ$,
- horizontal wavenumber spectra suggests an effective resolution of about 20 km.
- (see figure \ref{fig3}).
+\section{The LLC numerical simulations}
+We use the output of two latitude-longitude polar cap (LLC)
+MITgcm realstic numerical simulations. The outputs
+ analyzed here, LLC2160 and LLC4320,
+are the 1/24$^\circ$ and 1/48$^\circ$  --- 90-vertical-levels simulations from
+a hierarchy of global
+forward numerical solutions on a LLC grid   \citep{forget_etal2015} that were
+spun up from a ECCO2 adjoint-method state estimate. Both simulations were forced by
+tides and 6-hourly surface fluxes. The LLC2160
+output spans two years from March 2011 to April 2013; the LLC4320 was spun up from
+the LLC2160 simulation, spanning one year from September 2011 to September 2012.
+The LLC4320 simulation is an extension of the then 5-month long output used by
+\citet{rocha_etal2016}.
 
@@
 tides that project on mesoscales to submesoscales
-\citep{rocha_etal2016}. Hence, tidal forcing fundamentally distinguishes our analysis
+\citep{rocha_etal2016}. Tidal forcing fundamentally distinguishes our analysis
 from recent modeling studies that investigated upper-ocean dynamics
-and its seasonality \citep{sasaki_etal2014,qiu_etal2014}.
+and its seasonality \citep{sasaki_etal2014,qiu_etal2014}. The LLC2160 and LLC4320
+simulations fully resolve internal tides and
+other hydrostatic internal waves and other submesoscale processes at scales as
+small as about 5 km.
 
-The region of focus is the northwest Pacific, in the vicinities of the Kuroshio
-Extension, where previous studies have
-shown strong seasonality in submesoscale turbulence.
-We analyze a sub-domain of the LLC2160 simulation of about 2000 km$^2$
-spanning 155-175$^\circ$E; 25-40$^\circ$N (figure \ref{fig1}a-b).
+To study seasonal variations in the upper-ocean dynamics, we focus on the northwest
+Pacific, in the vicinities of the Kuroshio
+Extension, where previous studies have suggested strong submesoscale seasonality.
+We analyze a sub-domain of the LLC4320 and LLC2160 simulations of about 2000 km$^2$
+spanning 155-175$^\circ$E; 25-40$^\circ$N (figure \ref{fig1}a). The stratification
+in this mesoscale-rich subtropical region undergoes a strong seasonal cycle: wintertime
+enhance small-scale turbulence de-stratify the upper ocean, yielding mixed-layers
+as deep as 300m. In late Spring/early Summer enhanced solar radiation and mixed-layer
+instabilities re-stratify the upper ocean, yielding mixed-layers as shallow as 40m.
+The stratification and its seasonality is well-captured by both LLC simulations.
+The excellent comparisons of model stratification to Argo climatology lends confidence
+to the present analysis (supplemental material).
 
-\section{Seasonality in surface velocity and near-surface stratification}
+\section{Statistics of surface lateral velocity tensor}
 
-Focusing on the submesoscale variability, we present estimates of bulk quantities
-associated with the lateral gradient of the horizontal velocity.
-The lateral gradients are calculated using a centered
+We present estimates of bulk quantities
+associated with the lateral velocity tensor
+\begin{equation}
+\nabla_h \bf{u}_h = \left[\begin{matrix} u_x & u_y\\ v_x&v_y \end{matrix}\right] \,.
+\end{equation}
+The components of the lateral velocity tensor are calculated using a centered
 second-order finite differences scheme. We then calculate
-the vertical vorticity $\zeta \equiv v_x - u_y$ and
-horizontal divergence $\delta \equiv u_x + v_y$.
-These quantities highlight the submesoscale structures in the flow
-and hint on their dynamics.
+the vertical vorticity $\zeta \equiv v_x - u_y$, lateral rate of strain
+$\alpha \equiv [(u_x-v_y)^2 + (u_y + v_x)^2]^{1/2}$, and horizontal divergence $\delta \equiv u_x + v_y$.
+These diagnostics highlight the submesoscale structures in the flow
+and their analysis hint on the governing dynamics in the upper ocean
+\citep{capet_etal2008a,shcherbina_etal2013}.
 
-Figures \ref{fig1}a-b show snapshots of vertical vorticity $\zeta$ in early Spring
-(April 15) and Fall (October 15). Clearly, a strong seasonality in vorticity
+Figures \ref{fig1}b-c show snapshots of vertical vorticity $\zeta$ in early Spring
+(April 15) and Fall (October 15) in the LLC4320 (1/48$^\circ$) simulation.
+The model solutions captures theseasonality in vorticity
 exists: large values of fine-grained vertical vorticity are
-observed in early Spring with maximum values as high as $0.9f$, where $f$ is
-the local planetary vorticity, and root-mean-square (RMS) of about 0.25. In early
-Fall, the situation is the opposite: the vertical vorticity is coarse-grained,
-and its local maximum and RMS is no larger than $0.3f$ and $0.1f$, respectively.
-Indeed, the seasonal cycle in vorticity is very strong (Figure~\ref{fig2}):
-the RMS vertical vorticity is about twice as large in late Winter/early Spring
-than it is in late Summer/early Fall. This picture is consistent with
-recent studies: shallow baroclinic instabilities energize the submesoscales
-in late Winter, drawing from the available potential energy stored in deep mixed
-layers \citep{sasaki_etal2014,callies_etal2015,callies_etal2016}.
+observed in early Spring with maximum values as high as $1.5f$, where $f$ is
+the local planetary vorticity, and root-mean-square (RMS) of about 0.4. In early
+Fall, the situation is the opposite: the vertical vorticity is relatively coarse-grained,
+and its local maximum and RMS are no larger than $0.5f$ and $0.2f$, respectively.
+Indeed, the seasonal cycle in vorticity is very strong (figure~\ref{fig2}):
+in both simulations, the RMS vertical vorticity and lateral strain rate are about twice as large in
+late Winter/early Spring
+than it is in late Summer/early Fall. Because the wintertime vorticity and strain
+are dominated by the smallest scales in the flow, increasing the resolution from
+1/24$^\circ$ to 1/48$^\circ$ significantly increases the wintertime RMS vorticity
+and strain by about 40$\%$.
 
-The bulk of the vertical vorticity is associated with subinertial flows
+The bulk of vertical vorticity and lateral strain rate is associated with subinertial flows
 ($T_f = 2\pi/f_0\approx 23.5$h, where $f_0$ is the inertial frequency at
-the mean latitude): daily-averaging the velocity fields reduces the vertical
-vorticity by less than 20$\%$ and the seasonal cycle remains strong (see
-green lines in figure \ref{fig2}). Most of this seasonal cycle is associated
-with submesoscale flows: smoothing the velocity field
-with a Hanning filter with cut-off scale of 100 km reduces the RMS vorticity by
-at least 40$\%$, yielding bulk statistics of velocity gradient roughly consistent
-with AVISO geostrophic velocities (see lines blue and black in figure \ref{fig2}).
+the mean latitude): daily-averaging the velocity fields reduces the RMS
+vorticity by less than 40$\%$ and the RMS strain by less than 10$\%$; the seasonal
+cycle remains strong (see
+dashed lines in figures \ref{fig2}a-b). Most of this seasonal cycle is associated
+with submesoscale flows: smoothing the velocity fields
+with a Hanning filter with cut-off scale of 100 km significantly reduces the RMS
+vorticity and strain. The reduction in variance is dramatic ($>$80$\%$) in winter
+yielding RMS vorticity and strain roughly consistent the ones diagnosed from
+AVISO gridded geostrophic velocities (compare blue lines with black lines in figures \ref{fig2}a-b).
+The overall picture that emerges is consistent with recent studies: shallow baroclinic
+instabilities energize the submesoscales in late Winter, drawing from the available
+potential energy stored in deep mixed layers \citep{sasaki_etal2014,callies_etal2015,callies_etal2016}.
 
-The horizontal divergent part of the flow is significant: the RMS is $\delta/f > 0.1$).
+The seasonal cycle of the horizontal divergent part of the flow, however, add complexity
+to the submesoscale-turbulence-centred scenario depicted by those invertigators.
+
+ is significant: the RMS is $\delta/f > 0.1$).
 Most of these horizontal divergent flows are supra-inertial and project
@@
 \end{figure*}
+
+\section{Seasonality in governing submesoscale dynamics}
 
@@
  \caption{Surface (horizontal) KE (a) and SSH variance wavenumber spectra (b)
- in the1/48$^\circ$ simulation. Solid lines
+ in the 1/48$^\circ$ simulation. Solid lines
  are spectra based on hourly snapshots, dashed lines are spectra based on daily-averaged
@@
 \section{Supplemental material 2: On the LLC simulations}
+
+The time-step was 45s for
+the LLC2160 simulation and 25s for the LLC4320 simulation. Details of the LLC simulations
+are provided in the supplemental material.
+
+ LLC hierarchy setup
+and control files are available online
+(at \texttt{http://mitgcm.org/viewvc/MITgcm/MITgcm$\_$contrib/
+llc$\_$hires}). For the LLC hierarchy, the sea-surface height is calculated
+using a linearized equation. While the nominal horizontal resolution  of the
+LLC2160 simulation is 1/24$^\circ$,
+ horizontal wavenumber spectra suggests an effective resolution of about 20 km.
+ (see figure \ref{fig3}).
 

### 2016-06-07 writeup/rocha_etal.tex
Updates abstract abd intro

@@
 
-\keypoints{\item Upper-ocean submesoscale (10-100 km) turbulence and innertia-gravity waves
+\keypoints{\item  Upper-ocean submesoscale (10-100 km) turbulence and inertia-gravity waves
                   undergo strong seasonal cycles that are out-of-phase.
@@
                   sea-surface height variability in late Winter/early Spring.
-            \item Inertia-gravity waves are summertime enhanced near the surface and
-                  dominate the surface variability in late Summer/early Fall.
+            \item Inertia-gravity waves dominate the the horizontal velocity and
+                   sea-surface height variability in late Summer/early Fall.
                   }
@@
 \begin{abstract}
-    We show that internal waves are a loud noise.
+    Two new high-resolution numerical simulations with embedded tides suggest a
+    strong modulation in the governing near-surface dynamics at submesoscales
+    (roughly 10-100 km). Consistent with recent studies, deep late Winter mixed layers
+    are prone to baroclinic instabilities and submesoscale turbulence
+    prevails in wintertime. Inertia-gravity waves, on the other hand, are enhanced near
+    the surface when the upper ocean in strongly stratified. Hence, in summertime,
+    inertia-gravity waves strongly dominate the surface kinetic energy and
+    sea-surface height variance at submesoscales.
 \end{abstract}
@@
 Recent interest in upper-ocean dynamics has focused on the strong seasonal
-cycle of shallow baroclinic instabilities and their role in submesoscale (1-100 km)
+cycle of shallow baroclinic instabilities and their role in submesoscale roughly 1-100 km)
 turbulence and mesoscale modulation \citep{sasaki_etal2014,qiu_etal2014,
@@
 to the near-surface variability at submesoscales \citep{richman_etal2012,
-buhler_etal2014,rocha_etal2016}.
+buhler_etal2014,rocha_etal2016}, but their seasonality has not been investigated.
 
-Using the output of two new primitive equation $1/24^\circ$ and $1/24^\circ$ global
- simulations with embedded tides, we show that the near-surface
-expression of inertia-gravity waves undergoes a strong seasonal cycle that is out-of-phase
-with the seasonal cycle of the balanced variability. This seasonal cycle is associated
-with changes in the upper-ocean
-stratification. Consistent with previous studies,
-deep late Winter mixed-layers are prone to
+Using the output of two new $1/24^\circ$ and $1/48^\circ$ global
+numerical simulations with embedded tides, we show that inertia-gravity waves undergo
+a strong seasonality near the surface in the Kuroshio Extension region.
+The seasonal cycle of inertia-gravity waves is out-of-phase
+with the seasonal cycle of submesoscale turbulence. Consistent with previous studies,
+deep late Winter mixed layers are prone to
 shallow baroclinic instabilities that are roughly in geostrophic balance
@@
 driving a mild seasonal modulation of the mesoscales \citep{sasaki_etal2014,qiu_etal2014}.
-The horizontally divergent part of the flow, however, peaks in late Summer/early Fall,
-when the upper-ocean is strongly stratified. Hence these results suggest a
-strong seasonal modulation of the dominant upper-ocean submesoscale dynamics:
-quasi-balanced turbulence in late Winter/early Spring and
-quasi-linear inertia-gravity waves in late Summer/early Fall.
+
+Inertia-gravity waves, on the other hand, peak in late Summer/early Fall,
+when the upper ocean is strongly stratified. Thus there exists a
+strong seasonal modulation in the dominant upper-ocean submesoscale dynamics:
+submesoscale turbulence dominates the upper-ocean dynamics in late Winter/early Spring,
+whereas quasi-linear inertia-gravity waves prevail in late Summer/early Fall.
+Because submesoscale turbulence is significantly diminished in late Summer/early Fall,
+the present results suggest that inertia-gravity waves account for most of the
+summertime submesoscale SSH variability.
 
@@
 to the present analysis (supplemental material).
+
+\begin{figure*}[ht]
+\begin{center}
+\hspace{-1.25cm}\includegraphics[width=.7\textwidth]{figs/fig1_1.pdf}\\
+\vspace{-.125cm}
+\includegraphics[width=.9\textwidth]{figs/fig1_2.pdf}
+ \caption{(a) The study region with the subregion where the LLC outputs are
+          analyzed. LLC 4320 (1/48$^\circ$) snapshots of surface vorticity (b and c) and a transect
+          of potential density at 165$^\circ$E (d and e). The snapshots were
+          taken at 00:00 UTC.}
+\vspace{-1.5cm}
+ \label{fig1}
+ \end{center}
+ \end{figure*}
 
@@
 stratification.
-\begin{figure*}[ht]
-\begin{center}
-\hspace{-1.25cm}\includegraphics[width=.7\textwidth]{figs/fig1_1.pdf}\\
-\vspace{-.125cm}
-\includegraphics[width=.9\textwidth]{figs/fig1_2.pdf}
- \caption{(a) The study region with the subregion where the LLC outputs are
-          analyzed. LLC 4320 (1/48$^\circ$) snapshots of surface vorticity (b and c) and a transect
-          of potential density at 165$^\circ$E (d and e). The snapshots were
-          taken at 00:00 UTC.}
-\vspace{-1.5cm}
- \label{fig1}
- \end{center}
- \end{figure*}
 

### 2016-06-07 writeup/rocha_etal.tex
Minor fixes on intro/abs

@@
     are prone to baroclinic instabilities and submesoscale turbulence
-    prevails in wintertime. Inertia-gravity waves, on the other hand, are enhanced near
-    the surface when the upper ocean in strongly stratified. Hence, in summertime,
-    inertia-gravity waves strongly dominate the surface kinetic energy and
-    sea-surface height variance at submesoscales.
+    prevails in late Winter/early Spring. Inertia-gravity waves, on the other hand, are enhanced near
+    the surface when the upper ocean in strongly stratified. As a consequence of
+    reduction of submesoscale turbulence and enhancement of intertia-gravity waves,
+    the latter strongly dominate the submesoscale surface kinetic energy and
+    sea-surface height variance in late Summer/early Fall.
 \end{abstract}
@@
 whereas quasi-linear inertia-gravity waves prevail in late Summer/early Fall.
-Because submesoscale turbulence is significantly diminished in late Summer/early Fall,
+Because submesoscale turbulence is severely diminished in late Summer/early Fall,
 the present results suggest that inertia-gravity waves account for most of the

### 2016-06-08 writeup/rocha_etal.tex
Sec 3 check

@@
 
-A key aspect of the simulation analyzed here is that it is forced by
+A key aspect of the LLC simulations is that there were forced by
 the 16 most significant tidal components.
-Barotropic tides interact with topography and generate internal
+Because barotropic tides interact with topography and generate internal
 tides that project on mesoscales to submesoscales
-\citep{rocha_etal2016}. Tidal forcing fundamentally distinguishes our analysis
+\citep[e.g., ][]{rocha_etal2016}, tidal forcing fundamentally distinguishes our analysis
 from recent modeling studies that investigated upper-ocean dynamics
-and its seasonality \citep{sasaki_etal2014,qiu_etal2014}. The LLC2160 and LLC4320
-simulations fully resolve internal tides and
-other hydrostatic internal waves and other submesoscale processes at scales as
-small as about 5 km.
+and its seasonality \citep{sasaki_etal2014,qiu_etal2014}. Details of the LLC simulations
+are provided in the supplemental material.
 
@@
 in this mesoscale-rich subtropical region undergoes a strong seasonal cycle: wintertime
-enhance small-scale turbulence de-stratify the upper ocean, yielding mixed-layers
+enhance small-scale turbulence de-stratify the upper ocean, yielding mixed layers
 as deep as 300m. In late Spring/early Summer enhanced solar radiation and mixed-layer
-instabilities re-stratify the upper ocean, yielding mixed-layers as shallow as 40m.
-The stratification and its seasonality is well-captured by both LLC simulations.
-The excellent comparisons of model stratification to Argo climatology lends confidence
+instabilities re-stratify the upper ocean, yielding mixed layers as shallow as 40m.
+The stratification and its seasonality is well-captured by both LLC simulations:
+the skillful comparisons of model stratification to Argo climatology lends confidence
 to the present analysis (supplemental material).
@@
 \section{Statistics of surface lateral velocity tensor}
-
-We present estimates of bulk quantities
-associated with the lateral velocity tensor
+To study the seasonality in the surface velocity, we calculate the lateral velocity tensor
 \begin{equation}
@@
 and their analysis hint on the governing dynamics in the upper ocean
-\citep{capet_etal2008a,shcherbina_etal2013}.
+\citep[e.g.,][]{capet_etal2008a,shcherbina_etal2013}.
 
@@
 (April 15) and Fall (October 15) in the LLC4320 (1/48$^\circ$) simulation.
-The model solutions captures theseasonality in vorticity
-exists: large values of fine-grained vertical vorticity are
-observed in early Spring with maximum values as high as $1.5f$, where $f$ is
+The model solutions depict seasonality in vorticity: large values of
+fine-grained vertical vorticity are observed in early Spring with maximum
+values as large as $4f$, where $f$ is
 the local planetary vorticity, and root-mean-square (RMS) of about 0.4. In early
@@
 and its local maximum and RMS are no larger than $0.5f$ and $0.2f$, respectively.
-Indeed, the seasonal cycle in vorticity is very strong (figure~\ref{fig2}):
-in both simulations, the RMS vertical vorticity and lateral strain rate are about twice as large in
+Indeed, the seasonal cycle of vorticity and rate of strain is very strong (figure~\ref{fig2}):
+in both simulations, the RMS vorticity and strain are about twice as large in
 late Winter/early Spring
-than it is in late Summer/early Fall. Because the wintertime vorticity and strain
+than in late Summer/early Fall. Because the wintertime vorticity and strain rate
 are dominated by the smallest scales in the flow, increasing the resolution from
@@
 and strain by about 40$\%$.
+
+ \begin{figure*}[ht]
+   \begin{center}
+     \includegraphics[width=.8\textwidth]{figs/fig2.pdf}
+  \caption{Time series of the root-mean-square (RMS) of vorticity (a) and
+  lateral rate of strain (b), and horizontal divergence (c) in the LLC outputs and gridded AVISO data.}
+  \label{fig2}
+  \end{center}
+\end{figure*}
 
@@
 with submesoscale flows: smoothing the velocity fields
-with a Hanning filter with cut-off scale of 100 km significantly reduces the RMS
-vorticity and strain. The reduction in variance is dramatic ($>$80$\%$) in winter
+with a Hanning filter with cut-off scale of 100 km dramatically reduces the RMS
+vorticity and strain. The reduction in variance is about 80$\%$ in winter,
 yielding RMS vorticity and strain roughly consistent the ones diagnosed from
@@
 
-The seasonal cycle of the horizontal divergent part of the flow, however, add complexity
-to the submesoscale-turbulence-centred scenario depicted by those invertigators.
+The seasonal cycle of the horizontal divergence, however, showcases the complexity
+of the upper-ocean annual variability. If submesoscale eddies and fronts dominated
+the near-surface variability all year, then the seasonal cycle of horizontal divergence,
+vertical vorticity, and lateral strain rate would be in phase \citep[e.g.,][]{sasaki_etal2014}.
+While there is a clear wintertime peak in divergence of daily-averaged velocity
+(see dashed lines in figure \ref{fig2}a; RMS divergence $\sim$0.22 f in the 1/48$^\circ$
+simulation), the hourly fields show
+a stronger enhancement of
+lateral divergence in late Summer/early Fall (RMS divergence $\sim$0.22 f the 1/48$^\circ$
+simulation). Because the 1/48$^\circ$ simulation better resolve
+smaller-scale submesoscale flows, the
+winter peak in divergence is nearly as strong the summertime's. Of course, part of
+these submesoscale fronts and eddies evolve relatively fast and there is no clear
+temporal and spatial scale separation between those motions and inertia-gravity waves:
+daily-averaging the velocity fields efficiently suppresses the summertime horizontally
+divergent flows, but also reduces the wintertime lateral divergence by about 50$\%$.
 
- is significant: the RMS is $\delta/f > 0.1$).
-Most of these horizontal divergent flows are supra-inertial and project
-on scales smaller than 100 km: daily-averaging or smoothing the velocity fields
- (100 km cutoff) reduces the horizontal divergence by about 70$\%$.
-Interestingly, the seasonal cycle in horizontal divergence is out-of-phase with the
-seasonal cycle of vertical vorticity: the horizontal divergence is about 30$\%$ larger in late
-Summer/early Fall than it is in late Winter/early Spring. The divergence associated
-with submesoscale subinertial eddies is small and its seasonal cycle is in phase
-with that of vorticity (see green lines).  The seasonal cycle of the horizontal
-divergence is roughly in phase with the seasonal cycle of the near-surface (0-50m)
-stratification.
-
- \begin{figure*}[ht]
-   \begin{center}
-     \includegraphics[width=.9\textwidth]{figs/fig2.pdf}
-  \caption{Time series of the root-mean-square (RMS) of vorticity (a) and
-  lateral rate of strain (b), and horizontal divergence (c) in the LLC outputs and gridded AVISO data.}
-  \label{fig2}
-  \end{center}
-\end{figure*}
+\section{Seasonality in governing submesoscale dynamics}
+The overall results of figure \ref{fig2}c strongly suggest that the
+summertime submesoscale surface variability stems from different dynamics than
+then wintertime variability. To better understand these differences, we calculate
+joint probability distributions...
 
@@
 \end{figure*}
-
-\section{Seasonality in governing submesoscale dynamics}
 
@@
 
-\section{Projection on time scales}
-To better understand the high-frequency/high-wavenumber flows  we calculate frequency
-spectra of surface KE (figure \ref{fig4}). To obtain statistically meaningful results,
-we average spectral estimates along a section at 165$^\circ$E.  There is about 50
-independent spectral realizations.
-
-The surface KE spectra of subinertial flows very roughly follow a $-2$ power-law.
-The high-frequency flows are split near-inertial flows, tidal signals and their
-higher harmonics. Hence these high-frequency flows are dominated by inertia-gravity
-waves. (Note that the inertial peak nearly merges with the
-diurnal tide peak.) There is statistically significant seasonality. The seasonal
-cycle of the sub-inertial motions (at least for periods between 5-20 days)
-is out-of-phase with the seasonal cycle of supra-inertial flows: the former flows
-are more energetic in April and the latter in October.
-
-Most of these high-frequency inertia-gravity waves project onto scales smaller
-than 100 km. The surface KE spectra of 100-km-smoothed velocity field is significantly
-suppressed in supra-inertial frequencies --- there are no statistically significant
-changes at sub-inertial frequencies (see dashed lines in figure \ref{fig4}).
+% \section{Projection on time scales}
+% To better understand the high-frequency/high-wavenumber flows  we calculate frequency
+% spectra of surface KE (figure \ref{fig4}). To obtain statistically meaningful results,
+% we average spectral estimates along a section at 165$^\circ$E.  There is about 50
+% independent spectral realizations.
+%
+% The surface KE spectra of subinertial flows very roughly follow a $-2$ power-law.
+% The high-frequency flows are split near-inertial flows, tidal signals and their
+% higher harmonics. Hence these high-frequency flows are dominated by inertia-gravity
+% waves. (Note that the inertial peak nearly merges with the
+% diurnal tide peak.) There is statistically significant seasonality. The seasonal
+% cycle of the sub-inertial motions (at least for periods between 5-20 days)
+% is out-of-phase with the seasonal cycle of supra-inertial flows: the former flows
+% are more energetic in April and the latter in October.
+%
+% Most of these high-frequency inertia-gravity waves project onto scales smaller
+% than 100 km. The surface KE spectra of 100-km-smoothed velocity field is significantly
+% suppressed in supra-inertial frequencies --- there are no statistically significant
+% changes at sub-inertial frequencies (see dashed lines in figure \ref{fig4}).
 

### 2016-06-09 writeup/rocha_etal.tex
Small fix in sec3
Significant improvement in the text

@@
 
-A key aspect of the LLC simulations is that there were forced by
+A key aspect of the LLC simulations is that they were forced by
 the 16 most significant tidal components.
@@
 \vspace{-.125cm}
-\includegraphics[width=.9\textwidth]{figs/fig1_2.pdf}
+\includegraphics[width=.75\textwidth]{figs/fig1_2.pdf}
  \caption{(a) The study region with the subregion where the LLC outputs are
@@
    \begin{center}
-     \includegraphics[width=.8\textwidth]{figs/fig2.pdf}
+     \includegraphics[width=.75\textwidth]{figs/fig2.pdf}
   \caption{Time series of the root-mean-square (RMS) of vorticity (a) and
@@
 divergent flows, but also reduces the wintertime lateral divergence by about 50$\%$.
+As for the vorticity and strain rate, most of the lateral divergence is associated
+with submesoscale flows: smoothing the the velocity fields with a 100km-cut-off
+suppresses more than 80$\%$ of the RMS divergence.
 
@@
 then wintertime variability. To better understand these differences, we calculate
-joint probability distributions...
+vorticity-divergence and vorticity-strain joint probability distributions (jPDF).
+
+The April vorticity-strain jPDF have a shape characteristic of submesoscale turbulence
+ (see figures \ref{fig3}f-g). The alignment of vorticity and strain $\alpha \sim \zeta$
+ with strong positive skewness are fingerprints of submesoscale fronts
+ \citep{shcherbina_etal2013,mcwilliams2016}. The shape of the vorticity-strain
+ jPDF is similar for hourly, daily-averaged, and residual velocities, although
+ the vorticity skewness reduces 1.4 to 1.13 from hourly to daily-averaged.
+ The April vorticity-divergence jPDFs, however, have a egg-like shape that which again highlight
+ the strong skewness in vorticity. Thus wintertime submesoscale surface velocity
+ is strongly dominated by submesoscale turbulence --- this is true even for
+ superinertial submesoscale currents.
+
+The October vorticity-strain jPDF show much weaker skewness (the vorticity skewness
+is 0.68 and 0.67 for hourly and daily-averaged velocities). The shape of the
+ vorticity-strain jPDF  appears to be a combination of two half-ellipses centered
+ about $\zeta=0$, one with a 45$^\circ$ slope (characteristic of submesocale fronts
+ that persist in Summer) and one with a very steep slope. The vorticity-divergence
+ jPDF is significantly different than the April one. In particular, the jPDF
+ for daily-averaged currents is still egg-shaped but hourly and residual are
+ ellipses about $\zeta=0$, likely associated with inertia-gravity waves.
+
+ Time series of vorticity and divergence PDFs (supplemental material) suggest
+ a strong oscillation between these two regimes. In late Winter/early Spring
+ the vorticity is strongly positively skewed, whereas the divergence is moderately
+ negatively skewed as predicted by frontogenesis \citep{capet_etal2008a,mcwilliams2016}
+. In late Summer/early Fall, the divergence is more stronger, but PDFs are much less skewed,
+consistent with linear inertia-gravity waves.
+
+\section{Projection onto horizontal scales}
+
+To better understand the projection of these flows onto different horizontal
+scales, we calculate wavenumber spectra of kinetic energy and sea-surface height
+(SSH) variance.  Before calculating the spectra, time-mean and and spatial linear
+trend were removed, and the ensuing fields were multiplied by a Hanning spectral ``window''.
+We discuss only spectra for the 1/48$^\circ$, which simply extends the 1/24$^\circ$
+towards smaller scales (supplemental material).
+
+Figure \ref{fig3}a depicts the horizontal wavenumber spectra of surface KE in
+April and October. At scales larger than 20km, the spectra based on hourly
+velocity snapshots are nearly indistinguishable within 95$\%$ confidence level
+(compare solid lines in figure \ref{fig3}a). Consistent with the results of
+\citet{rocha_etal2016} who analyzed a 5-month output of the LLC4320 simulation in Drake Passage,
+there is significant high-frequency variabilty at submesoscales. Daily-averaging
+the velocity field suppresses spatial variability at scales smaller than about 250
+km both in April and October (compare solid lines against dashed lines in figure
+\ref{fig3}a). But this suppression is more dramatic in October, when the horizontally
+divergent flows peak. At scales
+smaller than 100 km, 39$\%$ of the surface KE in April is accounted for by super-inertial
+flows as opposed to 79$\%$ in October. The seasonality of subinertial submesoscale flows
+is much more dramatic than the seasonal cycle of the total flow. This is consistent
+with the results of \citet{sasaki_etal2014}, which are
+based on daily-averaged velocity fields of a different model without tidal forcing
+(P. Klein, personal communication).
+
+These high-frequency submesoscale flows significantly project on the sea-surface.
+In October, when the inertia-gravity waves are stronger near the surface and submesoscale
+turbulence is weak, there is a dramatic difference between the spectra
+based on hourly and daily-averaged SSH: at scales smaller than 100 km, the spectra
+of hourly SSH roughly follows a $-2$ power-law, whereas the spectra of daily-averaged
+SSH roughly follows a $-5$ power-law. At scales
+smaller than 100 km, 33$\%$ of the SSH variance in April is accounted for by suprainertial
+flows as opposed to 83$\%$ in October! The out-of-phase seasonal cycle of submesoscale
+turbulence and near-surface inertia-gravity waves conspire to yield very weak seasonality
+in the spectra of KE and SSH variance based on hourly field.
+
+\section{Summary and Conclusion}
+Our main finding is that
+near-surface upper-ocean inertia-gravity waves, undergo
+a strong seasonal cycle that is out-of-phase with the seasonal cycle of
+the submesoscale turbulence. \cite{dasaro1978} showed that linear internal waves
+velocity in the mixed-layer strongly depends on the density jump at the mixed-layer
+base. In Summer, the shallow mixed-layer overlays a strong seasonal pycnocline,
+and thus the internal waves projection onto the mixed-layer may be stronger
+according to \cite{dasaro1978}'s argumets. An
+alternative explanation is that the shape baroclinic modes experience significant
+seasonal modulation, with significant summertime peak in their surface amplitude.
+If the internal wave source has weak seasonal dependence, then a strong
+seasonal cycle in the near-surface expression of internal waves is expected.
+Internal tide generation is unlikely to vary significant seasonally, and therefore
+it is plausible to expect a strong seasonality in the the near-surface expression of
+ internal tides and
+other small-scale internal waves generated through internal tide interactions.
+Near-inertial wave generation peaks in Winter, but those waves project on large
+horizontal scales. \textcolor{red}{(REFS)}.
+
+Our work adds to recent studies that presented modeling \cite{sasaki_etal2014}
+and observational \citep{callies_etal2015,buckingham_etal2016} evidence of vigorous seasonality in
+submesoscale turbulence. We suggest that the dominance at inertia-gravity waves in
+Summer is a consequence of both suppresion of submesoscale turbulence and
+enhancement of inertia-gravity waves due to restratification of
+the upper ocean. Of course, our study focuses on a single patch of ocean in the
+vicinities of the Kuroshio Extension, which may be typical of mesoscale-rich
+subtropical regions, but not
+representative of other regions such as eastern boundary currents
+and the middle of the subtropical gyre: care must be taken in overgeneralizing
+these results.  We plan to report on the geographic variability of submesoscale
+ seasonality  in a future study.
+
+%The effects of smaller-scale/higher-frequency
+%``sub-submesoscale'' flows on the submesoscale surface velocity and SSH variability
+%are presently unknown.
+
+% That the  surface velocity and SSH submesoscale variability may be
+% dominated by ageostrophic flows in
+% Summer/Fall has consequences for the interpretation of data from
+% the future SWOT and COMPIRA altimeter missions,
+% which will deliver SSH measurements at submesoscales. To the extent that
+% high-frequency flows are dominated by spatially incoherent internal tides and other
+% internal waves, it may be very difficult (if not impossible) to
+% separate SSH submesoscale variability associated with geostrophic motions
+% from high-frequency, ageostrophic flows. Thus, previous claims that
+% one will be able to easily obtain submesoscale surface
+% geostrophic velocities and monitor such seasonal cycle on global scales
+% \citep{sasaki_etal2014,qiu_etal2014} are overstated.
 
@@
   \begin{center}
-    \includegraphics[width=.9\textwidth]{figs/fig3.pdf}
+    \includegraphics[width=.75\textwidth]{figs/fig3.pdf}
  \caption{Seasonal variation of joint probability distributions: vorticity vs.
@@
 
-\section{Projection on horizontal scales}
-
-To better understand the projection of these flows onto different horizontal
-scales, we calculate wavenumber spectra of kinetic energy and sea-surface height
-(SSH) variance. For simplicity we calculate zonal wavenumber spectra at various
-latitudes. The two-year time averaged and spatial linear trend were
-removed, and the ensuing fields were Hanning spectral ``window''. Statistical significance
-is obtained by averaging over latitude and time. Using conservative temporal and
-spatial decorrelation estimate of 10 days and 250 km, we obtain roughly 50
-independent realizations of the spectra.
-
-Figure \ref{fig3}a depicts the horizontal wavenumber spectra of surface KE in
-April and October. At scales larger than 20km, the spectra based on hourly
-velocity snapshots are nearly indistinguishable within 95$\%$ confidence level
-(compare solid lines in figure \ref{fig3}a). Consistent with the results of
-\citet{rocha_etal2016} who analyzed the LLC4320 output in Drake Passage,
-there is significant high-frequency variabilty at submesoscales. Daily-averaging
-the velocity field suppresses spatial variability at scales smaller than about 250
-km both in April and October (compare solid lines against dashed lines in figure
-\ref{fig3}a). But this suppression is more dramatic in October, when the horizontally
-divergent flows peak. At scales
-smaller than 100 km, 40$\%$ of the surface KE in April is accounted for by supra-inertial
-flows as opposed to 77$\%$ in October. The seasonality of subinertial submesoscale flows
-is significant, consistent with the results of \citet{sasaki_etal2014}, which are
-based on snapshots of a model without tidal forcing.
-
-These high-frequency submesoscale flows significantly project on the sea-surface.
-In October, when the horizontally divergent flows peak and the submesoscale
-balanced flows are minimum, there is a dramatic difference between the spectra
-based on hourly and daily-averaged SSH: at scales smaller than 100 km, the spectra
-of hourly SSH roughly follows a $-2$ power-law, whereas the spectra of daily-averaged
-SSH roughly follows a $-5$ power-law. At scales
-smaller than 100 km, 44$\%$ of the surface KE in April is accounted for by suprainertial
-flows as opposed to 86$\%$ in October! The out-of-phase seasonal cycle of subinertial
-and suprainertial flows conspire to yield nearly insignificant seasonality in the spectra
-of hourly SSH.
+\begin{figure*}[ht]
+  \begin{center}
+    \includegraphics[width=.9\textwidth]{figs/fig4.pdf}
+ \caption{Surface (horizontal) KE (a) and SSH variance wavenumber spectra (b)
+ in the 1/48$^\circ$ simulation. Solid lines
+ are spectra based on hourly snapshots, dashed lines are spectra based on daily-averaged
+ fields.}
+ \label{fig4}
+ \end{center}
+\end{figure*}
 
@@
 % contrast with the results in Sasaki et al.
-
-\begin{figure*}[ht]
-  \begin{center}
-    \includegraphics[width=.9\textwidth]{figs/fig4.pdf}
- \caption{Surface (horizontal) KE (a) and SSH variance wavenumber spectra (b)
- in the 1/48$^\circ$ simulation. Solid lines
- are spectra based on hourly snapshots, dashed lines are spectra based on daily-averaged
- fields.}
- \label{fig3}
- \end{center}
-\end{figure*}
 
@@
 %\end{figure}
-
-\section{Summary and Discussion}
-Our main finding is that
-upper-ocean divergent flows, dominated by inertia-gravity waves, undergo
-a strong seasonal cycle that is out-of-phase with the seasonal cycle of
-the quasi-balanced flows. Therefore, the present results suggest significant
-seasonal modulation of the dominant upper-ocean dynamics at submesoscales
-(10-100 km): quasi-balanced turbulence driven by shallow baroclinic instabilities
-in late Winter/early Spring
-and inertia-gravity waves dominate in late Summer/early Fall.
-
-That the  surface velocity and SSH submesoscale variability may be
-dominated by ageostrophic flows in
-Summer/Fall has consequences for the interpretation of data from
-the future SWOT and COMPIRA altimeter missions,
-which will deliver SSH measurements at submesoscales. To the extent that
-high-frequency flows are dominated by spatially incoherent internal tides and other
-internal waves, it may be very difficult (if not impossible) to
-separate SSH submesoscale variability associated with geostrophic motions
-from high-frequency, ageostrophic flows. Thus, previous claims that
-one will be able to easily obtain submesoscale surface
-geostrophic velocities and monitor such seasonal cycle on global scales
-\citep{sasaki_etal2014,qiu_etal2014} are overstated.
-
-The present analysis focuses on a single patch of ocean in the
-vicinities of the Kuroshio Extension, not
-representative of other regions such as eastern boundary currents
-and the middle of the subtropical gyre: care must be taken in overgeneralizing
-these results.  The effects of smaller-scale/higher-frequency
-``sub-submesoscale'' flows on the submesoscale surface velocity and SSH variability
-are presently unknown.
 

### 2016-06-09 writeup/rocha_etal.tex
Change the text accordingly
Updates bib

@@
 \begin{center}
-\hspace{-1.25cm}\includegraphics[width=.7\textwidth]{figs/fig1_1.pdf}\\
+\hspace{-1.25cm}\includegraphics[width=.6\textwidth]{figs/fig1_1.pdf}\\
 \vspace{-.125cm}
-\includegraphics[width=.75\textwidth]{figs/fig1_2.pdf}
+\includegraphics[width=.65\textwidth]{figs/fig1_2.pdf}
  \caption{(a) The study region with the subregion where the LLC outputs are
@@
    \begin{center}
-     \includegraphics[width=.75\textwidth]{figs/fig2.pdf}
+     \includegraphics[width=.65\textwidth]{figs/fig2.pdf}
   \caption{Time series of the root-mean-square (RMS) of vorticity (a) and
@@
 and thus the internal waves projection onto the mixed-layer may be stronger
-according to \cite{dasaro1978}'s argumets. An
+according to \cite{dasaro1978}'s arguments. An
 alternative explanation is that the shape baroclinic modes experience significant
@@
 seasonal cycle in the near-surface expression of internal waves is expected.
-Internal tide generation is unlikely to vary significant seasonally, and therefore
-it is plausible to expect a strong seasonality in the the near-surface expression of
- internal tides and
+Internal tide generation is unlikely to vary significantlly seasonally \citep[e.g.,][]{alford2003},
+and therefore it is plausible to expect a strong seasonality in the the near-surface
+expression of internal tides and
 other small-scale internal waves generated through internal tide interactions.
 Near-inertial wave generation peaks in Winter, but those waves project on large
-horizontal scales. \textcolor{red}{(REFS)}.
+horizontal scales \citep[e.g, ][]{qi_etal1995}.
 
@@
   \begin{center}
-    \includegraphics[width=.75\textwidth]{figs/fig3.pdf}
+    \includegraphics[width=.6\textwidth]{figs/fig3.pdf}
  \caption{Seasonal variation of joint probability distributions: vorticity vs.
-          divergence (a through e) and vorticity vs. strain rate (f through j).
+          divergence (a through e), vorticity vs. strain rate (f through l).
+          and vorticity vs. laplacian of sea-surface height (m through r).
           Dashed lines in (a) through (e) represent $\delta = \frac{\omega}{f_0}
@@
   \begin{center}
-    \includegraphics[width=.9\textwidth]{figs/fig4.pdf}
+    \includegraphics[width=.75\textwidth]{figs/fig4.pdf}
  \caption{Surface (horizontal) KE (a) and SSH variance wavenumber spectra (b)
@@
 %\end{figure}
+
+%
+%  ACKNOWLEDGMENTS
+
+\begin{acknowledgments}
+We thank the MITgcm community and our colleagues at the NASA Advanced
+Supercomputing (NAS) Division for their awesome support.
+This research was funded by NASA and NSF.
+\end{acknowledgments}
+
+\bibliographystyle{agufull08}
+\bibliography{rocha_etal}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%  ACKNOWLEDGMENTS
-
-\begin{acknowledgments}
-We thank the MITgcm community and our colleagues at the NASA Advanced
-Supercomputing (NAS) Division for their awesome support.
-This research was funded by NASA and NSF.
-\end{acknowledgments}
 
@@
 %\begin{thebibliography}{}
-
-\bibliographystyle{agufull08}
-\bibliography{rocha_etal}
 

### 2016-06-09 writeup/rocha_etal.tex
Updates title and legend

@@
 % Shorter version of title entered in capital letters:
-\titlerunninghead{Seasonality in governing upper-ocean dynamics at submesoscales}
+\titlerunninghead{Seasonality in governing dynamics at submesoscales}
 
@@
           \zeta$ where $\omega = 2\pi/$(12 h, 3 h) and $f_0$ is the inertial frequency
-          in the middle of the domain (32.5$^\circ$ N). Dashed lines in (f) through (j)
-          represent pure strain flow $\alpha = \zeta$.}
+          in the middle of the domain (32.5$^\circ$ N). Dashed lines in (f) through (l)
+          represent pure strain flow $\alpha = \zeta$. Dashed lines in (m) through
+          (o) represent geostrophic flow $\zeta = \frac{g}{f}\nabla_h^2 h$.}
  \label{fig3}

### 2016-06-09 writeup/rocha_etal.tex
Fixes title
Fixes legend

@@
 
-\title{Seasonality in governing upper-ocean dynamics at submesoscales}
+\title{Seasonality in governing dynamics at submesoscales}
 %
@@
           represent pure strain flow $\alpha = \zeta$. Dashed lines in (m) through
-          (o) represent geostrophic flow $\zeta = \frac{g}{f}\nabla_h^2 h$.}
+          (o) represent geostrophic flow $\zeta = \frac{g}{f}\nabla_h^2 \eta$.}
  \label{fig3}

### 2016-06-10 writeup/rocha_etal.tex
Makes various small edits + adds appendix

@@
     prevails in late Winter/early Spring. Inertia-gravity waves, on the other hand, are enhanced near
-    the surface when the upper ocean in strongly stratified. As a consequence of
-    reduction of submesoscale turbulence and enhancement of intertia-gravity waves,
-    the latter strongly dominate the submesoscale surface kinetic energy and
-    sea-surface height variance in late Summer/early Fall.
+    the surface when the upper ocean in strongly stratified. Thus the summertime
+    re-stratification weakens submesoscale turbulence and enhances intertia-gravity
+     waves near the surface. In the Kuroshio Extension region,
+    inertia-gravity waves strongly dominate the surface kinetic energy and
+    sea-surface height variance in late Summer/early Fall at scales as large as 80 km!
 \end{abstract}
@@
 a strong seasonality near the surface in the Kuroshio Extension region.
-The seasonal cycle of inertia-gravity waves is out-of-phase
+Interestingly, the seasonal cycle of inertia-gravity waves is out-of-phase
 with the seasonal cycle of submesoscale turbulence. Consistent with previous studies,
@@
 a hierarchy of global
-forward numerical solutions on a LLC grid   \citep{forget_etal2015} that were
+forward Massachusetts Institute of Technlogy general circulation model (MITgcm)
+numerical solutions on a LLC grid   \citep{forget_etal2015} that were
 spun up from a ECCO2 adjoint-method state estimate. Both simulations were forced by
@@
 Pacific, in the vicinities of the Kuroshio
-Extension, where previous studies have suggested strong submesoscale seasonality.
+Extension, where previous studies have suggested strong submesoscale seasonality
+\citep{sasaki_etal2014,qiu_etal2014}.
 We analyze a sub-domain of the LLC4320 and LLC2160 simulations of about 2000 km$^2$
@@
 values as large as $4f$, where $f$ is
-the local planetary vorticity, and root-mean-square (RMS) of about 0.4. In early
+the local planetary vorticity, and root-mean-square (RMS) of about 0.4 f. In early
 Fall, the situation is the opposite: the vertical vorticity is relatively coarse-grained,
@@
 vorticity and strain. The reduction in variance is about 80$\%$ in winter,
-yielding RMS vorticity and strain roughly consistent the ones diagnosed from
+yielding RMS vorticity and strain roughly consistent with the diagnostics from
 AVISO gridded geostrophic velocities (compare blue lines with black lines in figures \ref{fig2}a-b).
@@
 these submesoscale fronts and eddies evolve relatively fast and there is no clear
-temporal and spatial scale separation between those motions and inertia-gravity waves:
+temporal and spatial scale separation between those motions and inertia-gravity waves
+\citep{mcwilliams2016}:
 daily-averaging the velocity fields efficiently suppresses the summertime horizontally
@@
 then wintertime variability. To better understand these differences, we calculate
-vorticity-divergence and vorticity-strain joint probability distributions (jPDF).
+joint probability distributions (jPDF) of
+vorticity-divergence, vorticity-strain, and vorticity-laplacian of sea-surface height.
 
@@
  is strongly dominated by submesoscale turbulence --- this is true even for
- superinertial submesoscale currents.
+ superinertial submesoscale currents. The hourly velocity fields are significantly
+ageostrophic as depicted by the jPDF of vorticity-laplacian of SSH
+(figure \ref{fig3}o); it is only the daily-averaged velocity that is approximately
+in geostrophic balance (figure \ref{fig3}l). While there is certainly much to be
+learned using quasi-geostrophic models to study submesoscale turbulence, claims that
+the latter is dominantly in geostrophic balance \cite{sasaki_etal2014,callies_etal2016}
+is overstated; it is only the daily-averaged fields that are to a large extent
+in  geostrophic balance (the jPDF of daily-averaged vorticity vs. laplacian of SSH
+is a ellipsis with large eccentricity with main axis tilted by 45$^\circ$; figure \ref{fig3}n).
 
@@
  ellipses about $\zeta=0$, likely associated with inertia-gravity waves.
+That the submesoscale dynamics in October is mainly unbalanced is clearly depicted
+in the shape of jPDF of vorticity-laplacian of SSH for hourly and
+residual fields
+(figure \ref{fig3}p and \ref{fig3}r), which is an ellipsis aligned in the vertical axis.
+Of course daily-averaging the model fields filter out the ageostrophic flows and
+the daily-averaged flow is essentially geostrophic.
 
@@
 scales, we calculate wavenumber spectra of kinetic energy and sea-surface height
-(SSH) variance.  Before calculating the spectra, time-mean and and spatial linear
-trend were removed, and the ensuing fields were multiplied by a Hanning spectral ``window''.
+(SSH) variance.  Before calculating the spectra, time-mean and spatial linear
+trends were removed, and the ensuing fields were multiplied by a Hanning spectral ``window''.
 We discuss only spectra for the 1/48$^\circ$, which simply extends the 1/24$^\circ$
@@
 flows as opposed to 79$\%$ in October. The seasonality of subinertial submesoscale flows
-is much more dramatic than the seasonal cycle of the total flow. This is consistent
+is more dramatic than the seasonal cycle of the total flow. This is consistent
 with the results of \citet{sasaki_etal2014}, which are
@@
 These high-frequency submesoscale flows significantly project on the sea-surface.
-In October, when the inertia-gravity waves are stronger near the surface and submesoscale
-turbulence is weak, there is a dramatic difference between the spectra
+In October, when the inertia-gravity waves are strongest the surface and submesoscale
+turbulence is weakest, there is a dramatic difference between the spectra
 based on hourly and daily-averaged SSH: at scales smaller than 100 km, the spectra
 of hourly SSH roughly follows a $-2$ power-law, whereas the spectra of daily-averaged
-SSH roughly follows a $-5$ power-law. At scales
-smaller than 100 km, 33$\%$ of the SSH variance in April is accounted for by suprainertial
-flows as opposed to 83$\%$ in October! The out-of-phase seasonal cycle of submesoscale
+SSH roughly follows a $-5$ power-law. At these scales, 33$\%$ of the SSH variance in April is accounted for by suprainertial
+flows as opposed to 83$\%$ in October! Curiously, the out-of-phase seasonal cycle of submesoscale
 turbulence and near-surface inertia-gravity waves conspire to yield very weak seasonality
@@
 Our main finding is that
-near-surface upper-ocean inertia-gravity waves, undergo
+near-surface upper-ocean inertia-gravity waves undergo
 a strong seasonal cycle that is out-of-phase with the seasonal cycle of
@@
 velocity in the mixed-layer strongly depends on the density jump at the mixed-layer
-base. In Summer, the shallow mixed-layer overlays a strong seasonal pycnocline,
+base, with largest mixed-layer velocities when the jump is strongest.
+In Summer, the shallow mixed-layer overlays a strong seasonal pycnocline,
 and thus the internal waves projection onto the mixed-layer may be stronger
@@
 alternative explanation is that the shape baroclinic modes experience significant
-seasonal modulation, with significant summertime peak in their surface amplitude.
+seasonal modulation, with significant summertime peak in their surface amplitude
+ (supplemental material).
 If the internal wave source has weak seasonal dependence, then a strong
 seasonal cycle in the near-surface expression of internal waves is expected.
-Internal tide generation is unlikely to vary significantlly seasonally \citep[e.g.,][]{alford2003},
+Internal tide generation does not show a seasonal modulation \citep[e.g.,][]{alford2003},
 and therefore it is plausible to expect a strong seasonality in the the near-surface
@@
 %  this should go in a supplementary material
-\clearpage
-
-\appendix
-
-\section{Supplemental material 1: Spectral errors}
-
-\section{Supplemental material 2: On the LLC simulations}
-
-The time-step was 45s for
-the LLC2160 simulation and 25s for the LLC4320 simulation. Details of the LLC simulations
-are provided in the supplemental material.
-
- LLC hierarchy setup
-and control files are available online
+%\clearpage
+
+%\section{Supplemental material 1: Spectral errors}
+
+\newpage
+\section*{Supplemental material 1: On the LLC simulations}
+
+Table \ref{tab:llc} show details of the LLC spin-up hierarchy.
+Both simulation analyzed in the present paper, LLC2160 and LLC4320,
+were forced by surface fluxes are from the 0.14$^\circ$ European Centre for
+Medium-range Weather Forecasting (ECMWF) atmospheric operational model analysis,
+ starting 2011 humidity and the 16 most significant  tidal components.
+The LLC control files are available online
 (at \texttt{http://mitgcm.org/viewvc/MITgcm/MITgcm$\_$contrib/
-llc$\_$hires}). For the LLC hierarchy, the sea-surface height is calculated
+llc$\_$hires}).
+
+For the LLC hierarchy, the sea-surface height is calculated
 using a linearized equation. While the nominal horizontal resolution  of the
-LLC2160 simulation is 1/24$^\circ$,
- horizontal wavenumber spectra suggests an effective resolution of about 20 km.
- (see figure \ref{fig3}).
-
-\subsection{Seasonality in the upper-ocean stratification}
+LLC4320 simulation is 1/48$^\circ$,
+horizontal wavenumber spectra suggests an effective resolution of about 8 km.
+(see figure \ref{fig3}), while wavenumber spectra suggests an effective resolution
+of 20 km for  LLC2160 simulation.
+
+\begin{table*}[ht]
+\label{tab:llc}
+\caption{\small The MITgcm Latitute-Longitude Cap (LLC) spin-up hierarchy. This study will analyze
+         the output of LLC2160 and LLC4320. Model fields are available hourly.
+         Adapted from \cite{rocha_etal2016}.}
+\begin{center}
+\begin{tabular}{ | c | c | c | c | c |}
+\hline
+Simulation & Resolution & Time-step & Period  & Tides \\ \hline
+ECCO2 Adjoint & $1/6^{^\circ}$ & 1200 s & January 2009 - December 2011 & No\\
+LLC 1080   & $1/12^{^\circ}$ & 90 s & January 2010 - July 2012 & No\\
+LLC 2160   &  $1/24^{^\circ}$ & 45 s  & January 2011 - April 2013 & Yes\\
+LLC 4320   &  $1/48^{^\circ}$ & 25 s & September 2011 - September 2012 & Yes \\
+\hline
+\end{tabular}
+\end{center}
+\end{table*}
+
+\subsection*{Spectra}
+Kinetic energy and sea-surface height variance wavenumber spectra for the LLC2160
+and LLC4320 simulations are shown in figure \ref{figS2_3}. As expected, and
+consistent with the submesoscale regional study by \cite{capet_etal2008a}, the
+higher-resolution simulation (LLC4320) essentially extends the spectra of the
+lower-resolution simulation (LLC2160) towards smaller scale, demonstrating that
+the higher resolution is resolving a wider array of scales. Between 20 and 100 km
+there is linewidth agreement between the spectra of the two simulations.
+
+\begin{figure*}[ht]
+   \begin{center}
+     \includegraphics[width=1.\textwidth]{figs/fig_s2_3.pdf}
+  \caption{A comparison between spectra of the two LLC simulations 1/24$^\circ$
+  (green) and 1/48$^\circ$ (purple). The statistical errorbars represent conservative
+  significance levels considering that there are only four independent relations
+  for the monthly spectra.}
+  \label{figS2_3}
+  \end{center}
+\end{figure*}
+
+\subsection*{Seasonality in statistics: probability density functions}
+\begin{figure*}[ht]
+   \begin{center}
+     \includegraphics[width=1.\textwidth]{figs/fig_s2_2.pdf}
+  \caption{Time dependence of probability density functions of vertical vorticity (a, b)
+  and horizontal divergence (c, d). Monthly averages in April and October are
+  also shown (e, f).}
+  \label{figS2_2}
+  \end{center}
+\end{figure*}
+
+\subsection*{Seasonality in the upper-ocean stratification}
+
+Figure \ref{figS2_1} shows a comparison between potential density sections
+at 165$^\circ$E from
+the Roemmich-Gilson Argo Climatology \citep[updated from][]{roemmich_gilson2009}.
+and the LLC2160 and LLC4320 simulations. Of course, the Argo climatology is smoother
+owing to lower resolution and long-term averaging. Nonetheless, simulations reproduce well the strong
+seasonality in the stratification. The upper ocean is weakly stratified in late Winter/early Fall
+and restratifies in Summer. Argo climatology and LLC simulations mixed-layer depths are consistent
+at all seasons (not shown).
 
@@
 
-\subsection{Spectra}
-
-\begin{figure*}[ht]
-   \begin{center}
-     \includegraphics[width=1.\textwidth]{figs/fig_s2_3.pdf}
-  \caption{A comparison between spectra of the two LLC simulations 1/24$^\circ$
-  (green) and 1/48$^\circ$ (purple). The statistical errorbars represent conservative
-  significance levels considering that there are only four independent relations
-  for the monthly spectra.}
-  \label{figS2_3}
-  \end{center}
-\end{figure*}
-
-\subsection{Seasonality in statistics: probability density functions}
-\begin{figure*}[ht]
-   \begin{center}
-     \includegraphics[width=1.\textwidth]{figs/fig_s2_2.pdf}
-  \caption{Time dependence of probability density functions of vertical vorticity (a, b)
-  and horizontal divergence (c, d). Monthly averages in April and October are
-  also shown (e, f).}
-  \label{figS2_2}
-  \end{center}
-\end{figure*}
-
-\subsection{How well does the model captures high-frequency modes?}
+\subsection*{How well does the model captures high-frequency modes?}
 To assess how well the LLC2160 represents the high-frequency variability,
@@
 
-\section{Supplemental material 3: Seasonality in pressure modes near-surface amplitude}
+\section*{Supplemental material 3: Seasonality in pressure modes near-surface amplitude}
 

### 2016-06-10 writeup/rocha_etal.tex
Update figure 1

@@
  \caption{(a) The study region with the subregion where the LLC outputs are
-          analyzed. LLC 4320 (1/48$^\circ$) snapshots of surface vorticity (b and c) and a transect
+          analyzed. Colors represent the topography and black lines are absolute
+          dynamic topography contours every 0.1 m. LLC 4320 (1/48$^\circ$) snapshots of surface vorticity (b and c) and a transect
           of potential density at 165$^\circ$E (d and e). The snapshots were
@@
 This research was funded by NASA and NSF.
+The altimeter products were produced by Ssalto/Duacs
+and distributed by Aviso, with support from Cnes (\texttt{http://www.aviso.altimetry.fr/duacs/})
 \end{acknowledgments}
@@
 
-\subsection*{How well does the model captures high-frequency modes?}
-To assess how well the LLC2160 represents the high-frequency variability,
+\subsection*{How well does the LLC simulations captures high-frequency modes?}
+To assess how well the LLC simulations represents the high-frequency variability,
 we compare the model near-surface velocity field against two available
@@
 
-Figure \ref{figA1} shows KE frequency spectra for the model and mooring data.
-At the uppermost common depth (20 m), very low frequencies (periods > 10 days)
+Figure \ref{figA1} shows KE frequency spectra for the LLC simulations and mooring data.
+At the uppermost common depth (20 m), very low frequencies (periods $>$ 10 days)
 and the inertial, diurnal, and semi-diurnal frequencies agree well --- the KESS
-data are more energetic at intermedia frequencies. At 40 m, the model and observations
+data are more energetic at intermedia frequencies. At 40 m, the LLC simulations and observations
 have consistent spectra at periods larger than about 6 h. The data from moorings
-are more energetic at higher frequencies and tidal harmonics are not observed,
+are more energetic at higher frequencies,
 except for the first two harmonics that are apparent in the KESS data at 40 m.
-Three reasons for the lack of tidal harmonics in the mooring data are: 1) The high-
+There are at least three reasons for the slightly more energetic super-inertial
+variability in the observations: 1) The high-
 frequency variability is dominated by high-frequency noise due to mooring vertical
-excursions; 2) The model does not have enough resolution to resolve the full spectrum
-of ultraviolet catastrophic wave-wave interactions; 3) The model is not forced at
-frequencies higher that 6h.
+excursions; 2) The model does not have enough resolution to resolve the full internal
+wave spectrum; 3) The model is not forced at frequencies higher that 6h.
 
@@
 \section*{Supplemental material 3: Seasonality in pressure modes near-surface amplitude}
+
+Figure \ref{figS3}a shows the domain-averaged potential density from the WOA 2013
+climatology in April and October. Clearly, the the changes to the stratification
+are dramatic, with the formation of a seasonal pycnocline in Summer. The strong
+seasonality of upper-ocean stratification yields a strong seasonality in the
+near-surface shape and amplitude of baroclinic pressure modes (\ref{figS3}b).
+In particular, the surface amplitude is much large in Summer. Thus the projection
+of baroclinic tides on the surface may vary significantly seasonally.
+
+\textcolor{red}{Perhaps include a model with a slab over a slowly varying
+pycnocline. The model is simple, but the solution is a bit messy and may
+scare GRL readers...}.
 

### 2016-06-10 writeup/rocha_etal.tex
Adds small technical detail

@@
 (SSH) variance.  Before calculating the spectra, time-mean and spatial linear
-trends were removed, and the ensuing fields were multiplied by a Hanning spectral ``window''.
-We discuss only spectra for the 1/48$^\circ$, which simply extends the 1/24$^\circ$
+trends were removed, and the ensuing fields were multiplied by a two-dimensional Hanning spectral ``window''.
+The two-dimensional spectra were averaged azimuthly \citep[e.g., ][]{rocha_etal2016} .We discuss only spectra for the 1/48$^\circ$, which simply extends the 1/24$^\circ$
 towards smaller scales (supplemental material).

### 2016-06-10 writeup/rocha_etal.tex
Bills small editorial fixes

@@
 whereas quasi-linear inertia-gravity waves prevail in late Summer/early Fall.
-Because submesoscale turbulence is severely diminished in late Summer/early Fall,
+Because submesoscale turbulence is weakest in late Summer/early Fall,
 the present results suggest that inertia-gravity waves account for most of the
@@
 The stratification and its seasonality is well-captured by both LLC simulations:
-the skillful comparisons of model stratification to Argo climatology lends confidence
-to the present analysis (supplemental material).
-
+the skillful comparisons of model stratification to Argo climatology is necessary
+for the analysis in this paper (supplemental material).
 \begin{figure*}[ht]
@@
 These diagnostics highlight the submesoscale structures in the flow
-and their analysis hint on the governing dynamics in the upper ocean
 \citep[e.g.,][]{capet_etal2008a,shcherbina_etal2013}.
@@
 \section{Seasonality in governing submesoscale dynamics}
-The overall results of figure \ref{fig2}c strongly suggest that the
-summertime submesoscale surface variability stems from different dynamics than
-then wintertime variability. To better understand these differences, we calculate
+The overall results of figure \ref{fig2}c shows that summertime submesoscale surface
+ variability stems from different dynamics than wintertime variability.
+ To better understand these differences, we calculate
 joint probability distributions (jPDF) of
@@
 learned using quasi-geostrophic models to study submesoscale turbulence, claims that
-the latter is dominantly in geostrophic balance \cite{sasaki_etal2014,callies_etal2016}
-is overstated; it is only the daily-averaged fields that are to a large extent
+the latter is dominantly in geostrophic balance
+are overstated \citep{sasaki_etal2014,callies_etal2016}; it is only the daily-averaged fields that are to a large extent
 in  geostrophic balance (the jPDF of daily-averaged vorticity vs. laplacian of SSH
-is a ellipsis with large eccentricity with main axis tilted by 45$^\circ$; figure \ref{fig3}n).
+is a ellipse with large eccentricity with main axis tilted by 45$^\circ$; figure \ref{fig3}n).
 
@@
 residual fields
-(figure \ref{fig3}p and \ref{fig3}r), which is an ellipsis aligned in the vertical axis.
+(figure \ref{fig3}p and \ref{fig3}r), which is an ellipse aligned in the vertical axis.
 Of course daily-averaging the model fields filter out the ageostrophic flows and
@@
 according to \cite{dasaro1978}'s arguments. An
-alternative explanation is that the shape baroclinic modes experience significant
-seasonal modulation, with significant summertime peak in their surface amplitude
+alternative explanation is that the shape of the baroclinic modes significantly
+change seasonally. In particular the surface amplitude of the baroclinic modes are
+largest when the upper ocean is strongly stratified in late Summer/ early Fall
  (supplemental material).
@@
 Internal tide generation does not show a seasonal modulation \citep[e.g.,][]{alford2003},
-and therefore it is plausible to expect a strong seasonality in the the near-surface
+and therefore it is plausible to expect a strong seasonality in the near-surface
 expression of internal tides and
@@
 submesoscale turbulence. We suggest that the dominance at inertia-gravity waves in
-Summer is a consequence of both suppresion of submesoscale turbulence and
+Summer is a consequence of both  of submesoscale turbulence and
 enhancement of inertia-gravity waves due to restratification of

### 2016-06-11 writeup/rocha_etal.tex
Fixes typos.

@@
                   sea-surface height variability in late Winter/early Spring.
-            \item Inertia-gravity waves dominate the the horizontal velocity and
+            \item Inertia-gravity waves dominate the horizontal velocity and
                    sea-surface height variability in late Summer/early Fall.
@@
 As for the vorticity and strain rate, most of the lateral divergence is associated
-with submesoscale flows: smoothing the the velocity fields with a 100km-cut-off
+with submesoscale flows: smoothing the velocity fields with a 100km-cut-off
 suppresses more than 80$\%$ of the RMS divergence.

### 2016-06-11 writeup/rocha_etal.tex
Fixes yet another typo

@@
 Recent interest in upper-ocean dynamics has focused on the strong seasonal
-cycle of shallow baroclinic instabilities and their role in submesoscale roughly 1-100 km)
+cycle of shallow baroclinic instabilities and their role in submesoscale (roughly 1-100 km)
 turbulence and mesoscale modulation \citep{sasaki_etal2014,qiu_etal2014,

### 2016-06-15 writeup/rocha_etal.tex
Adds new ref

@@
 and observational \citep{callies_etal2015,buckingham_etal2016} evidence of vigorous seasonality in
-submesoscale turbulence. We suggest that the dominance at inertia-gravity waves in
-Summer is a consequence of both  of submesoscale turbulence and
+submesoscale turbulence. We conjecture that the dominance at inertia-gravity waves in
+Summer \citep{callies_etal2015} is a consequence of both  of submesoscale turbulence and
 enhancement of inertia-gravity waves due to restratification of
@@
   \caption{The seasonal variability of the WOA 2013 stratification averaged  over
-  the domain and the associated gravest three pressure modes. Only the upper 450 m is shown.}
+  the domain and the associated three gravest pressure modes. Only the upper 450 m is shown.}
   \label{figS3}

### 2016-06-27 writeup/rocha_etal.tex
Fixes weird sentence
Sarahz edits.

@@
 % Shorter version of title entered in capital letters:
-\titlerunninghead{Seasonality in governing dynamics at submesoscales}
+\titlerunninghead{Seasonality at submesoscales in the Kuroshio Extension}
 
@@
 
-\title{Seasonality in governing dynamics at submesoscales}
+\title{Seasonality in governing dynamics at submesoscales in the Kuroshio Extension}
 %
@@
     strong modulation in the governing near-surface dynamics at submesoscales
-    (roughly 10-100 km). Consistent with recent studies, deep late Winter mixed layers
-    are prone to baroclinic instabilities and submesoscale turbulence
-    prevails in late Winter/early Spring. Inertia-gravity waves, on the other hand, are enhanced near
-    the surface when the upper ocean in strongly stratified. Thus the summertime
+    (roughly 10-100 km) in the Kuroshio Extension. Consistent with recent studies, deep late Winter mixed layers
+    are prone to baroclinic instabilities, and submesoscale turbulence
+    prevails in late Winter/early Spring. Inertia-gravity waves are enhanced near
+    the surface when the upper ocean is strongly stratified. Thus the summertime
     re-stratification weakens submesoscale turbulence and enhances intertia-gravity
@@
     inertia-gravity waves strongly dominate the surface kinetic energy and
-    sea-surface height variance in late Summer/early Fall at scales as large as 80 km!
+    sea-surface height variance in late Summer/early Fall at scales as large as 80 km.
 \end{abstract}
@@
 
-Using the output of two new $1/24^\circ$ and $1/48^\circ$ global
+Using the output of two $1/24^\circ$ and $1/48^\circ$ global
 numerical simulations with embedded tides, we show that inertia-gravity waves undergo
@@
 We use the output of two latitude-longitude polar cap (LLC)
-MITgcm realstic numerical simulations. The outputs
- analyzed here, LLC2160 and LLC4320,
-are the 1/24$^\circ$ and 1/48$^\circ$  --- 90-vertical-levels simulations from
-a hierarchy of global
-forward Massachusetts Institute of Technlogy general circulation model (MITgcm)
+MITgcm realistic numerical simulations. The outputs
+analyzed here, LLC2160 (1/24$^\circ$) and LLC4320 (1/48$^\circ$)  have
+90-vertical-levels simulations from a hierarchy of global
+forward Massachusetts Institute of Technology general circulation model (MITgcm)
 numerical solutions on a LLC grid   \citep{forget_etal2015} that were
@@
 the LLC2160 simulation, spanning one year from September 2011 to September 2012.
-The LLC4320 simulation is an extension of the then 5-month long output used by
+The LLC4320 simulation is an extension of the 5-month long output used by
 \citet{rocha_etal2016}.
@@
 in this mesoscale-rich subtropical region undergoes a strong seasonal cycle: wintertime
-enhance small-scale turbulence de-stratify the upper ocean, yielding mixed layers
+enhancement of small-scale turbulence de-stratifies the upper ocean, yielding mixed layers
 as deep as 300m. In late Spring/early Summer enhanced solar radiation and mixed-layer
 instabilities re-stratify the upper ocean, yielding mixed layers as shallow as 40m.
-The stratification and its seasonality is well-captured by both LLC simulations:
-the skillful comparisons of model stratification to Argo climatology is necessary
-for the analysis in this paper (supplemental material).
+Very importantly, the stratification and its seasonality are well-captured by both LLC simulations:
+a comparison with Argo climatology shows that both simulations skillfully represent the Kuroshio
+Extension stratification and its seasonal variability  (supplemental material).
+
 \begin{figure*}[ht]
@@
  \caption{(a) The study region with the subregion where the LLC outputs are
-          analyzed. Colors represent the topography and black lines are absolute
-          dynamic topography contours every 0.1 m. LLC 4320 (1/48$^\circ$) snapshots of surface vorticity (b and c) and a transect
+          analyzed. Colors represent the topography and white lines are contours of absolute
+          dynamic topography every 0.1 m. LLC 4320 (1/48$^\circ$) snapshots of surface vorticity (b and c) and a transect
           of potential density at 165$^\circ$E (d and e). The snapshots were
@@
 values as large as $4f$, where $f$ is
-the local planetary vorticity, and root-mean-square (RMS) of about 0.4 f. In early
+the local planetary vorticity, and root-mean-square (RMS) of about $0.4f$. In early
 Fall, the situation is the opposite: the vertical vorticity is relatively coarse-grained,
-and its local maximum and RMS are no larger than $0.5f$ and $0.2f$, respectively.
-Indeed, the seasonal cycle of vorticity and rate of strain is very strong (figure~\ref{fig2}):
+its local maximum and RMS is smaller than $0.5f$.
+Indeed, the seasonal cycle of vorticity and rate of strain are strong (figure~\ref{fig2}):
 in both simulations, the RMS vorticity and strain are about twice as large in
@@
 
-The bulk of vertical vorticity and lateral strain rate is associated with subinertial flows
+The bulk of vertical vorticity and lateral strain rate are associated with subinertial flows
 ($T_f = 2\pi/f_0\approx 23.5$h, where $f_0$ is the inertial frequency at
-the mean latitude): daily-averaging the velocity fields reduces the RMS
+the mean latitude): daily-averaging the velocity fields suppresses superinertial
+motions and reduces the RMS
 vorticity by less than 40$\%$ and the RMS strain by less than 10$\%$; the seasonal
@@
 While there is a clear wintertime peak in divergence of daily-averaged velocity
-(see dashed lines in figure \ref{fig2}a; RMS divergence $\sim$0.22 f in the 1/48$^\circ$
+(see dashed lines in figure \ref{fig2}c; RMS divergence $\sim0.22 f$ in the 1/48$^\circ$
 simulation), the hourly fields show
 a stronger enhancement of
-lateral divergence in late Summer/early Fall (RMS divergence $\sim$0.22 f the 1/48$^\circ$
-simulation). Because the 1/48$^\circ$ simulation better resolve
-smaller-scale submesoscale flows, the
-winter peak in divergence is nearly as strong the summertime's. Of course, part of
-these submesoscale fronts and eddies evolve relatively fast and there is no clear
+lateral divergence in late Summer/early Fall (RMS divergence $\sim0.22 f$ the 1/48$^\circ$
+simulation). Because the 1/48$^\circ$ simulation better resolves
+smaller-scale submesoscale flows, a secondary RMS divergence peak in winter  is
+nearly as strong as the summertime's. Submesoscale fronts and eddies evolve
+relatively fast and there is no clear
 temporal and spatial scale separation between those motions and inertia-gravity waves
@@
 divergent flows, but also reduces the wintertime lateral divergence by about 50$\%$.
-As for the vorticity and strain rate, most of the lateral divergence is associated
+The vorticity and strain rate (Figure \ref{fig2}a-b) show that most of the lateral divergence is associated
 with submesoscale flows: smoothing the velocity fields with a 100km-cut-off
@@
 joint probability distributions (jPDF) of
-vorticity-divergence, vorticity-strain, and vorticity-laplacian of sea-surface height.
-
-The April vorticity-strain jPDF have a shape characteristic of submesoscale turbulence
+vorticity-divergence, vorticity-strain, and vorticity-Laplacian of sea-surface height.
+
+The April vorticity-strain jPDF has a shape characteristic of submesoscale turbulence
  (see figures \ref{fig3}f-g). The alignment of vorticity and strain $\alpha \sim \zeta$
@@
  the vorticity skewness reduces 1.4 to 1.13 from hourly to daily-averaged.
- The April vorticity-divergence jPDFs, however, have a egg-like shape that which again highlight
+ The April vorticity-divergence jPDFs, however, egg-shaped which again highlights
  the strong skewness in vorticity. Thus wintertime submesoscale surface velocity
@@
  superinertial submesoscale currents. The hourly velocity fields are significantly
-ageostrophic as depicted by the jPDF of vorticity-laplacian of SSH
+ageostrophic as depicted by the jPDF of vorticity-Laplacian of SSH
 (figure \ref{fig3}o); it is only the daily-averaged velocity that is approximately
@@
 learned using quasi-geostrophic models to study submesoscale turbulence, claims that
-the latter is dominantly in geostrophic balance
-are overstated \citep{sasaki_etal2014,callies_etal2016}; it is only the daily-averaged fields that are to a large extent
-in  geostrophic balance (the jPDF of daily-averaged vorticity vs. laplacian of SSH
-is a ellipse with large eccentricity with main axis tilted by 45$^\circ$; figure \ref{fig3}n).
-
-The October vorticity-strain jPDF show much weaker skewness (the vorticity skewness
+the submesoscale turbulence is dominantly in geostrophic balance
+are overstated \citep{sasaki_etal2014,callies_etal2016}; only the daily-averaged fields
+are largely
+in  geostrophic balance (the jPDF of daily-averaged vorticity vs. Laplacian of SSH
+is an ellipse with large eccentricity with main axis tilted by 45$^\circ$; figure \ref{fig3}n).
+
+The October vorticity-strain jPDF shows much weaker skewness (the vorticity skewness
 is 0.68 and 0.67 for hourly and daily-averaged velocities). The shape of the
@@
 jPDF is significantly different than the April one. In particular, the jPDF
- for daily-averaged currents is still egg-shaped but hourly and residual are
+for daily-averaged currents is still egg-shaped but hourly and residual fields are
 ellipses about $\zeta=0$, likely associated with inertia-gravity waves.
 That the submesoscale dynamics in October is mainly unbalanced is clearly depicted
-in the shape of jPDF of vorticity-laplacian of SSH for hourly and
+in the shape of jPDF of vorticity-Laplacian of SSH for hourly and
 residual fields
 (figure \ref{fig3}p and \ref{fig3}r), which is an ellipse aligned in the vertical axis.
-Of course daily-averaging the model fields filter out the ageostrophic flows and
+Of course daily-averaging the model suppresses the ageostrophic flows and
 the daily-averaged flow is essentially geostrophic.
@@
 (SSH) variance.  Before calculating the spectra, time-mean and spatial linear
-trends were removed, and the ensuing fields were multiplied by a two-dimensional Hanning spectral ``window''.
-The two-dimensional spectra were averaged azimuthly \citep[e.g., ][]{rocha_etal2016} .We discuss only spectra for the 1/48$^\circ$, which simply extends the 1/24$^\circ$
+trends were removed, and the resulting fields were multiplied by a two-dimensional Hanning ``spectral window''.
+The two-dimensional spectra were averaged azimuthly \citep[e.g., ][]{rocha_etal2016} .
+We discuss only spectra for the 1/48$^\circ$, which extends the 1/24$^\circ$
 towards smaller scales (supplemental material).
 
-Figure \ref{fig3}a depicts the horizontal wavenumber spectra of surface KE in
+Figure \ref{fig4}a depicts the horizontal wavenumber spectra of surface KE in
 April and October. At scales larger than 20km, the spectra based on hourly
@@
 \section{Summary and Conclusion}
-Our main finding is that
+Our main finding is that in tidally forced global simulations the
 near-surface upper-ocean inertia-gravity waves undergo
 a strong seasonal cycle that is out-of-phase with the seasonal cycle of
-the submesoscale turbulence. \cite{dasaro1978} showed that linear internal waves
-velocity in the mixed-layer strongly depends on the density jump at the mixed-layer
-base, with largest mixed-layer velocities when the jump is strongest.
-In Summer, the shallow mixed-layer overlays a strong seasonal pycnocline,
+submesoscale turbulence. \cite{dasaro1978} showed that velocity of linear internal waves
+in the mixed-layer strongly depends on the density jump at the mixed-layer
+base, with largest mixed layer velocities when the jump is strongest.
+In Summer, the shallow mixed layer overlays a strong seasonal pycnocline,
 and thus the internal waves projection onto the mixed-layer may be stronger
 according to \cite{dasaro1978}'s arguments. An
-alternative explanation is that the shape of the baroclinic modes significantly
-change seasonally. In particular the surface amplitude of the baroclinic modes are
-largest when the upper ocean is strongly stratified in late Summer/ early Fall
+alternative explanation is that the shape of the baroclinic modes
+change seasonally. In particular the baroclinic modes are
+significantly more surface intensified in late Summer/early Fall
  (supplemental material).
@@
 submesoscale turbulence. We conjecture that the dominance at inertia-gravity waves in
-Summer \citep{callies_etal2015} is a consequence of both  of submesoscale turbulence and
-enhancement of inertia-gravity waves due to restratification of
+Summer \citep{callies_etal2015} is a consequence both of suppresion of submesoscale turbulence and
+enhancement of inertia-gravity waves due to re-stratification of
 the upper ocean. Of course, our study focuses on a single patch of ocean in the
-vicinities of the Kuroshio Extension, which may be typical of mesoscale-rich
+vicinity of the Kuroshio Extension, which may be typical of mesoscale-rich
 subtropical regions, but not
-representative of other regions such as eastern boundary currents
-and the middle of the subtropical gyre: care must be taken in overgeneralizing
+representative of other regions such as low eddy kinetic energy eastern boundary currents
+and the middle of the subtropical gyre: care must be taken to avoid overgeneralizing
 these results.  We plan to report on the geographic variability of submesoscale
@@
   \begin{center}
-    \includegraphics[width=.6\textwidth]{figs/fig3.pdf}
- \caption{Seasonal variation of joint probability distributions: vorticity vs.
-          divergence (a through e), vorticity vs. strain rate (f through l).
-          and vorticity vs. laplacian of sea-surface height (m through r).
-          Dashed lines in (a) through (e) represent $\delta = \frac{\omega}{f_0}
-          \zeta$ where $\omega = 2\pi/$(12 h, 3 h) and $f_0$ is the inertial frequency
-          in the middle of the domain (32.5$^\circ$ N). Dashed lines in (f) through (l)
-          represent pure strain flow $\alpha = \zeta$. Dashed lines in (m) through
-          (o) represent geostrophic flow $\zeta = \frac{g}{f}\nabla_h^2 \eta$.}
+    \includegraphics[width=1.\textwidth]{figs/fig3.pdf}
+ \caption{Seasonal variation of joint probability distributions:  vorticity vs. strain rate (a through f).
+          and vorticity vs. Laplacian of sea-surface height (g through l) in April (a through c;
+          g  through i) and October (d through f;
+          j  through m).
+          Dashed lines in (a) through (e) represent pure strain flow $\alpha = \pm\zeta$,
+          characteristic of fronts. Dashed lines in (f) through
+          (l) represent geostrophic flow $\zeta = \frac{g}{f}\nabla_h^2 \eta$.}
  \label{fig3}

### 2016-06-27 writeup/rocha_etal.tex
Minor corrections.

@@
     re-stratification weakens submesoscale turbulence and enhances intertia-gravity
-     waves near the surface. In the Kuroshio Extension region,
+     waves near the surface. In the Kuroshio Extension,
     inertia-gravity waves strongly dominate the surface kinetic energy and
-    sea-surface height variance in late Summer/early Fall at scales as large as 80 km.
+    sea-surface height variance in late Summer/early Fall at submesoscales.
 \end{abstract}
@@
 
-\section{Statistics of surface lateral velocity tensor}
+\section{Statistics of the surface lateral velocity tensor}
 To study the seasonality in the surface velocity, we calculate the lateral velocity tensor
@@
 The altimeter products were produced by Ssalto/Duacs
-and distributed by Aviso, with support from Cnes (\texttt{http://www.aviso.altimetry.fr/duacs/})
+and distributed by Aviso, with support from CNES (\texttt{http://www.aviso.altimetry.fr/duacs/}).
 \end{acknowledgments}
@@
 
-\subsection*{How well does the LLC simulations captures high-frequency modes?}
+\subsection*{How well do the LLC simulations captures high-frequency modes?}
 To assess how well the LLC simulations represents the high-frequency variability,

### 2016-06-28 writeup/rocha_etal.tex
Teriz edits.

@@
     strong modulation in the governing near-surface dynamics at submesoscales
-    (roughly 10-100 km) in the Kuroshio Extension. Consistent with recent studies, deep late Winter mixed layers
+    (roughly 10-100 km) in the Kuroshio Extension. Consistent with recent studies, deep late-Winter mixed layers
     are prone to baroclinic instabilities, and submesoscale turbulence
@@
 We use the output of two latitude-longitude polar cap (LLC)
-MITgcm realistic numerical simulations. The outputs
+realistic numerical simulations. The outputs
 analyzed here, LLC2160 (1/24$^\circ$) and LLC4320 (1/48$^\circ$)  have
-90-vertical-levels simulations from a hierarchy of global
+90-vertical-levels and are
 forward Massachusetts Institute of Technology general circulation model (MITgcm)
@@
 A key aspect of the LLC simulations is that they were forced by
-the 16 most significant tidal components.
+the 16 most-significant tidal components.
 Because barotropic tides interact with topography and generate internal
-tides that project on mesoscales to submesoscales
+tides that project onto mesoscales to submesoscales
 \citep[e.g., ][]{rocha_etal2016}, tidal forcing fundamentally distinguishes our analysis
@@
 To study seasonal variations in the upper-ocean dynamics, we focus on the northwest
-Pacific, in the vicinities of the Kuroshio
+Pacific, in the vicinity of the Kuroshio
 Extension, where previous studies have suggested strong submesoscale seasonality
@@
 spanning 155-175$^\circ$E; 25-40$^\circ$N (figure \ref{fig1}a). The stratification
-in this mesoscale-rich subtropical region undergoes a strong seasonal cycle: wintertime
-enhancement of small-scale turbulence de-stratifies the upper ocean, yielding mixed layers
+in this mesoscale-rich subtropical region undergoes a vigorous seasonal cycle: wintertime-enhanced
+small-scale turbulence de-stratifies the upper ocean, yielding mixed layers
 as deep as 300m. In late Spring/early Summer enhanced solar radiation and mixed-layer
 instabilities re-stratify the upper ocean, yielding mixed layers as shallow as 40m.
-Very importantly, the stratification and its seasonality are well-captured by both LLC simulations:
+Fundamentally, the stratification and its seasonality are well-captured by both LLC simulations:
 a comparison with Argo climatology shows that both simulations skillfully represent the Kuroshio
@@
 The components of the lateral velocity tensor are calculated using a centered
-second-order finite differences scheme. We then calculate
+second-order finite difference scheme. We then diagnose
 the vertical vorticity $\zeta \equiv v_x - u_y$, lateral rate of strain
@@
 yielding RMS vorticity and strain roughly consistent with the diagnostics from
-AVISO gridded geostrophic velocities (compare blue lines with black lines in figures \ref{fig2}a-b).
+AVISO gridded geostrophic velocities (compare red lines to black lines in figures \ref{fig2}a-b).
 The overall picture that emerges is consistent with recent studies: shallow baroclinic
@@
 smaller-scale submesoscale flows, a secondary RMS divergence peak in winter  is
-nearly as strong as the summertime's. Submesoscale fronts and eddies evolve
-relatively fast and there is no clear
+nearly as strong as in Summer. Submesoscale fronts and eddies evolve
+relatively fast, and there is no clear
 temporal and spatial scale separation between those motions and inertia-gravity waves
@@
 \section{Seasonality in governing submesoscale dynamics}
-The overall results of figure \ref{fig2}c shows that summertime submesoscale surface
- variability stems from different dynamics than wintertime variability.
+The overall results of figure \ref{fig2}c suggest that submesoscale surface
+ variability stems from different dynamics in Summer than Winter.
  To better understand these differences, we calculate
 joint probability distributions (jPDF) of
-vorticity-divergence, vorticity-strain, and vorticity-Laplacian of sea-surface height.
+vorticity-strain and vorticity-Laplacian of sea-surface height (\ref{fig3}).
 
 The April vorticity-strain jPDF has a shape characteristic of submesoscale turbulence
- (see figures \ref{fig3}f-g). The alignment of vorticity and strain $\alpha \sim \zeta$
+ (see figures \ref{fig3}a-c). The alignment of vorticity and strain $\alpha \sim \pm\zeta$
  with strong positive skewness are fingerprints of submesoscale fronts
@@
  the vorticity skewness reduces 1.4 to 1.13 from hourly to daily-averaged.
- The April vorticity-divergence jPDFs, however, egg-shaped which again highlights
- the strong skewness in vorticity. Thus wintertime submesoscale surface velocity
- is strongly dominated by submesoscale turbulence --- this is true even for
+ Thus wintertime submesoscale surface velocity is strongly dominated by submesoscale turbulence --- this is true even for
  superinertial submesoscale currents. The hourly velocity fields are significantly
 ageostrophic as depicted by the jPDF of vorticity-Laplacian of SSH
-(figure \ref{fig3}o); it is only the daily-averaged velocity that is approximately
-in geostrophic balance (figure \ref{fig3}l). While there is certainly much to be
+(figure \ref{fig3}g). While there is certainly much to be
 learned using quasi-geostrophic models to study submesoscale turbulence, claims that
@@
 in  geostrophic balance (the jPDF of daily-averaged vorticity vs. Laplacian of SSH
-is an ellipse with large eccentricity with main axis tilted by 45$^\circ$; figure \ref{fig3}n).
+is an ellipse with large eccentricity with main axis tilted by 45$^\circ$; figure \ref{fig3}h).
+
+\begin{figure*}[ht]
+  \begin{center}
+    \includegraphics[width=1.\textwidth]{figs/fig3.pdf}
+ \caption{Seasonal variation of joint probability distributions:  vorticity vs. strain rate (a through f).
+          and vorticity vs. Laplacian of sea-surface height (g through l) in April (a through c;
+          g  through i) and October (d through f;
+          j  through m).
+          Dashed lines in (a) through (e) represent pure strain flow $\alpha = \pm\zeta$,
+          characteristic of fronts. Dashed lines in (f) through
+          (l) represent geostrophic flow $\zeta = \frac{g}{f}\nabla_h^2 \eta$.}
+ \label{fig3}
+ \end{center}
+\end{figure*}
 
@@
 about $\zeta=0$, one with a 45$^\circ$ slope (characteristic of submesocale fronts
-that persist in Summer) and one with a very steep slope. The vorticity-divergence
-jPDF is significantly different than the April one. In particular, the jPDF
-for daily-averaged currents is still egg-shaped but hourly and residual fields are
-ellipses about $\zeta=0$, likely associated with inertia-gravity waves.
+that persist in Summer) and one with a very steep slope.
 That the submesoscale dynamics in October is mainly unbalanced is clearly depicted
@@
 residual fields
-(figure \ref{fig3}p and \ref{fig3}r), which is an ellipse aligned in the vertical axis.
-Of course daily-averaging the model suppresses the ageostrophic flows and
-the daily-averaged flow is essentially geostrophic.
+(figure \ref{fig3}j and \ref{fig3}k), which is an ellipse aligned in the vertical axis.
+Of course daily-averaging the model suppresses the ageostrophic flows, and
+the daily-averaged flow is essentially geostrophic as depicted by the 45$^\circ$-tilted
+ellipse in the vorticity-Laplacian of SSH jPDF.
 
@@
  negatively skewed as predicted by frontogenesis \citep{capet_etal2008a,mcwilliams2016}
-. In late Summer/early Fall, the divergence is more stronger, but PDFs are much less skewed,
+. In late Summer/early Fall, the divergence is stronger, but PDFs are much less skewed,
 consistent with linear inertia-gravity waves.
@@
 
+\begin{figure*}[ht]
+  \begin{center}
+    \includegraphics[width=.75\textwidth]{figs/fig4.pdf}
+ \caption{Surface (horizontal) KE (a) and SSH variance wavenumber spectra (b)
+ in the 1/48$^\circ$ simulation. Solid lines
+ are spectra based on hourly snapshots, dashed lines are spectra based on daily-averaged
+ fields.}
+ \label{fig4}
+ \end{center}
+\end{figure*}
+
 Figure \ref{fig4}a depicts the horizontal wavenumber spectra of surface KE in
@@
 velocity snapshots are nearly indistinguishable within 95$\%$ confidence level
-(compare solid lines in figure \ref{fig3}a). Consistent with the results of
+(compare solid lines in figure \ref{fig4}a). Consistent with the results of
 \citet{rocha_etal2016} who analyzed a 5-month output of the LLC4320 simulation in Drake Passage,
@@
 km both in April and October (compare solid lines against dashed lines in figure
-\ref{fig3}a). But this suppression is more dramatic in October, when the horizontally
+\ref{fig4}a). But this suppression is dramatic in October, when the horizontally
 divergent flows peak. At scales
@@
 In October, when the inertia-gravity waves are strongest the surface and submesoscale
-turbulence is weakest, there is a dramatic difference between the spectra
+turbulence is weakest. There is a dramatic difference between the spectra
 based on hourly and daily-averaged SSH: at scales smaller than 100 km, the spectra
@@
 turbulence and near-surface inertia-gravity waves conspire to yield very weak seasonality
-in the spectra of KE and SSH variance based on hourly field.
+in the spectra of KE and SSH variance based on hourly fields.
 
@@
 and observational \citep{callies_etal2015,buckingham_etal2016} evidence of vigorous seasonality in
-submesoscale turbulence. We conjecture that the dominance at inertia-gravity waves in
-Summer \citep{callies_etal2015} is a consequence both of suppresion of submesoscale turbulence and
+submesoscale turbulence. We conjecture that the summertime dominance of inertia-gravity waves
+ \citep{callies_etal2015} is a consequence both of suppression of submesoscale turbulence and
 enhancement of inertia-gravity waves due to re-stratification of
@@
 representative of other regions such as low eddy kinetic energy eastern boundary currents
-and the middle of the subtropical gyre: care must be taken to avoid overgeneralizing
+and the middle of the subtropical gyre; care must be taken to avoid overgeneralizing
 these results.  We plan to report on the geographic variability of submesoscale
@@
 % \citep{sasaki_etal2014,qiu_etal2014} are overstated.
-
-\begin{figure*}[ht]
-  \begin{center}
-    \includegraphics[width=1.\textwidth]{figs/fig3.pdf}
- \caption{Seasonal variation of joint probability distributions:  vorticity vs. strain rate (a through f).
-          and vorticity vs. Laplacian of sea-surface height (g through l) in April (a through c;
-          g  through i) and October (d through f;
-          j  through m).
-          Dashed lines in (a) through (e) represent pure strain flow $\alpha = \pm\zeta$,
-          characteristic of fronts. Dashed lines in (f) through
-          (l) represent geostrophic flow $\zeta = \frac{g}{f}\nabla_h^2 \eta$.}
- \label{fig3}
- \end{center}
-\end{figure*}
-
-\begin{figure*}[ht]
-  \begin{center}
-    \includegraphics[width=.75\textwidth]{figs/fig4.pdf}
- \caption{Surface (horizontal) KE (a) and SSH variance wavenumber spectra (b)
- in the 1/48$^\circ$ simulation. Solid lines
- are spectra based on hourly snapshots, dashed lines are spectra based on daily-averaged
- fields.}
- \label{fig4}
- \end{center}
-\end{figure*}
 

### 2016-06-28 writeup/rocha_etal.tex
Teriz edits in the supplemental material.

@@
           g  through i) and October (d through f;
-          j  through m).
+          j  through l).
           Dashed lines in (a) through (e) represent pure strain flow $\alpha = \pm\zeta$,
@@
 Medium-range Weather Forecasting (ECMWF) atmospheric operational model analysis,
- starting 2011 humidity and the 16 most significant  tidal components.
+ starting 2011 and the 16 most significant  tidal components.
 The LLC control files are available online
@@
 horizontal wavenumber spectra suggests an effective resolution of about 8 km.
-(see figure \ref{fig3}), while wavenumber spectra suggests an effective resolution
+(see figure \ref{fig4}), and wavenumber spectra suggests an effective resolution
 of 20 km for  LLC2160 simulation.
@@
 \label{tab:llc}
-\caption{\small The MITgcm Latitute-Longitude Cap (LLC) spin-up hierarchy. This study will analyze
+\caption{\small The MITgcm Latitute-Longitude Cap (LLC) spin-up hierarchy. This study analyzes
          the output of LLC2160 and LLC4320. Model fields are available hourly.
@@
 
-\subsection*{How well do the LLC simulations captures high-frequency modes?}
+\subsection*{How well do the LLC simulations capture high-frequency modes?}
 To assess how well the LLC simulations represents the high-frequency variability,
@@
 and the inertial, diurnal, and semi-diurnal frequencies agree well --- the KESS
-data are more energetic at intermedia frequencies. At 40 m, the LLC simulations and observations
-have consistent spectra at periods larger than about 6 h. The data from moorings
+data are more energetic at intermediate frequencies. At 40 m, the LLC simulations and observations
+have consistent spectra at periods larger than about 6 h. The data from the moorings
 are more energetic at higher frequencies,
@@
 excursions; 2) The model does not have enough resolution to resolve the full internal
-wave spectrum; 3) The model is not forced at frequencies higher that 6h.
+wave spectrum; 3) The model is not forced at frequencies higher than 6h.
 
@@
 Figure \ref{figS3}a shows the domain-averaged potential density from the WOA 2013
-climatology in April and October. Clearly, the the changes to the stratification
+climatology in April and October. Clearly, the changes to the stratification
 are dramatic, with the formation of a seasonal pycnocline in Summer. The strong
@@
 
-\textcolor{red}{Perhaps include a model with a slab over a slowly varying
-pycnocline. The model is simple, but the solution is a bit messy and may
-scare GRL readers...}.
+\textcolor{red}{Perhaps include analytical results from a model with a slab over a slowly varying
+pycnocline.}
 

### 2016-06-28 writeup/rocha_etal.tex
Small edits

@@
      waves near the surface. In the Kuroshio Extension,
-    inertia-gravity waves strongly dominate the surface kinetic energy and
-    sea-surface height variance in late Summer/early Fall at submesoscales.
+    inertia-gravity waves strongly dominate the submesoscale surface kinetic energy and
+    sea-surface height variance in late Summer/early Fall.
 \end{abstract}
@@
 realistic numerical simulations. The outputs
-analyzed here, LLC2160 (1/24$^\circ$) and LLC4320 (1/48$^\circ$)  have
-90-vertical-levels and are
+analyzed here, LLC2160 (1/24$^\circ$) and LLC4320 (1/48$^\circ$),   are
 forward Massachusetts Institute of Technology general circulation model (MITgcm)
-numerical solutions on a LLC grid   \citep{forget_etal2015} that were
-spun up from a ECCO2 adjoint-method state estimate. Both simulations were forced by
+numerical solutions on a LLC grid  \citep{forget_etal2015} with
+90-vertical-levels. The coarser-resolution LLC simulation was
+spun up from a ECCO2 adjoint-method state estimate, which synthetized millions
+of observations from 2009 through 2011. Both simulations were forced by
 tides and 6-hourly surface fluxes. The LLC2160
@@
 instabilities re-stratify the upper ocean, yielding mixed layers as shallow as 40m.
-Fundamentally, the stratification and its seasonality are well-captured by both LLC simulations:
+Fundamentally, the upper-ocean density structure is well-captured by both LLC simulations:
 a comparison with Argo climatology shows that both simulations skillfully represent the Kuroshio
@@
 than in late Summer/early Fall. Because the wintertime vorticity and strain rate
-are dominated by the smallest scales in the flow, increasing the resolution from
+are dominated by the smallest scales in the flow (the KE spectra is shallower than
+a $-3$ power-law in Winter), increasing the resolution from
 1/24$^\circ$ to 1/48$^\circ$ significantly increases the wintertime RMS vorticity
@@
   \caption{Time series of the root-mean-square (RMS) of vorticity (a) and
-  lateral rate of strain (b), and horizontal divergence (c) in the LLC outputs and gridded AVISO data.}
+  rate of strain (b), and horizontal divergence (c) in the LLC outputs and gridded AVISO data.}
   \label{fig2}
@@
 AVISO gridded geostrophic velocities (compare red lines to black lines in figures \ref{fig2}a-b).
-The overall picture that emerges is consistent with recent studies: shallow baroclinic
+The picture that emerges is consistent with recent studies: shallow baroclinic
 instabilities energize the submesoscales in late Winter, drawing from the available
@@
  variability stems from different dynamics in Summer than Winter.
- To better understand these differences, we calculate
+ To characterize these differences, we calculate
 joint probability distributions (jPDF) of
@@
 
-To better understand the projection of these flows onto different horizontal
+To better quantify the projection of these flows onto different horizontal
 scales, we calculate wavenumber spectra of kinetic energy and sea-surface height
@@
 km both in April and October (compare solid lines against dashed lines in figure
-\ref{fig4}a). But this suppression is dramatic in October, when the horizontally
-divergent flows peak. At scales
+\ref{fig4}a). But this suppression is dramatic in October, when the inertia-gravity waves peak. At scales
 smaller than 100 km, 39$\%$ of the surface KE in April is accounted for by super-inertial
@@
 of hourly SSH roughly follows a $-2$ power-law, whereas the spectra of daily-averaged
-SSH roughly follows a $-5$ power-law. At these scales, 33$\%$ of the SSH variance in April is accounted for by suprainertial
+SSH roughly follows a $-5$ power-law. At these scales, 33$\%$ of the SSH variance in April is accounted for by super-inertial
 flows as opposed to 83$\%$ in October! Curiously, the out-of-phase seasonal cycle of submesoscale
@@
 vicinity of the Kuroshio Extension, which may be typical of mesoscale-rich
-subtropical regions, but not
+subtropical regions, but is unlikely to be 
 representative of other regions such as low eddy kinetic energy eastern boundary currents

### 2016-06-28 writeup/rocha_etal.tex
Small edit and change fig size

@@
 \begin{center}
-\hspace{-1.25cm}\includegraphics[width=.6\textwidth]{figs/fig1_1.pdf}\\
+\hspace{-1.25cm}\includegraphics[width=.75\textwidth]{figs/fig1_1.pdf}\\
 \vspace{-.125cm}
-\includegraphics[width=.65\textwidth]{figs/fig1_2.pdf}
+\includegraphics[width=.75\textwidth]{figs/fig1_2.pdf}
  \caption{(a) The study region with the subregion where the LLC outputs are
@@
    \begin{center}
-     \includegraphics[width=.65\textwidth]{figs/fig2.pdf}
+     \includegraphics[width=.75\textwidth]{figs/fig2.pdf}
   \caption{Time series of the root-mean-square (RMS) of vorticity (a) and
@@
 
-The bulk of vertical vorticity and lateral strain rate are associated with subinertial flows
+The bulk of vertical vorticity and strain rate are associated with subinertial flows
 ($T_f = 2\pi/f_0\approx 23.5$h, where $f_0$ is the inertial frequency at
@@
   \begin{center}
-    \includegraphics[width=.75\textwidth]{figs/fig4.pdf}
+    \includegraphics[width=.9\textwidth]{figs/fig4.pdf}
  \caption{Surface (horizontal) KE (a) and SSH variance wavenumber spectra (b)
@@
 
-These high-frequency submesoscale flows significantly project on the sea-surface.
+The inertia-gravity waves significantly project on the sea-surface.
 In October, when the inertia-gravity waves are strongest the surface and submesoscale

### 2016-06-28 writeup/rocha_etal.tex
Small edit.
Small edits

@@
 ageostrophic as depicted by the jPDF of vorticity-Laplacian of SSH
-(figure \ref{fig3}g). While there is certainly much to be
-learned using quasi-geostrophic models to study submesoscale turbulence, claims that
-the submesoscale turbulence is dominantly in geostrophic balance
-are overstated \citep{sasaki_etal2014,callies_etal2016}; only the daily-averaged fields
+(figure \ref{fig3}g). Even in April, only the daily-averaged fields
 are largely
 in  geostrophic balance (the jPDF of daily-averaged vorticity vs. Laplacian of SSH
-is an ellipse with large eccentricity with main axis tilted by 45$^\circ$; figure \ref{fig3}h).
+is an ellipse with large eccentricity and main axis tilted by 45$^\circ$; figure \ref{fig3}h).
+
+%While there is certainly much to be
+%learned using quasi-geostrophic models to study submesoscale turbulence, claims that
+%the submesoscale turbulence is dominantly in geostrophic balance
+%are overstated \citep{sasaki_etal2014,callies_etal2016};
 

### 2016-06-28 writeup/rocha_etal.tex
Small edits

@@
 
-\authors{Cesar et al. \altaffilmark{1}}
+\authors{Cesar B. Rocha et al. \altaffilmark{1}}
 % Eric Brown,\altaffilmark{1,2} Rick Williams,\altaffilmark{3}
@@
     are prone to baroclinic instabilities, and submesoscale turbulence
-    prevails in late Winter/early Spring. Inertia-gravity waves are enhanced near
-    the surface when the upper ocean is strongly stratified. Thus the summertime
-    re-stratification weakens submesoscale turbulence and enhances intertia-gravity
+    prevails in late Winter/early Spring. While summertime
+    re-stratification weakens submesoscale turbulence, it also enhances intertia-gravity
      waves near the surface. In the Kuroshio Extension,
@@
 enhancement of inertia-gravity waves due to re-stratification of
-the upper ocean. Of course, our study focuses on a single patch of ocean in the
+the upper ocean. Of course, this study on a single patch of ocean in the
 vicinity of the Kuroshio Extension, which may be typical of mesoscale-rich

### 2016-06-28 writeup/rocha_etal.tex
Small edits

@@
 Our main finding is that in tidally forced global simulations the
-near-surface upper-ocean inertia-gravity waves undergo
+near-surface upper-ocean inertia-gravity waves in the Kuroshio Extension undergo
 a strong seasonal cycle that is out-of-phase with the seasonal cycle of

### 2016-06-29 writeup/rocha_etal.tex
Some minor fixes by Sarah

--- 2016-06-28 writeup/rocha_etal.tex
@@
     prevails in late Winter/early Spring. While summertime
-    re-stratification weakens submesoscale turbulence, it also enhances intertia-gravity
+    re-stratification weakens submesoscale turbulence, it also enhances inertia-gravity
      waves near the surface. In the Kuroshio Extension,
@@
 
-Using the output of two $1/24^\circ$ and $1/48^\circ$ global
+Using the output of $1/24^\circ$ and $1/48^\circ$ global
 numerical simulations with embedded tides, we show that inertia-gravity waves undergo
@@
 90-vertical-levels. The coarser-resolution LLC simulation was
-spun up from a ECCO2 adjoint-method state estimate, which synthetized millions
+spun up from an ECCO2 adjoint-method state estimate, constrained to millions
 of observations from 2009 through 2011. Both simulations were forced by
@@
 the local planetary vorticity, and root-mean-square (RMS) of about $0.4f$. In early
-Fall, the situation is the opposite: the vertical vorticity is relatively coarse-grained,
-its local maximum and RMS is smaller than $0.5f$.
+Fall, the situation is the opposite: the vertical vorticity is relatively coarse-grained;
+its local maximum and RMS are both smaller than $0.5f$.
 Indeed, the seasonal cycle of vorticity and rate of strain are strong (figure~\ref{fig2}):
@@
  jPDF is similar for hourly, daily-averaged, and residual velocities, although
- the vorticity skewness reduces 1.4 to 1.13 from hourly to daily-averaged.
- Thus wintertime submesoscale surface velocity is strongly dominated by submesoscale turbulence --- this is true even for
+ the vorticity skewness reduces from 1.4 to 1.13 from hourly to daily-averaged.
+ The April results are characteristic of Winter, indicating that wintertime submesoscale
+ surface velocity is strongly dominated by submesoscale turbulence --- this is true even for
  superinertial submesoscale currents. The hourly velocity fields are significantly
@@
 that persist in Summer) and one with a very steep slope.
-That the submesoscale dynamics in October is mainly unbalanced is clearly depicted
+That the submesoscale dynamics in October are mainly unbalanced is clearly depicted
 in the shape of jPDF of vorticity-Laplacian of SSH for hourly and
@@
 The two-dimensional spectra were averaged azimuthly \citep[e.g., ][]{rocha_etal2016} .
-We discuss only spectra for the 1/48$^\circ$, which extends the 1/24$^\circ$
+We discuss only spectra for the 1/48$^\circ$ simulation, which extends the 1/24$^\circ$ simulation
 towards smaller scales (supplemental material).
@@
 \section{Summary and Conclusion}
-Our main finding is that in tidally forced global simulations the
-near-surface upper-ocean inertia-gravity waves undergo
+Our main finding is that in global simulations with embedded tides the
+near-surface upper-ocean inertia-gravity waves in the Kuroshio Extension undergo
 a strong seasonal cycle that is out-of-phase with the seasonal cycle of
@@
 alternative explanation is that the shape of the baroclinic modes
-change seasonally. In particular the baroclinic modes are
+changes seasonally. In particular the baroclinic modes are
 significantly more surface intensified in late Summer/early Fall
@@
 
-Our work adds to recent studies that presented modeling \cite{sasaki_etal2014}
+Our work adds to recent studies that presented modeling \citep{sasaki_etal2014}
 and observational \citep{callies_etal2015,buckingham_etal2016} evidence of vigorous seasonality in
@@
 enhancement of inertia-gravity waves due to re-stratification of
-the upper ocean. Of course, this study on a single patch of ocean in the
+the upper ocean. Of course, this study focuses on a single patch of ocean in the
 vicinity of the Kuroshio Extension, which may be typical of mesoscale-rich
@@
 the higher resolution is resolving a wider array of scales. Between 20 and 100 km
-there is linewidth agreement between the spectra of the two simulations.
+the spectra of the two simulations are visually indistinguishable.
 

### 2016-07-04 writeup/rocha_etal.tex
Teriz last fixes.

@@
 % Shorter version of title entered in capital letters:
-\titlerunninghead{Seasonality at submesoscales in the Kuroshio Extension}
+\titlerunninghead{Seasonality at submesoscales}
 
@@
 
-\title{Seasonality in governing dynamics at submesoscales in the Kuroshio Extension}
+\title{Seasonality in governing submesoscale dynamics in the Kuroshio Extension}
 %
@@
            \item  Submesoscale turbulence dominates the horizontal velocity and
-                  sea-surface height variability in late Winter/early Spring.
+                  sea-surface height variability in late winter/early spring.
             \item Inertia-gravity waves dominate the horizontal velocity and
-                   sea-surface height variability in late Summer/early Fall.
+                   sea-surface height variability in late summer/early fall.
                   }
@@
     strong modulation in the governing near-surface dynamics at submesoscales
-    (roughly 10-100 km) in the Kuroshio Extension. Consistent with recent studies, deep late-Winter mixed layers
+    (roughly 10-100 km) in the Kuroshio Extension. Consistent with recent studies, deep late-winter mixed layers
     are prone to baroclinic instabilities, and submesoscale turbulence
-    prevails in late Winter/early Spring. While summertime
+    prevails in late winter/early spring. While summertime
     re-stratification weakens submesoscale turbulence, it also enhances inertia-gravity
@@
     inertia-gravity waves strongly dominate the submesoscale surface kinetic energy and
-    sea-surface height variance in late Summer/early Fall.
+    sea-surface height variance in late summer/early fall.
 \end{abstract}
@@
 with the seasonal cycle of submesoscale turbulence. Consistent with previous studies,
-deep late Winter mixed layers are prone to
+deep late-winter mixed layers are prone to
 shallow baroclinic instabilities that are roughly in geostrophic balance
@@
 
-Inertia-gravity waves, on the other hand, peak in late Summer/early Fall,
+Inertia-gravity waves, on the other hand, peak in late summer/early fall,
 when the upper ocean is strongly stratified. Thus there exists a
 strong seasonal modulation in the dominant upper-ocean submesoscale dynamics:
-submesoscale turbulence dominates the upper-ocean dynamics in late Winter/early Spring,
-whereas quasi-linear inertia-gravity waves prevail in late Summer/early Fall.
-Because submesoscale turbulence is weakest in late Summer/early Fall,
+submesoscale turbulence dominates the upper-ocean dynamics in late winter/early ,
+whereas quasi-linear inertia-gravity waves prevail in late summer/early fall.
+Because submesoscale turbulence is weakest in late summer/early fall,
 the present results suggest that inertia-gravity waves account for most of the
@@
 We analyze a sub-domain of the LLC4320 and LLC2160 simulations of about 2000 km$^2$
-spanning 155-175$^\circ$E; 25-40$^\circ$N (figure \ref{fig1}a). The stratification
+spanning 155-175$^\circ$E; 25-40$^\circ$N winter \ref{fig1}a). The stratification
 in this mesoscale-rich subtropical region undergoes a vigorous seasonal cycle: wintertime-enhanced
 small-scale turbulence de-stratifies the upper ocean, yielding mixed layers
-as deep as 300m. In late Spring/early Summer enhanced solar radiation and mixed-layer
+as deep as 300m. In late spring/early summer enhanced solar radiation and mixed-layer
 instabilities re-stratify the upper ocean, yielding mixed layers as shallow as 40m.
@@
           analyzed. Colors represent the topography and white lines are contours of absolute
-          dynamic topography every 0.1 m. LLC 4320 (1/48$^\circ$) snapshots of surface vorticity (b and c) and a transect
+          dynamic topography every 0.1 m from AVISO. LLC 4320 (1/48$^\circ$) snapshots of surface vorticity (b and c) and transects
           of potential density at 165$^\circ$E (d and e). The snapshots were
@@
 
-Figures \ref{fig1}b-c show snapshots of vertical vorticity $\zeta$ in early Spring
-(April 15) and Fall (October 15) in the LLC4320 (1/48$^\circ$) simulation.
+Figures \ref{fig1}b-c show snapshots of vertical vorticity $\zeta$ in early spring
+(April 15) and fall (October 15) in the LLC4320 (1/48$^\circ$) simulation.
 The model solutions depict seasonality in vorticity: large values of
-fine-grained vertical vorticity are observed in early Spring with maximum
+fine-grained vertical vorticity are observed in early spring with maximum
 values as large as $4f$, where $f$ is
 the local planetary vorticity, and root-mean-square (RMS) of about $0.4f$. In early
-Fall, the situation is the opposite: the vertical vorticity is relatively coarse-grained;
+fall, the situation is the opposite: the vertical vorticity is relatively coarse-grained;
 its local maximum and RMS are both smaller than $0.5f$.
-Indeed, the seasonal cycle of vorticity and rate of strain are strong (figure~\ref{fig2}):
+Indeed, the seasonal cycle of vorticity and rate of strain are strong winter~\ref{fig2}):
 in both simulations, the RMS vorticity and strain are about twice as large in
-late Winter/early Spring
-than in late Summer/early Fall. Because the wintertime vorticity and strain rate
+late winter/early spring
+than in late summer/early fall. Because the wintertime vorticity and strain rate
 are dominated by the smallest scales in the flow (the KE spectra is shallower than
-a $-3$ power-law in Winter), increasing the resolution from
+a $-3$ power-law in winter), increasing the resolution from
 1/24$^\circ$ to 1/48$^\circ$ significantly increases the wintertime RMS vorticity
@@
 The picture that emerges is consistent with recent studies: shallow baroclinic
-instabilities energize the submesoscales in late Winter, drawing from the available
+instabilities energize the submesoscales in late winter, drawing from the available
 potential energy stored in deep mixed layers \citep{sasaki_etal2014,callies_etal2015,callies_etal2016}.
@@
 a stronger enhancement of
-lateral divergence in late Summer/early Fall (RMS divergence $\sim0.22 f$ the 1/48$^\circ$
+lateral divergence in late summer/early fall (RMS divergence $\sim0.22 f$ the 1/48$^\circ$
 simulation). Because the 1/48$^\circ$ simulation better resolves
 smaller-scale submesoscale flows, a secondary RMS divergence peak in winter  is
-nearly as strong as in Summer. Submesoscale fronts and eddies evolve
+nearly as strong as in summer. Submesoscale fronts and eddies evolve
 relatively fast, and there is no clear
@@
 divergent flows, but also reduces the wintertime lateral divergence by about 50$\%$.
-The vorticity and strain rate (Figure \ref{fig2}a-b) show that most of the lateral divergence is associated
+The vorticity and strain rate winter \ref{fig2}a-b) show that most of the lateral divergence is associated
 with submesoscale flows: smoothing the velocity fields with a 100km-cut-off
@@
 The overall results of figure \ref{fig2}c suggest that submesoscale surface
- variability stems from different dynamics in Summer than Winter.
+ variability stems from different dynamics in summer than winter.
  To characterize these differences, we calculate
@@
  the vorticity skewness reduces from 1.4 to 1.13 from hourly to daily-averaged.
- The April results are characteristic of Winter, indicating that wintertime submesoscale
+ The April results are characteristic of winter, indicating that wintertime submesoscale
  surface velocity is strongly dominated by submesoscale turbulence --- this is true even for
@@
 ageostrophic as depicted by the jPDF of vorticity-Laplacian of SSH
-(figure \ref{fig3}g). Even in April, only the daily-averaged fields
+winter \ref{fig3}g). Even in April, only the daily-averaged fields
 are largely
@@
 about $\zeta=0$, one with a 45$^\circ$ slope (characteristic of submesocale fronts
-that persist in Summer) and one with a very steep slope.
+that persist in summer) and one with a very steep slope.
 That the submesoscale dynamics in October are mainly unbalanced is clearly depicted
@@
 residual fields
-(figure \ref{fig3}j and \ref{fig3}k), which is an ellipse aligned in the vertical axis.
+winter \ref{fig3}j and \ref{fig3}k), which is an ellipse aligned in the vertical axis.
 Of course daily-averaging the model suppresses the ageostrophic flows, and
@@
  Time series of vorticity and divergence PDFs (supplemental material) suggest
- a strong oscillation between these two regimes. In late Winter/early Spring
+ a strong oscillation between these two regimes. In late winter/early spring
  the vorticity is strongly positively skewed, whereas the divergence is moderately
  negatively skewed as predicted by frontogenesis \citep{capet_etal2008a,mcwilliams2016}
-. In late Summer/early Fall, the divergence is stronger, but PDFs are much less skewed,
+. In late summer/early fall, the divergence is stronger, but PDFs are much less skewed,
 consistent with linear inertia-gravity waves.
@@
 Our main finding is that in global simulations with embedded tides the
-near-surface upper-ocean inertia-gravity waves in the Kuroshio Extension undergo
+near-surface inertia-gravity waves in the Kuroshio Extension undergo
 a strong seasonal cycle that is out-of-phase with the seasonal cycle of
@@
 base, with largest mixed layer velocities when the jump is strongest.
-In Summer, the shallow mixed layer overlays a strong seasonal pycnocline,
+In summer, the shallow mixed layer overlays a strong seasonal pycnocline,
 and thus the internal waves projection onto the mixed-layer may be stronger
@@
 changes seasonally. In particular the baroclinic modes are
-significantly more surface intensified in late Summer/early Fall
+significantly more surface intensified in late summer/early fall
  (supplemental material).
@@
 other small-scale internal waves generated through internal tide interactions.
-Near-inertial wave generation peaks in Winter, but those waves project on large
+Near-inertial wave generation peaks in winter, but those waves project on large
 horizontal scales \citep[e.g, ][]{qi_etal1995}.
@@
 % dominated by ageostrophic flows in
-% Summer/Fall has consequences for the interpretation of data from
+% summer/fall has consequences for the interpretation of data from
 % the future SWOT and COMPIRA altimeter missions,
@@
 % To better understand the high-frequency/high-wavenumber flows  we calculate frequency
-% spectra of surface KE (figure \ref{fig4}). To obtain statistically meaningful results,
+% spectra of surface KE winter \ref{fig4}). To obtain statistically meaningful results,
 % we average spectral estimates along a section at 165$^\circ$E.  There is about 50
@@
 \newpage
-\section*{Supplemental material 1: On the LLC simulations}
-
-Table \ref{tab:llc} show details of the LLC spin-up hierarchy.
-Both simulation analyzed in the present paper, LLC2160 and LLC4320,
-were forced by surface fluxes are from the 0.14$^\circ$ European Centre for
+\section*{Details of the LLC simulations}
+
+Table \ref{tab:llc} shows details of the LLC spin-up hierarchy.
+Both simulations analyzed in the present paper, LLC2160 and LLC4320,
+were forced by surface fluxes from the 0.14$^\circ$ European Centre for
 Medium-range Weather Forecasting (ECMWF) atmospheric operational model analysis,
- starting 2011 and the 16 most significant  tidal components.
+ starting in 2011, and the 16 most significant  tidal components.
 The LLC control files are available online
@@
 of 20 km for  LLC2160 simulation.
+
+The spin-up of the upper-ocean in the LLC4320 (1/48$^\circ$) simulation is
+a couple of days, as depicted the time series of surface horizontal divergence
+in Figure 2b of the main text (see initial rapid increase in RMS divergence in
+September 2011). To avoid the spin-up stage, we chose October to characterize the
+late summer/early fall. The late winter/early spring waves chosen to be 6 months
+out-of-phase with late summer/early fall (hence April).
 
@@
 
-\subsection*{Seasonality in statistics: probability density functions}
+\subsection*{Seasonality in probability density functions}
+
+Probability density functions (pdfs) of vertical vorticity and horizontal divergence
+ at the surface (Figure \ref{figS2_2}) depicts the strong seasonality in the
+ submesoscale dynamics. Time series of vorticity pdf indicate that
+ the development of positive skewness in winter (Figures \ref{figS2_2}a-b) both in
+ hourly ($\sim 1.4$) and daily-averaged ($\sim1.1$) fields. The skewness
+ in divergence for hourly fields is very small year-round ($\sim -0.24$), suggesting the
+ prevalence of inertia-gravity waves. Filtering out the waves by daily-averaging
+ the velocity field increases the skewness in divergence, particularly in winter ($\sim-0.64$). Both
+positive vorticity (cyclonic) skewness and negative divergence (convergence) skewness
+are characteristics of submesoscale turbulence \citep{mcwilliams2016}.
+
 \begin{figure*}[ht]
@@
 owing to lower resolution and long-term averaging. Nonetheless, simulations reproduce well the strong
-seasonality in the stratification. The upper ocean is weakly stratified in late Winter/early Fall
-and restratifies in Summer. Argo climatology and LLC simulations mixed-layer depths are consistent
+seasonality in the stratification. The upper ocean is weakly stratified in late winter/early fall
+and restratifies in summer. Argo climatology and LLC simulations mixed-layer depths are consistent
 at all seasons (not shown).
@@
   \label{figS2_1}
+  \end{center}
+\end{figure*}
+
+\subsection*{Depth-dependence}
+Figure \ref{figS2_2} shows the depth-dependence of the  root-mean-square of the velocity
+gradient lateral tensor for the LLC4320 simulation. The strong seasonality, as depicted
+at the surface in Figure 2 of the main manuscript, is confined to the mixed layer, which
+varies from 50 m in summer to about 350 m in winter.
+
+\begin{figure*}[ht]
+   \begin{center}
+     \includegraphics[width=.9\textwidth]{figs/fig_s2_2.png}
+  \caption{Depth-dependence of root-mean-square of vertical vorticity (a,d), rate of lateral strain (b,e)
+  and horizontal divergence (c,f) for hourly (a,b,c) and daily-averaged (d,e,f,) fields.}
+  \label{figS2_2}
   \end{center}
@@
 
-\section*{Supplemental material 3: Seasonality in pressure modes near-surface amplitude}
+\section*{Seasonality in pressure modes near-surface amplitude}
 
@@
 climatology in April and October. Clearly, the changes to the stratification
-are dramatic, with the formation of a seasonal pycnocline in Summer. The strong
+are dramatic, with the formation of a seasonal pycnocline in summer. The strong
 seasonality of upper-ocean stratification yields a strong seasonality in the
 near-surface shape and amplitude of baroclinic pressure modes (\ref{figS3}b).
-In particular, the surface amplitude is much large in Summer. Thus the projection
+In particular, the surface amplitude is much large in summer. Thus the projection
 of baroclinic tides on the surface may vary significantly seasonally.
 
-\textcolor{red}{Perhaps include analytical results from a model with a slab over a slowly varying
-pycnocline.}
+%\textcolor{red}{Perhaps include analytical results from a model with a slab over a slowly varying
+%pycnocline.}
 

### 2016-07-04 writeup/rocha_etal.tex
Updates stratification figure

@@
 seasonality in the stratification. The upper ocean is weakly stratified in late winter/early fall
-and restratifies in summer. Argo climatology and LLC simulations mixed-layer depths are consistent
+and re-stratifies in summer. Argo climatology and LLC simulations mixed-layer depths are consistent
 at all seasons (not shown).
@@
           climatology (a, b) and the LLC simulations 1/24$^\circ$ (c, d)
-          and  1/48$^\circ$ (e, f). \textcolor{red}{Need to re-plot  the 1/24$^\circ$
-          figures once extraction (currently down) becomes available}.}
+          and  1/48$^\circ$ (e, f).}
   \label{figS2_1}

### 2016-07-04 writeup/rocha_etal.tex
Minor edit

@@
 representative of other regions such as low eddy kinetic energy eastern boundary currents
-and the middle of the subtropical gyre; care must be taken to avoid overgeneralizing
-these results.  We plan to report on the geographic variability of submesoscale
+and the middle of the subtropical gyre.  We plan to report on the geographic variability of submesoscale
  seasonality in a future study.

### 2016-07-06 writeup/rocha_etal.tex
Fixes small ambiguity.

@@
 \section{Statistics of the surface lateral velocity tensor}
-To study the seasonality in the surface velocity, we calculate the lateral velocity tensor
+To study the seasonality in the surface velocity, we calculate the lateral velocity gradient tensor
 \begin{equation}
@@
 \end{equation}
-The components of the lateral velocity tensor are calculated using a centered
+The components of the lateral velocity gradient tensor are calculated using a centered
 second-order finite difference scheme. We then diagnose

### 2016-07-08 writeup/rocha_etal.tex
Small fix.
Adds Bilz edits

@@
 
-\title{Seasonality in governing submesoscale dynamics in the Kuroshio Extension}
+\title{Seasonality of submesoscale dynamics in the Kuroshio Extension}
 %
@@
 \begin{abstract}
-    Two new high-resolution numerical simulations with embedded tides suggest a
+    Two new high-resolution numerical simulations with embedded tides show a
     strong modulation in the governing near-surface dynamics at submesoscales
@@
 callies_etal2015, thompson_etal2016,buckingham_etal2016}. Contemporary studies
-have also suggested that inertia-gravity waves contribute significantly
+have also shown that inertia-gravity waves contribute significantly
 to the near-surface variability at submesoscales \citep{richman_etal2012,
@@
 submesoscale turbulence dominates the upper-ocean dynamics in late winter/early ,
-whereas quasi-linear inertia-gravity waves prevail in late summer/early fall.
+whereas submesoscale inertia-gravity waves prevail in late summer/early fall.
 Because submesoscale turbulence is weakest in late summer/early fall,
-the present results suggest that inertia-gravity waves account for most of the
+the present results indicate that inertia-gravity waves account for most of the
 summertime submesoscale SSH variability.
@@
 the LLC2160 simulation, spanning one year from September 2011 to September 2012.
-The LLC4320 simulation is an extension of the 5-month long output used by
+The LLC4320 simulation is an extension of the 3-month long output used by
 \citet{rocha_etal2016}.
@@
 \citep[e.g., ][]{rocha_etal2016}, tidal forcing fundamentally distinguishes our analysis
-from recent modeling studies that investigated upper-ocean dynamics
-and its seasonality \citep{sasaki_etal2014,qiu_etal2014}. Details of the LLC simulations
+from other modeling studies of seasonality \citep{sasaki_etal2014,qiu_etal2014}. Details of the LLC simulations
 are provided in the supplemental material.
@@
 Pacific, in the vicinity of the Kuroshio
-Extension, where previous studies have suggested strong submesoscale seasonality
+Extension, where previous studies have presented strong mesoscale and submesoscale seasonality
 \citep{sasaki_etal2014,qiu_etal2014}.
@@
 \begin{equation}
-\nabla_h \bf{u}_h = \left[\begin{matrix} u_x & u_y\\ v_x&v_y \end{matrix}\right] \,.
+\left[\begin{matrix} u_x & u_y\\ v_x&v_y \end{matrix}\right]
 \end{equation}
-The components of the lateral velocity gradient tensor are calculated using a centered
+using a centered
 second-order finite difference scheme. We then diagnose
-the vertical vorticity $\zeta \equiv v_x - u_y$, lateral rate of strain
-$\alpha \equiv [(u_x-v_y)^2 + (u_y + v_x)^2]^{1/2}$, and horizontal divergence $\delta \equiv u_x + v_y$.
+the vertical vorticity, and horizontal divergence
+\begin{equation}
+\zeta \equiv v_x - u_y\, ,
+\end{equation}
+lateral rate of strain
+\begin{equation}
+  \alpha \equiv [(u_x-v_y)^2 + (u_y + v_x)^2]^{1/2}\,,\\
+\end{equation}
+and horizontal divergence
+\begin{equation}
+\delta \equiv u_x + v_y\, .
+\end{equation}
 These diagnostics highlight the submesoscale structures in the flow
@@
 its local maximum and RMS are both smaller than $0.5f$.
-Indeed, the seasonal cycle of vorticity and rate of strain are strong winter~\ref{fig2}):
+Indeed, the vorticity and rate of strain are strong in winter (Figure \ref{fig2}):
 in both simulations, the RMS vorticity and strain are about twice as large in
@@
 a $-3$ power-law in winter), increasing the resolution from
-1/24$^\circ$ to 1/48$^\circ$ significantly increases the wintertime RMS vorticity
+1/24$^\circ$ to 1/48$^\circ$ increases the wintertime RMS vorticity
 and strain by about 40$\%$.
@@
 divergent flows, but also reduces the wintertime lateral divergence by about 50$\%$.
-The vorticity and strain rate winter \ref{fig2}a-b) show that most of the lateral divergence is associated
+Figure \ref{fig2}c) also shows that most of the lateral divergence is associated
 with submesoscale flows: smoothing the velocity fields with a 100km-cut-off
@@
 \section{Seasonality in governing submesoscale dynamics}
-The overall results of figure \ref{fig2}c suggest that submesoscale surface
- variability stems from different dynamics in summer than winter.
+The results of figure \ref{fig2}a-c show that submesoscale surface
+ variability stems from different dynamics in summer than in winter.
  To characterize these differences, we calculate
 joint probability distributions (jPDF) of
-vorticity-strain and vorticity-Laplacian of sea-surface height (\ref{fig3}).
+vorticity-strain and vorticity-Laplacian of sea-surface height (Figure \ref{fig3}).
 
@@
 
- Time series of vorticity and divergence PDFs (supplemental material) suggest
+ Time series of PDFs of vorticity and divergence (supplemental material) show
  a strong oscillation between these two regimes. In late winter/early spring
@@
 Figure \ref{fig4}a depicts the horizontal wavenumber spectra of surface KE in
-April and October. At scales larger than 20km, the spectra based on hourly
-velocity snapshots are nearly indistinguishable within 95$\%$ confidence level
-(compare solid lines in figure \ref{fig4}a). Consistent with the results of
+April and October. At scales larger than 20km, the April and October spectra based on hourly (the solid curves)
+velocity snapshots are nearly indistinguishable from each other within 95$\%$ confidence level.
+Consistent with the results of
 \citet{rocha_etal2016} who analyzed a 5-month output of the LLC4320 simulation in Drake Passage,
@@
 The inertia-gravity waves significantly project on the sea-surface.
-In October, when the inertia-gravity waves are strongest the surface and submesoscale
-turbulence is weakest. There is a dramatic difference between the spectra
-based on hourly and daily-averaged SSH: at scales smaller than 100 km, the spectra
+There is a dramatic difference between the spectra
+based on hourly and daily-averaged SSH in October (Figure \ref{4}b): at scales smaller than 100 km, the spectra
 of hourly SSH roughly follows a $-2$ power-law, whereas the spectra of daily-averaged
@@
 \section{Summary and Conclusion}
-Our main finding is that in global simulations with embedded tides the
+Our work adds to recent studies that presented modeling \citep{sasaki_etal2014}
+and observational \citep{callies_etal2015,buckingham_etal2016} evidence of vigorous seasonality in
+submesoscale turbulence. Our main finding is that in global simulations with embedded tides the
 near-surface inertia-gravity waves in the Kuroshio Extension undergo
 a strong seasonal cycle that is out-of-phase with the seasonal cycle of
-submesoscale turbulence. \cite{dasaro1978} showed that velocity of linear internal waves
+submesoscale turbulence. We conjecture that the summertime dominance of inertia-gravity waves
+ \citep{callies_etal2015} is a consequence both of suppression of submesoscale turbulence and
+enhancement of inertia-gravity waves due to re-stratification of
+the upper ocean.
+
+\cite{dasaro1978} showed that velocity of linear internal waves
 in the mixed-layer strongly depends on the density jump at the mixed-layer
@@
 Internal tide generation does not show a seasonal modulation \citep[e.g.,][]{alford2003},
-and therefore it is plausible to expect a strong seasonality in the near-surface
+and therefore one expects a strong seasonality in the near-surface
 expression of internal tides and
@@
 
-Our work adds to recent studies that presented modeling \citep{sasaki_etal2014}
-and observational \citep{callies_etal2015,buckingham_etal2016} evidence of vigorous seasonality in
-submesoscale turbulence. We conjecture that the summertime dominance of inertia-gravity waves
- \citep{callies_etal2015} is a consequence both of suppression of submesoscale turbulence and
-enhancement of inertia-gravity waves due to re-stratification of
-the upper ocean. Of course, this study focuses on a single patch of ocean in the
+An important caveat is that this study focuses on a single patch of ocean in the
 vicinity of the Kuroshio Extension, which may be typical of mesoscale-rich

### 2016-07-08 writeup/rocha_etal.tex
Updates figure3

@@
 
+\begin{figure*}[ht]
+  \begin{center}
+    \includegraphics[width=.95\textwidth]{figs/fig3.pdf}
+ \caption{Seasonal variation of joint probability distributions:  vorticity vs. strain rate (a through d).
+          and vorticity vs. Laplacian of sea-surface height (e through h) in April (a, b,
+          e, f) and October (c, d, g, h).
+          Dashed lines in (a) through (d) represent pure strain flow $\alpha = \pm\zeta$,
+          characteristic of fronts. Dashed lines in (e) through
+          (h) represent geostrophic flow $\zeta = \frac{g}{f}\nabla_h^2 \eta$.}
+ \label{fig3}
+ \end{center}
+\end{figure*}
+
 \section{Seasonality in governing submesoscale dynamics}
@@
 
-\begin{figure*}[ht]
-  \begin{center}
-    \includegraphics[width=1.\textwidth]{figs/fig3.pdf}
- \caption{Seasonal variation of joint probability distributions:  vorticity vs. strain rate (a through f).
-          and vorticity vs. Laplacian of sea-surface height (g through l) in April (a through c;
-          g  through i) and October (d through f;
-          j  through l).
-          Dashed lines in (a) through (e) represent pure strain flow $\alpha = \pm\zeta$,
-          characteristic of fronts. Dashed lines in (f) through
-          (l) represent geostrophic flow $\zeta = \frac{g}{f}\nabla_h^2 \eta$.}
- \label{fig3}
- \end{center}
-\end{figure*}
-
 The October vorticity-strain jPDF shows much weaker skewness (the vorticity skewness
@@
   \begin{center}
-    \includegraphics[width=.9\textwidth]{figs/fig4.pdf}
+    \includegraphics[width=.95\textwidth]{figs/fig4.pdf}
  \caption{Surface (horizontal) KE (a) and SSH variance wavenumber spectra (b)

### 2016-07-08 writeup/rocha_etal.tex
Fixes some figure referencing

@@
 The April vorticity-strain jPDF has a shape characteristic of submesoscale turbulence
- (see figures \ref{fig3}a-c). The alignment of vorticity and strain $\alpha \sim \pm\zeta$
+ (Figures \ref{fig3}a-b). The alignment of vorticity and strain $\alpha \sim \pm\zeta$
  with strong positive skewness are fingerprints of submesoscale fronts
  \citep{shcherbina_etal2013,mcwilliams2016}. The shape of the vorticity-strain
- jPDF is similar for hourly, daily-averaged, and residual velocities, although
+ jPDF is similar for hourly, daily-averaged, although
  the vorticity skewness reduces from 1.4 to 1.13 from hourly to daily-averaged.
@@
 ageostrophic as depicted by the jPDF of vorticity-Laplacian of SSH
-winter \ref{fig3}g). Even in April, only the daily-averaged fields
+winter (Figure \ref{fig3}e). Even in April, only the daily-averaged fields
 are largely
 in  geostrophic balance (the jPDF of daily-averaged vorticity vs. Laplacian of SSH
-is an ellipse with large eccentricity and main axis tilted by 45$^\circ$; figure \ref{fig3}h).
-
-%While there is certainly much to be
-%learned using quasi-geostrophic models to study submesoscale turbulence, claims that
-%the submesoscale turbulence is dominantly in geostrophic balance
-%are overstated \citep{sasaki_etal2014,callies_etal2016};
+is an ellipse with large eccentricity and main axis tilted by 45$^\circ$; Figure \ref{fig3}f).
 
@@
 That the submesoscale dynamics in October are mainly unbalanced is clearly depicted
-in the shape of jPDF of vorticity-Laplacian of SSH for hourly and
-residual fields
-winter \ref{fig3}j and \ref{fig3}k), which is an ellipse aligned in the vertical axis.
+in the shape of jPDF of vorticity-Laplacian of SSH for hourly fields
+winter (Figure \ref{fig3}g), which is an ellipse aligned in the vertical axis.
 Of course daily-averaging the model suppresses the ageostrophic flows, and

### 2016-07-08 writeup/rocha_etal.tex
Fix legend
Fixes typo.

@@
 
-The inertia-gravity waves significantly project on the sea-surface.
+The inertia-gravity waves project on the sea-surface.
 There is a dramatic difference between the spectra
-based on hourly and daily-averaged SSH in October (Figure \ref{4}b): at scales smaller than 100 km, the spectra
+based on hourly and daily-averaged SSH in October (Figure \ref{fig4}b): at scales smaller than 100 km, the spectra
 of hourly SSH roughly follows a $-2$ power-law, whereas the spectra of daily-averaged
@@
   (green) and 1/48$^\circ$ (purple). The statistical errorbars represent conservative
-  significance levels considering that there are only four independent relations
-  for the monthly spectra.}
+  95$\%$ significance levels.}
   \label{figS2_3}

### 2016-07-08 writeup/rocha_etal.tex
Fix typo
Remove one adverb

@@
                   sea-surface height variability in late winter/early spring.
-            \item Inertia-gravity waves dominate the horizontal velocity and
+            \item Submesoscale inertia-gravity waves dominate the horizontal velocity and
                    sea-surface height variability in late summer/early fall.
@@
 flows as opposed to 83$\%$ in October! Curiously, the out-of-phase seasonal cycle of submesoscale
-turbulence and near-surface inertia-gravity waves conspire to yield very weak seasonality
+turbulence and near-surface inertia-gravity waves conspire to yield weak seasonality
 in the spectra of KE and SSH variance based on hourly fields.

### 2016-07-08 writeup/rocha_etal.tex
adds reference
Tiny change.

@@
   \begin{center}
-    \includegraphics[width=.95\textwidth]{figs/fig4.pdf}
+    \includegraphics[width=1.\textwidth]{figs/fig4.pdf}
  \caption{Surface (horizontal) KE (a) and SSH variance wavenumber spectra (b)
@@
 and observational \citep{callies_etal2015,buckingham_etal2016} evidence of vigorous seasonality in
-submesoscale turbulence. Our main finding is that in global simulations with embedded tides the
+submesoscale turbulence. In particular, our main finding is that in global simulations with embedded tides the
 near-surface inertia-gravity waves in the Kuroshio Extension undergo
@@
 other small-scale internal waves generated through internal tide interactions.
-Near-inertial wave generation peaks in winter, but those waves project on large
-horizontal scales \citep[e.g, ][]{qi_etal1995}.
+Near-inertial wave generation peaks in winter \citep[e.g.,][]{alford2003}, but
+those waves project on large horizontal scales \citep[e.g, ][]{qi_etal1995}.
 

### 2016-07-08 writeup/rocha_etal.tex
Fixes typo

@@
 An important caveat is that this study focuses on a single patch of ocean in the
-vicinity of the Kuroshio Extension, which may be typical of mesoscale-rich
-subtropical regions, but is unlikely to be
+vicinity of the Kuroshio Extension. This may be a typical of mesoscale-rich
+subtropical region, but it is unlikely to be
 representative of other regions such as low eddy kinetic energy eastern boundary currents

### 2016-07-09 writeup/rocha_etal.tex
Minor changes

@@
 
-\authors{Cesar B. Rocha et al. \altaffilmark{1}}
+\authors{Cesar B. Rocha\altaffilmark{1}, Sarah T. Gille\altaffilmark{1},
+        Teresa K. Chereskin\altaffilmark{1}, and Dimitris Menemenlis\altaffilmark{2}}
 % Eric Brown,\altaffilmark{1,2} Rick Williams,\altaffilmark{3}
@@
 \altaffiltext{1}{Scripps Institution of Oceanography, University of California San Diego, La Jolla, CA, USA}
+\altaffiltext{2}{Earth Sciences Division, Jet Propulsion Laboratory, California Institute of Technology, Pasadena, CA, USA}
 
@@
     prevails in late winter/early spring. While summertime
-    re-stratification weakens submesoscale turbulence, it also enhances inertia-gravity
+    re-stratification weakens submesoscale turbulence, it also enhances submesoscale inertia-gravity
      waves near the surface. In the Kuroshio Extension,
@@
           e, f) and October (c, d, g, h).
-          Dashed lines in (a) through (d) represent pure strain flow $\alpha = \pm\zeta$,
+          Dashed lines in (a) through (d) represent one dimensional shear flow $\alpha = \pm\zeta$,
           characteristic of fronts. Dashed lines in (e) through
@@
 \begin{acknowledgments}
-We thank the MITgcm community and our colleagues at the NASA Advanced
+ William R. Young provided feedback on the first draft. We thank the MITgcm community and our colleagues at the NASA Advanced
 Supercomputing (NAS) Division for their awesome support.
-This research was funded by NASA and NSF.
-The altimeter products were produced by Ssalto/Duacs
+This research was funded by NSF (OCE1357047) and NASA (NNX13AE44G,NNX13AE85G,NNX16AH67G).
+Access to LLC outputs can be obtained directly from D. Menemenlis. The altimeter products were produced by Ssalto/Duacs
 and distributed by Aviso, with support from CNES (\texttt{http://www.aviso.altimetry.fr/duacs/}).

### 2016-07-10 writeup/rocha_etal.tex
Supplemental material now in a different file

@@
 %Corresponding author mailing address and e-mail address:
-%\authoraddr{Corresponding author: A. B. Smith,
-%Department of Hydrology and Water Resources, University of
-%Arizona, Harshbarger Building 11, Tucson, AZ 85721, USA.
-%(a.b.smith@hwr.arizona.edu)}
+\authoraddr{Corresponding author: Cesar B. Rocha,
+Scripps Institution of Oceanography, University of California San Diego, La Jolla, CA, USA.
+(crocha@ucsd.edu)}
 
@@
 \begin{acknowledgments}
- William R. Young provided feedback on the first draft. We thank the MITgcm community and our colleagues at the NASA Advanced
+ William R. Young provided helpful feedback on the first draft. We thank the MITgcm community and our colleagues at the NASA Advanced
 Supercomputing (NAS) Division for their awesome support.
 This research was funded by NSF (OCE1357047) and NASA (NNX13AE44G,NNX13AE85G,NNX16AH67G).
-Access to LLC outputs can be obtained directly from D. Menemenlis. The altimeter products were produced by Ssalto/Duacs
+The LLC outputs can be obtained from D. Menemenlis (menemenlis@jpl.nasa.gov). The altimeter products were produced by Ssalto/Duacs
 and distributed by Aviso, with support from CNES (\texttt{http://www.aviso.altimetry.fr/duacs/}).
@@
 %\section{Supplemental material 1: Spectral errors}
-
-\newpage
-\section*{Details of the LLC simulations}
-
-Table \ref{tab:llc} shows details of the LLC spin-up hierarchy.
-Both simulations analyzed in the present paper, LLC2160 and LLC4320,
-were forced by surface fluxes from the 0.14$^\circ$ European Centre for
-Medium-range Weather Forecasting (ECMWF) atmospheric operational model analysis,
- starting in 2011, and the 16 most significant  tidal components.
-The LLC control files are available online
-(at \texttt{http://mitgcm.org/viewvc/MITgcm/MITgcm$\_$contrib/
-llc$\_$hires}).
-
-For the LLC hierarchy, the sea-surface height is calculated
-using a linearized equation. While the nominal horizontal resolution  of the
-LLC4320 simulation is 1/48$^\circ$,
-horizontal wavenumber spectra suggests an effective resolution of about 8 km.
-(see figure \ref{fig4}), and wavenumber spectra suggests an effective resolution
-of 20 km for  LLC2160 simulation.
-
-The spin-up of the upper-ocean in the LLC4320 (1/48$^\circ$) simulation is
-a couple of days, as depicted the time series of surface horizontal divergence
-in Figure 2b of the main text (see initial rapid increase in RMS divergence in
-September 2011). To avoid the spin-up stage, we chose October to characterize the
-late summer/early fall. The late winter/early spring waves chosen to be 6 months
-out-of-phase with late summer/early fall (hence April).
-
-\begin{table*}[ht]
-\label{tab:llc}
-\caption{\small The MITgcm Latitute-Longitude Cap (LLC) spin-up hierarchy. This study analyzes
-         the output of LLC2160 and LLC4320. Model fields are available hourly.
-         Adapted from \cite{rocha_etal2016}.}
-\begin{center}
-\begin{tabular}{ | c | c | c | c | c |}
-\hline
-Simulation & Resolution & Time-step & Period  & Tides \\ \hline
-ECCO2 Adjoint & $1/6^{^\circ}$ & 1200 s & January 2009 - December 2011 & No\\
-LLC 1080   & $1/12^{^\circ}$ & 90 s & January 2010 - July 2012 & No\\
-LLC 2160   &  $1/24^{^\circ}$ & 45 s  & January 2011 - April 2013 & Yes\\
-LLC 4320   &  $1/48^{^\circ}$ & 25 s & September 2011 - September 2012 & Yes \\
-\hline
-\end{tabular}
-\end{center}
-\end{table*}
-
-\subsection*{Spectra}
-Kinetic energy and sea-surface height variance wavenumber spectra for the LLC2160
-and LLC4320 simulations are shown in figure \ref{figS2_3}. As expected, and
-consistent with the submesoscale regional study by \cite{capet_etal2008a}, the
-higher-resolution simulation (LLC4320) essentially extends the spectra of the
-lower-resolution simulation (LLC2160) towards smaller scale, demonstrating that
-the higher resolution is resolving a wider array of scales. Between 20 and 100 km
-the spectra of the two simulations are visually indistinguishable.
-
-\begin{figure*}[ht]
-   \begin{center}
-     \includegraphics[width=1.\textwidth]{figs/fig_s2_3.pdf}
-  \caption{A comparison between spectra of the two LLC simulations 1/24$^\circ$
-  (green) and 1/48$^\circ$ (purple). The statistical errorbars represent conservative
-  95$\%$ significance levels.}
-  \label{figS2_3}
-  \end{center}
-\end{figure*}
-
-\subsection*{Seasonality in probability density functions}
-
-Probability density functions (pdfs) of vertical vorticity and horizontal divergence
- at the surface (Figure \ref{figS2_2}) depicts the strong seasonality in the
- submesoscale dynamics. Time series of vorticity pdf indicate that
- the development of positive skewness in winter (Figures \ref{figS2_2}a-b) both in
- hourly ($\sim 1.4$) and daily-averaged ($\sim1.1$) fields. The skewness
- in divergence for hourly fields is very small year-round ($\sim -0.24$), suggesting the
- prevalence of inertia-gravity waves. Filtering out the waves by daily-averaging
- the velocity field increases the skewness in divergence, particularly in winter ($\sim-0.64$). Both
-positive vorticity (cyclonic) skewness and negative divergence (convergence) skewness
-are characteristics of submesoscale turbulence \citep{mcwilliams2016}.
-
-\begin{figure*}[ht]
-   \begin{center}
-     \includegraphics[width=1.\textwidth]{figs/fig_s2_2.pdf}
-  \caption{Time dependence of probability density functions of vertical vorticity (a, b)
-  and horizontal divergence (c, d). Monthly averages in April and October are
-  also shown (e, f).}
-  \label{figS2_2}
-  \end{center}
-\end{figure*}
-
-\subsection*{Seasonality in the upper-ocean stratification}
-
-Figure \ref{figS2_1} shows a comparison between potential density sections
-at 165$^\circ$E from
-the Roemmich-Gilson Argo Climatology \citep[updated from][]{roemmich_gilson2009}.
-and the LLC2160 and LLC4320 simulations. Of course, the Argo climatology is smoother
-owing to lower resolution and long-term averaging. Nonetheless, simulations reproduce well the strong
-seasonality in the stratification. The upper ocean is weakly stratified in late winter/early fall
-and re-stratifies in summer. Argo climatology and LLC simulations mixed-layer depths are consistent
-at all seasons (not shown).
-
-\begin{figure*}[ht]
-   \begin{center}
-     \includegraphics[width=1.\textwidth]{figs/fig_s2_1.pdf}
-  \caption{Potential density: a comparison between Roemmich-Gilson Argo
-          climatology (a, b) and the LLC simulations 1/24$^\circ$ (c, d)
-          and  1/48$^\circ$ (e, f).}
-  \label{figS2_1}
-  \end{center}
-\end{figure*}
-
-\subsection*{Depth-dependence}
-Figure \ref{figS2_2} shows the depth-dependence of the  root-mean-square of the velocity
-gradient lateral tensor for the LLC4320 simulation. The strong seasonality, as depicted
-at the surface in Figure 2 of the main manuscript, is confined to the mixed layer, which
-varies from 50 m in summer to about 350 m in winter.
-
-\begin{figure*}[ht]
-   \begin{center}
-     \includegraphics[width=.9\textwidth]{figs/fig_s2_2.png}
-  \caption{Depth-dependence of root-mean-square of vertical vorticity (a,d), rate of lateral strain (b,e)
-  and horizontal divergence (c,f) for hourly (a,b,c) and daily-averaged (d,e,f,) fields.}
-  \label{figS2_2}
-  \end{center}
-\end{figure*}
-
-\subsection*{How well do the LLC simulations capture high-frequency modes?}
-To assess how well the LLC simulations represents the high-frequency variability,
-we compare the model near-surface velocity field against two available
-mooring data: KEO (the data is available online at \texttt{http://www.pmel.noaa.gov/ocs/KEO})
-and KESS mooring $\#$7 (the data is available online at \texttt{http://uskess.whoi.edu}).
-For proper comparison, we use model data closest to these two moorings (west of
-the region considered in the present study). Focusing on high-frequencies, we
-calculate frequency spectral estimates every month. There are about 48 independent
-spectral realizations.
-
-Figure \ref{figA1} shows KE frequency spectra for the LLC simulations and mooring data.
-At the uppermost common depth (20 m), very low frequencies (periods $>$ 10 days)
-and the inertial, diurnal, and semi-diurnal frequencies agree well --- the KESS
-data are more energetic at intermediate frequencies. At 40 m, the LLC simulations and observations
-have consistent spectra at periods larger than about 6 h. The data from the moorings
-are more energetic at higher frequencies,
-except for the first two harmonics that are apparent in the KESS data at 40 m.
-There are at least three reasons for the slightly more energetic super-inertial
-variability in the observations: 1) The high-
-frequency variability is dominated by high-frequency noise due to mooring vertical
-excursions; 2) The model does not have enough resolution to resolve the full internal
-wave spectrum; 3) The model is not forced at frequencies higher than 6h.
-
-\begin{figure*}[ht]
-   \begin{center}
-     \includegraphics[width=.9\textwidth]{figs/figA1.pdf}
-  \caption{A comparison between frequency spectra of LLC simulations and two
-  available moored current meter records.}
-  \label{figA1}
-  \end{center}
-\end{figure*}
-
-\section*{Seasonality in pressure modes near-surface amplitude}
-
-Figure \ref{figS3}a shows the domain-averaged potential density from the WOA 2013
-climatology in April and October. Clearly, the changes to the stratification
-are dramatic, with the formation of a seasonal pycnocline in summer. The strong
-seasonality of upper-ocean stratification yields a strong seasonality in the
-near-surface shape and amplitude of baroclinic pressure modes (\ref{figS3}b).
-In particular, the surface amplitude is much large in summer. Thus the projection
-of baroclinic tides on the surface may vary significantly seasonally.
-
-%\textcolor{red}{Perhaps include analytical results from a model with a slab over a slowly varying
-%pycnocline.}
-
-\begin{figure*}[ht]
-   \begin{center}
-     \includegraphics[width=1.\textwidth]{figs/fig_s3.pdf}
-  \caption{The seasonal variability of the WOA 2013 stratification averaged  over
-  the domain and the associated three gravest pressure modes. Only the upper 450 m is shown.}
-  \label{figS3}
-  \end{center}
-\end{figure*}
 

### 2016-07-10 writeup/rocha_etal.tex
Small fixes.

@@
 
-\altaffiltext{1}{Scripps Institution of Oceanography, University of California San Diego, La Jolla, CA, USA}
-\altaffiltext{2}{Earth Sciences Division, Jet Propulsion Laboratory, California Institute of Technology, Pasadena, CA, USA}
+\altaffiltext{1}{Scripps Institution of Oceanography, University of California San Diego, La Jolla, CA, USA.}
+\altaffiltext{2}{Earth Sciences Division, Jet Propulsion Laboratory, California Institute of Technology, Pasadena, CA, USA.}
 
@@
 strong seasonal modulation in the dominant upper-ocean submesoscale dynamics:
-submesoscale turbulence dominates the upper-ocean dynamics in late winter/early ,
+submesoscale turbulence dominates the upper-ocean dynamics in late winter/early,
 whereas submesoscale inertia-gravity waves prevail in late summer/early fall.
@@
 We analyze a sub-domain of the LLC4320 and LLC2160 simulations of about 2000 km$^2$
-spanning 155-175$^\circ$E; 25-40$^\circ$N winter \ref{fig1}a). The stratification
+spanning 155-175$^\circ$E; 25-40$^\circ$N (Figure \ref{fig1}). The stratification
 in this mesoscale-rich subtropical region undergoes a vigorous seasonal cycle: wintertime-enhanced
@@
 second-order finite difference scheme. We then diagnose
-the vertical vorticity, and horizontal divergence
+the vertical vorticity
 \begin{equation}
@@
 its local maximum and RMS are both smaller than $0.5f$.
-Indeed, the vorticity and rate of strain are strong in winter (Figure \ref{fig2}):
-in both simulations, the RMS vorticity and strain are about twice as large in
+Indeed, the vorticity and rate of strain are strongest in wintertime (Figure \ref{fig2}):
+in both simulations, the RMS vorticity and strain rate are about twice as large in
 late winter/early spring
@@
 ($T_f = 2\pi/f_0\approx 23.5$h, where $f_0$ is the inertial frequency at
-the mean latitude): daily-averaging the velocity fields suppresses superinertial
+the mean latitude): daily-averaging the velocity fields suppresses super-inertial
 motions and reduces the RMS
-vorticity by less than 40$\%$ and the RMS strain by less than 10$\%$; the seasonal
+vorticity by 40$\%$ and the RMS strain by 10$\%$; the seasonal
 cycle remains strong (see
@@
 with a Hanning filter with cut-off scale of 100 km dramatically reduces the RMS
-vorticity and strain. The reduction in variance is about 80$\%$ in winter,
-yielding RMS vorticity and strain roughly consistent with the diagnostics from
+vorticity and strain rate. The reduction in variance is about 80$\%$ in winter,
+yielding RMS vorticity and strain  rate roughly consistent with the diagnostics from
 AVISO gridded geostrophic velocities (compare red lines to black lines in figures \ref{fig2}a-b).
@@
 divergent flows, but also reduces the wintertime lateral divergence by about 50$\%$.
-Figure \ref{fig2}c) also shows that most of the lateral divergence is associated
+Figure \ref{fig2}c also shows that most of the lateral divergence is associated
 with submesoscale flows: smoothing the velocity fields with a 100km-cut-off
@@
  The April results are characteristic of winter, indicating that wintertime submesoscale
- surface velocity is strongly dominated by submesoscale turbulence --- this is true even for
- superinertial submesoscale currents. The hourly velocity fields are significantly
+ surface velocity is strongly dominated by submesoscale turbulence.
+ The hourly velocity fields are significantly
 ageostrophic as depicted by the jPDF of vorticity-Laplacian of SSH
@@
 winter (Figure \ref{fig3}g), which is an ellipse aligned in the vertical axis.
-Of course daily-averaging the model suppresses the ageostrophic flows, and
+Of course daily-averaging the model suppresses the ageostrophic, super-inertial flows, and
 the daily-averaged flow is essentially geostrophic as depicted by the 45$^\circ$-tilted
@@
  the vorticity is strongly positively skewed, whereas the divergence is moderately
- negatively skewed as predicted by frontogenesis \citep{capet_etal2008a,mcwilliams2016}
+ negatively skewed as predicted by frontogenesis \citep[e.g., ][]{capet_etal2008a,mcwilliams2016}
 . In late summer/early fall, the divergence is stronger, but PDFs are much less skewed,
@@
 Consistent with the results of
-\citet{rocha_etal2016} who analyzed a 5-month output of the LLC4320 simulation in Drake Passage,
+\citet{rocha_etal2016}, who analyzed a 5-month output of the LLC4320 simulation in Drake Passage,
 there is significant high-frequency variabilty at submesoscales. Daily-averaging
@@
 flows as opposed to 83$\%$ in October! Curiously, the out-of-phase seasonal cycle of submesoscale
-turbulence and near-surface inertia-gravity waves conspire to yield weak seasonality
+turbulence and near-surface submesoscale inertia-gravity waves conspire to yield weak seasonality
 in the spectra of KE and SSH variance based on hourly fields.
@@
 submesoscale turbulence. In particular, our main finding is that in global simulations with embedded tides the
-near-surface inertia-gravity waves in the Kuroshio Extension undergo
+near-surface submesoscale inertia-gravity waves in the Kuroshio Extension undergo
 a strong seasonal cycle that is out-of-phase with the seasonal cycle of
@@
 
-\cite{dasaro1978} showed that velocity of linear internal waves
+\cite{dasaro1978} showed that the velocity of linear internal waves
 in the mixed-layer strongly depends on the density jump at the mixed-layer
@@
 expression of internal tides and
-other small-scale internal waves generated through internal tide interactions.
+other submesoscale inertia-gravity waves generated through internal tide interactions.
 Near-inertial wave generation peaks in winter \citep[e.g.,][]{alford2003}, but

### 2016-07-10 writeup/rocha_etal.tex
Small fixes

@@
 Pacific, in the vicinity of the Kuroshio
-Extension, where previous studies have presented strong mesoscale and submesoscale seasonality
+Extension, where previous studies have shown strong mesoscale and submesoscale seasonality
 \citep{sasaki_etal2014,qiu_etal2014}.
@@
     \includegraphics[width=.95\textwidth]{figs/fig3.pdf}
- \caption{Seasonal variation of joint probability distributions:  vorticity vs. strain rate (a through d).
+ \caption{Seasonal variation of joint probability distributions:  vorticity vs. strain rate (a through d),
           and vorticity vs. Laplacian of sea-surface height (e through h) in April (a, b,
@@
 \section{Summary and Conclusion}
-Our work adds to recent studies that presented modeling \citep{sasaki_etal2014}
+Our work adds to recent studies that present modeling \citep{sasaki_etal2014}
 and observational \citep{callies_etal2015,buckingham_etal2016} evidence of vigorous seasonality in

### 2016-07-11 writeup/rocha_etal.tex
Fixes some minor stuff

@@
     Two new high-resolution numerical simulations with embedded tides show a
-    strong modulation in the governing near-surface dynamics at submesoscales
+    strong modulation of near-surface dynamics at submesoscales
     (roughly 10-100 km) in the Kuroshio Extension. Consistent with recent studies, deep late-winter mixed layers
@@
 cycle of shallow baroclinic instabilities and their role in submesoscale (roughly 1-100 km)
-turbulence and mesoscale modulation \citep{sasaki_etal2014,qiu_etal2014,
+turbulence and mesoscale (roughly 100-300 km) modulation \citep{sasaki_etal2014,qiu_etal2014,
 callies_etal2015, thompson_etal2016,buckingham_etal2016}. Contemporary studies
@@
 when the upper ocean is strongly stratified. Thus there exists a
-strong seasonal modulation in the dominant upper-ocean submesoscale dynamics:
+strong seasonal modulation of upper-ocean submesoscale dynamics:
 submesoscale turbulence dominates the upper-ocean dynamics in late winter/early,
@@
 The LLC4320 simulation is an extension of the 3-month long output used by
-\citet{rocha_etal2016}.
+\citet{rocha_etal2016}.  Details of the LLC simulations are provided in the supplemental material.
 
-A key aspect of the LLC simulations is that they were forced by
+A key aspect of the LLC2160 and LLC4320 simulations is that they were forced by
 the 16 most-significant tidal components.
@@
 \citep[e.g., ][]{rocha_etal2016}, tidal forcing fundamentally distinguishes our analysis
-from other modeling studies of seasonality \citep{sasaki_etal2014,qiu_etal2014}. Details of the LLC simulations
-are provided in the supplemental material.
+from other modeling studies of seasonality \citep{sasaki_etal2014,qiu_etal2014}.
 
@@
 cycle remains strong (see
-dashed lines in figures \ref{fig2}a-b). Most of this seasonal cycle is associated
+dashed lines in figures \ref{fig2}a-b). Indeed, most of this seasonal cycle is associated
 with submesoscale flows: smoothing the velocity fields
@@
 instabilities energize the submesoscales in late winter, drawing from the available
-potential energy stored in deep mixed layers \citep{sasaki_etal2014,callies_etal2015,callies_etal2016}.
+potential energy stored in large lateral buoyancy gradients in deep mixed layers \citep{sasaki_etal2014,callies_etal2015,callies_etal2016}.
 
@@
 
-\section{Seasonality in governing submesoscale dynamics}
+\section{Seasonality of submesoscale dynamics}
 The results of figure \ref{fig2}a-c show that submesoscale surface
@@
 Consistent with the results of
-\citet{rocha_etal2016}, who analyzed a 5-month output of the LLC4320 simulation in Drake Passage,
+\citet{rocha_etal2016}, who analyzed a 3-month output of the LLC4320 simulation in Drake Passage,
 there is significant high-frequency variabilty at submesoscales. Daily-averaging
@@
 \cite{dasaro1978} showed that the velocity of linear internal waves
-in the mixed-layer strongly depends on the density jump at the mixed-layer
+in the mixed layer strongly depends on the density jump at the mixed-layer
 base, with largest mixed layer velocities when the jump is strongest.
@@
 \bibliographystyle{agufull08}
-\bibliography{rocha_etal}
+%\bibliography{rocha_etal}
 
@@
 % Text
-% \end{thebibliography}
-%
+\begin{thebibliography}{16}
+\providecommand{\natexlab}[1]{#1}
+\expandafter\ifx\csname urlstyle\endcsname\relax
+  \providecommand{\doi}[1]{doi:\discretionary{}{}{}#1}\else
+  \providecommand{\doi}{doi:\discretionary{}{}{}\begingroup
+  \urlstyle{rm}\Url}\fi
+
+\bibitem[{\textit{Alford}(2003)}]{alford2003}
+Alford, M.~H. (2003), {Redistribution of energy available for ocean mixing by
+  long-range propagation of internal wave}, \textit{Nature},
+  \textit{423}(6936), 159--162.
+
+\bibitem[{\textit{Buckingham et~al.}(2016)\textit{Buckingham, Naveira~Garabato,
+  Thompson, Brannigan, Lazar, Marshall, George~Nurser, Damerell, Heywood, and
+  Belcher}}]{buckingham_etal2016}
+Buckingham, C.~E., A.~C. Naveira~Garabato, A.~F. Thompson, L.~Brannigan,
+  A.~Lazar, D.~P. Marshall, A.~George~Nurser, G.~Damerell, K.~J. Heywood, and
+  S.~E. Belcher (2016), {Seasonality of submesoscale flows in the ocean surface
+  boundary layer}, \textit{Geophysical Research Letters}, \textit{43}(5),
+  2118--2126.
+
+\bibitem[{\textit{B{\"u}hler et~al.}(2014)\textit{B{\"u}hler, Callies, and
+  Ferrari}}]{buhler_etal2014}
+B{\"u}hler, O., J.~Callies, and R.~Ferrari (2014), {Wave--vortex decomposition
+  of one-dimensional ship-track data}, \textit{Journal of Fluid Mechanics},
+  \textit{756}, 1007--1026.
+
+\bibitem[{\textit{Callies et~al.}(2015)\textit{Callies, Ferrari, Klymak, and
+  Gula}}]{callies_etal2015}
+Callies, J., R.~Ferrari, J.~M. Klymak, and J.~Gula (2015), {Seasonality in
+  submesoscale turbulence}, \textit{Nature communications}, \textit{6}.
+
+\bibitem[{\textit{Callies et~al.}(2016)\textit{Callies, Flierl, Ferrari, and
+  Fox-Kemper}}]{callies_etal2016}
+Callies, J., G.~Flierl, R.~Ferrari, and B.~Fox-Kemper (2016), {The role of
+  mixed-layer instabilities in submesoscale turbulence}, \textit{Journal of
+  Fluid Mechanics}, \textit{788}, 5--41.
+
+\bibitem[{\textit{Capet et~al.}(2008)\textit{Capet, McWilliams, Molemaker, and
+  Shchepetkin}}]{capet_etal2008a}
+Capet, X., J.~C. McWilliams, M.~J. Molemaker, and A.~Shchepetkin (2008),
+  {Mesoscale to submesoscale transition in the California Current System. Part
+  I: Flow structure, eddy flux, and observational tests}, \textit{Journal of
+  Physical Oceanography}, \textit{38}(1), 29--43.
+
+\bibitem[{\textit{D'Asaro}(1978)}]{dasaro1978}
+D'Asaro, E.~A. (1978), {Mixed layer velocities induced by internal wave},
+  \textit{Journal of Geophysical Research: Oceans}, \textit{83}(C5),
+  2437--2438.
+
+\bibitem[{\textit{Forget et~al.}(2015)\textit{Forget, Campin, Heimbach, Hill,
+  Ponte, and Wunsch}}]{forget_etal2015}
+Forget, G., J.-M. Campin, P.~Heimbach, C.~Hill, R.~Ponte, and C.~Wunsch (2015),
+  {ECCO version 4: an integrated framework for non-linear inverse modeling and
+  global ocean state estimation}, \textit{Geosci. Model Dev.}, \textit{8},
+  3071--3104.
+
+\bibitem[{\textit{McWilliams}(2016)}]{mcwilliams2016}
+McWilliams, J.~C. (2016), {Submesoscale currents in the ocean}, in
+  \textit{Proc. R. Soc. A}, vol. 472, p. 20160117, The Royal Society.
+
+\bibitem[{\textit{Qi et~al.}(1995)\textit{Qi, De~Szoeke, Paulson, and
+  Eriksen}}]{qi_etal1995}
+Qi, H., R.~A. De~Szoeke, C.~A. Paulson, and C.~C. Eriksen (1995), {The
+  structure of near-inertial waves during ocean storms}, \textit{Journal of
+  Physical Oceanography}, \textit{25}, 2853--2871.
+
+\bibitem[{\textit{Qiu et~al.}(2014)\textit{Qiu, Chen, Klein, Sasaki, and
+  Sasai}}]{qiu_etal2014}
+Qiu, B., S.~Chen, P.~Klein, H.~Sasaki, and Y.~Sasai (2014), {Seasonal mesoscale
+  and submesoscale eddy variability along the North Pacific Subtropical
+  Countercurrent}, \textit{Journal of Physical Oceanography}, \textit{44}(12),
+  3079--3098.
+
+\bibitem[{\textit{Richman et~al.}(2012)\textit{Richman, Arbic, Shriver,
+  Metzger, and Wallcraft}}]{richman_etal2012}
+Richman, J.~G., B.~K. Arbic, J.~F. Shriver, E.~J. Metzger, and A.~J. Wallcraft
+  (2012), {Inferring dynamics from the wavenumber spectra of an eddying global
+  ocean model with embedded tides}, \textit{Journal of Geophysical Research:
+  Oceans}, \textit{117}(C12).
+
+\bibitem[{\textit{Rocha et~al.}(2016)\textit{Rocha, Chereskin, Gille, and
+  Menemenlis}}]{rocha_etal2016}
+Rocha, C.~B., T.~K. Chereskin, S.~T. Gille, and D.~Menemenlis (2016),
+  {Mesoscale to submesoscale wavenumber spectra in Drake Passage},
+  \textit{Journal of Physical Oceanography}, \textit{46}, 601--620.
+
+\bibitem[{\textit{Sasaki et~al.}(2014)\textit{Sasaki, Klein, Qiu, and
+  Sasai}}]{sasaki_etal2014}
+Sasaki, H., P.~Klein, B.~Qiu, and Y.~Sasai (2014), {Impact of oceanic-scale
+  interactions on the seasonal modulation of ocean dynamics by the atmosphere},
+  \textit{Nature communications}, \textit{5}.
+
+\bibitem[{\textit{Shcherbina et~al.}(2013)\textit{Shcherbina, D'Asaro, Lee,
+  Klymak, Molemaker, and McWilliams}}]{shcherbina_etal2013}
+Shcherbina, A.~Y., E.~A. D'Asaro, C.~M. Lee, J.~M. Klymak, M.~J. Molemaker, and
+  J.~C. McWilliams (2013), {Statistics of vertical vorticity, divergence, and
+  strain in a developed submesoscale turbulence field}, \textit{Geophysical
+  Research Letters}, \textit{40}(17), 4706--4711.
+
+\bibitem[{\textit{Thompson et~al.}(2016)\textit{Thompson, Lazar, Buckingham,
+  Naveira~Garabato, Damerell, and Heywood}}]{thompson_etal2016}
+Thompson, A.~F., A.~Lazar, C.~Buckingham, A.~C. Naveira~Garabato, G.~M.
+  Damerell, and K.~J. Heywood (2016), {Open-ocean submesoscale motions: A full
+  seasonal cycle of mixed layer instabilities from gliders}, \textit{Journal of
+  Physical Oceanography}, \textit{46}(4), 1285--1307.
+
+\end{thebibliography}
+
 % Or,

### 2016-07-11 writeup/rocha_etal_supporting-info.tex
Adds new references

--- 2016-07-10 writeup/rocha_etal.tex
@@
 % AGUtmpl.tex: this template file is for articles formatted with LaTeX2e,
-% Modified July 2014
+% Modified November 2013
 %
@@
 %
-% FOR FIGURES, DO NOT USE \psfrag or \subfigure.
+% FOR FIGURES, DO NOT USE \psfrag
 %
@@
 % ef    EARTH'S FUTURE
-% ea    EARTH AND SPACE SCIENCE
 %
@@
 
-%\documentclass[draft,grl]{agutex2015}
-\documentclass[grl]{agutex2015}
+\documentclass[draft,grl]{agutexSI}
 
-% To create numbered lines:
-
-% If you don't already have lineno.sty, you can download it from
-% http://www.ctan.org/tex-archive/macros/latex/contrib/ednotes/
-% (or search the internet for lineno.sty ctan), available at TeX Archive Network (CTAN).
-% Take care that you always use the latest version.
-
-% To activate the commands, uncomment \usepackage{lineno}
-% and \linenumbers*[1]command, below:
-
-% \usepackage{lineno}
-% \linenumbers*[1]
-%  To add line numbers to lines with equations:
-%  \begin{linenomath*}
-%  \begin{equation}
-%  \end{equation}
-%  \end{linenomath*}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Figures and Tables
+%
+%  SUPPORTING INFORMATION TEMPLATE
+%
+%% ------------------------------------------------------------------------ %%
 %
 %
+%Please use this template when formatting and submitting your Supporting Information.
+
+%This template serves as both a “table of contents” for the supporting information for your article and as a summary of files.
+%
+%
+%OVERVIEW
+%
+%Please note that all supporting information will be peer reviewed with your manuscript.
+%In general, the purpose of the supporting information is to enable authors to provide and archive auxiliary information such as data %tables, method information, figures, video, or computer software, in digital formats so that other scientists can use it.
+%The key criteria are that the data:
+% 1. supplement the main scientific conclusions of the paper but are not essential to the conclusions (with the exception of
+%    including %data so the experiment can be reproducible);
+% 2. are likely to be usable or used by other scientists working in the field;
+% 3. are described with sufficient precision that other scientists can understand them, and
+% 4. are not exe files.
+%
+%USING THIS TEMPLATE
+%
+%***All references should be included in the reference list of the main paper so that they can be indexed, linked, and counted as citations.  The reference section does not count toward length limits.
+%
+%All Supporting text and figures should be included in this document.
+Insert supporting information content into each appropriate section of the template.
+%Figures and tables should appear above each caption.
+To add additional captions, simply copy and paste each sample caption as needed.
+
+%Tables may be included, but can also be uploaded separately, especially if they are larger than 1 page, or if necessary for retaining table formatting.
+Data sets, large tables, movie files, and audio files should be uploaded separately, following AGU naming conventions.
+Include their captions in this document and list the file name with the caption.
+You will be prompted to upload these files on the Upload Files tab during the submission process, using file type “Supporting Information (SI)”
+
+%IMPORTANT NOTE ON FIGURES AND TABLES
+% Placeholders for figures and tables appear after the \end{article} command, after references.
 % DO NOT USE \psfrag or \subfigure commands.
 %
-%
 %  Uncomment the following command to include .eps files
-%  (comment out this line for draft format):
-%\usepackage[dvipdf]{graphicx}
 \usepackage{graphicx}
-
-% CR added this
-\usepackage{color}
-\usepackage{amsmath}
-
 %
@@
 %   when using Draft:
-%\setkeys{Gin}{draft=false}
+\setkeys{Gin}{draft=false}
 %
@@
 %
-% See how to enter figures and tables at the end of the article, after
-% references.
 %
@@
 
-% Author names in capital letters:
 \authorrunninghead{ROCHA ET AL.}
@@
 
+% Author names in capital letters:
+%\authorrunninghead{BALES ET AL.}
+
+% Shorter version of title entered in capital letters:
+%\titlerunninghead{SHORT TITLE}
+
 %Corresponding author mailing address and e-mail address:
+%\authoraddr{Corresponding author: A. B. Smith,
+%Department of Hydrology and Water Resources, University of
+%Arizona, Harshbarger Building 11, Tucson, AZ 85721, USA.
+%(a.b.smith@hwr.arizona.edu)}
 \authoraddr{Corresponding author: Cesar B. Rocha,
@@
 
-\title{Seasonality of submesoscale dynamics in the Kuroshio Extension}
+%\includegraphics{agu_pubart-white_reduced.eps}
+
+% \altaffilmark will produce footnote;
+% matching \altaffiltext will appear at bottom of page.
+
+\authors{Cesar B. Rocha\altaffilmark{1}, Sarah T. Gille\altaffilmark{1},
+        Teresa K. Chereskin\altaffilmark{1}, and Dimitris Menemenlis\altaffilmark{2}}
+% Eric Brown,\altaffilmark{1,2} Rick Williams,\altaffilmark{3}
+% John B. McDougall\altaffilmark{4}, and S. Visconti\altaffilmark{5}}
+
+\altaffiltext{1}{Scripps Institution of Oceanography, University of California San Diego, La Jolla, CA, USA}
+\altaffiltext{2}{Earth Sciences Division, Jet Propulsion Laboratory, California Institute of Technology, Pasadena, CA, USA}
+
 %
-% e.g., \title{Terrestrial ring current:
-% Origin, formation, and decay $\alpha\beta\Gamma\Delta$}
+% e.g., \title{Supporting Information for "Terrestrial ring current:
+% Origin, formation, and decay $\alpha\beta\Gamma\Delta$"}
 %
+%DOI: 10.1002/%insert paper number here%
 
@@
 
-\authors{Cesar B. Rocha\altaffilmark{1}, Sarah T. Gille\altaffilmark{1},
-        Teresa K. Chereskin\altaffilmark{1}, and Dimitris Menemenlis\altaffilmark{2}}
+% \authors{A. B. Smith,\altaffilmark{1}
 % Eric Brown,\altaffilmark{1,2} Rick Williams,\altaffilmark{3}
@@
 
-\altaffiltext{1}{Scripps Institution of Oceanography, University of California San Diego, La Jolla, CA, USA.}
-\altaffiltext{2}{Earth Sciences Division, Jet Propulsion Laboratory, California Institute of Technology, Pasadena, CA, USA.}
+%\altaffiltext{1}{Department of Hydrology and Water Resources,
+%University of Arizona, Tucson, Arizona, USA.}
 
@@
 %Infrastrutture Civili, Politecnico di Torino, Turin, Italy.}
-
-%% ------------------------------------------------------------------------ %%
-%
-%  KEYPOINTS
-%
-%% ------------------------------------------------------------------------ %%
-
-% Key points are 1 to 3 points that the author provides,
-% that are 100 characters or less, that are ultimately published
-% with the article.
-%% for example:
-% \keypoints{\item Here is the first keypoint. what happens if it is a
-% long keypoint, like this one. We want to see this wrap please.
-% \item This is the second.
-% \item And here is the third keypoint
-% }
-
-\keypoints{\item  Upper-ocean submesoscale (10-100 km) turbulence and inertia-gravity waves
-                  undergo strong seasonal cycles that are out-of-phase.
-           \item  Submesoscale turbulence dominates the horizontal velocity and
-                  sea-surface height variability in late winter/early spring.
-            \item Submesoscale inertia-gravity waves dominate the horizontal velocity and
-                   sea-surface height variability in late summer/early fall.
-                  }
-
-%% Keypoints will print underneath the abstract.
-
-%% ------------------------------------------------------------------------ %%
-%
-%  ABSTRACT
-%
-%% ------------------------------------------------------------------------ %%
-
-% >> Do NOT include any \begin...\end commands within
-% >> the body of the abstract.
-
-\begin{abstract}
-    Two new high-resolution numerical simulations with embedded tides show a
-    strong modulation in the governing near-surface dynamics at submesoscales
-    (roughly 10-100 km) in the Kuroshio Extension. Consistent with recent studies, deep late-winter mixed layers
-    are prone to baroclinic instabilities, and submesoscale turbulence
-    prevails in late winter/early spring. While summertime
-    re-stratification weakens submesoscale turbulence, it also enhances submesoscale inertia-gravity
-     waves near the surface. In the Kuroshio Extension,
-    inertia-gravity waves strongly dominate the submesoscale surface kinetic energy and
-    sea-surface height variance in late summer/early fall.
-\end{abstract}
 
@@
 
-\section{Introduction}
+\noindent\textbf{Contents of this file}
+%%%Remove or add items as needed%%%
+\begin{enumerate}
+\item Text S1 with details of the LLC simulations.
+\item Figure S1 with spectra of LLC2160 and LLC4320 simulations.
+\item Table S1 with details about the LLC spin-up.
+\item Figure S2 with time series of vorticity and divergence probability density
+      functions.
+\item Figure S3: root-mean-squares of vorticity, strain rate, and divergence
+      as a function of depth.
+\item Figure S4 with comparison of Argo climatological upper-ocean stratification
+      and the LLC simulations
+\item Text S5 with details about the comparison of model and mooring kinetic energy
+            frequency spectra.
+\item Figure S5: comparison of upper-ocean kinetic energy
+      frequency spectra with spectra from available moored current meters.
+\item Figure S6 showing the seasonal variations of the near-surface amplitude
+      of the vertical modes.
+%if Tables are larger than 1 page, upload as separate excel file
+\end{enumerate}
 
-Recent interest in upper-ocean dynamics has focused on the strong seasonal
-cycle of shallow baroclinic instabilities and their role in submesoscale (roughly 1-100 km)
-turbulence and mesoscale modulation \citep{sasaki_etal2014,qiu_etal2014,
-callies_etal2015, thompson_etal2016,buckingham_etal2016}. Contemporary studies
-have also shown that inertia-gravity waves contribute significantly
-to the near-surface variability at submesoscales \citep{richman_etal2012,
-buhler_etal2014,rocha_etal2016}, but their seasonality has not been investigated.
+\noindent\textbf{Introduction}
+%Type or paste your text here. The introduction gives a brief overview of the supporting information. You should include information %about as many of the following as possible (when appropriate):
+% 1. a general overview of the kind of data files;
+% 2. information about when and how the data were collected or created;
+% 3. a general description of processing steps used;
+% 4. any known imperfections or anomalies in the data.
 
-Using the output of $1/24^\circ$ and $1/48^\circ$ global
-numerical simulations with embedded tides, we show that inertia-gravity waves undergo
-a strong seasonality near the surface in the Kuroshio Extension region.
-Interestingly, the seasonal cycle of inertia-gravity waves is out-of-phase
-with the seasonal cycle of submesoscale turbulence. Consistent with previous studies,
-deep late-winter mixed layers are prone to
-shallow baroclinic instabilities that are roughly in geostrophic balance
-and flux energy upscale \citep{sasaki_etal2014,callies_etal2016},
-driving a mild seasonal modulation of the mesoscales \citep{sasaki_etal2014,qiu_etal2014}.
+This document contains supporting information with details regarding the model
+simulations, comparisons with observations, and supplemental figures.
 
-Inertia-gravity waves, on the other hand, peak in late summer/early fall,
-when the upper ocean is strongly stratified. Thus there exists a
-strong seasonal modulation in the dominant upper-ocean submesoscale dynamics:
-submesoscale turbulence dominates the upper-ocean dynamics in late winter/early,
-whereas submesoscale inertia-gravity waves prevail in late summer/early fall.
-Because submesoscale turbulence is weakest in late summer/early fall,
-the present results indicate that inertia-gravity waves account for most of the
-summertime submesoscale SSH variability.
-
-\section{The LLC numerical simulations}
-We use the output of two latitude-longitude polar cap (LLC)
-realistic numerical simulations. The outputs
-analyzed here, LLC2160 (1/24$^\circ$) and LLC4320 (1/48$^\circ$),   are
-forward Massachusetts Institute of Technology general circulation model (MITgcm)
-numerical solutions on a LLC grid  \citep{forget_etal2015} with
-90-vertical-levels. The coarser-resolution LLC simulation was
-spun up from an ECCO2 adjoint-method state estimate, constrained to millions
-of observations from 2009 through 2011. Both simulations were forced by
-tides and 6-hourly surface fluxes. The LLC2160
-output spans two years from March 2011 to April 2013; the LLC4320 was spun up from
-the LLC2160 simulation, spanning one year from September 2011 to September 2012.
-The LLC4320 simulation is an extension of the 3-month long output used by
-\citet{rocha_etal2016}.
-
-A key aspect of the LLC simulations is that they were forced by
-the 16 most-significant tidal components.
-Because barotropic tides interact with topography and generate internal
-tides that project onto mesoscales to submesoscales
-\citep[e.g., ][]{rocha_etal2016}, tidal forcing fundamentally distinguishes our analysis
-from other modeling studies of seasonality \citep{sasaki_etal2014,qiu_etal2014}. Details of the LLC simulations
-are provided in the supplemental material.
-
-To study seasonal variations in the upper-ocean dynamics, we focus on the northwest
-Pacific, in the vicinity of the Kuroshio
-Extension, where previous studies have shown strong mesoscale and submesoscale seasonality
-\citep{sasaki_etal2014,qiu_etal2014}.
-We analyze a sub-domain of the LLC4320 and LLC2160 simulations of about 2000 km$^2$
-spanning 155-175$^\circ$E; 25-40$^\circ$N (Figure \ref{fig1}). The stratification
-in this mesoscale-rich subtropical region undergoes a vigorous seasonal cycle: wintertime-enhanced
-small-scale turbulence de-stratifies the upper ocean, yielding mixed layers
-as deep as 300m. In late spring/early summer enhanced solar radiation and mixed-layer
-instabilities re-stratify the upper ocean, yielding mixed layers as shallow as 40m.
-Fundamentally, the upper-ocean density structure is well-captured by both LLC simulations:
-a comparison with Argo climatology shows that both simulations skillfully represent the Kuroshio
-Extension stratification and its seasonal variability  (supplemental material).
-
-\begin{figure*}[ht]
-\begin{center}
-\hspace{-1.25cm}\includegraphics[width=.75\textwidth]{figs/fig1_1.pdf}\\
-\vspace{-.125cm}
-\includegraphics[width=.75\textwidth]{figs/fig1_2.pdf}
- \caption{(a) The study region with the subregion where the LLC outputs are
-          analyzed. Colors represent the topography and white lines are contours of absolute
-          dynamic topography every 0.1 m from AVISO. LLC 4320 (1/48$^\circ$) snapshots of surface vorticity (b and c) and transects
-          of potential density at 165$^\circ$E (d and e). The snapshots were
-          taken at 00:00 UTC.}
-\vspace{-1.5cm}
- \label{fig1}
- \end{center}
- \end{figure*}
-
-\section{Statistics of the surface lateral velocity tensor}
-To study the seasonality in the surface velocity, we calculate the lateral velocity gradient tensor
-\begin{equation}
-\left[\begin{matrix} u_x & u_y\\ v_x&v_y \end{matrix}\right]
-\end{equation}
-using a centered
-second-order finite difference scheme. We then diagnose
-the vertical vorticity
-\begin{equation}
-\zeta \equiv v_x - u_y\, ,
-\end{equation}
-lateral rate of strain
-\begin{equation}
-  \alpha \equiv [(u_x-v_y)^2 + (u_y + v_x)^2]^{1/2}\,,\\
-\end{equation}
-and horizontal divergence
-\begin{equation}
-\delta \equiv u_x + v_y\, .
-\end{equation}
-These diagnostics highlight the submesoscale structures in the flow
-\citep[e.g.,][]{capet_etal2008a,shcherbina_etal2013}.
-
-Figures \ref{fig1}b-c show snapshots of vertical vorticity $\zeta$ in early spring
-(April 15) and fall (October 15) in the LLC4320 (1/48$^\circ$) simulation.
-The model solutions depict seasonality in vorticity: large values of
-fine-grained vertical vorticity are observed in early spring with maximum
-values as large as $4f$, where $f$ is
-the local planetary vorticity, and root-mean-square (RMS) of about $0.4f$. In early
-fall, the situation is the opposite: the vertical vorticity is relatively coarse-grained;
-its local maximum and RMS are both smaller than $0.5f$.
-Indeed, the vorticity and rate of strain are strongest in wintertime (Figure \ref{fig2}):
-in both simulations, the RMS vorticity and strain rate are about twice as large in
-late winter/early spring
-than in late summer/early fall. Because the wintertime vorticity and strain rate
-are dominated by the smallest scales in the flow (the KE spectra is shallower than
-a $-3$ power-law in winter), increasing the resolution from
-1/24$^\circ$ to 1/48$^\circ$ increases the wintertime RMS vorticity
-and strain by about 40$\%$.
-
- \begin{figure*}[ht]
-   \begin{center}
-     \includegraphics[width=.75\textwidth]{figs/fig2.pdf}
-  \caption{Time series of the root-mean-square (RMS) of vorticity (a) and
-  rate of strain (b), and horizontal divergence (c) in the LLC outputs and gridded AVISO data.}
-  \label{fig2}
-  \end{center}
-\end{figure*}
-
-The bulk of vertical vorticity and strain rate are associated with subinertial flows
-($T_f = 2\pi/f_0\approx 23.5$h, where $f_0$ is the inertial frequency at
-the mean latitude): daily-averaging the velocity fields suppresses super-inertial
-motions and reduces the RMS
-vorticity by 40$\%$ and the RMS strain by 10$\%$; the seasonal
-cycle remains strong (see
-dashed lines in figures \ref{fig2}a-b). Most of this seasonal cycle is associated
-with submesoscale flows: smoothing the velocity fields
-with a Hanning filter with cut-off scale of 100 km dramatically reduces the RMS
-vorticity and strain rate. The reduction in variance is about 80$\%$ in winter,
-yielding RMS vorticity and strain  rate roughly consistent with the diagnostics from
-AVISO gridded geostrophic velocities (compare red lines to black lines in figures \ref{fig2}a-b).
-The picture that emerges is consistent with recent studies: shallow baroclinic
-instabilities energize the submesoscales in late winter, drawing from the available
-potential energy stored in deep mixed layers \citep{sasaki_etal2014,callies_etal2015,callies_etal2016}.
-
-The seasonal cycle of the horizontal divergence, however, showcases the complexity
-of the upper-ocean annual variability. If submesoscale eddies and fronts dominated
-the near-surface variability all year, then the seasonal cycle of horizontal divergence,
-vertical vorticity, and lateral strain rate would be in phase \citep[e.g.,][]{sasaki_etal2014}.
-While there is a clear wintertime peak in divergence of daily-averaged velocity
-(see dashed lines in figure \ref{fig2}c; RMS divergence $\sim0.22 f$ in the 1/48$^\circ$
-simulation), the hourly fields show
-a stronger enhancement of
-lateral divergence in late summer/early fall (RMS divergence $\sim0.22 f$ the 1/48$^\circ$
-simulation). Because the 1/48$^\circ$ simulation better resolves
-smaller-scale submesoscale flows, a secondary RMS divergence peak in winter  is
-nearly as strong as in summer. Submesoscale fronts and eddies evolve
-relatively fast, and there is no clear
-temporal and spatial scale separation between those motions and inertia-gravity waves
-\citep{mcwilliams2016}:
-daily-averaging the velocity fields efficiently suppresses the summertime horizontally
-divergent flows, but also reduces the wintertime lateral divergence by about 50$\%$.
-Figure \ref{fig2}c also shows that most of the lateral divergence is associated
-with submesoscale flows: smoothing the velocity fields with a 100km-cut-off
-suppresses more than 80$\%$ of the RMS divergence.
-
-\begin{figure*}[ht]
-  \begin{center}
-    \includegraphics[width=.95\textwidth]{figs/fig3.pdf}
- \caption{Seasonal variation of joint probability distributions:  vorticity vs. strain rate (a through d),
-          and vorticity vs. Laplacian of sea-surface height (e through h) in April (a, b,
-          e, f) and October (c, d, g, h).
-          Dashed lines in (a) through (d) represent one dimensional shear flow $\alpha = \pm\zeta$,
-          characteristic of fronts. Dashed lines in (e) through
-          (h) represent geostrophic flow $\zeta = \frac{g}{f}\nabla_h^2 \eta$.}
- \label{fig3}
- \end{center}
-\end{figure*}
-
-\section{Seasonality in governing submesoscale dynamics}
-The results of figure \ref{fig2}a-c show that submesoscale surface
- variability stems from different dynamics in summer than in winter.
- To characterize these differences, we calculate
-joint probability distributions (jPDF) of
-vorticity-strain and vorticity-Laplacian of sea-surface height (Figure \ref{fig3}).
-
-The April vorticity-strain jPDF has a shape characteristic of submesoscale turbulence
- (Figures \ref{fig3}a-b). The alignment of vorticity and strain $\alpha \sim \pm\zeta$
- with strong positive skewness are fingerprints of submesoscale fronts
- \citep{shcherbina_etal2013,mcwilliams2016}. The shape of the vorticity-strain
- jPDF is similar for hourly, daily-averaged, although
- the vorticity skewness reduces from 1.4 to 1.13 from hourly to daily-averaged.
- The April results are characteristic of winter, indicating that wintertime submesoscale
- surface velocity is strongly dominated by submesoscale turbulence.
- The hourly velocity fields are significantly
-ageostrophic as depicted by the jPDF of vorticity-Laplacian of SSH
-winter (Figure \ref{fig3}e). Even in April, only the daily-averaged fields
-are largely
-in  geostrophic balance (the jPDF of daily-averaged vorticity vs. Laplacian of SSH
-is an ellipse with large eccentricity and main axis tilted by 45$^\circ$; Figure \ref{fig3}f).
-
-The October vorticity-strain jPDF shows much weaker skewness (the vorticity skewness
-is 0.68 and 0.67 for hourly and daily-averaged velocities). The shape of the
-vorticity-strain jPDF  appears to be a combination of two half-ellipses centered
-about $\zeta=0$, one with a 45$^\circ$ slope (characteristic of submesocale fronts
-that persist in summer) and one with a very steep slope.
-That the submesoscale dynamics in October are mainly unbalanced is clearly depicted
-in the shape of jPDF of vorticity-Laplacian of SSH for hourly fields
-winter (Figure \ref{fig3}g), which is an ellipse aligned in the vertical axis.
-Of course daily-averaging the model suppresses the ageostrophic, super-inertial flows, and
-the daily-averaged flow is essentially geostrophic as depicted by the 45$^\circ$-tilted
-ellipse in the vorticity-Laplacian of SSH jPDF.
-
- Time series of PDFs of vorticity and divergence (supplemental material) show
- a strong oscillation between these two regimes. In late winter/early spring
- the vorticity is strongly positively skewed, whereas the divergence is moderately
- negatively skewed as predicted by frontogenesis \citep[e.g., ][]{capet_etal2008a,mcwilliams2016}
-. In late summer/early fall, the divergence is stronger, but PDFs are much less skewed,
-consistent with linear inertia-gravity waves.
-
-\section{Projection onto horizontal scales}
-
-To better quantify the projection of these flows onto different horizontal
-scales, we calculate wavenumber spectra of kinetic energy and sea-surface height
-(SSH) variance.  Before calculating the spectra, time-mean and spatial linear
-trends were removed, and the resulting fields were multiplied by a two-dimensional Hanning ``spectral window''.
-The two-dimensional spectra were averaged azimuthly \citep[e.g., ][]{rocha_etal2016} .
-We discuss only spectra for the 1/48$^\circ$ simulation, which extends the 1/24$^\circ$ simulation
-towards smaller scales (supplemental material).
-
-\begin{figure*}[ht]
-  \begin{center}
-    \includegraphics[width=1.\textwidth]{figs/fig4.pdf}
- \caption{Surface (horizontal) KE (a) and SSH variance wavenumber spectra (b)
- in the 1/48$^\circ$ simulation. Solid lines
- are spectra based on hourly snapshots, dashed lines are spectra based on daily-averaged
- fields.}
- \label{fig4}
- \end{center}
-\end{figure*}
-
-Figure \ref{fig4}a depicts the horizontal wavenumber spectra of surface KE in
-April and October. At scales larger than 20km, the April and October spectra based on hourly (the solid curves)
-velocity snapshots are nearly indistinguishable from each other within 95$\%$ confidence level.
-Consistent with the results of
-\citet{rocha_etal2016}, who analyzed a 5-month output of the LLC4320 simulation in Drake Passage,
-there is significant high-frequency variabilty at submesoscales. Daily-averaging
-the velocity field suppresses spatial variability at scales smaller than about 250
-km both in April and October (compare solid lines against dashed lines in figure
-\ref{fig4}a). But this suppression is dramatic in October, when the inertia-gravity waves peak. At scales
-smaller than 100 km, 39$\%$ of the surface KE in April is accounted for by super-inertial
-flows as opposed to 79$\%$ in October. The seasonality of subinertial submesoscale flows
-is more dramatic than the seasonal cycle of the total flow. This is consistent
-with the results of \citet{sasaki_etal2014}, which are
-based on daily-averaged velocity fields of a different model without tidal forcing
-(P. Klein, personal communication).
-
-The inertia-gravity waves project on the sea-surface.
-There is a dramatic difference between the spectra
-based on hourly and daily-averaged SSH in October (Figure \ref{fig4}b): at scales smaller than 100 km, the spectra
-of hourly SSH roughly follows a $-2$ power-law, whereas the spectra of daily-averaged
-SSH roughly follows a $-5$ power-law. At these scales, 33$\%$ of the SSH variance in April is accounted for by super-inertial
-flows as opposed to 83$\%$ in October! Curiously, the out-of-phase seasonal cycle of submesoscale
-turbulence and near-surface submesoscale inertia-gravity waves conspire to yield weak seasonality
-in the spectra of KE and SSH variance based on hourly fields.
-
-\section{Summary and Conclusion}
-Our work adds to recent studies that present modeling \citep{sasaki_etal2014}
-and observational \citep{callies_etal2015,buckingham_etal2016} evidence of vigorous seasonality in
-submesoscale turbulence. In particular, our main finding is that in global simulations with embedded tides the
-near-surface submesoscale inertia-gravity waves in the Kuroshio Extension undergo
-a strong seasonal cycle that is out-of-phase with the seasonal cycle of
-submesoscale turbulence. We conjecture that the summertime dominance of inertia-gravity waves
- \citep{callies_etal2015} is a consequence both of suppression of submesoscale turbulence and
-enhancement of inertia-gravity waves due to re-stratification of
-the upper ocean.
-
-\cite{dasaro1978} showed that the velocity of linear internal waves
-in the mixed-layer strongly depends on the density jump at the mixed-layer
-base, with largest mixed layer velocities when the jump is strongest.
-In summer, the shallow mixed layer overlays a strong seasonal pycnocline,
-and thus the internal waves projection onto the mixed-layer may be stronger
-according to \cite{dasaro1978}'s arguments. An
-alternative explanation is that the shape of the baroclinic modes
-changes seasonally. In particular the baroclinic modes are
-significantly more surface intensified in late summer/early fall
- (supplemental material).
-If the internal wave source has weak seasonal dependence, then a strong
-seasonal cycle in the near-surface expression of internal waves is expected.
-Internal tide generation does not show a seasonal modulation \citep[e.g.,][]{alford2003},
-and therefore one expects a strong seasonality in the near-surface
-expression of internal tides and
-other submesoscale inertia-gravity waves generated through internal tide interactions.
-Near-inertial wave generation peaks in winter \citep[e.g.,][]{alford2003}, but
-those waves project on large horizontal scales \citep[e.g, ][]{qi_etal1995}.
-
-An important caveat is that this study focuses on a single patch of ocean in the
-vicinity of the Kuroshio Extension. This may be a typical of mesoscale-rich
-subtropical region, but it is unlikely to be
-representative of other regions such as low eddy kinetic energy eastern boundary currents
-and the middle of the subtropical gyre.  We plan to report on the geographic variability of submesoscale
- seasonality in a future study.
-
-%The effects of smaller-scale/higher-frequency
-%``sub-submesoscale'' flows on the submesoscale surface velocity and SSH variability
-%are presently unknown.
-
-% That the  surface velocity and SSH submesoscale variability may be
-% dominated by ageostrophic flows in
-% summer/fall has consequences for the interpretation of data from
-% the future SWOT and COMPIRA altimeter missions,
-% which will deliver SSH measurements at submesoscales. To the extent that
-% high-frequency flows are dominated by spatially incoherent internal tides and other
-% internal waves, it may be very difficult (if not impossible) to
-% separate SSH submesoscale variability associated with geostrophic motions
-% from high-frequency, ageostrophic flows. Thus, previous claims that
-% one will be able to easily obtain submesoscale surface
-% geostrophic velocities and monitor such seasonal cycle on global scales
-% \citep{sasaki_etal2014,qiu_etal2014} are overstated.
-
-% Here we argue that those divergent motions are likely inertia-gravity waves.
-% We can show a couple of spectra and argue that they roughly satisfy polarization
-% relations.
-%
-% We can also try and show some observations here. Any cruises with ADCP data.
-% While it may be hard to have enough data for the errorbars to be small, a
-% rough consistency may be better than nothing.
-%
-% Perhaps a mooring data should show strong seasonal modulation at high-frequencies?
-%
-% We must be able to present some observational evidence that the model is not
-% misleading at high-frequencies.
-
-% Here we argue that high-frequency (supra-daily) flows significantly project
-% on the sea-surface, and thus estimation of submesoscale (10-100 km) geostrophic
-% velocity from sea-surface height (SSH) is not warranted. Calculating spectral
-% KE fluxes from both velocity and geostrophic velocity (from SSH) would
-% contrast with the results in Sasaki et al.
-
-% \section{Projection on time scales}
-% To better understand the high-frequency/high-wavenumber flows  we calculate frequency
-% spectra of surface KE winter \ref{fig4}). To obtain statistically meaningful results,
-% we average spectral estimates along a section at 165$^\circ$E.  There is about 50
-% independent spectral realizations.
-%
-% The surface KE spectra of subinertial flows very roughly follow a $-2$ power-law.
-% The high-frequency flows are split near-inertial flows, tidal signals and their
-% higher harmonics. Hence these high-frequency flows are dominated by inertia-gravity
-% waves. (Note that the inertial peak nearly merges with the
-% diurnal tide peak.) There is statistically significant seasonality. The seasonal
-% cycle of the sub-inertial motions (at least for periods between 5-20 days)
-% is out-of-phase with the seasonal cycle of supra-inertial flows: the former flows
-% are more energetic in April and the latter in October.
-%
-% Most of these high-frequency inertia-gravity waves project onto scales smaller
-% than 100 km. The surface KE spectra of 100-km-smoothed velocity field is significantly
-% suppressed in supra-inertial frequencies --- there are no statistically significant
-% changes at sub-inertial frequencies (see dashed lines in figure \ref{fig4}).
-
-% say something about relative energy in smoother vs. unsmoothed fields.
-% also say something about comparison with observations.
-
-%\begin{figure}[ht]
-%  \begin{center}
-%    \includegraphics[width=.45\textwidth]{figs/fig4.pdf}
-% \caption{Surface (horizontal) KE frequency spectra. Solid lines
-% are spectra based on the 1/24$^\circ$ fields, dashed line are spectra
-% based on 100-km-smoothed fields.}
-% \label{fig4}
-% \end{center}
-%\end{figure}
-
-%
-%  ACKNOWLEDGMENTS
-
-\begin{acknowledgments}
- William R. Young provided helpful feedback on the first draft. We thank the MITgcm community and our colleagues at the NASA Advanced
-Supercomputing (NAS) Division for their awesome support.
-This research was funded by NSF (OCE1357047) and NASA (NNX13AE44G,NNX13AE85G,NNX16AH67G).
-The LLC outputs can be obtained from D. Menemenlis (menemenlis@jpl.nasa.gov). The altimeter products were produced by Ssalto/Duacs
-and distributed by Aviso, with support from CNES (\texttt{http://www.aviso.altimetry.fr/duacs/}).
-\end{acknowledgments}
-
-\bibliographystyle{agufull08}
-\bibliography{rocha_etal}
-
-%  this should go in a supplementary material
 %\clearpage
 
-%\section{Supplemental material 1: Spectral errors}
+%Delete all unused file types below. Copy/paste for multiples of each file type as needed.
+\noindent\textbf{S1: Detail of the LLC simulations}
+
+Table \ref{tab:llc} shows details of the LLC spin-up hierarchy.
+Both simulations analyzed in the present paper, LLC2160 and LLC4320,
+were forced by surface fluxes from the 0.14$^\circ$ European Centre for
+Medium-range Weather Forecasting (ECMWF) atmospheric operational model analysis,
+ starting in 2011, and the 16 most significant  tidal components.
+The LLC control files are available online
+(at \texttt{http://mitgcm.org/viewvc/MITgcm/MITgcm$\_$contrib/
+llc$\_$hires}).
+
+For the LLC hierarchy, the sea-surface height is calculated
+using a linearized equation. While the nominal horizontal resolution  of the
+LLC4320 and LLC2160 simulations are 1/48$^\circ$ and 1/24$^\circ$, r
+horizontal wavenumber spectra suggests an effective resolution of about 8 km and 20 km,
+ respectively (see figure S1).
+
+The spin-up of the upper-ocean in the LLC4320 (1/48$^\circ$) simulation is
+a couple of days, as depicted the time series of surface horizontal divergence
+in Figure 2b of the main text (see initial rapid increase in RMS divergence in
+September 2011). To avoid the spin-up stage, we chose October to characterize the
+late summer/early fall. The late winter/early spring waves chosen to be 6 months
+out-of-phase with late summer/early fall (hence April).
+
+\noindent\textbf{S5: How well do the LLC simulations capture high-frequency modes?}
+
+To assess how well the LLC simulations represents the high-frequency variability,
+we compare the model near-surface velocity field against two available
+mooring data: KEO (the data is available online at \texttt{http://www.pmel.noaa.gov/ocs/KEO})
+and KESS mooring $\#$7 (the data is available online at \texttt{http://uskess.whoi.edu}).
+For proper comparison, we use model data closest to these two moorings (west of
+the region considered in the present study). Focusing on high-frequencies, we
+calculate frequency spectral estimates every month. There are about 48 independent
+spectral realizations.
+
+Figure \ref{figA1} shows KE frequency spectra for the LLC simulations and mooring data.
+At the uppermost common depth (20 m), very low frequencies (periods $>$ 10 days)
+and the inertial, diurnal, and semi-diurnal frequencies agree well --- the KESS
+data are more energetic at intermediate frequencies. At 40 m, the LLC simulations and observations
+have consistent spectra at periods larger than about 6 h. The data from the moorings
+are more energetic at higher frequencies,
+except for the first two harmonics that are apparent in the KESS data at 40 m.
+There are at least three reasons for the slightly more energetic super-inertial
+variability in the observations: 1) The high-
+frequency variability is dominated by high-frequency noise due to mooring vertical
+excursions; 2) The model does not have enough resolution to resolve the full internal
+wave spectrum; 3) The model is not forced at frequencies higher than 6h.
+
+%\noindent\textbf{Data Set S1.} %Type or paste caption here.
+%upload your dataset(s) to AGU's journal submission site and select "Supporting Information (SI)" as the file type. Following naming %convention: ds01.
+
+%Repeat for any additional Supporting data sets
+
+%\noindent\textbf{Movie S1.} %Type or paste caption here.
+%upload your movie(s) to AGU's journal submission site and select, "Supporting Information %(SI)" as the file type. Following naming convention: ms01.
+
+%Repeat any additional Supporting movies
+
+%\noindent\textbf{Audio S1.} %Type or paste caption here.
+%upload your audio file(s) to AGU's journal submission site and select "Supporting Information %(SI)" as the file type. Following naming convention: auds01.
+
+%Repeat for any additional Supporting audio files
 
 %%% End of body of article:
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Optional Appendix goes here
-%
-% \appendix resets counters and redefines section heads
-% but doesn't print anything.
-% After typing \appendix
-%
-%\section{Here Is Appendix Title}
-% will show
-% Appendix A: Here Is Appendix Title
-%
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %
-% Optional Glossary or Notation section, goes here
+% Optional Notation section goes here
 %
-%%%%%%%%%%%%%%
-% Glossary is only allowed in Reviews of Geophysics
-% \section*{Glossary}
-% \paragraph{Term}
-% Term Definition here
-%
-%%%%%%%%%%%%%%
 % Notation -- End each entry with a period.
@@
 % AGU does not want a .bib or a .bbl file. Please copy in the contents of your .bbl file here.
+\bibliographystyle{agufull08}
+%\bibliography{rocha_etal}
+
+\begin{thebibliography}{3}
+\providecommand{\natexlab}[1]{#1}
+\expandafter\ifx\csname urlstyle\endcsname\relax
+  \providecommand{\doi}[1]{doi:\discretionary{}{}{}#1}\else
+  \providecommand{\doi}{doi:\discretionary{}{}{}\begingroup
+  \urlstyle{rm}\Url}\fi
+
+  \bibitem[{\textit{Levitus et~al.}(2013)\textit{Levitus, Antonov, Baranova,
+    Boyer, Coleman, Garcia, Grodsky, Johnson, Locarnini, Mishonov
+    et~al.}}]{levitus_etal2013}
+  Levitus, S., J.~Antonov, O.~K. Baranova, T.~Boyer, C.~Coleman, H.~Garcia,
+    A.~Grodsky, D.~Johnson, R.~Locarnini, A.~V. Mishonov, et~al. (2013), The
+    world ocean database, \textit{Data Science Journal}, \textit{12}(0),
+    WDS229--WDS234.
+
+\bibitem[{\textit{Rocha et~al.}(2016)\textit{Rocha, Chereskin, Gille, and
+  Menemenlis}}]{rocha_etal2016}
+Rocha, C.~B., T.~K. Chereskin, S.~T. Gille, and D.~Menemenlis (2016),
+  {Mesoscale to submesoscale wavenumber spectra in Drake Passage},
+  \textit{Journal of Physical Oceanography}, \textit{46}, 601--620.
+
+\bibitem[{\textit{Roemmich and Gilson}(2009)}]{roemmich_gilson2009}
+  Roemmich, D., and J.~Gilson (2009), {The 2004--2008 mean and annual cycle of
+    temperature, salinity, and steric height in the global ocean from the Argo
+    Program}, \textit{Progress in Oceanography}, \textit{82}(2), 81--100.
+
+\end{thebibliography}
 
@@
 \end{article}
+\clearpage
+
+% Delete all unused file types below. Copy/paste for multiples of each file type as needed.
+
+% enter figures and tables here:
 %
-%
-%% Enter Figures and Tables here:
-%
-% DO NOT USE \psfrag or \subfigure commands.
-%
-% Figure captions go below the figure.
-% Table titles go above tables; all other caption information
-%  should be placed in footnotes below the table.
-%
-%----------------
 % EXAMPLE FIGURE
-%
- %\begin{figure}
- %\noindent\includegraphics[width=20pc]{samplefigure.eps}
+% ---------------
+% \begin{figure}
+%\setfigurenum{S1} %%Change number for each figure
+% \noindent\includegraphics[width=20pc]{samplefigure.eps}
 %\caption{Caption text here}
@@
  %\end{figure}
-%
+
+ \begin{figure}
+    \begin{center}
+      \setfigurenum{S1}
+      \includegraphics[width=40pc]{figs/fig_s2_3.pdf}
+   \caption{A comparison between spectra of the two LLC simulations 1/24$^\circ$
+   (green) and 1/48$^\circ$ (purple). The statistical errorbars represent conservative
+   95$\%$ significance levels.  As expected the
+   higher-resolution simulation (LLC4320) essentially extends the spectra of the
+   lower-resolution simulation (LLC2160) towards smaller scale, demonstrating that
+   the higher resolution is resolving a wider array of scales. Between 20 and 100 km
+   the spectra of the two simulations are visually indistinguishable.}
+   \label{figS2_3}
+   \end{center}
+ \end{figure}
+
+ \begin{figure}
+    \begin{center}
+      \setfigurenum{S2}
+      \includegraphics[width=40pc]{figs/fig_s2_2.pdf}
+   \caption{Time dependence of probability density functions of vertical vorticity (a, b)
+   and horizontal divergence (c, d). Monthly averages in April and October are
+   also shown (e, f). The time series of vorticity PDF indicate that
+   the development of positive skewness in winter (Figures \ref{figS2_2}a-b) both in
+   hourly ($\sim 1.4$) and daily-averaged ($\sim1.1$) fields. The skewness
+   in divergence for hourly fields is very small year-round ($\sim -0.24$), suggesting the
+   prevalence of inertia-gravity waves. Filtering out the waves by daily-averaging
+   the velocity field increases the skewness in divergence, particularly in winter ($\sim-0.64$). Both
+  positive vorticity (cyclonic) skewness and negative divergence (convergence) skewness
+  are characteristics of submesoscale turbulence.}
+   \label{figS2_2}
+   \end{center}
+ \end{figure}
+
+ \begin{figure}
+    \begin{center}
+      \setfigurenum{S3}
+      \includegraphics[width=40pc]{figs/fig_s2_2.png}
+   \caption{Depth-dependence of root-mean-square of vertical vorticity (a,d), rate of lateral strain (b,e)
+   and horizontal divergence (c,f) for hourly (a,b,c) and daily-averaged (d,e,f,) fields.
+   The strong seasonality, as depicted
+   at the surface in Figure 2 of the main manuscript, is confined to the mixed layer, which
+   varies from 50 m in summer to about 350 m in winter.}
+   \label{figS2_2}
+   \end{center}
+ \end{figure}
+
+ \begin{figure}
+    \begin{center}
+      \setfigurenum{S4}
+      \includegraphics[width=40pc]{figs/fig_s2_1.pdf}
+   \caption{Potential density: a comparison between Roemmich-Gilson Argo
+           climatology [updates from \cite{roemmich_gilson2009}] (a, b) and the LLC simulations 1/24$^\circ$ (c, d)
+           and  1/48$^\circ$ (e, f). Of course, the Argo climatology is smoother
+           owing to lower resolution and long-term averaging. Nonetheless, simulations reproduce well the strong
+           seasonality in the stratification. The upper ocean is weakly stratified in late winter/early fall
+           and re-stratifies in summer. Argo climatology and LLC simulations mixed-layer depths are consistent
+           at all seasons (not shown). The Argo climatology is available online at
+           \texttt{http://sio-argo.ucsd.edu/RG\_Climatology.html}.}
+   \label{figS2_1}
+   \end{center}
+ \end{figure}
+
+ \begin{figure}
+    \begin{center}
+      \setfigurenum{S5}
+      \includegraphics[width=40pc]{figs/figA1.pdf}
+   \caption{A comparison between frequency spectra of LLC simulations and two
+   available moored current meter records.}
+   \label{figA1}
+   \end{center}
+ \end{figure}
+
+ \begin{figure}
+    \begin{center}
+      \setfigurenum{S6}
+      \includegraphics[width=40pc]{figs/fig_s3.pdf}
+   \caption{The seasonal variability of the WOA 2013 \citep{levitus_etal2013} stratification averaged  over
+   the domain and the associated three gravest pressure modes. Only the upper 450 m is shown.
+   Clearly, the changes to the stratification
+   are dramatic, with the formation of a seasonal pycnocline in summer. The strong
+   seasonality of upper-ocean stratification yields a strong seasonality in the
+   near-surface shape and amplitude of baroclinic pressure modes.
+   In particular, the surface amplitude is much large in summer. Thus the projection
+   of baroclinic tides on the surface may vary significantly seasonally. The WOA data is
+   available online at \texttt{https://www.nodc.noaa.gov/OC5/woa13/}.}
+   \label{figS3}
+   \end{center}
+ \end{figure}
+
+ \begin{table}
+ \label{tab:llc}
+ \settablenum{S1}
+ \caption{\small The MITgcm Latitute-Longitude Cap (LLC) spin-up hierarchy. This study analyzes
+          the output of LLC2160 and LLC4320. Model fields are available hourly.
+          Adapted from \cite{rocha_etal2016}.}
+ \begin{center}
+ \begin{tabular}{ | c | c | c | c | c |}
+ \hline
+ Simulation & Resolution & Time-step & Period  & Tides \\ \hline
+ ECCO2 Adjoint & $1/6^{^\circ}$ & 1200 s & January 2009 - December 2011 & No\\
+ LLC 1080   & $1/12^{^\circ}$ & 90 s & January 2010 - July 2012 & No\\
+ LLC 2160   &  $1/24^{^\circ}$ & 45 s  & January 2011 - April 2013 & Yes\\
+ LLC 4320   &  $1/48^{^\circ}$ & 25 s & September 2011 - September 2012 & Yes \\
+ \hline
+ \end{tabular}
+ \end{center}
+ \end{table}
+
 % ---------------
@@
 %\begin{table}
+%\settablenum{S1} %%Change number for each table
 %\caption{Time of the Transition Between Phase 1 and Phase 2\tablenotemark{a}}
@@
 %\end{table}
-
-% See below for how to make sideways figures or tables.
+% ---------------
+%
+% EXAMPLE LARGE TABLE (UPLOADED SEPARATELY)
+%\begin{table}
+%\settablenum{S1} %%Change number for each table
+%\caption{Time of the Transition Between Phase 1 and Phase 2\tablenotemark{a}}
+%\end{table}
 

### 2016-07-13 writeup/rocha_etal.tex
Dimitris' comments

@@
 \keypoints{\item  Upper-ocean submesoscale (10-100 km) turbulence and inertia-gravity waves
-                  undergo strong seasonal cycles that are out-of-phase.
+                  undergo strong seasonal cycles that are out of phase.
            \item  Submesoscale turbulence dominates the horizontal velocity and
@@
 a strong seasonality near the surface in the Kuroshio Extension region.
-Interestingly, the seasonal cycle of inertia-gravity waves is out-of-phase
-with the seasonal cycle of submesoscale turbulence. Consistent with previous studies,
+Interestingly, the seasonal cycle of inertia-gravity waves is out of phase
+with the seasonal cycle of submesoscale turbulence. On one hand, consistent with previous studies,
 deep late-winter mixed layers are prone to
@@
 90-vertical-levels. The coarser-resolution LLC simulation was
-spun up from an ECCO2 adjoint-method state estimate, constrained to millions
+spun up from an Estimating the Circulation and Climate of the Ocean, Phase II (ECCO2)
+adjoint-method state estimate, constrained to millions
 of observations from 2009 through 2011. Both simulations were forced by
-tides and 6-hourly surface fluxes. The LLC2160
+tides and 6-hourly surface atmospheric fields. The LLC2160
 output spans two years from March 2011 to April 2013; the LLC4320 was spun up from
@@
 small-scale turbulence de-stratifies the upper ocean, yielding mixed layers
-as deep as 300m. In late spring/early summer enhanced solar radiation and mixed-layer
-instabilities re-stratify the upper ocean, yielding mixed layers as shallow as 40m.
+as deep as 300 m. In late spring/early summer enhanced solar radiation and mixed-layer
+instabilities re-stratify the upper ocean, yielding mixed layers as shallow as 40 m.
 Fundamentally, the upper-ocean density structure is well-captured by both LLC simulations:
@@
 are dominated by the smallest scales in the flow (the KE spectra is shallower than
-a $-3$ power-law in winter), increasing the resolution from
+a $-3$ power law in winter), increasing the resolution from
 1/24$^\circ$ to 1/48$^\circ$ increases the wintertime RMS vorticity
@@
      \includegraphics[width=.75\textwidth]{figs/fig2.pdf}
-  \caption{Time series of the root-mean-square (RMS) of vorticity (a) and
-  rate of strain (b), and horizontal divergence (c) in the LLC outputs and gridded AVISO data.}
+  \caption{Time series of the root-mean-square (RMS) of (a) vorticity,
+  (b) rate of strain, and (c) horizontal divergence in the LLC outputs and gridded AVISO data.}
   \label{fig2}
@@
 While there is a clear wintertime peak in divergence of daily-averaged velocity
-(see dashed lines in figure \ref{fig2}c; RMS divergence $\sim0.22 f$ in the 1/48$^\circ$
+(see dashed lines in figure \ref{fig2}c; RMS divergence $\sim0.1 f$ in the 1/48$^\circ$
 simulation), the hourly fields show
@@
 Figure \ref{fig2}c also shows that most of the lateral divergence is associated
-with submesoscale flows: smoothing the velocity fields with a 100km-cut-off
+with submesoscale flows: smoothing the velocity fields with a 100-km-cutoff
 suppresses more than 80$\%$ of the RMS divergence.
@@
     \includegraphics[width=.95\textwidth]{figs/fig3.pdf}
- \caption{Seasonal variation of joint probability distributions:  vorticity vs. strain rate (a through d),
-          and vorticity vs. Laplacian of sea-surface height (e through h) in April (a, b,
+ \caption{Seasonal variation of joint probability distributions:  vorticity vs.\ strain rate (a through d),
+          and vorticity vs.\ Laplacian of sea-surface height (e through h) in April (a, b,
           e, f) and October (c, d, g, h).
@@
  \citep{shcherbina_etal2013,mcwilliams2016}. The shape of the vorticity-strain
- jPDF is similar for hourly, daily-averaged, although
+ jPDF is similar for hourly and daily-averaged fields, although
  the vorticity skewness reduces from 1.4 to 1.13 from hourly to daily-averaged.
@@
 are largely
-in  geostrophic balance (the jPDF of daily-averaged vorticity vs. Laplacian of SSH
+in  geostrophic balance (the jPDF of daily-averaged vorticity vs.\\ Laplacian of SSH
 is an ellipse with large eccentricity and main axis tilted by 45$^\circ$; Figure \ref{fig3}f).
@@
 in the shape of jPDF of vorticity-Laplacian of SSH for hourly fields
-winter (Figure \ref{fig3}g), which is an ellipse aligned in the vertical axis.
-Of course daily-averaging the model suppresses the ageostrophic, super-inertial flows, and
+ (Figure \ref{fig3}g), which is an ellipse aligned in the vertical axis.
+Daily-averaging the model suppresses the ageostrophic, super-inertial flows, and, therefore,
 the daily-averaged flow is essentially geostrophic as depicted by the 45$^\circ$-tilted
@@
 trends were removed, and the resulting fields were multiplied by a two-dimensional Hanning ``spectral window''.
-The two-dimensional spectra were averaged azimuthly \citep[e.g., ][]{rocha_etal2016} .
+The two-dimensional spectra were averaged azimuthaly \citep[e.g., ][]{rocha_etal2016} .
 We discuss only spectra for the 1/48$^\circ$ simulation, which extends the 1/24$^\circ$ simulation
@@
  in the 1/48$^\circ$ simulation. Solid lines
- are spectra based on hourly snapshots, dashed lines are spectra based on daily-averaged
+ are spectra based on hourly snapshots, and dashed lines are spectra based on daily-averaged
  fields.}
@@
 Figure \ref{fig4}a depicts the horizontal wavenumber spectra of surface KE in
-April and October. At scales larger than 20km, the April and October spectra based on hourly (the solid curves)
-velocity snapshots are nearly indistinguishable from each other within 95$\%$ confidence level.
+April and October. At scales larger than 20km, the April and October spectra based on hourly
+velocity snapshots (solid lines) are nearly indistinguishable from each other within 95$\%$ confidence level.
 Consistent with the results of
@@
 the velocity field suppresses spatial variability at scales smaller than about 250
-km both in April and October (compare solid lines against dashed lines in figure
+km, both in April and October (compare solid lines against dashed lines in Figure
 \ref{fig4}a). But this suppression is dramatic in October, when the inertia-gravity waves peak. At scales
@@
 An important caveat is that this study focuses on a single patch of ocean in the
-vicinity of the Kuroshio Extension. This may be a typical of mesoscale-rich
+vicinity of the Kuroshio Extension. This may be typical of a mesoscale-rich
 subtropical region, but it is unlikely to be
-representative of other regions such as low eddy kinetic energy eastern boundary currents
+representative of other regions such as low-eddy-kinetic-energy eastern boundary currents
 and the middle of the subtropical gyre.  We plan to report on the geographic variability of submesoscale
@@
 
-% say something about relative energy in smoother vs. unsmoothed fields.
+% say something about relative energy in smoother vs.\\ unsmoothed fields.
 % also say something about comparison with observations.
@@
 This research was funded by NSF (OCE1357047) and NASA (NNX13AE44G,NNX13AE85G,NNX16AH67G).
-The LLC outputs can be obtained from D. Menemenlis (menemenlis@jpl.nasa.gov). The altimeter products were produced by Ssalto/Duacs
+The LLC output can be obtained from the ECCO project (\texttt{http://ecco2.org/llc$\_$hires}). The altimeter products were produced by Ssalto/Duacs
 and distributed by Aviso, with support from CNES (\texttt{http://www.aviso.altimetry.fr/duacs/}).

### 2016-07-13 writeup/rocha_etal_supporting-info.tex
Small fix
Dimtris' comments

@@
  starting in 2011, and the 16 most significant  tidal components.
-The LLC control files are available online
+The LLC code and model configuration files are available online
 (at \texttt{http://mitgcm.org/viewvc/MITgcm/MITgcm$\_$contrib/
@@
 except for the first two harmonics that are apparent in the KESS data at 40 m.
-There are at least three reasons for the slightly more energetic super-inertial
+There are at least two reasons for the slightly more energetic super-inertial
 variability in the observations: 1) The high-
 frequency variability is dominated by high-frequency noise due to mooring vertical
-excursions; 2) The model does not have enough resolution to resolve the full internal
-wave spectrum; 3) The model is not forced at frequencies higher than 6h.
+excursions; and 2) The model does not have enough resolution to resolve the full internal
+wave spectrum.
 
@@
       \includegraphics[width=40pc]{figs/fig_s2_3.pdf}
-   \caption{A comparison between spectra of the two LLC simulations 1/24$^\circ$
-   (green) and 1/48$^\circ$ (purple). The statistical errorbars represent conservative
+   \caption{Comparison between spectra of the 1/24$^\circ$
+   (green) and 1/48$^\circ$ (purple) LLC simulations. The statistical error bars represent conservative
    95$\%$ significance levels.  As expected the
    higher-resolution simulation (LLC4320) essentially extends the spectra of the
-   lower-resolution simulation (LLC2160) towards smaller scale, demonstrating that
+   lower-resolution simulation (LLC2160) towards smaller scales, demonstrating that
    the higher resolution is resolving a wider array of scales. Between 20 and 100 km
@@
    \caption{Potential density: a comparison between Roemmich-Gilson Argo
-           climatology [updates from \cite{roemmich_gilson2009}] (a, b) and the LLC simulations 1/24$^\circ$ (c, d)
-           and  1/48$^\circ$ (e, f). Of course, the Argo climatology is smoother
-           owing to lower resolution and long-term averaging. Nonetheless, simulations reproduce well the strong
+           climatology [updates from \cite{roemmich_gilson2009}] (a, b) and the 1/24$^\circ$ (c, d)
+           and  1/48$^\circ$ (e, f) LLC simulations. Of course, the Argo climatology is smoother,
+           owing to lower resolution and long-term averaging. Nonetheless, simulations reproduce the strong
            seasonality in the stratification. The upper ocean is weakly stratified in late winter/early fall
            and re-stratifies in summer. Argo climatology and LLC simulations mixed-layer depths are consistent
-           at all seasons (not shown). The Argo climatology is available online at
+            during all seasons (not shown). The Argo climatology is available online at
            \texttt{http://sio-argo.ucsd.edu/RG\_Climatology.html}.}
@@
       \includegraphics[width=40pc]{figs/figA1.pdf}
-   \caption{A comparison between frequency spectra of LLC simulations and two
+   \caption{Comparison between frequency spectra of LLC simulations and two
    available moored current meter records.}
@@
    near-surface shape and amplitude of baroclinic pressure modes.
-   In particular, the surface amplitude is much large in summer. Thus the projection
-   of baroclinic tides on the surface may vary significantly seasonally. The WOA data is
+   In particular, the surface amplitude is much larger in summer. Thus the projection
+   of baroclinic tides on the surface may have significant seasonal variations. The WOA data is
    available online at \texttt{https://www.nodc.noaa.gov/OC5/woa13/}.}
@@
  \hline
- Simulation & Resolution & Time-step & Period  & Tides \\ \hline
- ECCO2 Adjoint & $1/6^{^\circ}$ & 1200 s & January 2009 - December 2011 & No\\
- LLC 1080   & $1/12^{^\circ}$ & 90 s & January 2010 - July 2012 & No\\
+ Simulation & Grid spacing & Time-step & Period  & Tides \\ \hline
+ ECCO2 Adjoint & 18 km & 1200 s & January 2009 - December 2011 & No\\
+ LLC 1080   & $1/12^{^\circ}$ & 90 s & January 2010 - June 2012 & Yes \\
  LLC 2160   &  $1/24^{^\circ}$ & 45 s  & January 2011 - April 2013 & Yes\\
- LLC 4320   &  $1/48^{^\circ}$ & 25 s & September 2011 - September 2012 & Yes \\
+ LLC 4320   &  $1/48^{^\circ}$ & 25 s & September 2011 - October 2012 & Yes \\
  \hline

### 2016-07-13 writeup/rocha_etal.tex
Adds two new MITgcm and ECCO2 references

@@
 analyzed here, LLC2160 (1/24$^\circ$) and LLC4320 (1/48$^\circ$),   are
-forward Massachusetts Institute of Technology general circulation model (MITgcm)
+forward Massachusetts Institute of Technology general circulation model \citep[MITgcm; ][]{marshall_etal1997}
 numerical solutions on a LLC grid  \citep{forget_etal2015} with
 90-vertical-levels. The coarser-resolution LLC simulation was
-spun up from an Estimating the Circulation and Climate of the Ocean, Phase II (ECCO2)
+spun up from an Estimating the Circulation and Climate of the Ocean, Phase II \citep[ECCO2; ][]{menemenlis_etal2008}
 adjoint-method state estimate, constrained to millions
@@
 % Text
-\begin{thebibliography}{16}
+\begin{thebibliography}{18}
 \providecommand{\natexlab}[1]{#1}
@@
 
+  \bibitem[{\textit{Marshall et~al.}(1997)\textit{Marshall, Hill, Perelman, and
+    Adcroft}}]{marshall_etal1997}
+  Marshall, J., C.~Hill, L.~Perelman, and A.~Adcroft (1997), Hydrostatic,
+    quasi-hydrostatic, and nonhydrostatic ocean modeling, \textit{Journal of
+    Geophysical Research: Oceans}, \textit{102}(C3), 5733--5752.
+
   \bibitem[{\textit{McWilliams}(2016)}]{mcwilliams2016}
@@
     \textit{Proc. R. Soc. A}, vol. 472, p. 20160117, The Royal Society.
+
+  \bibitem[{\textit{Menemenlis et~al.}(2008)\textit{Menemenlis, Campin, Heimbach,
+    Hill, Lee, Nguyen, Schodlok, and Zhang}}]{menemenlis_etal2008}
+  Menemenlis, D., J.-M. Campin, P.~Heimbach, C.~Hill, T.~Lee, A.~Nguyen,
+    M.~Schodlok, and H.~Zhang (2008), Ecco2: High resolution global ocean and sea
+    ice data synthesis, \textit{Mercator Ocean Quarterly Newsletter},
+    \textit{31}, 13--21.
 

### 2016-07-14 writeup/rocha_etal.tex
Fixes small typos

@@
 of observations from 2009 through 2011. Both simulations were forced by
-tides and 6-hourly surface atmospheric fields. The LLC2160
+tides hourly and by 6-hourly surface atmospheric fields. The LLC2160
 output spans two years from March 2011 to April 2013; the LLC4320 was spun up from
@@
 
-\section{Statistics of the surface lateral velocity tensor}
+\section{Statistics of the surface lateral velocity gradient tensor}
 To study the seasonality in the surface velocity, we calculate the lateral velocity gradient tensor
@@
 \section{Summary and Conclusion}
-Our work adds to recent studies that present modeling \citep{sasaki_etal2014}
+Our work adds to recent modeling \citep{sasaki_etal2014}
 and observational \citep{callies_etal2015,buckingham_etal2016} evidence of vigorous seasonality in
@@
 Near-inertial wave generation peaks in winter \citep[e.g.,][]{alford2003}, but
-those waves project on large horizontal scales \citep[e.g, ][]{qi_etal1995}.
+those waves project onto large horizontal scales \citep[e.g, ][]{qi_etal1995}.
 

### 2016-07-14 writeup/rocha_etal.tex
Fixes tiny typo.

@@
 near-surface submesoscale inertia-gravity waves in the Kuroshio Extension undergo
-a strong seasonal cycle that is out-of-phase with the seasonal cycle of
+a strong seasonal cycle that is out of phase with the seasonal cycle of
 submesoscale turbulence. We conjecture that the summertime dominance of inertia-gravity waves

### 2016-07-15 writeup/rocha_etal.tex
Adds link to repo in the acknowledgement

@@
 The LLC output can be obtained from the ECCO project (\texttt{http://ecco2.org/llc$\_$hires}). The altimeter products were produced by Ssalto/Duacs
-and distributed by Aviso, with support from CNES (\texttt{http://www.aviso.altimetry.fr/duacs/}).
+and distributed by AVISO, with support from CNES (\texttt{http://www.aviso.altimetry.fr/duacs/}).
+Codes and output files are available online at the project repository
+ \texttt{https://github.com/crocha700/UpperOceanSeasonality}.
 \end{acknowledgments}

### 2016-07-15 writeup/rocha_etal.tex
Fixes typo

@@
 \cite{dasaro1978} showed that the velocity of linear internal waves
-in the mixed layer strongly depends on the density jump at the mixed-layer
+in the mixed layer strongly depends on the density jump at the mixed layer
 base, with largest mixed layer velocities when the jump is strongest.
 In summer, the shallow mixed layer overlays a strong seasonal pycnocline,
-and thus the internal waves projection onto the mixed-layer may be stronger
+and thus the internal waves projection onto the mixed layer may be stronger
 according to \cite{dasaro1978}'s arguments. An

### 2016-07-15 writeup/rocha_etal.tex
Adds thanks to Greg
Adds parenthesis

@@
 \begin{acknowledgments}
- William R. Young provided helpful feedback on the first draft. We thank the MITgcm community and our colleagues at the NASA Advanced
+ William R. Young provided helpful feedback on the first draft. Greg Wagner first
+ suggested that the near-surface shape of the baroclinic modes may change seasonally.
+ We thank the MITgcm community and our colleagues at the NASA Advanced
 Supercomputing (NAS) Division for their awesome support.
@@
 Codes and output files are available online at the project repository
- \texttt{https://github.com/crocha700/UpperOceanSeasonality}.
+ (\texttt{https://github.com/crocha700/UpperOceanSeasonality}).
 \end{acknowledgments}

### 2016-07-15 writeup/rocha_etal.tex
Small changes

@@
 the present results indicate that inertia-gravity waves account for most of the
-summertime submesoscale SSH variability.
+summertime submesoscale sea-surface height variability.
 
 \section{The LLC numerical simulations}
-We use the output of two latitude-longitude polar cap (LLC)
+We use results of two latitude-longitude polar cap (LLC)
 realistic numerical simulations. The outputs
@@
 output spans two years from March 2011 to April 2013; the LLC4320 was spun up from
-the LLC2160 simulation, spanning one year from September 2011 to September 2012.
+the LLC2160 simulation, spanning one year from September 2011 to October 2012.
 The LLC4320 simulation is an extension of the 3-month long output used by
@@
 We analyze a sub-domain of the LLC4320 and LLC2160 simulations of about 2000 km$^2$
-spanning 155-175$^\circ$E; 25-40$^\circ$N (Figure \ref{fig1}). The stratification
+spanning 155-175$^\circ$E; 25-40$^\circ$N (Figure \ref{fig1}a). The stratification
 in this mesoscale-rich subtropical region undergoes a vigorous seasonal cycle: wintertime-enhanced
 small-scale turbulence de-stratifies the upper ocean, yielding mixed layers
-as deep as 300 m. In late spring/early summer enhanced solar radiation and mixed-layer
-instabilities re-stratify the upper ocean, yielding mixed layers as shallow as 40 m.
+as deep as 300 m (Figure \ref{fig1}d). In late spring/early summer enhanced solar radiation and mixed-layer
+instabilities re-stratify the upper ocean, yielding mixed layers as shallow as 40 m (Figure \ref{fig1}e).
 Fundamentally, the upper-ocean density structure is well-captured by both LLC simulations:
 a comparison with Argo climatology shows that both simulations skillfully represent the Kuroshio
-Extension stratification and its seasonal variability  (supplemental material).
+Extension stratification and its seasonal variability  (supporting information).
 
@@
 with submesoscale flows: smoothing the velocity fields
-with a Hanning filter with cut-off scale of 100 km dramatically reduces the RMS
+with a Hanning filter with cut off scale of 100 km dramatically reduces the RMS
 vorticity and strain rate. The reduction in variance is about 80$\%$ in winter,
@@
  surface velocity is strongly dominated by submesoscale turbulence.
- The hourly velocity fields are significantly
-ageostrophic as depicted by the jPDF of vorticity-Laplacian of SSH
-winter (Figure \ref{fig3}e). Even in April, only the daily-averaged fields
+ The hourly velocity and sea-surface heigh fields have an important
+ageostrophic component as depicted by the jPDF of vorticity-Laplacian of SSH
+winter (Figure \ref{fig3}e). In other words, even in April, when submesoscale turbulence
+prevails, only the daily-averaged fields
 are largely
-in  geostrophic balance (the jPDF of daily-averaged vorticity vs.\\ Laplacian of SSH
+in  geostrophic balance (the jPDF of daily-averaged vorticity vs.\\ Laplacian of sea-surface height
 is an ellipse with large eccentricity and main axis tilted by 45$^\circ$; Figure \ref{fig3}f).
@@
 that persist in summer) and one with a very steep slope.
-That the submesoscale dynamics in October are mainly unbalanced is clearly depicted
-in the shape of jPDF of vorticity-Laplacian of SSH for hourly fields
+That the submesoscale dynamics in October are mainly ageostrophic is clearly depicted
+in the shape of jPDF of vorticity-Laplacian of sea-surface height for hourly fields
  (Figure \ref{fig3}g), which is an ellipse aligned in the vertical axis.
@@
 the daily-averaged flow is essentially geostrophic as depicted by the 45$^\circ$-tilted
-ellipse in the vorticity-Laplacian of SSH jPDF.
-
- Time series of PDFs of vorticity and divergence (supplemental material) show
+ellipse in the vorticity-Laplacian of sea-surface height jPDF.
+
+ Time series of PDFs of vorticity and divergence (supporting information) show
  a strong oscillation between these two regimes. In late winter/early spring
@@
 scales, we calculate wavenumber spectra of kinetic energy and sea-surface height
-(SSH) variance.  Before calculating the spectra, time-mean and spatial linear
+ variance.  Before calculating the spectra, time-mean and spatial linear
 trends were removed, and the resulting fields were multiplied by a two-dimensional Hanning ``spectral window''.
@@
 We discuss only spectra for the 1/48$^\circ$ simulation, which extends the 1/24$^\circ$ simulation
-towards smaller scales (supplemental material).
+towards smaller scales (supporting information).
 
@@
     \includegraphics[width=1.\textwidth]{figs/fig4.pdf}
- \caption{Surface (horizontal) KE (a) and SSH variance wavenumber spectra (b)
+ \caption{Surface (horizontal) kinetic energy (a) and sea-surface height variance wavenumber spectra (b)
  in the 1/48$^\circ$ simulation. Solid lines
@@
 Figure \ref{fig4}a depicts the horizontal wavenumber spectra of surface KE in
-April and October. At scales larger than 20km, the April and October spectra based on hourly
+April and October. At scales larger than 20 km, the April and October spectra based on hourly
 velocity snapshots (solid lines) are nearly indistinguishable from each other within 95$\%$ confidence level.
@@
 flows as opposed to 79$\%$ in October. The seasonality of subinertial submesoscale flows
-is more dramatic than the seasonal cycle of the total flow. This is consistent
+is strong, consistent
 with the results of \citet{sasaki_etal2014}, which are
@@
 There is a dramatic difference between the spectra
-based on hourly and daily-averaged SSH in October (Figure \ref{fig4}b): at scales smaller than 100 km, the spectra
-of hourly SSH roughly follows a $-2$ power-law, whereas the spectra of daily-averaged
-SSH roughly follows a $-5$ power-law. At these scales, 33$\%$ of the SSH variance in April is accounted for by super-inertial
+based on hourly and daily-averaged sea-surface height in October (Figure \ref{fig4}b): at scales smaller than 100 km, the spectra
+of hourly sea-surface height roughly follows a $-2$ power-law, whereas the spectra of daily-averaged
+sea-surface height roughly follows a $-5$ power-law. At these scales, 33$\%$ of the sea-surface height variance in April is accounted for by super-inertial
 flows as opposed to 83$\%$ in October! Curiously, the out-of-phase seasonal cycle of submesoscale
 turbulence and near-surface submesoscale inertia-gravity waves conspire to yield weak seasonality
-in the spectra of KE and SSH variance based on hourly fields.
+in the spectra of KE and sea-surface height variance based on hourly fields.
 
@@
 significantly more surface intensified in late summer/early fall
- (supplemental material).
+ (supporting information).
 If the internal wave source has weak seasonal dependence, then a strong

### 2016-07-15 writeup/rocha_etal.tex
Fixes capitalization of ECCO2 in Menemenlis reference

@@
   Menemenlis, D., J.-M. Campin, P.~Heimbach, C.~Hill, T.~Lee, A.~Nguyen,
-    M.~Schodlok, and H.~Zhang (2008), Ecco2: High resolution global ocean and sea
+    M.~Schodlok, and H.~Zhang (2008), ECCO2: High resolution global ocean and sea
     ice data synthesis, \textit{Mercator Ocean Quarterly Newsletter},

### 2016-08-14 writeup/RochatEtal_ResponseToReviewers.tex
Starts ResponseToReviewers document

\documentclass[11pt]{article}

%% WRY has commented out some unused packages %%
%% If needed, activate these by uncommenting
\usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
%\geometry{letterpaper}                   % ... or a4paper or a5paper or ...
\geometry{a4paper,left=2.5cm,right=2.5cm,top=2.5cm,bottom=2.5cm}
%\geometry{landscape}                % Activate for rotated page geometry
%\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent

\usepackage{natbib}

%for figures
%\usepackage{graphicx}

\usepackage{color}
\definecolor{mygreen}{RGB}{28,172,0} % color values Red, Green, Blue
\definecolor{mylilas}{RGB}{170,55,241}
%% for graphics this one is also OK:
\usepackage{epsfig}

%% AMS mathsymbols are enabled with
\usepackage{amssymb,amsmath}

%% more options in enumerate
\usepackage{enumerate}
%\usepackage{enumitem}

%% insert code
\usepackage{listings}

\usepackage[utf8]{inputenc}
\usepackage{multicol}
\setlength{\columnsep}{1cm}

\usepackage{hyperref}

\usepackage{fourier, heuristica}
\usepackage{array, booktabs}
\usepackage{graphicx}
\usepackage[x11names]{xcolor}
\usepackage{colortbl}
\usepackage{caption}
\DeclareCaptionFont{blue}{\color{LightSteelBlue3}}

\newcommand{\foo}{\color{LightSteelBlue3}\makebox[0pt]{\textbullet}\hskip-0.5pt\vrule width 1pt\hspace{\labelsep}}

\newcommand{\R}[1]{\textbf{R}\textbf{#1}}
\newcommand{\ie}{i.e.}

\newcommand{\sZ}{\mathsf{Z}}
\newcommand{\sE}{\mathsf{E}}
%\newcommand{\sA}{\mathsf{A}}

\title{{\normalfont Response to reviewers of the manuscript \texttt{2016GL070261} }
       ``Seasonality of submesoscale
       dynamics \\in the Kuroshio Extension''}
\author{Cesar B. Rocha,
        Sarah T. Gille, Teresa K. Chereskin, \& Dimitris Menemenlis}
\date{}
\begin{document}

%\include{symbols}

\maketitle

\section{Reviewer \#1}

{\it This is a timely investigation into the submesoscale sea surface height (SSH) and
velocity variations in the Kuroshio Extension region using global high-resolution
OGCM simulations. The main conclusion of the study that the surface ocean submesoscale
turbulence and inertia-gravity waves modulate out-of-phase seasonally, is novel and
important. Overall, the manuscript is well presented and clearly written and I
recommend its publication in GRL after the following comments are addressed.}\\

We thank reviewer \#1 for the speedy review. Below we address the reviewer’s
comments and suggestions.

\begin{enumerate}

\item {\it In the bottom paragraph on P.10, the authors emphasized that "At scales larger
than 20km, the April and October spectra based on hourly velocity snapshots are
nearly indistinguishable from each other". This emphasis appears to be contradictory
to the statements in other places of the manuscript that the seasonality of subinertial
submesoscale flows is strong (e.g., those relating to Figure 1b-c and Figure 2).}

\item {\it Does the mesoscale eddy variability in the Kuroshio Extension region of the model
have a similar energy level as those in the along-track altimeter data? This point
is worth mentioning as it impacts on the relative importance between the instability-induced
turbulence versus intia-gravity waves.}

\item {\it P.4, line 2 from bottom: Missing a word after "early".}

\item {\it P.6, line 5-7: Should October 15 be part of "late spring/early summer"?}

\item {\it P.7, line 12: Change "KE spectra" to "KE spectrum"?}

\item {\it P.8, bottom: The black line in Figure 2c shows a non-zero divergence value.
      Shouldn't the gridded SSH data be non-divergent by definition?}

\item {\it P.11, line 4-6: Figure 4a reveals that daily averaging not only reduces the
      kinetic energy level for scales smaller than 250km, but also those longer
      than 500km. What part of the large-scale flow signals are being removed?}

\end{enumerate}

%
% Reviewer 2
%

\section{Reviewer \#2}
{\it In this paper, the authors describe seasonality of submesoscale (10-100 km)
activities in Kuroshio Extension region by using outputs from their new high
resolution numerical simulations. Authors found that submesoscale turbulence
dominates in late winter and early spring because of mixed layer instabilities, and
inertial-internal waves in late summer and early fall because of internal tides and
stratification near the surface. This new simulation, which includes tides and a high
frequency wind forcing, and authors' attempts to analyze surface submesoscale
turbulence and near-inertial waves are potentially important to oceanographic
community. However, some discussions seem weak, possibly because it is very
difficult to separate sub- and super-inertial motions in the submeso dynamics as
authors mentioned.}\\

We thank reviewer \#2 for the speedy review. Below we address the reviewer’s
comments and suggestions.

\subsection{Major points}

\begin{enumerate}

  \item {\it First of all, generation mechanism of submesoscale inertial-gravity waves in summer
        and early fall is unclear to me. In the paper, authors attribute the mechanism to
        internal tides (X-12, section 6, last sentence) and emphasize using the tidal forcing in
        their model (X-5, section 2, 2nd paragraph). However, authors do not explain or show
        any interaction mechanism of internal tides (please at least add some reference). With
        the fixed frequencies of tidal constituents and existence of the critical latitude for the
        diurnal tides, I am not sure what mechanism can represent the surface near-inertial
        wave generation within the whole analyzed region by internal tides (with the
        horizontal scales of 10-100km). If authors compare Figures 2 and 4 with those
        estimated from the non-tidal forcing run in the Supplemental material, authors'
        conclusion could be strengthened.}

  \item {\it I also wonder why authors ignore wind-induced near-inertial waves (X-13, the first
        line). It is well studied that wind-induced near-inertial motion in mixed layer in back
        ground flow fields disperse quickly and propagate into the ocean interior (e.g., Young
        and Ben-Jelloul 1997, J. Mar. Res., 55, 735-66). The distribution of those waves
        depends on the distribution of vorticity (or Laplacian of vorticity) of background
        flows (Klein and Llewellyn Smith 2001, J. Mar. Res., 59, 697-723; Klein et al. 2004, Q.
        J. R. Meteorol. Soc., 130, 1153-1166). Therefore, those wind-induced near-inertial
        waves could have the horizontal scales of 10-100km. Note that those studies are based
        on geostrophically balanced or quasigeostrophic turbulent field. However, authors
        may need to study interactions between submeso dynamics regime (Rossby number
        and Richardson number are close to 1) and near-inertial waves in winter and spring
        cases where I guess that non-linear terms may play a role. I don't think that authors
        need to explain all internal wave dynamics in this GRL paper. I would rather suggest
        to remove all discussions related to submesoscale inertial-gravity waves and focus on
        describing differences between October and April.}

  \item {\it Finally, I don't understand why authors define 10-100 km as a submesoscale and used
        this scale for the spatial filtering (X-7, section 3). Authors should discuss Figure 4 first
        and then give a reason why they define 10-100 km as a submesoscale and use this
        scale for the filtering. In X-11, section 5, author point out that a part of the dynamics
        around 250 km scale should be explained by near-inertial waves. Then, I wonder why
        authors decide to discuss internal waves in 10-100 km scale.}
\end{enumerate}

\subsection{Minor points}

\begin{enumerate}

  \item {\it X-4, section 1, 2nd paragraph and X-13, section 6 last: Authors claimed that they
        analyze the Kuroshio Extension region. However, their study region does not
        include between $~$140E and $~$155E where the Kuroshio Extension is strongest in
        Figure 1(a). Please explain why you exclude this region in the paper. Since those
        longitudes includes the Izu-Ogasawara ridge (internal tide generation) and
        formation region of Subtropical mode water (strong mixed layer front), I would
        expect that including these latitudes would strengthen authors conclusions. I
        would also like to point out that positions of KEO and KESS moorings in the
        supplemental material are outside of the study region.}

  \item {\it X-4, section 1, 3rd paragraph: “submesoscale inertial-gravity waves”. I feel that this
        nomenclature is too sudden. It might be better to write “inertial-gravity waves
        with a horizontal scales of 10-100 km (hereafter submesoscale inertial-gravity
        waves)”?}

  \item {\it X-9, section 4, 2nd paragraph: I wonder what we could learn from the sentence “In
          other words, even in April, when submesoscale turbulence prevails, only the
          daily-averaged fields are largely in geostrophic balance”. This suggests that the
          balanced and upward cascade motion (originated from the submesoscale
          instability) scenario in Sasaki et al., (2014, nature comm. 5) could be wrong
          because they use the daily averaged data (X-11, section 5, last of 1st paragraph)?
          Or it just means that the balanced regime is actually the submeso-dynamics
          regime?}

  \item {\it X-10, section 4, 2nd paragraph, the last sentence: Please add some reference.}

\end{enumerate}

\section{Reviewer \# 3}
        {\it This manuscript provides model-based evidence of a seasonal cycle in submesoscale motions
         in the western North Pacific (Kuroshio Extension). The manuscript summarizes the results of
         a study in which two years of moderate-resolution (1/24th degree or 4.6 km) and one year of
         high-resolution (1/48th degree or 2.3 km) model output were examined for evidence of
         submesoscale activity. Measures of submesoscale activity include vorticity, divergence and
         strain rate and the authors demonstrate that these quantities are enhanced in winter. Kinetic
         energy and sea surface height (SSH) spectra also show seasonal trends. These results are
         consistent with an existing suite of studies reporting seasonality to submesoscale flows.

         What is relevant for the present study and distinguishes it from previous studies is that the
         authors report an increase in internal wave activity during summer and which (they suggest)
         projects onto the SSH variance. Because of the relevance of this study to challenges facing
         the upcoming Surface Water Ocean Topography (SWOT) mission—which aims at measuring
         submesoscale flows from space from nearly-balanced flows—this study will be of interest to
         readers of GRL. The major error that should be addressed prior to publication is a more direct
         connection of divergence and increased SSH variance to inertia-gravity waves. At present,
         this connection is entirely speculative since any ageostrophic flows can give rise to
        divergence and SSH variance. Edits should be made to the text and figures, as well.}\\

        We thank  Christian Buckingham for the speedy and thorough review. Below we address
                  Buckingham's comments and suggestions.

\subsection{Major comments}

\begin{enumerate}

  \item {\it The authors must clearly demonstrate that increased inertia-gravity waves (IGWs) are
        responsible for the seasonal signal in horizontal divergence and SSH spectra at high
        wavenumbers. The authors argue that this is so but it is not demonstrated with sufficient
        clarity. It is true that a portion of this energy/variance will be attributed to internal waves
        modified by rotation but other flows give rise to divergence (and thus, SSH variance). Thus,
        the seasonal changes reported by the authors might be the result of seasonal changes in these
        other flows. Given that this is the main message of the paper, the authors must make this
        modification.}

  \item {\it There is very little introduction in this manuscript. The authors must address this
          shortcoming prior to publication. The obvious method of addressing this issue is to discuss
          the seasonality of IGWs in the context of SWOT and why this might matter. The introduction
          should, therefore, place this study matter?”}

  \item {\it There needs to be a discussion about the effective resolution of your model. The fact that
        the model has a resolution of dx = 2.3 km and 4.6 km does not mean it can resolve oceanic
        phenomena at this scale. Its effective horizontal resolution is closer to 8 times the grid
        resolution (Soufflet et al. 2013). This corresponds to 18 km and 35 km for LLC4320 and
        LLC2160 output, respectively. [A more exact value could be obtained by estimating zonal
        and meridional spectra (of u, v or SSH) and identifying the high wavenumbers / low
        wavelengths at which the spectral energy falls precipitously.] This means the model is not
        resolving a number of submesoscale coherent vortices (SCVs) that may occur as a result of
        ML baroclinic instability; these have diameters close to 5 km. So the follow-up questions
        from a dynamical point-of-view become (1) where does the unresolved energy come from or
        go to? and (2) how does this affect your results? These questions need to be addressed or at
        least discussed if the main topic of your paper is submesoscale dynamics.}

 \item {\it If the authors could estimate the magnitude of the sea surface height signature within a
        particular band as a function of season that might be helpful. For example, the authors could
        integrate SSH spectra (Figure 4b) from hourly fields between 10 km and 33 km (i.e.,
        wavenumber = 3 x 10-2 km-1 which appears to be the cross-over point for the two curves) to
        obtain SSH variances within a particular wavenumber band. Additionally, under the
        assumption that SSH variance in winter is dominated by balanced motions while SSH
        variance in summer is the sum of balanced + unbalanced motions, the difference between
        these two quantities yields a bound1 for the SSH variance due to unbalanced motions.}

\end{enumerate}

\subsection{Minor comments}

\begin{enumerate}

  \item {\it (Section 1, Paragraph 1) The authors use the term ``submesoscale'' and ``mesoscale'' as
        descriptive of motions having (1-100-km) and (100-300 km) horizontal scales,
        respectively. I find these definitions misleading.

        One way to define mesoscale and submesoscale is to define what scales of motion a
        particular sensor will resolve. While this would not be optimal, it is objective. This is
        probably the motivation for the definition used by Callies and Ferrari (2013), since the
        swath altimeter (as initially proposed) would resolve scales greater than 1 km but
        smaller than the present-day, profiling satellite altimeters – i.e., eddy-like features
        with diameters greater than 80-100 km (see Chelton et al. 2011, Progress).
        Unfortunately, this definition has the adverse effect of changing for each sensor.

        A better definition of the submesoscale might be founded on dynamical arguments.
        Wunsch and Stammer (1998) introduce the geostrophic equations and note that the
        lower limit for which this approximation is strictly valid is ~30 km. This conveniently
        defines the lower limit of the mesoscale regime. I would therefore define the
        submesoscale regime as anything smaller than this but one for which the hydrostatic
        approximation remains valid (e.g., 1 km).

      The motivation for wanting to use the term ``submesoscale'' in the title is clear –
      submesoscale motions are a hot topic within the community and it catches individuals’
      eyes. This is fine. But what must be made clear within the manuscript is what scales of
      motion and what force balances (i.e., dynamical regimes) are being described.}

  \item {\it (Abstract) The abstract could be rewritten to emphasize its connection to SWOT.}

  \item {\it (Section 1, last paragraph) Adding “with implications for SWOT” at the end of the
          introduction would also help this article.}

  \item {\it (Section 1, Paragraph 1) The authors mention a suite of papers documenting
        seasonality at the submesoscale but fail missed a few: Ostrovskii (1995) and
        Brannigan et al. (2015). These are observation- and model-based studies, resp.}

  \item {\it (Section 1, Paragraph 2) Model resolutions: it would be helpful for readers to place in
        parentheses nominal horizontal resolutions corresponding to these spatial resolutions.}

  \item {\it (Section 1, Paragraph 3) What types of horizontal and temporal scales for IGWs are
        we talking about here? How would this fall onto the Garrett-Munk (GM) internal
        wave spectrum? (Recall: the GM spectrum contains both balanced and unbalanced
        flows, not just internal waves, despite that the name suggests this.)}

 \item {\it Discretization of the underlying equations of motion. How do you expect the
        combination of vertical and horizontal resolution modifies internal wave properties
        that exist in your model? Might you be attributing characteristics to these waves that
        would be different in the real ocean?}

\item {\it There should be some discussion of the seasonality of the mixed layer deformation
      radius. Buckingham et al. (2016) point out that all modeling and observational studies
      to date necessarily introduce seasonally varying energy by virtue of the deformation
      radius changing with season and falling below the grid resolution in certain seasons.
      This might be worth a comment.}

\item {\it (Section 2, Paragraph 1) ``The LLC4320 simulation is an extension of the 3-month
      long output used by Rocha et al. (2016).'' I think this is good to mention, here, so
      please keep. In contrast, in (Section 5, Paragraph 2) the authors note a consistency
      between the results in the Kuroshio Extension and those in the Drake Passage. This is
      not a consistency as it is a different location, different time. The author has already
      made the reader aware of the previous study so there is no reason to cite again.}

\item {\it (Section 5, Paragraph 1) eliminate ``spectral'' after Hanning window. Also, no need for
        quotation marks.}

\item {\it (Section 5, Paragraph 1) ``the projection of these flows onto different horizontal scales
            ...'' Have you thought about computing scale-dependent vorticity, strain rate and
            divergence? This is really what you would be doing
            aforementioned flows into spectral space.}

\item {\it (Section 5, Paragraph 3) Remove the exclamation mark.}

\item {\it (Section 4, first paragraph) I understand the motivation for examining the Laplacian of
      sea surface height but the average reader will not understand the connection.}

\item {\it (Section 4, second paragraph) Last sentence. This is good. I understand this but to
        make this legible for the general audience it should be explained why we expect a near
        one-to-one line for the joint-PDF of vorticity and the Laplacian of SSH.}

\item {\it (Section 4, last paragraph) ``consistent with linear inertia-gravity waves''.
        It would be
        helpful to have a reference after this statement.}

\item {\it (Section 4) I find “jPDF” odd. Try “joint-PDF” or “PDF” in general.}

\item {\it (Section 4, last paragraph) ``...whereas divergence is moderately, negatively skewed as
        predicted by ...'' Specifically mention convergence/downwelling.}

\item {\it (Figure 1) ** Place panels (d) and (e) before panels (b) and (c) since this is the order
        in which you refer to them in the manuscript. ** Difficult to see contours in (d) and
        (e). ** Consider overlaying mixed layer depth on these transects since it is mentioned
        in the text.}

\item {\it (Figure 2) Is the mean of the strain rate zero? Is the mean of the divergence zero? The
        variance should be used rather than the root-mean-squared (RMS).}

\item {\it (Figure 2) What is the sensitivity of the estimates to the stencil type? See Arbic et al.
        paper describing this sensitivity for satellite altimetry.}

\item {\it (Section 3 and Figures 2 \& 3) At what depth are you computing these quantities?}

\item {\it (Section 2, last paragraph) It is not just solar radiation that enters the surface buoyancy
      flux calculation.}

\item {\it (Section 2, last paragraph) ``as shallow as 40 m'' – I bet that the mixed layer depths get
      even shallower than this. At midlatitudes, 20-30 m is not uncommon.}

\item {\it 24. (Section 3, Paragraph 1) ``These diagnostic highlight the submesoscale structures in
      the flow.'' It might help the authors to motivate their calculations using the following:

      Any horizontal velocity field, uh, can be expressed in terms of vorticity, divergence,
      strain rate and mean flow. This can be seen by expressing uh as a Taylor series
      expansion and then decomposing the velocity gradient into symmetric and anti-
      symmetric parts [Landau and Lifshitz, 1987]. Because vortices, vertical motion and
      fronts are ubiquitous at the submesoscale, this is a natural decomposition and
      variances of these quantities are apt descriptors of submesoscale turbulence.}

\item {\it (Figure 3) Would be nice to include a third row that contains the collapsed (i.e., one-
      dimensional) PDFs of normalized vorticity to illustrate the skewness of hourly and
      daily-averaged flow. I believe you already have this information in the supplementary
      material but would be nice to illustrate to the reader. Also, in your normalization of
      vorticity, are you accounting for the fact that the Coriolis parameter changes with
      latitude? Your domain spans considerable changes in f and the relevant dynamical
      parameter is the gradient Rossby number.}

\item {\it (Figure 4b) Strictly speaking, the label on the y-axis should not be SSH variance. SSH
      variance would be the integral of this curve. The y-axis should be labelled as the
      power spectral density of sea surface height (SSH), or SSH Spectra.}

\item {\it The authors may find the work of Brüggeman and Eden (2015) helpful in addressing
      some of the aforementioned questions.}

\end{enumerate}

\end{document}

### 2016-08-15 writeup/RochatEtal_ResponseToReviewers.tex
Updates reviews

@@
 %\newcommand{\sA}{\mathsf{A}}
+
+\newcommand{\bdp}{\begin{description}}
+\newcommand{\edp}{\end{description}}
 
@@
 
-\item {\it (Section 4) I find “jPDF” odd. Try “joint-PDF” or “PDF” in general.}
+\item {\it (Section 4) I find “jPDF” odd. Try “joint-PDF” or “PDF” in general.}\\
+
+We changed jPDF with joint-PDF.
 
@@
 \item {\it 24. (Section 3, Paragraph 1) ``These diagnostic highlight the submesoscale structures in
-      the flow.'' It might help the authors to motivate their calculations using the following:
+      the flow.'' It might help the authors to motivate their calculations using the following:}\\
 
-      Any horizontal velocity field, uh, can be expressed in terms of vorticity, divergence,
+\begin{description}
+      {\it Any horizontal velocity field, uh, can be expressed in terms of vorticity, divergence,
       strain rate and mean flow. This can be seen by expressing uh as a Taylor series
@@
       variances of these quantities are apt descriptors of submesoscale turbulence.}
+\end{description}
 
@@
       variance would be the integral of this curve. The y-axis should be labelled as the
-      power spectral density of sea surface height (SSH), or SSH Spectra.}
+      power spectral density of sea surface height (SSH), or SSH Spectra.}\\
+
+      \bdp
+      We labelled the y-axis of the SSH variance spectrum as \textit{SSH variance density}
+      --- short for \textit{SSH variance spectral density} --- not \textit{SSH variance} as the reviewer
+      suggests.  We chose to use the short term to avoid cluttering the label
+      since there is no ambiguity because we include the
+      units of SSH variance per unit wavenumber.
+      We disagree that the label of the y-axis should be \textit{power spectrum of SSH}
+      or \textit{SSH spectra} [{\it sic}]. On physical grounds, \textit{power spectrum} is a misleading term,
+      because SSH variance density has nothing to do with \textit{power}.\footnote{The
+      term \textit{power spectrum} was for the \textit{frequency} spectrum of some quantities.}
+      And ``SSH [wavenumber] spectrum'' is the plot of SSH variance spectral density
+      as a function of wavenumber, not the quantity in the y-axis.\footnote{This confusion is similar
+      to the pervasive axis mislabelling in plots of probability distribution function (PDF)
+      --- the correct label of the y-axis is \textit{probability density}
+      not PDF.}
+      \edp
 

### 2016-08-15 writeup/rocha_etal.tex
Starts changing text for review1
Updates reviews

@@
 
-% \usepackage{lineno}
-% \linenumbers*[1]
-%  To add line numbers to lines with equations:
+%\usepackage{lineno}
+%\linenumbers*[1]
+%To add line numbers to lines with equations:
 %  \begin{linenomath*}
@@
 
-The April vorticity-strain jPDF has a shape characteristic of submesoscale turbulence
+The April vorticity-strain joint-PDF has a shape characteristic of submesoscale turbulence
  (Figures \ref{fig3}a-b). The alignment of vorticity and strain $\alpha \sim \pm\zeta$
@@
 \begin{acknowledgments}
- William R. Young provided helpful feedback on the first draft. Greg Wagner first
+ William R. Young and the referees --- Christian Buckingham and two anonymous ---
+  provided constructive and helpful comments. Greg Wagner first
  suggested that the near-surface shape of the baroclinic modes may change seasonally.

### 2016-08-15 writeup/rocha_etal.tex
Start drafting response

@@
 strong seasonal modulation of upper-ocean submesoscale dynamics:
-submesoscale turbulence dominates the upper-ocean dynamics in late winter/early,
+submesoscale turbulence dominates the upper-ocean dynamics in late winter/early fall,
 whereas submesoscale inertia-gravity waves prevail in late summer/early fall.
@@
 small-scale turbulence de-stratifies the upper ocean, yielding mixed layers
-as deep as 300 m (Figure \ref{fig1}d). In late spring/early summer enhanced solar radiation and mixed-layer
-instabilities re-stratify the upper ocean, yielding mixed layers as shallow as 40 m (Figure \ref{fig1}e).
+as deep as 300 m (Figure \ref{fig1}d). Summertime re-stratification yields mixed layers shallower
+than 40 m (Figure \ref{fig1}e).
 Fundamentally, the upper-ocean density structure is well-captured by both LLC simulations:
@@
 than in late summer/early fall. Because the wintertime vorticity and strain rate
-are dominated by the smallest scales in the flow (the KE spectra is shallower than
+are dominated by the smallest scales in the flow (the KE spectrum is shallower than
 a $-3$ power law in winter), increasing the resolution from

### 2016-08-15 writeup/RochatEtal_ResponseToReviewers.tex
Start drafting response
Minor addition
Fixes typos.

@@
 is worth mentioning as it impacts on the relative importance between the instability-induced
-turbulence versus intia-gravity waves.}
+turbulence versus inertia-gravity waves.}
 
@@
 
+  Fixed.
+
 \item {\it P.6, line 5-7: Should October 15 be part of "late spring/early summer"?}
 
+  We re-wrote that sentence to avoid ambiguity.
+
 \item {\it P.7, line 12: Change "KE spectra" to "KE spectrum"?}
+
+Fixed.
 
@@
 
+  The convergence of meridians (the $\beta$-effect) accounts for the
+  small divergence depicted by the black line in Figure 2c.
+
 \item {\it P.11, line 4-6: Figure 4a reveals that daily averaging not only reduces the
@@
       than 500km. What part of the large-scale flow signals are being removed?}
+
+  Barotropic tides.
 
@@
 \item {\it (Section 5, Paragraph 1) eliminate ``spectral'' after Hanning window. Also, no need for
-        quotation marks.}
+        quotation marks.}\\
+        OK.
 
@@
 
-\item {\it (Section 5, Paragraph 3) Remove the exclamation mark.}
+\item {\it (Section 5, Paragraph 3) Remove the exclamation mark.}\\
+
+    Removed.
 
@@
       flux calculation.}
+
+      We re-wrote that sentence.
 

### 2016-08-15 writeup/rocha_etal.tex
Fixes typos.

@@
  variance.  Before calculating the spectra, time-mean and spatial linear
-trends were removed, and the resulting fields were multiplied by a two-dimensional Hanning ``spectral window''.
+trends were removed, and the resulting fields were multiplied by a two-dimensional Hanning window.
 The two-dimensional spectra were averaged azimuthaly \citep[e.g., ][]{rocha_etal2016} .
@@
 sea-surface height roughly follows a $-5$ power-law. At these scales, 33$\%$ of the sea-surface height variance in April is accounted for by super-inertial
-flows as opposed to 83$\%$ in October! Curiously, the out-of-phase seasonal cycle of submesoscale
+flows as opposed to 83$\%$ in October. Curiously, the out-of-phase seasonal cycle of submesoscale
 turbulence and near-surface submesoscale inertia-gravity waves conspire to yield weak seasonality

### 2016-08-15 Argo/argo_kuroshio.py
Adds code for plotting argo climatology

@@ -1,0 +1,278 @@
+
+import matplotlib.pyplot as plt
+from mpl_toolkits.basemap import Basemap
+import scipy as sp
+import numpy as np
+import glob
+import pandas
+import seawater as sw
+import matplotlib.colors as mcolors
+from netCDF4 import Dataset
+import cmocean
+
+
+plt.close('all')
+
+#
+# settings
+#
+
+## mapping
+lonplot = (155,175)
+latplot = (25, 40)
+#
+# Dataset: Argo climatology
+#
+argopath='RG_ArgoClim_p5xp5_099_2015_annual.nc'
+
+
+argo = Dataset(argopath)
+
+lona = np.array(argo.variables['LONGITUDE'][:])
+lata = np.array(argo.variables['LATITUDE'][:])
+
+lat = lata[(lata > latplot[0]) & (lata < latplot[1])]
+lon = lona[(lona > lonplot[0]) & (lona < lonplot[1])]
+
+flat = (lata > latplot[0]) & (lata < latplot[1])
+flon = (lona > lonplot[0]) & (lona < lonplot[1])
+
+
+Tm =  np.array(argo.variables['ARGO_TEMPERATURE_MEAN'][:,flat,flon])
+Sm =  np.array(argo.variables['ARGO_SALINITY_MEAN'][:,flat,flon])
+Pm =  np.array(argo.variables['PRESSURE'][:])
+
+Ta =  np.array(argo.variables['ARGO_TEMPERATURE_ANNUAL_ANOMALY'][:,:,flat,flon])
+Sa =  np.array(argo.variables['ARGO_SALINITY_ANNUAL_ANOMALY'][:,:,flat,flon])
+
+lona = lona[(lona > lonplot[0]) & (lona < lonplot[1]+7)]
+lata = lata[(lata > latplot[0]) & (lata < latplot[1])]
+
+T = Tm[np.newaxis,...]+Ta
+S = Sm[np.newaxis,...]+Sa
+
+
+# A transect at 165
+ilon = 19
+
+lon = lon[ilon]
+
+T = T[...,ilon]
+S = S[...,ilon]
+
+
+D = np.empty_like(S)
+
+for i in range(12):
+    D[i] = sw.pden(S[i],T[i],Pm[...,np.newaxis],pr=0)-1000
+
+
+
+cs = np.array([22.,22.5,23.,23.5,24.,24.5,25.,25.5,26.,26.5])
+
+
[…]

### 2016-08-15 writeup/rocha_etal.tex
Adds figure 4

@@
           analyzed. Colors represent the topography and white lines are contours of absolute
-          dynamic topography every 0.1 m from AVISO. LLC 4320 (1/48$^\circ$) snapshots of surface vorticity (b and c) and transects
-          of potential density at 165$^\circ$E (d and e). The snapshots were
+          dynamic topography every 0.1 m from AVISO. LLC 4320 (1/48$^\circ$) snapshots of
+          transects
+          of potential density at 165$^\circ$E (b and c) and  surface vorticity (d and e) . The snapshots were
           taken at 00:00 UTC.}
@@
   \begin{center}
-    \includegraphics[width=1.\textwidth]{figs/fig4.pdf}
+    \includegraphics[width=.9\textwidth]{figs/fig4.pdf}
+ \caption{LLC4320 Wavenumber-frequency spectrum of SSH variance in (a) April and (b) October.
+          (c) The ratio between (b) and (a) --- the ratio is masked in regions
+          with SSH variance density smaller than $10^{-2}$ m$^2$/[cpkm $\times$ cph].
+          The light gray shaded region depicts the dispersion relations for inertia-gravity
+          waves from mode 1 through mode 4 across the latitudinal domain; horizontal dashed lines represent the semi-diurnal
+          lunar tidal frequency (M$_2$), and the inertial frequency at mid-latitude
+          (f$_{32.5}$) and its first harmonic (2f$_{32.5}$).}
+ \label{fig4}
+ \end{center}
+\end{figure*}
+
+\begin{figure*}[ht]
+  \begin{center}
+    \includegraphics[width=.85\textwidth]{figs/fig5.pdf}
  \caption{Surface (horizontal) kinetic energy (a) and sea-surface height variance wavenumber spectra (b)
@@
  fields.}
- \label{fig4}
+ \label{fig5}
  \end{center}
@@
 
-Figure \ref{fig4}a depicts the horizontal wavenumber spectra of surface KE in
+Figure \ref{fig5}a depicts the horizontal wavenumber spectra of surface KE in
 April and October. At scales larger than 20 km, the April and October spectra based on hourly
@@
 km, both in April and October (compare solid lines against dashed lines in Figure
-\ref{fig4}a). But this suppression is dramatic in October, when the inertia-gravity waves peak. At scales
+\ref{fig5}a). But this suppression is dramatic in October, when the inertia-gravity waves peak. At scales
 smaller than 100 km, 39$\%$ of the surface KE in April is accounted for by super-inertial
@@
 There is a dramatic difference between the spectra
-based on hourly and daily-averaged sea-surface height in October (Figure \ref{fig4}b): at scales smaller than 100 km, the spectra
+based on hourly and daily-averaged sea-surface height in October (Figure \ref{fig5}b): at scales smaller than 100 km, the spectra
 of hourly sea-surface height roughly follows a $-2$ power-law, whereas the spectra of daily-averaged
@@
 % To better understand the high-frequency/high-wavenumber flows  we calculate frequency
-% spectra of surface KE winter \ref{fig4}). To obtain statistically meaningful results,
+% spectra of surface KE winter \ref{fig5}). To obtain statistically meaningful results,
 % we average spectral estimates along a section at 165$^\circ$E.  There is about 50
@@
 % suppressed in supra-inertial frequencies --- there are no statistically significant
-% changes at sub-inertial frequencies (see dashed lines in figure \ref{fig4}).
+% changes at sub-inertial frequencies (see dashed lines in figure \ref{fig5}).
 

### 2016-08-16 writeup/rocha_etal.tex
Starts drafting text for figure 4

@@
  To characterize these differences, we calculate
-joint probability distributions (jPDF) of
+joint probability distributions (joint-PDF) of
 vorticity-strain and vorticity-Laplacian of sea-surface height (Figure \ref{fig3}).
@@
  \citep{shcherbina_etal2013,mcwilliams2016}. The shape of the vorticity-strain
- jPDF is similar for hourly and daily-averaged fields, although
+ joint-PDF is similar for hourly and daily-averaged fields, although
  the vorticity skewness reduces from 1.4 to 1.13 from hourly to daily-averaged.
@@
  The hourly velocity and sea-surface heigh fields have an important
-ageostrophic component as depicted by the jPDF of vorticity-Laplacian of SSH
+ageostrophic component as depicted by the joint-PDF of vorticity-Laplacian of SSH
 winter (Figure \ref{fig3}e). In other words, even in April, when submesoscale turbulence
@@
 are largely
-in  geostrophic balance (the jPDF of daily-averaged vorticity vs.\\ Laplacian of sea-surface height
+in  geostrophic balance (the joint-PDF of daily-averaged vorticity vs.\\ Laplacian of sea-surface height
 is an ellipse with large eccentricity and main axis tilted by 45$^\circ$; Figure \ref{fig3}f).
 
-The October vorticity-strain jPDF shows much weaker skewness (the vorticity skewness
+The October vorticity-strain joint-PDF shows much weaker skewness (the vorticity skewness
 is 0.68 and 0.67 for hourly and daily-averaged velocities). The shape of the
-vorticity-strain jPDF  appears to be a combination of two half-ellipses centered
+vorticity-strain joint-PDF  appears to be a combination of two half-ellipses centered
 about $\zeta=0$, one with a 45$^\circ$ slope (characteristic of submesocale fronts
@@
 That the submesoscale dynamics in October are mainly ageostrophic is clearly depicted
-in the shape of jPDF of vorticity-Laplacian of sea-surface height for hourly fields
+in the shape of joint-PDF of vorticity-Laplacian of sea-surface height for hourly fields
  (Figure \ref{fig3}g), which is an ellipse aligned in the vertical axis.
@@
 the daily-averaged flow is essentially geostrophic as depicted by the 45$^\circ$-tilted
-ellipse in the vorticity-Laplacian of sea-surface height jPDF.
+ellipse in the vorticity-Laplacian of sea-surface height joint-PDF.
 
@@
 
+To confirm that inertia-gravity undergo a vigorous seasonal cycle near the surface,
+we calculate wavenumber-frequency spectrum of SSH variance. Focusing on the high-frequency
+content, we compute the wavenumber-frequency spectrum  every 10 days and average
+the results to obtain a spectral estimate. Before calculating the each spectrum we
+remove linear trends and multiply the data by a three-dimensional Hanning window.
+The spectral estimate is then azimuthally-averaged in wavenumber.
+
+Figures \ref{fig4}a-b show the wavenumber-frequency of SSH variance for the LLC4320
+simulation. Mesoscales flows and the lunar semi-diurnal tide contain most of the
+SSH variance. But the is significant SSH variance at submesoscales, both subinertial
+and super-inertial. Subinertial SSH variance at submesoscales is larger in April, whereas
+super-inertial SSH variance at submesoscales is larger in October.  Figure \ref{fig4}c,
+which shows the ratio between the two spectra, clearly depicts the out-of-phase
+seasonality of super-inertial and subinertial submesoscale flows. More importantly,
+inertia-gravity
+waves account for most of the enhancement of super-inertial SSH variance in October.
+
 \section{Projection onto horizontal scales}
 
-To better quantify the projection of these flows onto different horizontal
+To quantify the projection of these flows onto different horizontal
 scales, we calculate wavenumber spectra of kinetic energy and sea-surface height

### 2016-08-16 writeup/rocha_etal.tex
Removes fool detail

@@
 \begin{center}
-\hspace{-1.25cm}\includegraphics[width=.75\textwidth]{figs/fig1_1.pdf}\\
+\hspace{-1.25cm}\includegraphics[width=.7\textwidth]{figs/fig1_1.pdf}\\
 \vspace{-.125cm}
-\includegraphics[width=.75\textwidth]{figs/fig1_2.pdf}
- \caption{(a) The study region with the subregion where the LLC outputs are
+\includegraphics[width=.7\textwidth]{figs/fig1_2.pdf}
+ \caption{\small (a) The study region with the subregion where the LLC outputs are
           analyzed. Colors represent the topography and white lines are contours of absolute
@@
    \begin{center}
-     \includegraphics[width=.75\textwidth]{figs/fig2.pdf}
-  \caption{Time series of the root-mean-square (RMS) of (a) vorticity,
+     \includegraphics[width=.65\textwidth]{figs/fig2.pdf}
+  \caption{\small Time series of the root-mean-square (RMS) of (a) vorticity,
   (b) rate of strain, and (c) horizontal divergence in the LLC outputs and gridded AVISO data.}
@@
 The picture that emerges is consistent with recent studies: shallow baroclinic
-instabilities energize the submesoscales in late winter, drawing from the available
+instabilities energize the submesoscales in winter, drawing from the available
 potential energy stored in large lateral buoyancy gradients in deep mixed layers \citep{sasaki_etal2014,callies_etal2015,callies_etal2016}.
@@
   \begin{center}
-    \includegraphics[width=.95\textwidth]{figs/fig3.pdf}
- \caption{Seasonal variation of joint probability distributions:  vorticity vs.\ strain rate (a through d),
+    \includegraphics[width=.75\textwidth]{figs/fig3.pdf}
+ \caption{\small Seasonal variation of joint probability distributions:  vorticity vs.\
+          strain rate (a through d),
           and vorticity vs.\ Laplacian of sea-surface height (e through h) in April (a, b,
@@
     \includegraphics[width=.9\textwidth]{figs/fig4.pdf}
- \caption{LLC4320 Wavenumber-frequency spectrum of SSH variance in (a) April and (b) October.
+ \caption{\small LLC4320 Wavenumber-frequency spectrum of SSH variance in (a) April and (b) October.
           (c) The ratio between (b) and (a) --- the ratio is masked in regions
@@
   \begin{center}
-    \includegraphics[width=.85\textwidth]{figs/fig5.pdf}
- \caption{Surface (horizontal) kinetic energy (a) and sea-surface height variance wavenumber spectra (b)
+    \includegraphics[width=.65\textwidth]{figs/fig5.pdf}
+ \caption{\small Surface (horizontal) kinetic energy (a) and sea-surface height variance wavenumber spectra (b)
  in the 1/48$^\circ$ simulation. Solid lines

### 2016-08-16 writeup/RochatEtal_ResponseToReviewers.tex
Fixes typo
Updates response to reviewers

@@
 
-  Barotropic tides.
+  \bdp
+  The large-scale super-inertial flows are barotropic tides. The new figure 4
+  supports this interpretation --- there is significant semi-diurnal lunar
+  tidal SSH variance (and KE) at scales larger than 500 km.
+  \edp
 
@@
 
+        \bdp
+          We changed the order of the panels.
+        \edp
+
 \item {\it (Figure 2) Is the mean of the strain rate zero? Is the mean of the divergence zero? The
@@
       units of SSH variance per unit wavenumber.
-      We disagree that the label of the y-axis should be \textit{power spectrum of SSH}
+      We disagree that the label of the y-axis should be \textit{power spectral density of SSH}
       or \textit{SSH spectra} [{\it sic}]. On physical grounds, \textit{power spectrum} is a misleading term,

### 2016-08-16 writeup/RochatEtal_ResponseToReviewers.tex
Updates response to reviewer

@@
       We disagree that the label of the y-axis should be \textit{power spectral density of SSH}
-      or \textit{SSH spectra} [{\it sic}]. On physical grounds, \textit{power spectrum} is a misleading term,
+      or \textit{SSH spectra} [{\it sic}]. On physical grounds, \textit{power} adds ambiguity,
       because SSH variance density has nothing to do with \textit{power}.\footnote{The
-      term \textit{power spectrum} was for the \textit{frequency} spectrum of some quantities.}
+      term \textit{power spectral density} was for the \textit{frequency} spectrum.}
       And ``SSH [wavenumber] spectrum'' is the plot of SSH variance spectral density
@@
       to the pervasive axis mislabelling in plots of probability distribution function (PDF)
-      --- the correct label of the y-axis is \textit{probability density}
+      --- the correct label is \textit{probability density}
       not PDF.}

### 2016-08-16 writeup/RochatEtal_ResponseToReviewers.tex
Adds numbering to fig4
Removes little dash

@@
 
-  We re-wrote that sentence to avoid ambiguity.
+  We rewrote that sentence to avoid ambiguity.
 
@@
 
-      We re-wrote that sentence.
+      We rewrote that sentence.
 
@@
       units of SSH variance per unit wavenumber.
-      We disagree that the label of the y-axis should be \textit{power spectral density of SSH}
+      We further disagree that the y-axis should be labelled \textit{power spectral density of SSH}
       or \textit{SSH spectra} [{\it sic}]. On physical grounds, \textit{power} adds ambiguity,

### 2016-08-16 writeup/RochatEtal_ResponseToReviewers.tex
Drafts response about the definition of submesoscales

@@
       eyes. This is fine. But what must be made clear within the manuscript is what scales of
-      motion and what force balances (i.e., dynamical regimes) are being described.}
+      motion and what force balances (i.e., dynamical regimes) are being described.}\\
+
+      \bdp
+          We respect the reviewer's opinion about the definition of submesoscales.
+          But there seems to be no clear definition in the literature --- the
+          Stammer $\&$ Wunsch paper that the reviewer states that their
+          definition of mesoscales is ``very rough'' and ``there is no generally
+          agreed upon definition''. There is also irritation that oceanographers
+          decided to use the term mesoscales for flows with dynamics analogous to
+          the meteorological synoptic scale.\\
+
+          Some investigators define mesoscales as flows with scales near the
+          1st baroclinic deformation radius. But there is significant confusion
+          owing to factors of  $2 \pi$ --- e.g., some investigators would use $R_1 = 30$ km,
+          others would argue for $L_1 = 2\pi \times 30 \approx 190$ km
+          (Wunsch \& Ferrari 2004; Ferrari \& Wunsch 2009).
+          Also, the scales of equilibrated baroclinic geostrophic turbulence are
+          larger than the deformation radius (e.g., Larichev \& Held 1995),
+          although linear theorists always deploy ``deformation radius'' as
+          the most relevant scale.\\
+
+          According
+          to our definition the mesoscales are the energy-containing scales of the
+          flow, the eddies that contain most of the eddy kinetic energy in the ocean
+          (Ferrari \& Wunsch 2009). On observational
+          grounds, this scale varies roughly from 100-300 km (from the centroid of KE wavenumber
+          spectrum from current altimeters for example). We then \textit{define}
+          the energetically subdominant scales, the scales smaller than about 100 km.
+          Our lower wavelength definition is limited by the effective resolution
+          of our higher resolution simulation ($\sim$10 km).
+          This definition is consistent with Callies \& Ferrari 2013 (they did not
+          define mesoscales/submesoscales based on scales that a sensor resolves
+          as the reviewer suggests).\\
+
+          We understand that our definition may not be perfect --- we may indulge some readers and
+          irritate others. But
+          besides consistency with our recent
+          work (Rocha et al. 2016), there are at least two reasons for sticking to
+          our definition: (i) this is the definition used by the SWOT community (Fu \& Ferrari
+              2008); incidentally current altimeters do not resolve
+              scales smaller then about 100 km; and (ii) our goal is determine the flows
+              (geostrophically balanced turbulence, inertia-gravity waves, etc) that
+              dominate the 10-100 km horizontal
+              scales and their seasonality. In any event,  readers that make to the first sentence of our
+              abstract will find our explicit definition of submesoscales.
+      \edp
 
@@
         quotation marks.}\\
+
         OK.

### 2016-08-16 writeup/RochatEtal_ResponseToReviewers.tex
Minor modifications

@@
           (Wunsch \& Ferrari 2004; Ferrari \& Wunsch 2009).
-          Also, the scales of equilibrated baroclinic geostrophic turbulence are
-          larger than the deformation radius (e.g., Larichev \& Held 1995),
-          although linear theorists always deploy ``deformation radius'' as
-          the most relevant scale.\\
+          To add yet another controversy, Larichev \& Held 1995 argue
+          that the deformation length is not the panacea as linear theorists
+          suggest. In baroclinic geostrophic turbulence, the bulk of the
+          energy production occurs on scales much larger than the deformation
+          radius; the energy-containing scales of equilibrated baroclinic geostrophic
+          turbulence are also larger than the deformation radius (e.g., Larichev \& Held 1995).\\
 
@@
               2008); incidentally current altimeters do not resolve
-              scales smaller then about 100 km; and (ii) our goal is determine the flows
+              scales smaller then about 100 km; and (ii) our goal is to determine the flows
               (geostrophically balanced turbulence, inertia-gravity waves, etc) that
@@
 
-        OK.
+        Removed.
 
@@
 
-\item {\it (Section 3 and Figures 2 \& 3) At what depth are you computing these quantities?}
+\item {\it (Section 3 and Figures 2 \& 3) At what depth are you computing these quantities?}\\
+
+  All quantities are computed at the surface as indicated in the title of the
+  section. We now explicitly indicate that in the caption of figures 2 \& 3.
 

### 2016-08-16 writeup/rocha_etal.tex
Minor modifications

@@
      \includegraphics[width=.65\textwidth]{figs/fig2.pdf}
-  \caption{\small Time series of the root-mean-square (RMS) of (a) vorticity,
+  \caption{\small Time series of the root-mean-square (RMS) of surface (a) vorticity,
   (b) rate of strain, and (c) horizontal divergence in the LLC outputs and gridded AVISO data.}
@@
     \includegraphics[width=.75\textwidth]{figs/fig3.pdf}
- \caption{\small Seasonal variation of joint probability distributions:  vorticity vs.\
+ \caption{\small Seasonal variation of joint probability distributions of
+          surface lateral velocity gradient tensor:  vorticity vs.\
           strain rate (a through d),

### 2016-08-16 writeup/RochatEtal_ResponseToReviewers.tex
Small fix

@@
           to our definition the mesoscales are the energy-containing scales of the
-          flow, the eddies that contain most of the eddy kinetic energy in the ocean
+          flow --- the eddies that contain most of the eddy kinetic energy in the ocean
           (Ferrari \& Wunsch 2009). On observational

### 2016-08-17 writeup/RochatEtal_ResponseToReviewers.tex
Small fix
Adds wavenumber spectrum to fig4, changes caption

@@
           agreed upon definition''. There is also irritation that oceanographers
-          decided to use the term mesoscales for flows with dynamics analogous to
-          the meteorological synoptic scale.\\
+          decided to term mesoscales the flows with dynamics analogous to
+          the meteorological synoptic scale motions.\\
 
@@
 
-          We understand that our definition may not be perfect --- we may indulge some readers and
+          We understand that our definition is imperfect --- we may amuse some readers but
           irritate others. But

### 2016-08-17 writeup/rocha_etal.tex
Adds wavenumber spectrum to fig4, changes caption

@@
 which shows the ratio between the two spectra, clearly depicts the out-of-phase
-seasonality of super-inertial and subinertial submesoscale flows. More importantly,
-inertia-gravity
-waves account for most of the enhancement of super-inertial SSH variance in October.
-
-\section{Projection onto horizontal scales}
-
-To quantify the projection of these flows onto different horizontal
-scales, we calculate wavenumber spectra of kinetic energy and sea-surface height
- variance.  Before calculating the spectra, time-mean and spatial linear
-trends were removed, and the resulting fields were multiplied by a two-dimensional Hanning window.
-The two-dimensional spectra were averaged azimuthaly \citep[e.g., ][]{rocha_etal2016} .
-We discuss only spectra for the 1/48$^\circ$ simulation, which extends the 1/24$^\circ$ simulation
-towards smaller scales (supporting information).
+seasonality of super-inertial and subinertial submesoscale flows.  At scales smaller
+than 100 km (submesoscales), there the SSH variance is 66$\%$ larger in October than in April.
+More importantly, most of this SSH variance increase is accounted for
+inertia-gravity waves, particularly modes two through four --- higher vertical
+modes are more sensitive to changes in stratification (supporting information).
+Similarly, the submesoscale superinertial surface KE is 63$\%$ larger in October
+than the April.
 
@@
   \begin{center}
-    \includegraphics[width=.9\textwidth]{figs/fig4.pdf}
+    \includegraphics[width=1.\textwidth]{figs/fig4.pdf}
  \caption{\small LLC4320 Wavenumber-frequency spectrum of SSH variance in (a) April and (b) October.
-          (c) The ratio between (b) and (a) --- the ratio is masked in regions
-          with SSH variance density smaller than $10^{-2}$ m$^2$/[cpkm $\times$ cph].
-          The light gray shaded region depicts the dispersion relations for inertia-gravity
+          (c) Wavenumber spectrum of SSH variance --- the integral of (a) and (b) over frequency.
+          In (a) and (b), the light gray shaded region depicts the dispersion relations for inertia-gravity
           waves from mode 1 through mode 4 across the latitudinal domain; horizontal dashed lines represent the semi-diurnal
           lunar tidal frequency (M$_2$), and the inertial frequency at mid-latitude
-          (f$_{32.5}$) and its first harmonic (2f$_{32.5}$).}
+          (f$_{32.5}$) and its first harmonic (2f$_{32.5}$). }
   \label{fig4}
@@
 
-\begin{figure*}[ht]
-  \begin{center}
-    \includegraphics[width=.65\textwidth]{figs/fig5.pdf}
- \caption{\small Surface (horizontal) kinetic energy (a) and sea-surface height variance wavenumber spectra (b)
- in the 1/48$^\circ$ simulation. Solid lines
- are spectra based on hourly snapshots, and dashed lines are spectra based on daily-averaged
- fields.}
- \label{fig5}
- \end{center}
-\end{figure*}
-
-Figure \ref{fig5}a depicts the horizontal wavenumber spectra of surface KE in
-April and October. At scales larger than 20 km, the April and October spectra based on hourly
-velocity snapshots (solid lines) are nearly indistinguishable from each other within 95$\%$ confidence level.
-Consistent with the results of
-\citet{rocha_etal2016}, who analyzed a 3-month output of the LLC4320 simulation in Drake Passage,
-there is significant high-frequency variabilty at submesoscales. Daily-averaging
-the velocity field suppresses spatial variability at scales smaller than about 250
-km, both in April and October (compare solid lines against dashed lines in Figure
-\ref{fig5}a). But this suppression is dramatic in October, when the inertia-gravity waves peak. At scales
-smaller than 100 km, 39$\%$ of the surface KE in April is accounted for by super-inertial
-flows as opposed to 79$\%$ in October. The seasonality of subinertial submesoscale flows
-is strong, consistent
-with the results of \citet{sasaki_etal2014}, which are
-based on daily-averaged velocity fields of a different model without tidal forcing
-(P. Klein, personal communication).
-
-The inertia-gravity waves project on the sea-surface.
-There is a dramatic difference between the spectra
-based on hourly and daily-averaged sea-surface height in October (Figure \ref{fig5}b): at scales smaller than 100 km, the spectra
-of hourly sea-surface height roughly follows a $-2$ power-law, whereas the spectra of daily-averaged
-sea-surface height roughly follows a $-5$ power-law. At these scales, 33$\%$ of the sea-surface height variance in April is accounted for by super-inertial
-flows as opposed to 83$\%$ in October. Curiously, the out-of-phase seasonal cycle of submesoscale
-turbulence and near-surface submesoscale inertia-gravity waves conspire to yield weak seasonality
-in the spectra of KE and sea-surface height variance based on hourly fields.
-
+%\section{Projection onto horizontal scales}
+%
+%To quantify the projection of these flows onto different horizontal
+%scales, we calculate wavenumber spectra of kinetic energy and sea-surface height
+% variance.  Before calculating the spectra, time-mean and spatial linear
+%trends were removed, and the resulting fields were multiplied by a two-dimensional Hanning window.
+%The two-dimensional spectra were averaged azimuthaly \citep[e.g., ][]{rocha_etal2016} .
+%We discuss only spectra for the 1/48$^\circ$ simulation, which extends the 1/24$^\circ$ simulation
+%towards smaller scales (supporting information).
+%
+%
+
+%
+%
+%\begin{figure*}[ht]
+%  \begin{center}
+%    \includegraphics[width=.65\textwidth]{figs/fig5.pdf}
+% \caption{\small Surface (horizontal) kinetic energy (a) and sea-surface height variance wavenumber spectra (b)
+% in the 1/48$^\circ$ simulation. Solid lines
+% are spectra based on hourly snapshots, and dashed lines are spectra based on daily-averaged
+% fields.}
+% \label{fig5}
+% \end{center}
+%\end{figure*}
+%
+%
+%Figure \ref{fig5}a depicts the horizontal wavenumber spectra of surface KE in
+%April and October. At scales larger than 20 km, the April and October spectra based on hourly
+%velocity snapshots (solid lines) are nearly indistinguishable from each other within 95$\%$ confidence level.
+%Consistent with the results of
+%\citet{rocha_etal2016}, who analyzed a 3-month output of the LLC4320 simulation in Drake Passage,
+%there is significant high-frequency variabilty at submesoscales. Daily-averaging
+%the velocity field suppresses spatial variability at scales smaller than about 250
+%km, both in April and October (compare solid lines against dashed lines in Figure
+%\ref{fig5}a). But this suppression is dramatic in October, when the inertia-gravity waves peak. At scales
+%smaller than 100 km, 39$\%$ of the surface KE in April is accounted for by super-inertial
+%flows as opposed to 79$\%$ in October. The seasonality of subinertial submesoscale flows
+%is strong, consistent
+%with the results of \citet{sasaki_etal2014}, which are
+%based on daily-averaged velocity fields of a different model without tidal forcing
+%(P. Klein, personal communication).
+%
+%The inertia-gravity waves project on the sea-surface.
+%There is a dramatic difference between the spectra
+%based on hourly and daily-averaged sea-surface height in October (Figure \ref{fig5}b): at scales smaller than 100 km, the spectra
+%of hourly sea-surface height roughly follows a $-2$ power-law, whereas the spectra of daily-averaged
+%sea-surface height roughly follows a $-5$ power-law. At these scales, 33$\%$ of the sea-surface height variance in April is accounted for by super-inertial
+%flows as opposed to 83$\%$ in October. Curiously, the out-of-phase seasonal cycle of submesoscale
+%turbulence and near-surface submesoscale inertia-gravity waves conspire to yield weak seasonality
+%in the spectra of KE and sea-surface height variance based on hourly fields.
+%
+%
 \section{Summary and Conclusion}

### 2016-08-17 writeup/rocha_etal_supporting-info.tex
Adds reference to ECCO project
Syncronize renaming
Fixes typo in SI
Updates supporting information

@@
       of the vertical modes.
+\item Figure S7 Wavenumber-frequency spectrum of KE for LLC4320 simulation.
 %if Tables are larger than 1 page, upload as separate excel file
@@
 (at \texttt{http://mitgcm.org/viewvc/MITgcm/MITgcm$\_$contrib/
-llc$\_$hires}).
+llc$\_$hires}). The model output can be obtained from  obtained from the ECCO project (\texttt{http://ecco2.org/llc$\_$hires}).
 
@@
       \setfigurenum{S1}
-      \includegraphics[width=40pc]{figs/fig_s2_3.pdf}
+      \includegraphics[width=40pc]{figs/figS1.pdf}
    \caption{Comparison between spectra of the 1/24$^\circ$
@@
       \setfigurenum{S2}
-      \includegraphics[width=40pc]{figs/fig_s2_2.pdf}
+      \includegraphics[width=40pc]{figs/figS2.pdf}
    \caption{Time dependence of probability density functions of vertical vorticity (a, b)
@@
       \setfigurenum{S3}
-      \includegraphics[width=40pc]{figs/fig_s2_2.png}
+      \includegraphics[width=40pc]{figs/figS3.png}
    \caption{Depth-dependence of root-mean-square of vertical vorticity (a,d), rate of lateral strain (b,e)
@@
       \setfigurenum{S4}
-      \includegraphics[width=40pc]{figs/fig_s2_1.pdf}
+      \includegraphics[width=40pc]{figs/figS4.pdf}
    \caption{Potential density: a comparison between Roemmich-Gilson Argo
@@
       \setfigurenum{S5}
-      \includegraphics[width=40pc]{figs/figA1.pdf}
+      \includegraphics[width=40pc]{figs/figS5.pdf}
    \caption{Comparison between frequency spectra of LLC simulations and two
@@
       \setfigurenum{S6}
-      \includegraphics[width=40pc]{figs/fig_s3.pdf}
+      \includegraphics[width=40pc]{figs/figS6.pdf}
    \caption{The seasonal variability of the WOA 2013 \citep{levitus_etal2013} stratification averaged  over
@@
    \label{figS3}
+   \end{center}
+ \end{figure}
+
+ \begin{figure}
+    \begin{center}
+      \setfigurenum{S7}
+      \includegraphics[width=40pc]{figs/figS7.pdf}
+   \caption{LLC4320 Wavenumber-frequency spectrum of KE in (a) April
+    and (b) October. (c) Wavenumber spectrum of SSH variance — the integral of (a)
+    and (b) over frequency. In (a) and (b), the light gray shaded region depicts the
+    dispersion relations for inertia-gravity waves from mode 1 through mode 4 across
+    the latitudinal domain; horizontal dashed lines represent the semi-diurnal lunar
+    tidal frequency (M2), and the inertial frequency at mid-latitude (f32.5) and its
+    first harmonic (2f32.5).}
+   \label{figS7}
    \end{center}

### 2016-08-17 writeup/rocha_etal.tex
Edits acknowledgements

@@
 
-\section{Statistics of the surface lateral velocity gradient tensor}
+\section{Bulk statistics of the surface lateral velocity gradient tensor}
 To study the seasonality in the surface velocity, we calculate the lateral velocity gradient tensor
@@
 
-\section{Seasonality of submesoscale dynamics}
+\section{Joint probability density functions}
+
 The results of figure \ref{fig2}a-c show that submesoscale surface
@@
 consistent with linear inertia-gravity waves.
+
+\section{Wavenumber-frequency spectrum}
 
@@
 \begin{acknowledgments}
- William R. Young and the referees --- Christian Buckingham and two anonymous ---
-  provided constructive and helpful comments. Greg Wagner first
+ William R. Young and the reviewers --- Christian Buckingham and two anonymous referees ---
+  provided exceptionally helpful and constructive comments. Greg Wagner first
  suggested that the near-surface shape of the baroclinic modes may change seasonally.

### 2016-08-17 writeup/rocha_etal.tex
Adds new references

@@
 %
-\section{Summary and Conclusion}
+\section{Discussion and conclusion}
 Our work adds to recent modeling \citep{sasaki_etal2014}
@@
 alternative explanation is that the shape of the baroclinic modes
-changes seasonally. In particular the baroclinic modes are
-significantly more surface intensified in late summer/early fall
- (supporting information).
+changes seasonally (supporting information).
+
 If the internal wave source has weak seasonal dependence, then a strong
@@
 other submesoscale inertia-gravity waves generated through internal tide interactions.
-Near-inertial wave generation peaks in winter \citep[e.g.,][]{alford2003}, but
-those waves project onto large horizontal scales \citep[e.g, ][]{qi_etal1995}.
+Windy near-inertial wave (NIW) generation is bursty and peaks in winter \citep[e.g.,][]{alford2003}.
+While NIWs are generated at large scales, mesoscale turbulence can
+refract and disperse those waves in few inertial periods, thereby generating
+submesoscale (10-100 km) near-inertial waves \citep{young_benjelloul1997}. But in our
+simulations the wavy seasonality in submesoscale surface KE and SSH variance is mostly accounted for
+by inertia-gravity waves with periods shorter than 12.4 h (Figure \ref{fig4}a-b).
 
@@
 \bibliographystyle{agufull08}
-%\bibliography{rocha_etal}
+\bibliography{rocha_etal}
 
@@
 % Text
-\begin{thebibliography}{18}
-\providecommand{\natexlab}[1]{#1}
-\expandafter\ifx\csname urlstyle\endcsname\relax
-  \providecommand{\doi}[1]{doi:\discretionary{}{}{}#1}\else
-  \providecommand{\doi}{doi:\discretionary{}{}{}\begingroup
-  \urlstyle{rm}\Url}\fi
-
-\bibitem[{\textit{Alford}(2003)}]{alford2003}
-Alford, M.~H. (2003), {Redistribution of energy available for ocean mixing by
-  long-range propagation of internal wave}, \textit{Nature},
-  \textit{423}(6936), 159--162.
-
-\bibitem[{\textit{Buckingham et~al.}(2016)\textit{Buckingham, Naveira~Garabato,
-  Thompson, Brannigan, Lazar, Marshall, George~Nurser, Damerell, Heywood, and
-  Belcher}}]{buckingham_etal2016}
-Buckingham, C.~E., A.~C. Naveira~Garabato, A.~F. Thompson, L.~Brannigan,
-  A.~Lazar, D.~P. Marshall, A.~George~Nurser, G.~Damerell, K.~J. Heywood, and
-  S.~E. Belcher (2016), {Seasonality of submesoscale flows in the ocean surface
-  boundary layer}, \textit{Geophysical Research Letters}, \textit{43}(5),
-  2118--2126.
-
-\bibitem[{\textit{B{\"u}hler et~al.}(2014)\textit{B{\"u}hler, Callies, and
-  Ferrari}}]{buhler_etal2014}
-B{\"u}hler, O., J.~Callies, and R.~Ferrari (2014), {Wave--vortex decomposition
-  of one-dimensional ship-track data}, \textit{Journal of Fluid Mechanics},
-  \textit{756}, 1007--1026.
-
-\bibitem[{\textit{Callies et~al.}(2015)\textit{Callies, Ferrari, Klymak, and
-  Gula}}]{callies_etal2015}
-Callies, J., R.~Ferrari, J.~M. Klymak, and J.~Gula (2015), {Seasonality in
-  submesoscale turbulence}, \textit{Nature communications}, \textit{6}.
-
-\bibitem[{\textit{Callies et~al.}(2016)\textit{Callies, Flierl, Ferrari, and
-  Fox-Kemper}}]{callies_etal2016}
-Callies, J., G.~Flierl, R.~Ferrari, and B.~Fox-Kemper (2016), {The role of
-  mixed-layer instabilities in submesoscale turbulence}, \textit{Journal of
-  Fluid Mechanics}, \textit{788}, 5--41.
-
-\bibitem[{\textit{Capet et~al.}(2008)\textit{Capet, McWilliams, Molemaker, and
-  Shchepetkin}}]{capet_etal2008a}
-Capet, X., J.~C. McWilliams, M.~J. Molemaker, and A.~Shchepetkin (2008),
-  {Mesoscale to submesoscale transition in the California Current System. Part
-  I: Flow structure, eddy flux, and observational tests}, \textit{Journal of
-  Physical Oceanography}, \textit{38}(1), 29--43.
-
-\bibitem[{\textit{D'Asaro}(1978)}]{dasaro1978}
-D'Asaro, E.~A. (1978), {Mixed layer velocities induced by internal wave},
-  \textit{Journal of Geophysical Research: Oceans}, \textit{83}(C5),
-  2437--2438.
-
-\bibitem[{\textit{Forget et~al.}(2015)\textit{Forget, Campin, Heimbach, Hill,
-  Ponte, and Wunsch}}]{forget_etal2015}
-Forget, G., J.-M. Campin, P.~Heimbach, C.~Hill, R.~Ponte, and C.~Wunsch (2015),
-  {ECCO version 4: an integrated framework for non-linear inverse modeling and
-  global ocean state estimation}, \textit{Geosci. Model Dev.}, \textit{8},
-  3071--3104.
-
-  \bibitem[{\textit{Marshall et~al.}(1997)\textit{Marshall, Hill, Perelman, and
-    Adcroft}}]{marshall_etal1997}
-  Marshall, J., C.~Hill, L.~Perelman, and A.~Adcroft (1997), Hydrostatic,
-    quasi-hydrostatic, and nonhydrostatic ocean modeling, \textit{Journal of
-    Geophysical Research: Oceans}, \textit{102}(C3), 5733--5752.
-
-  \bibitem[{\textit{McWilliams}(2016)}]{mcwilliams2016}
-  McWilliams, J.~C. (2016), {Submesoscale currents in the ocean}, in
-    \textit{Proc. R. Soc. A}, vol. 472, p. 20160117, The Royal Society.
-
-  \bibitem[{\textit{Menemenlis et~al.}(2008)\textit{Menemenlis, Campin, Heimbach,
-    Hill, Lee, Nguyen, Schodlok, and Zhang}}]{menemenlis_etal2008}
-  Menemenlis, D., J.-M. Campin, P.~Heimbach, C.~Hill, T.~Lee, A.~Nguyen,
-    M.~Schodlok, and H.~Zhang (2008), ECCO2: High resolution global ocean and sea
-    ice data synthesis, \textit{Mercator Ocean Quarterly Newsletter},
-    \textit{31}, 13--21.
-
-\bibitem[{\textit{Qi et~al.}(1995)\textit{Qi, De~Szoeke, Paulson, and
-  Eriksen}}]{qi_etal1995}
-Qi, H., R.~A. De~Szoeke, C.~A. Paulson, and C.~C. Eriksen (1995), {The
-  structure of near-inertial waves during ocean storms}, \textit{Journal of
-  Physical Oceanography}, \textit{25}, 2853--2871.
-
-\bibitem[{\textit{Qiu et~al.}(2014)\textit{Qiu, Chen, Klein, Sasaki, and
-  Sasai}}]{qiu_etal2014}
-Qiu, B., S.~Chen, P.~Klein, H.~Sasaki, and Y.~Sasai (2014), {Seasonal mesoscale
-  and submesoscale eddy variability along the North Pacific Subtropical
-  Countercurrent}, \textit{Journal of Physical Oceanography}, \textit{44}(12),
-  3079--3098.
-
-\bibitem[{\textit{Richman et~al.}(2012)\textit{Richman, Arbic, Shriver,
-  Metzger, and Wallcraft}}]{richman_etal2012}
-Richman, J.~G., B.~K. Arbic, J.~F. Shriver, E.~J. Metzger, and A.~J. Wallcraft
-  (2012), {Inferring dynamics from the wavenumber spectra of an eddying global
-  ocean model with embedded tides}, \textit{Journal of Geophysical Research:
-  Oceans}, \textit{117}(C12).
-
-\bibitem[{\textit{Rocha et~al.}(2016)\textit{Rocha, Chereskin, Gille, and
-  Menemenlis}}]{rocha_etal2016}
-Rocha, C.~B., T.~K. Chereskin, S.~T. Gille, and D.~Menemenlis (2016),
-  {Mesoscale to submesoscale wavenumber spectra in Drake Passage},
-  \textit{Journal of Physical Oceanography}, \textit{46}, 601--620.
-
-\bibitem[{\textit{Sasaki et~al.}(2014)\textit{Sasaki, Klein, Qiu, and
-  Sasai}}]{sasaki_etal2014}
-Sasaki, H., P.~Klein, B.~Qiu, and Y.~Sasai (2014), {Impact of oceanic-scale
-  interactions on the seasonal modulation of ocean dynamics by the atmosphere},
-  \textit{Nature communications}, \textit{5}.
-
-\bibitem[{\textit{Shcherbina et~al.}(2013)\textit{Shcherbina, D'Asaro, Lee,
-  Klymak, Molemaker, and McWilliams}}]{shcherbina_etal2013}
-Shcherbina, A.~Y., E.~A. D'Asaro, C.~M. Lee, J.~M. Klymak, M.~J. Molemaker, and
-  J.~C. McWilliams (2013), {Statistics of vertical vorticity, divergence, and
-  strain in a developed submesoscale turbulence field}, \textit{Geophysical
-  Research Letters}, \textit{40}(17), 4706--4711.
-
-\bibitem[{\textit{Thompson et~al.}(2016)\textit{Thompson, Lazar, Buckingham,
-  Naveira~Garabato, Damerell, and Heywood}}]{thompson_etal2016}
-Thompson, A.~F., A.~Lazar, C.~Buckingham, A.~C. Naveira~Garabato, G.~M.
-  Damerell, and K.~J. Heywood (2016), {Open-ocean submesoscale motions: A full
-  seasonal cycle of mixed layer instabilities from gliders}, \textit{Journal of
-  Physical Oceanography}, \textit{46}(4), 1285--1307.
-
-\end{thebibliography}
-
 % Or,

### 2016-08-17 writeup/rocha_etal.tex
Abbreviates IGWs

@@
 
-\keypoints{\item  Upper-ocean submesoscale (10-100 km) turbulence and inertia-gravity waves
+\keypoints{\item  Upper-ocean submesoscale (10-100 km) turbulence and inertia-gravity
+                  waves
                   undergo strong seasonal cycles that are out of phase.
@@
                   sea-surface height variability in late winter/early spring.
-            \item Submesoscale inertia-gravity waves dominate the horizontal velocity and
-                   sea-surface height variability in late summer/early fall.
+            \item Submesoscale inertia-gravity waves dominate the horizontal velocity
+                   and sea-surface height variability in late summer/early fall.
                   }
@@
 turbulence and mesoscale (roughly 100-300 km) modulation \citep{sasaki_etal2014,qiu_etal2014,
-callies_etal2015, thompson_etal2016,buckingham_etal2016}. Contemporary studies
-have also shown that inertia-gravity waves contribute significantly
+brannigan_etal2015,callies_etal2015, thompson_etal2016,buckingham_etal2016}. Contemporary studies
+have also shown that inertia-gravity waves (IGWs) contribute significantly
 to the near-surface variability at submesoscales \citep{richman_etal2012,
@@
 
+% now SWOT
+
 Using the output of $1/24^\circ$ and $1/48^\circ$ global
-numerical simulations with embedded tides, we show that inertia-gravity waves undergo
+numerical simulations with embedded tides, we show that IGWs undergo
 a strong seasonality near the surface in the Kuroshio Extension region.
-Interestingly, the seasonal cycle of inertia-gravity waves is out of phase
+Interestingly, the seasonal cycle of IGWs is out of phase
 with the seasonal cycle of submesoscale turbulence. On one hand, consistent with previous studies,
@@
 
-Inertia-gravity waves, on the other hand, peak in late summer/early fall,
+IGWs, on the other hand, peak in late summer/early fall,
 when the upper ocean is strongly stratified. Thus there exists a
@@
 submesoscale turbulence dominates the upper-ocean dynamics in late winter/early fall,
-whereas submesoscale inertia-gravity waves prevail in late summer/early fall.
+whereas submesoscale IGWs prevail in late summer/early fall.
 Because submesoscale turbulence is weakest in late summer/early fall,
-the present results indicate that inertia-gravity waves account for most of the
+the present results indicate that IGWs account for most of the
 summertime submesoscale sea-surface height variability.
@@
 These diagnostics highlight the submesoscale structures in the flow
-\citep[e.g.,][]{capet_etal2008a,shcherbina_etal2013}.
+\citep[e.g.,][]{shcherbina_etal2013}.
 
@@
 relatively fast, and there is no clear
-temporal and spatial scale separation between those motions and inertia-gravity waves
+temporal and spatial scale separation between those motions and IGWs
 \citep{mcwilliams2016}:
@@
  the vorticity is strongly positively skewed, whereas the divergence is moderately
- negatively skewed as predicted by frontogenesis \citep[e.g., ][]{capet_etal2008a,mcwilliams2016}
+ negatively skewed as predicted by frontogenesis \citep[e.g., ][]{mcwilliams2016}
 . In late summer/early fall, the divergence is stronger, but PDFs are much less skewed,
-consistent with linear inertia-gravity waves.
+consistent with linear IGWs.
 
@@
 More importantly, most of this SSH variance increase is accounted for
-inertia-gravity waves, particularly modes two through four --- higher vertical
+IGWs, particularly modes two through four --- higher vertical
 modes are more sensitive to changes in stratification (supporting information).
@@
 %km, both in April and October (compare solid lines against dashed lines in Figure
-%\ref{fig5}a). But this suppression is dramatic in October, when the inertia-gravity waves peak. At scales
+%\ref{fig5}a). But this suppression is dramatic in October, when the IGWs peak. At scales
 %smaller than 100 km, 39$\%$ of the surface KE in April is accounted for by super-inertial
@@
 %
-%The inertia-gravity waves project on the sea-surface.
+%The IGWs project on the sea-surface.
 %There is a dramatic difference between the spectra
@@
 %flows as opposed to 83$\%$ in October. Curiously, the out-of-phase seasonal cycle of submesoscale
-%turbulence and near-surface submesoscale inertia-gravity waves conspire to yield weak seasonality
+%turbulence and near-surface submesoscale IGWs conspire to yield weak seasonality
 %in the spectra of KE and sea-surface height variance based on hourly fields.
@@
 submesoscale turbulence. In particular, our main finding is that in global simulations with embedded tides the
-near-surface submesoscale inertia-gravity waves in the Kuroshio Extension undergo
+near-surface submesoscale IGWs in the Kuroshio Extension undergo
 a strong seasonal cycle that is out of phase with the seasonal cycle of
-submesoscale turbulence. We conjecture that the summertime dominance of inertia-gravity waves
+submesoscale turbulence. We conjecture that the summertime dominance of IGWs
  \citep{callies_etal2015} is a consequence both of suppression of submesoscale turbulence and
-enhancement of inertia-gravity waves due to re-stratification of
+enhancement of IGWs due to re-stratification of
 the upper ocean.
@@
 
-If the internal wave source has weak seasonal dependence, then a strong
-seasonal cycle in the near-surface expression of internal waves is expected.
-Internal tide generation does not show a seasonal modulation \citep[e.g.,][]{alford2003},
+Internal tide energy flux does not show a seasonal modulation \citep[e.g.,][]{alford2003},
 and therefore one expects a strong seasonality in the near-surface
 expression of internal tides and
-other submesoscale inertia-gravity waves generated through internal tide interactions.
-Windy near-inertial wave (NIW) generation is bursty and peaks in winter \citep[e.g.,][]{alford2003}.
+higher frequency IGWs generated through internal tide interactions.
+Windy near-inertial wave (NIW) generation is bursty and peaks in early winter, but
+NIW energy appears to have similar energy level in October and April
+\citep{alford_etal2016}.
 While NIWs are generated at large scales, mesoscale turbulence can
 refract and disperse those waves in few inertial periods, thereby generating
-submesoscale (10-100 km) near-inertial waves \citep{young_benjelloul1997}. But in our
-simulations the wavy seasonality in submesoscale surface KE and SSH variance is mostly accounted for
-by inertia-gravity waves with periods shorter than 12.4 h (Figure \ref{fig4}a-b).
+submesoscale (10-100 km) near-inertial waves \citep{alford_etal2016}. But in our
+simulations the inertia-gravity-wave seasonality in submesoscale surface KE and SSH
+variance is mostly accounted for
+by higher frequency waves  (Figure \ref{fig4}a-b), likely because submesoscale
+NIWs quickly propagate into the interior and have a small projection onto SSH.
 
@@
 
-% Here we argue that those divergent motions are likely inertia-gravity waves.
+% Here we argue that those divergent motions are likely IGWs.
 % We can show a couple of spectra and argue that they roughly satisfy polarization
@@
 %
-% Most of these high-frequency inertia-gravity waves project onto scales smaller
+% Most of these high-frequency IGWs project onto scales smaller
 % than 100 km. The surface KE spectra of 100-km-smoothed velocity field is significantly

### 2016-08-17 writeup/rocha_etal.tex
Adds contex to conclusion

@@
 Internal tide energy flux does not show a seasonal modulation \citep[e.g.,][]{alford2003},
-and therefore one expects a strong seasonality in the near-surface
+hence one expects a strong seasonality in the near-surface
 expression of internal tides and
-higher frequency IGWs generated through internal tide interactions.
+higher frequency IGWs if the near-surface stratification varies strongly.
 Windy near-inertial wave (NIW) generation is bursty and peaks in early winter, but
@@
 by higher frequency waves  (Figure \ref{fig4}a-b), likely because submesoscale
-NIWs quickly propagate into the interior and have a small projection onto SSH.
+NIWs quickly propagate into the interior and weakly project onto SSH.
 
+In the context of the SWOT mission, our results suggest that if one is interested
+only in the geostrophic flow, so that surface velocity can be diagnosed
+from SSH, then the noise-to-signal ratio will have a strong seasonality.
 An important caveat is that this study focuses on a single patch of ocean in the

### 2016-08-17 writeup/rocha_etal.tex
Small edits in conclusion

@@
 refract and disperse those waves in few inertial periods, thereby generating
-submesoscale (10-100 km) near-inertial waves \citep{alford_etal2016}. But in our
-simulations the inertia-gravity-wave seasonality in submesoscale surface KE and SSH
+submesoscale (10-100 km) near-inertial waves \citep[][and references therein]{alford_etal2016}.
+But in our simulations the inertia-gravity-wave seasonality in submesoscale surface KE and SSH
 variance is mostly accounted for
@@
 only in the geostrophic flow, so that surface velocity can be diagnosed
-from SSH, then the noise-to-signal ratio will have a strong seasonality.
-An important caveat is that this study focuses on a single patch of ocean in the
-vicinity of the Kuroshio Extension. This may be typical of a mesoscale-rich
-subtropical region, but it is unlikely to be
-representative of other regions such as low-eddy-kinetic-energy eastern boundary currents
-and the middle of the subtropical gyre.  We plan to report on the geographic variability of submesoscale
+from SSH, then the noise-to-signal ratio will have a strong seasonality owing
+to the presence of incoherent IGWs.
+An important caveat is that the Kuroshio Extension may be typical of a mesoscale-rich
+subtropical regions, but it is unlikely to be
+representative of other regions such as low-eddy-kinetic-energy eastern boundary
+currents.  We plan to report on the geographic variability of submesoscale
 seasonality in a future study.

### 2016-08-18 writeup/rocha_etal.tex
Fixes legend

@@
 \section{Wavenumber-frequency spectrum}
-
-To confirm that inertia-gravity undergo a vigorous seasonal cycle near the surface,
+To confirm that submesoscale IGWs undergo a vigorous seasonal cycle near the surface,
 we calculate wavenumber-frequency spectrum of SSH variance. Focusing on the high-frequency
@@
 remove linear trends and multiply the data by a three-dimensional Hanning window.
-The spectral estimate is then azimuthally-averaged in wavenumber.
+We azimuthally-average the spectrum in wavenumber.
 
@@
 and super-inertial. Subinertial SSH variance at submesoscales is larger in April, whereas
-super-inertial SSH variance at submesoscales is larger in October.  Figure \ref{fig4}c,
-which shows the ratio between the two spectra, clearly depicts the out-of-phase
-seasonality of super-inertial and subinertial submesoscale flows.  At scales smaller
+super-inertial SSH variance at submesoscales is larger in October. At scales smaller
 than 100 km (submesoscales), there the SSH variance is 66$\%$ larger in October than in April.
-More importantly, most of this SSH variance increase is accounted for
-IGWs, particularly modes two through four --- higher vertical
-modes are more sensitive to changes in stratification (supporting information).
+IGWs, particularly modes two through four, account for most of the SSH variance
+increase in summer, --- higher vertical
+modes are more sensitive to changes in near-surface stratification (supporting information).
 Similarly, the submesoscale superinertial surface KE is 63$\%$ larger in October
-than the April.
+than the April (supporting information).
+
+Figure \ref{fig4}c shows the integral over frequency
+of wavenumber-frequency spectra in figures \ref{fig4}a-b. The
+SSH wavenumber spectrum approximately follows a -2 power-law at submesoscales,
+both in April and October (solid lines in figure \ref{fig4}c)  --- there is
+small difference between the spectra owing to the phase cancellation between the
+seasonal cycle of submesoscale turbulence and inertia-gravity waves. The sub-inertial flow
+has steeper submesoscale SSH variance wavenumber spectrum (dashed lines in figure \ref{fig4}c).
+ The difference between the SSH variance wavenumber spectrum for all frequencies and
+ sub-inertial frequencies is dramatic in October when super-inertial flows account for
+ about 80\% of the SSH variance and KE at submesoscales. We obtain similar results by computing
+SSH variance and KE wavenumber spectra directly from hourly and daily-averaged
+SSH velocity fields.
 
@@
 submesoscale (10-100 km) near-inertial waves \citep[][and references therein]{alford_etal2016}.
-But in our simulations the inertia-gravity-wave seasonality in submesoscale surface KE and SSH
-variance is mostly accounted for
-by higher frequency waves  (Figure \ref{fig4}a-b), likely because submesoscale
+But in our simulations the higher frequency inertia-gravity-wave
+account for most of the seasonality in super-inertial surface KE and SSH
+(Figure \ref{fig4}a-b), likely because submesoscale
 NIWs quickly propagate into the interior and weakly project onto SSH.

### 2016-08-18 writeup/rocha_etal.tex
Starts drafting SWOT relavant paragraph

@@
 % now SWOT
+The upcoming Surface and Water Ocean Topography (SWOT) mission will measure
+sea-surface height (SSH) on global scale with resolution of about 20 km, nearly
+ten time higher than current nadir-looking altimeters. Thus the study of the
+partition between geostrophic and ageostrophic flows across the submesoscales and
+the quantification of their seasonality seasonality has an important application
+to planning of the SWOT
+mission. In particular, the presence of inertia-gravity waves may be consider
+noise if one wants to diagnose surface currents from SSH.
 

### 2016-08-18 writeup/RochatEtal_ResponseToReviewers.tex
Adds response to GM comment

@@
         wave spectrum? (Recall: the GM spectrum contains both balanced and unbalanced
-        flows, not just internal waves, despite that the name suggests this.)}
+        flows, not just internal waves, despite that the name suggests this.)}\\
+
+        \bdp
+          We explicit state that those are submesoscale IGWs. Using
+          our definition of submesoscales are IGWs
+          with horizontal scales about 10-100 km\footnote{Bill Young suggested the
+          term submesoscale IGWs.}; IGWs have periods  The GM spectrum is not accurate
+          close to the surface (most of our results concern surface currents and SSH,
+          and it does not account for tidal and inertial peaks. We believe that the
+          new plot of the wavenumber-frequency spectrum of the SSH
+          variance (new Figure 4) --- following the reviewer's suggestion ---
+          will more clearly convey the idea that there is no horizontal scale
+          separation between submesoscale IGWs and submesoscale turbulence.\\
+
+          If the reviewer is deploying ``balanced'' as a short to ``geostrophically
+          balanced'', then the comment about the GM spectrum containing both balanced
+          and unbalanced flows is egregiously incorrect.  GM used linear theory
+          of IGW to synthesize observations, mostly mooring data --- they
+          fitted an analytic model to (super-inertial) frequency spectra and
+          different vertical modes (see Walter Munk's chapter
+          in EPO 1981 and references therein).
+          One could argue
+          that other flows project onto similar temporal scales, but those motions
+          are super-inertial (e.g., stratified turbulence), not (sub-inertial)
+          balanced flows.
+        \edp
 
@@
       because SSH variance density has nothing to do with \textit{power}.\footnote{The
-      term \textit{power spectral density} was for the \textit{frequency} spectrum.}
+      term \textit{power spectral density}  appears to have been introduced
+      for frequency spectrum of various quantities in signal processing (Rob Pinkel, personal
+      communication).}
       And ``SSH [wavenumber] spectrum'' is the plot of SSH variance spectral density
@@
       to the pervasive axis mislabelling in plots of probability distribution function (PDF)
-      --- the correct label is \textit{probability density}
-      not PDF.}
+      --- the best label is \textit{probability density} not PDF.}
       \edp

### 2016-08-18 writeup/RochatEtal_ResponseToReviewers.tex
Polished intro

@@
 to the statements in other places of the manuscript that the seasonality of subinertial
-submesoscale flows is strong (e.g., those relating to Figure 1b-c and Figure 2).}
+submesoscale flows is strong (e.g., those relating to Figure 1b-c and Figure 2).}\\
+
+    \bdp
+        We agree that the emphasis on the similarity of hourly spectra can appear
+        out of contradictory if placed out of context. Those spectra (now Figure 4c
+        and supporting information) are similar because both submesoscale sub-inertial
+        and super-inertial flows undergo a strong seasonal cycle that is out of
+        phase --- there is a phase cancellation and the spectra based on hourly
+        spectra is indistinguishable within errorbars. In any event, we removed
+        that sentence and we believe the new Figure 4 will more efficiently convey
+        the phase cancellation point.
+    \edp
 
@@
 
+  \bdp
     The convergence of meridians (the $\beta$-effect) accounts for the
     small divergence depicted by the black line in Figure 2c.
+  \edp
 
@@
           the seasonality of IGWs in the context of SWOT and why this might matter. The introduction
-          should, therefore, place this study matter?”}
+          should, therefore, place your study in the greater context and answer the question,
+          ``Why does this study matter?''}\\
+
+          \bdp
+              We agree that the planning of SWOT is perhaps the most immediate
+              application of our results. We have added a paragraph dedicated to SWOT (new
+              paragraph 2 of section 1). We have also added a discussion about the
+              implication for SWOT in section 6. Thanks for making us provide a better
+              context for out study.
+          \edp
 
@@
         seasonality at the submesoscale but fail missed a few: Ostrovskii (1995) and
-        Brannigan et al. (2015). These are observation- and model-based studies, resp.}
+        Brannigan et al. (2015). These are observation- and model-based studies, resp.}\\
+
+        Thanks for pointing that out.
 
@@
 
-          If the reviewer is deploying ``balanced'' as a short to ``geostrophically
+          If the reviewer is using ``balanced'' as a short to ``geostrophically
           balanced'', then the comment about the GM spectrum containing both balanced
@@
 
-We changed jPDF with joint-PDF.
+We changed jPDF with joint-PDF; it does look much better.
 
@@
       We further disagree that the y-axis should be labelled \textit{power spectral density of SSH}
-      or \textit{SSH spectra} [{\it sic}]. On physical grounds, \textit{power} adds ambiguity,
+      or \textit{SSH spectra} [{\it sic}]. On physical grounds, \textit{power} adds
+      unnecessary ambiguity,
       because SSH variance density has nothing to do with \textit{power}.\footnote{The

### 2016-08-18 writeup/rocha_etal.tex
Cut some clutter
Polished intro

--- 2016-08-18 writeup/rocha_etal.tex
@@
 % now SWOT
+Characterizing the partition between geostrophic and ageostrophic flows across
+the submesoscales and their seasonality has immediate applications for the planning
+of the upcoming Surface and Water Ocean Topography (SWOT) satellite mission.
+SWOT will measure
+sea-surface height (SSH) with about 15 km resolution, thereby providing the
+first global SSH measurements at submesoscales \citep{fu_ubelmann2014}. A
+challenge for SWOT is the presence of ageostrophic flows that straddle the submesoscales and project
+onto SSH \citep[e.g., ][]{richman_etal2012} --- the
+nuisance\footnote{``One man's noise is
+another man's signal'' (Walter Munk).} may upset the diagnose
+of submesoscale surface geostrophic velocities.
 
@@
 Interestingly, the seasonal cycle of IGWs is out of phase
-with the seasonal cycle of submesoscale turbulence. On one hand, consistent with previous studies,
+with the seasonal cycle of submesoscale turbulence. Consistent with previous studies,
 deep late-winter mixed layers are prone to
@@
 
-IGWs, on the other hand, peak in late summer/early fall,
+Submesoscale IGWs, however, peak in late summer/early fall,
 when the upper ocean is strongly stratified. Thus there exists a
@@
 the present results indicate that IGWs account for most of the
-summertime submesoscale sea-surface height variability.
+summertime submesoscale sea-surface height variability with implications for SWOT.
 
@@
 The LLC4320 simulation is an extension of the 3-month long output used by
-\citet{rocha_etal2016}.  Details of the LLC simulations are provided in the supplemental material.
+\citet{rocha_etal2016} --- for details, see supplemental material.
 
@@
   \begin{center}
-    \includegraphics[width=1.\textwidth]{figs/fig4.pdf}
- \caption{\small LLC4320 Wavenumber-frequency spectrum of SSH variance in (a) April and (b) October.
+    \includegraphics[width=.9\textwidth]{figs/fig4.pdf}
+ \caption{\small LLC4320 wavenumber-frequency spectrum of SSH variance in (a) April and (b) October.
           (c) Wavenumber spectrum of SSH variance --- the integral of (a) and (b) over frequency.
@@
 
-In the context of the SWOT mission, our results suggest that if one is interested
+In the context of SWOT, our results suggest that if one is interested
 only in the geostrophic flow, so that surface velocity can be diagnosed

### 2016-08-18 writeup/rocha_etal.tex
More additions to abstract

@@
     inertia-gravity waves strongly dominate the submesoscale surface kinetic energy and
-    sea-surface height variance in late summer/early fall.
+    sea-surface height (SSH) variance in late summer/early fall, with implications for
+    the accuracy of surface geostrophic currents diagnosed from SSH measured by
+    high-resolution
+    satellite altimeters.
 \end{abstract}
@@
 from SSH, then the noise-to-signal ratio will have a strong seasonality owing
-to the presence of incoherent IGWs.
+to the presence of incoherent IGWs (coherent internal tides may be removed
+efficiently).
 An important caveat is that the Kuroshio Extension may be typical of a mesoscale-rich

### 2016-08-18 writeup/RochatEtal_ResponseToReviewers.tex
Nothings changes

@@
           irritate others. But
-          besides consistency with our recent
+          besides unsettled definitons by our community abd consistency with our recent
           work (Rocha et al. 2016), there are at least two reasons for sticking to
@@
 
-  \item {\it (Abstract) The abstract could be rewritten to emphasize its connection to SWOT.}
+  \item {\it (Abstract) The abstract could be rewritten to emphasize its connection to SWOT.}\\
+
+        \bdp
+          We now describe that there are implications for the accuracy of high-resolution
+          altimeters.
+        \edp
 
@@
           introduction would also help this article.}
+
+          \bdp
+            Added.
+          \edp
 
@@
         parentheses nominal horizontal resolutions corresponding to these spatial resolutions.}
+
+        \bdp
+          We have added both nominal and effective resolutions.
+        \edp
 
@@
       not a consistency as it is a different location, different time. The author has already
-      made the reader aware of the previous study so there is no reason to cite again.}
+      made the reader aware of the previous study so there is no reason to cite again.}\\
+
+        Removed.
 
@@
             ...'' Have you thought about computing scale-dependent vorticity, strain rate and
-            divergence? This is really what you would be doing
-            aforementioned flows into spectral space.}
+            divergence? This is really what you would be doing if you did a breakdown
+            of the aforementioned flows into spectral space.}
+
+            \bdp
+                Following the reviewer's suggestion, we have replaced the old section
+                5 with a discussion of the wavenumber-frequency spectrum of SSH variance.
+                To better illustrate the projection of different flows across
+                different horizontal scales, figure 5c shows the integral of the
+                wavenumber-frequency spectrum over frequencies --- the result is
+                very similar to calculating the wavenumber spectrum directly from
+                hourly and daily-averaged fields (old figure 4b).
+            \edp
 
@@
 
+      \bdp
+      Any reader familiar with basic geostrophy concepts is likely to understand the connection.
+      \edp
+
 \item {\it (Section 4, second paragraph) Last sentence. This is good. I understand this but to
@@
         one-to-one line for the joint-PDF of vorticity and the Laplacian of SSH.}
+
+      As above.
 
@@
         \bdp
-          We changed the order of the panels.
+          We changed the order of the panels. All figures are provided in high-quality
+          resolution. We avoid plotting the mixed-layer depth because it is noisy,
+          hindering an important detail we would like the judicious reader to note,
+          particularly in figure 1c: the wiggles in the mixed-layer base
+          indicate internal waves.
         \edp
@@
 
-\begin{description}
-      {\it Any horizontal velocity field, uh, can be expressed in terms of vorticity, divergence,
+      {\it ``Any horizontal velocity field, uh, can be expressed in terms of vorticity, divergence,
       strain rate and mean flow. This can be seen by expressing uh as a Taylor series
@@
       fronts are ubiquitous at the submesoscale, this is a natural decomposition and
-      variances of these quantities are apt descriptors of submesoscale turbulence.}
-\end{description}
+      variances of these quantities are apt descriptors of submesoscale turbulence.''}\\
+
+      \bdp
+        Adding the elementary motivation above is yet another controversy: some
+        readers may appreciate it but others may find it condescending.
+        We argue that most physical oceanographers interested in our paper will
+        likely know this material. Even readers from other areas, who did not take
+        basic fluid mechanics, will have the intuition that if the flow peaks
+        at mesoscales (energy-containing), then second-order statistics of the
+        velocity gradient will better highlight the submesoscales than the
+        KE (second-order statistic of the velocity).
+      \edp
 
@@
       parameter is the gradient Rossby number.}
+
+      \bdp
+        Added. Yes, as we mention in the text, we have normalized by those
+        quantities by the \textit{local} inertial frequency, not the inertial
+        frequency at mid-latitude.
+      \edp
 

### 2016-08-18 writeup/rocha_etal.tex
Nothings changes

@@
 realistic numerical simulations. The outputs
-analyzed here, LLC2160 (1/24$^\circ$) and LLC4320 (1/48$^\circ$),   are
+analyzed here, LLC2160 (nominal resolution 1/24$^\circ$$\approx$4.7 km;
+effective resolution $\approx$ 20 km) and LLC4320
+(nominal resolution 1/48$^\circ$$\approx$2.3 km;
+effective resolution $\approx$8 km),   are
 forward Massachusetts Institute of Technology general circulation model \citep[MITgcm; ][]{marshall_etal1997}

### 2016-08-18 writeup/RochatEtal_ResponseToReviewers.tex
Lots of minor additions

@@
 \item {\it (Section 2, last paragraph) ``as shallow as 40 m'' – I bet that the mixed layer depths get
-      even shallower than this. At midlatitudes, 20-30 m is not uncommon.}
+      even shallower than this. At midlatitudes, 20-30 m is not uncommon.}\\
+
+      We rewrote that sentence.
 
@@
       \bdp
-        Added. Yes, as we mention in the text, we have normalized by those
+        We have experimented with a new figure 3, including a third row to show the
+        vorticity PDFs. But the extra plots add little extra, if any, information
+        and take a lot of space. For what it is worth, we explicitly give the relevant statistic
+        (skewness) for each month, etc.  As the reviewer mentions, the collapsed
+        PDFs for vorticity and divergence are shown in Figure S2, which is a plot
+        that compares April/October and hourly/daily-averaged in a single plot.\\
+
+        Yes, as we mention in the text, we have normalized by those
         quantities by the \textit{local} inertial frequency, not the inertial
@@
 \item {\it The authors may find the work of Brüggeman and Eden (2015) helpful in addressing
-      some of the aforementioned questions.}
+      some of the aforementioned questions.}\\
+
+      Thanks for the suggestion.
 

### 2016-08-19 writeup/RochatEtal_ResponseToReviewers.tex
Fixes typo
Fixes remaining typo
Adds many responses to reviewer 2

@@
         spectra is indistinguishable within errorbars. In any event, we removed
-        that sentence and we believe the new Figure 4 will more efficiently convey
-        the phase cancellation point.
+        that sentence and we believe the new Figure 4 will convey
+        the phase cancellation idea more efficiently.
     \edp
@@
         estimated from the non-tidal forcing run in the Supplemental material, authors'
-        conclusion could be strengthened.}
+        conclusion could be strengthened.}\\
+
+        \bdp
+          The generation mechanism generation of internal
+          waves are changes in the wind (NIWs) and barotropic tides sloshing over
+          topographic features --- we are not claiming that there exists a different
+          generation mechanism
+          in summer compared to winter; neither are we claiming that surface NIWs
+          are generated by internal tides. What is different is the near-surface
+          expression of IGWs, which is very sensitive to the near-surface stratification
+          (D'saro 1978). This is particular true for higher vertical-mode waves
+          (supplemental material).\\
+
+          To reference potential mechanisms of how energy can quickly be transferred to
+          high wavenumbers by interaction with mesoscale-submesoscale turbulence and
+           PSI,  we added new references to the discussion (Ponte \& Klein 2015;
+          Mackinon \& Winters 2004; Alford et al 2016). \\
+
+          The simulation not forced with tides (ECCO2 Adjoint) has much coarser
+          horizontal resolution and barely resolves the submesoscales; its grid
+          spacing is 18 km, while its effective resolution larger than 70 km.
+      \edp
 
@@
         to remove all discussions related to submesoscale inertial-gravity waves and focus on
-        describing differences between October and April.}
+        describing differences between October and April.}\\
+
+        \bdp
+            We are now citing the recent review on NIWs by Alford et al. (2016) that contain
+            a section dedicated to mesoscale-turbulence-NIW interactions, citing all the
+            aforementioned papers. We believe the new discussion together with Figure 4
+            and Figure S7 will help clear any ambiguity. Those figures unambiguously
+            show that most of the IGWs with horizontal scales between 10-100 km
+            are mostly super-inertial (not near-inertial); the refracted/dispersed  NIWs
+            propagate into the interior and only weakly project onto SSH.
+        \edp
 
@@
         authors decide to discuss internal waves in 10-100 km scale.}
+
+        \bdp
+            Following reviewer's \#3 suggestion, we have added paragraphs about SWOT
+            to the introduction and conclusion. This adds context to the manuscript
+            and help justify why we focus only on 10-100 km scales. Our definition
+            of submesoscales have pros and cons (see reply to reviewer's 3
+            minor comment \#1) but is consistent with the definition used by
+            (at least part of) the SWOT community (Fu\& Ferrari 2008). Giving the
+            definition, which we present in the first sentence of the introduction,
+            then there is no ambiguity in using 100 km as a cut-off scale for the filter.
+            We prefer to present bulk statistics first, and then discuss details of
+            different scales --- this makes even more sense now with the new wavenumber-frequency
+            analysis.
+        \edp
+
 \end{enumerate}
@@
         would also like to point out that positions of KEO and KESS moorings in the
-        supplemental material are outside of the study region.}
+        supplemental material are outside of the study region.}\\
+
+        \bdp
+             The upper bound on
+             the size of the domain is chosen for lateral homogeneity purposes,
+             necessary for the spectral analysis. We decided to focus in
+             this subdomain to avoid the strong lateral inhomogeneities associated
+             with the Kuroshio upstream and near its separation from Japan's coast.
+             We have consider larger and smaller domains and found insiginificant
+             quantitative differences.\\
+
+             The supporting information explicitly states that both KEO and KESS
+             mooring were outside our chosen domain. We analyzed data from those
+             mooring simply to assess the skill of numerical model in representing
+             high-frequency modes. As discussed in the supporting information, the
+             comparison uses model output in the grid points closest
+             to the mooring.
+        \edp
 
@@
         with a horizontal scales of 10-100 km (hereafter submesoscale inertial-gravity
-        waves)”?}
+        waves)”?}\\
+
+        \bdp
+          Definition included --- thanks for pointing that out.
+        \edp
 
@@
 
+          \bdp
+              This simply means that there is significant ageostrophic components,
+              even in April. Of course, daily-averaging the fields filters most of
+              ageostrophic. This does not imply that Sasaki et al's mechanism is
+              incorrect; it does suggests, however, a more complicated picture, with
+              likely small energy leak towards small scales. In any event, to avoid ambiguity, we
+              have rewritten that sentence and excluded repeated information.
+          \edp
+
   \item {\it X-10, section 4, 2nd paragraph, the last sentence: Please add some reference.}
-
+          The sentence has been rewritten.
 \end{enumerate}

### 2016-08-19 writeup/rocha_etal.tex
Adds many responses to reviewer 2

@@
 
-Submesoscale IGWs, however, peak in late summer/early fall,
+IGWs with horizontal scales between 10-100 km (hereafter submesoscale IGWs),
+however, peak in late summer/early fall,
 when the upper ocean is strongly stratified. Thus there exists a
@@
 ageostrophic component as depicted by the joint-PDF of vorticity-Laplacian of SSH
-winter (Figure \ref{fig3}e). In other words, even in April, when submesoscale turbulence
-prevails, only the daily-averaged fields
+winter (Figure \ref{fig3}e). The daily-averaged fields
 are largely
-in  geostrophic balance (the joint-PDF of daily-averaged vorticity vs.\\ Laplacian of sea-surface height
-is an ellipse with large eccentricity and main axis tilted by 45$^\circ$; Figure \ref{fig3}f).
+in  geostrophic balance (Figure \ref{fig3}f).
 
@@
 a strong seasonal cycle that is out of phase with the seasonal cycle of
-submesoscale turbulence. We conjecture that the summertime dominance of IGWs
- \citep{callies_etal2015} is a consequence both of suppression of submesoscale turbulence and
-enhancement of IGWs due to re-stratification of
-the upper ocean.
+submesoscale turbulence. IGW energy and SSH variance are significantly spread across
+the submesoscales, likely through mesoscale refraction and dispersion \citep[e.g.,][and references there in]{ponte_klein2015,alford_etal2016}
+ and parametric
+subharmonic instability \citep[e.g., ][]{mackinnon_winters2005}. In the LLC simultions,
+super-inertial IGWs account for most of the surface submesoscale KE and SSH variance;
+near-inertial IGWs project onto larger scales.
+
+ %We further conjecture that the summertime dominance of IGWs
+ %\citep{callies_etal2015} is a consequence both of suppression of submesoscale turbulence and
+%enhancement of IGWs due to re-stratification of
+%the upper ocean.
 

### 2016-08-20 writeup/RochatEtal_ResponseToReviewers.tex
Small fixes

@@
 
-          The simulation not forced with tides (ECCO2 Adjoint) has much coarser
+          The simulation without tides (ECCO2 Adjoint) has much coarser
           horizontal resolution and barely resolves the submesoscales; its grid
-          spacing is 18 km, while its effective resolution larger than 70 km.
+          spacing is 18 km, while its effective resolution larger than 60 km.
       \edp
@@
         other flows. Given that this is the main message of the paper, the authors must make this
-        modification.}
+        modification.}\\
+
+        \bdp
+            We have included wavenumber-frequency
+            spectra of SSH variance (new figure 4) and KE (figure S7). Plotting bounds for the
+            dispersion relationship of free IGWs, as suggested by the reviewer, unambiguously
+            help us make that IGWs account for most of the super-inertial variability at
+            scales between 10-100 km. Thanks for the excellent suggestion!
+        \edp
 
@@
           irritate others. But
-          besides unsettled definitons by our community abd consistency with our recent
+          besides unsettled definitons by our community and consistency with our recent
           work (Rocha et al. 2016), there are at least two reasons for sticking to
@@
           balanced'', then the comment about the GM spectrum containing both balanced
-          and unbalanced flows is egregiously incorrect.  GM used linear theory
+          and unbalanced flows is incorrect.  GM used linear theory
           of IGW to synthesize observations, mostly mooring data --- they

### 2016-08-20 writeup/rocha_etal.tex
Small fixes

@@
 \section{Discussion and conclusion}
-Our work adds to recent modeling \citep{sasaki_etal2014}
-and observational \citep{callies_etal2015,buckingham_etal2016} evidence of vigorous seasonality in
+Our work adds to recent modeling \citep[e.g., ][]{sasaki_etal2014}
+and observational \citep[e.g., ]{callies_etal2015,buckingham_etal2016} evidence of vigorous seasonality in
 submesoscale turbulence. In particular, our main finding is that in global simulations with embedded tides the

### 2016-08-20 writeup/rocha_etal.tex
Fixes typo

--- 2016-08-19 writeup/rocha_etal.tex
@@
     strong modulation of near-surface dynamics at submesoscales
-    (roughly 10-100 km) in the Kuroshio Extension. Consistent with recent studies, deep late-winter mixed layers
+    (10-100 km) in the Kuroshio Extension. Consistent with recent studies, deep late-winter mixed layers
     are prone to baroclinic instabilities, and submesoscale turbulence
@@
 % now SWOT
-Characterizing the partition between geostrophic and ageostrophic flows across
+The characterization of the partition between geostrophic and ageostrophic flows across
 the submesoscales and their seasonality has immediate applications for the planning
-of the upcoming Surface and Water Ocean Topography (SWOT) satellite mission.
+of the Surface and Water Ocean Topography (SWOT) satellite mission.
 SWOT will measure
@@
 \section{Discussion and conclusion}
-Our work adds to recent modeling \citep{sasaki_etal2014}
-and observational \citep{callies_etal2015,buckingham_etal2016} evidence of vigorous seasonality in
+Our work adds to recent modeling \citep[e.g., ][]{sasaki_etal2014}
+and observational \citep[e.g., ][]{callies_etal2015,buckingham_etal2016} evidence of vigorous seasonality in
 submesoscale turbulence. In particular, our main finding is that in global simulations with embedded tides the
@@
 submesoscale turbulence. IGW energy and SSH variance are significantly spread across
-the submesoscales, likely through mesoscale refraction and dispersion \citep[e.g.,][and references there in]{ponte_klein2015,alford_etal2016}
+the submesoscales, likely through mesoscale refraction and dispersion \citep[e.g.,][and references therein]{ponte_klein2015,alford_etal2016}
  and parametric
-subharmonic instability \citep[e.g., ][]{mackinnon_winters2005}. In the LLC simultions,
+sub-harmonic instability \citep[e.g., ][]{mackinnon_winters2005}. In the LLC simultions,
 super-inertial IGWs account for most of the surface submesoscale KE and SSH variance;
@@
 in the mixed layer strongly depends on the density jump at the mixed layer
-base, with largest mixed layer velocities when the jump is strongest.
+base, with largest velocities when the jump is strongest.
 In summer, the shallow mixed layer overlays a strong seasonal pycnocline,
@@
 account for most of the seasonality in super-inertial surface KE and SSH
-(Figure \ref{fig4}a-b), likely because submesoscale
+(Figure \ref{fig4}a-b), possibly because submesoscale
 NIWs quickly propagate into the interior and weakly project onto SSH.
@@
 from SSH, then the noise-to-signal ratio will have a strong seasonality owing
-to the presence of incoherent IGWs (coherent internal tides may be removed
-efficiently).
+to the presence of incoherent IGWs --- the coherent fraction of internal tides may be removed
+efficiently.
 An important caveat is that the Kuroshio Extension may be typical of a mesoscale-rich
@@
  William R. Young and the reviewers --- Christian Buckingham and two anonymous referees ---
-  provided exceptionally helpful and constructive comments. Greg Wagner first
- suggested that the near-surface shape of the baroclinic modes may change seasonally.
+  provided exceptionally helpful and constructive comments. Greg Wagner
+ suggested that the near-surface shape of the baroclinic modes changes seasonally.
  We thank the MITgcm community and our colleagues at the NASA Advanced

### 2016-08-20 writeup/rocha_etal.tex
Small fix

@@
  and parametric
-sub-harmonic instability \citep[e.g., ][]{mackinnon_winters2005}. In the LLC simultions,
+sub-harmonic instability \citep[e.g., ][]{mackinnon_winters2005}.
+In the LLC simultions,
 super-inertial IGWs account for most of the surface submesoscale KE and SSH variance;
-near-inertial IGWs project onto larger scales.
+near-inertial IGWs project onto larger scales, possibly because submesoscale
+NIWs quickly propagate into the interior and weakly project onto SSH.
 
@@
 higher frequency IGWs if the near-surface stratification varies strongly.
-Windy near-inertial wave (NIW) generation is bursty and peaks in early winter, but
-NIW energy appears to have similar energy level in October and April
-\citep{alford_etal2016}.
-While NIWs are generated at large scales, mesoscale turbulence can
-refract and disperse those waves in few inertial periods, thereby generating
-submesoscale (10-100 km) near-inertial waves \citep[][and references therein]{alford_etal2016}.
-But in our simulations the higher frequency inertia-gravity-wave
+Windy near-inertial wave (NIW) generation is bursty and peaks in early winter and
+NIW energy is similar energy level in October and April
+\citep{alford_etal2016}, but in our simulations the higher frequency inertia-gravity-wave
 account for most of the seasonality in super-inertial surface KE and SSH
-(Figure \ref{fig4}a-b), possibly because submesoscale
-NIWs quickly propagate into the interior and weakly project onto SSH.
+(Figure \ref{fig4}a-b),
 

### 2016-08-20 writeup/rocha_etal.tex
Fixes font of frequencies

@@
 The LLC4320 simulation is an extension of the 3-month long output used by
-\citet{rocha_etal2016} --- for details, see supplemental material.
+\citet{rocha_etal2016} --- for details, see supporting information.
 
@@
 The bulk of vertical vorticity and strain rate are associated with subinertial flows
-($T_f = 2\pi/f_0\approx 23.5$h, where $f_0$ is the inertial frequency at
-the mean latitude): daily-averaging the velocity fields suppresses super-inertial
+($T_{32.5} = 2\pi/f_{32.5}\approx 22.3$h, where $f_{32.5}$ is the inertial frequency at
+the  mid-latitude $32.5^\circ$N): daily-averaging the velocity fields suppresses super-inertial
 motions and reduces the RMS
@@
           waves from mode 1 through mode 4 across the latitudinal domain; horizontal dashed lines represent the semi-diurnal
-          lunar tidal frequency (M$_2$), and the inertial frequency at mid-latitude
-          (f$_{32.5}$) and its first harmonic (2f$_{32.5}$). }
+          lunar tidal frequency ($M_2$), and the inertial frequency at mid-latitude
+          ($f_{32.5}$) and its first harmonic (2$f_{32.5}$). }
   \label{fig4}

### 2016-08-22 writeup/rocha_etal.tex
Minor edits

@@
   \begin{center}
-    \includegraphics[width=.9\textwidth]{figs/fig4.pdf}
+    \includegraphics[width=.8\textwidth]{figs/fig4.pdf}
  \caption{\small LLC4320 wavenumber-frequency spectrum of SSH variance in (a) April and (b) October.
@@
 from SSH, then the noise-to-signal ratio will have a strong seasonality owing
-to the presence of incoherent IGWs --- the coherent fraction of internal tides may be removed
+to the presence of IGWs --- the coherent fraction of internal tides may be removed
 efficiently.

### 2016-08-22 writeup/rocha_etal_supporting-info.tex
Removes annoying of course
Minor edits

@@
            climatology [updates from \cite{roemmich_gilson2009}] (a, b) and the 1/24$^\circ$ (c, d)
-           and  1/48$^\circ$ (e, f) LLC simulations. Of course, the Argo climatology is smoother,
+           and  1/48$^\circ$ (e, f) LLC simulations. The Argo climatology is smoother
            owing to lower resolution and long-term averaging. Nonetheless, simulations reproduce the strong
@@
     the latitudinal domain; horizontal dashed lines represent the semi-diurnal lunar
-    tidal frequency (M2), and the inertial frequency at mid-latitude (f32.5) and its
-    first harmonic (2f32.5).}
+    tidal frequency (M2), and the inertial frequency at mid-latitude ($f_{32.5}$) and its
+    first harmonic ($2f_{32.5}$).}
    \label{figS7}

### 2016-08-22 writeup/rocha_etal.tex
Adds little table and minor edits

@@
 
-\keypoints{\item  Upper-ocean submesoscale (10-100 km) turbulence and inertia-gravity
+\keypoints{\item  Upper-ocean submesoscale (here 10-100 km) turbulence and inertia-gravity
                   waves
@@
 of the Surface and Water Ocean Topography (SWOT) satellite mission.
-SWOT will measure
+SWOT will provide
 sea-surface height (SSH) with about 15 km resolution, thereby providing the
@@
 nuisance\footnote{``One man's noise is
-another man's signal'' (Walter Munk).} may upset the diagnose
+another man's signal'' (Walter Munk).} may upset the estimation
 of submesoscale surface geostrophic velocities.
@@
 Because submesoscale turbulence is weakest in late summer/early fall,
-the present results indicate that IGWs account for most of the
+IGWs account for most of the
 summertime submesoscale sea-surface height variability with implications for SWOT.
@@
 numerical solutions on a LLC grid  \citep{forget_etal2015} with
-90-vertical-levels. The coarser-resolution LLC simulation was
+90 vertical levels. The coarser-resolution LLC simulation was
 spun up from an Estimating the Circulation and Climate of the Ocean, Phase II \citep[ECCO2; ][]{menemenlis_etal2008}
@@
 small-scale turbulence de-stratifies the upper ocean, yielding mixed layers
-as deep as 300 m (Figure \ref{fig1}d). Summertime re-stratification yields mixed layers shallower
-than 40 m (Figure \ref{fig1}e).
+as deep as 300 m (Figure \ref{fig1}b). Summertime re-stratification yields mixed layers shallower
+than 40 m (Figure \ref{fig1}c).
 Fundamentally, the upper-ocean density structure is well-captured by both LLC simulations:
@@
 temporal and spatial scale separation between those motions and IGWs
-\citep{mcwilliams2016}:
+\citep[][]{mcwilliams2016}:
 daily-averaging the velocity fields efficiently suppresses the summertime horizontally
@@
 \section{Wavenumber-frequency spectrum}
-To confirm that submesoscale IGWs undergo a vigorous seasonal cycle near the surface,
-we calculate wavenumber-frequency spectrum of SSH variance. Focusing on the high-frequency
+To confirm that submesoscale IGWs account for most of the super-inertial SSH variance
+and surface KE, we calculate wavenumber-frequency spectrum of SSH variance. Focusing on
+the high-frequency
 content, we compute the wavenumber-frequency spectrum  every 10 days and average
-the results to obtain a spectral estimate. Before calculating the each spectrum we
+the results to obtain a spectral estimate. Before calculating the discrete Fourier
+transform, we
 remove linear trends and multiply the data by a three-dimensional Hanning window.
@@
 simulation. Mesoscales flows and the lunar semi-diurnal tide contain most of the
-SSH variance. But the is significant SSH variance at submesoscales, both subinertial
-and super-inertial. Subinertial SSH variance at submesoscales is larger in April, whereas
+SSH variance. But there is significant SSH variance at submesoscales, both sub-inertial
+and super-inertial. Sub-inertial SSH variance at submesoscales is larger in April, whereas
 super-inertial SSH variance at submesoscales is larger in October. At scales smaller
@@
 IGWs, particularly modes two through four, account for most of the SSH variance
-increase in summer, --- higher vertical
+increase in summer --- higher vertical
 modes are more sensitive to changes in near-surface stratification (supporting information).
-Similarly, the submesoscale superinertial surface KE is 63$\%$ larger in October
+In October, super-inertial IGWs account for a SSH standard deviation of about 4.9 cm,
+comparable to the April sub-inertial SSH standard deviation of $\sim$5.2 cm (Table \ref{tab1}).
+The
+Similarly, the submesoscale super-inertial surface KE is 63$\%$ larger in October
 than the April (supporting information).
 
+% table
+\begin{table}
+\label{tab1}
+\caption{The square root of the mean-square SSH at submesoscales (10-100 km)
+        estimated from the wavenumber-frequency spectrum of SSH variance in figure \ref{fig4}.
+        }
+  \begin{center}
+    \begin{tabular}{  l | c | c }
+    \hline
+    & April & October\\
+    \hline
+  sub-inertial ($<0.8f$)   & 5.2 cm & 2.0 cm  \\
+  super-inertial ($>1.2f$) & 2.9 cm  & 4.9 cm\\
+  \hline
+    \end{tabular}
+  \end{center}
+\end{table}
+
 Figure \ref{fig4}c shows the integral over frequency
-of wavenumber-frequency spectra in figures \ref{fig4}a-b. The
-SSH wavenumber spectrum approximately follows a -2 power-law at submesoscales,
-both in April and October (solid lines in figure \ref{fig4}c)  --- there is
+of wavenumber-frequency spectra in figures \ref{fig4}a-b. ,
+Both in April and October, the
+SSH wavenumber spectrum approximately follows a -2 power-law at submesoscales
+(solid lines in figure \ref{fig4}c)  --- there is
 small difference between the spectra owing to the phase cancellation between the
@@
 higher frequency IGWs if the near-surface stratification varies strongly.
-Windy near-inertial wave (NIW) generation is bursty and peaks in early winter and
+Windy near-inertial wave (NIW) generation is bursty and peaks in early winter, but
 NIW energy is similar energy level in October and April
-\citep{alford_etal2016}, but in our simulations the higher frequency inertia-gravity-wave
+\citep{alford_etal2016}. Furthermore, in our simulations the higher frequency inertia-gravity-wave
 account for most of the seasonality in super-inertial surface KE and SSH

### 2016-08-23 writeup/RochatEtal_ResponseToReviewers.tex
Fixes typo
Adds little table and minor edits
Small edits

@@
         \bdp
-          Definition included --- thanks for pointing that out.
+          We included the definition --- thanks for pointing that out.
         \edp
@@
           Or it just means that the balanced regime is actually the submeso-dynamics
-          regime?}
+          regime?}\\
 
@@
               implication for SWOT in section 6. Thanks for making us provide a better
-              context for out study.
+              context for our study.
           \edp
@@
         go to? and (2) how does this affect your results? These questions need to be addressed or at
-        least discussed if the main topic of your paper is submesoscale dynamics.}
+        least discussed if the main topic of your paper is submesoscale dynamics.}\\
+
+        \bdp
+            We are aware of the this issue --- the effective resolution based on the KE spectrum,
+            was already mentioned
+            in the supporting information to the first draft.  In
+            fact, the smallest scale discussed in our paper (10 km) was set by the effective
+            resolution of highest-resolution simulation. To be more explicit, we have now included
+            the effective resolution together with the nominal resolutions in the main text (Section 2).
+
+            The ``unresolved energy'' does not come from anywhere, it is unresolved.
+            To the extend that 1-km-scale are associated with a forward
+            energy transfer, then their effects on larger scales (10-100 km) is likely insignificant.
+            Also important to this discussion is that higher resolution simulations will resolve
+            1-km-scale eddies and high-wavenumber IGWs that are generated through wave-wave
+            interactions and wave-vorticity interactions.
+        \edp
 
@@
         variance in summer is the sum of balanced + unbalanced motions, the difference between
-        these two quantities yields a bound1 for the SSH variance due to unbalanced motions.}
+        these two quantities yields a bound1 for the SSH variance due to unbalanced motions.}\\
+
+        \bdp
+            We have integrated the wavenumber-frequency
+            spectrum of SSH (Figure 4) to estimate variance in a particular frequency
+            bands.  We have included a table with the standard
+            deviation of SSH for super-inertial and sub-inertial flows at scales
+            smaller than 100 km in different seasons. Thanks for the great suggestion!
+        \edp
 
@@
           We explicit state that those are submesoscale IGWs. Using
-          our definition of submesoscales are IGWs
+          our definition of submesoscales, these are IGWs
           with horizontal scales about 10-100 km\footnote{Bill Young suggested the
@@
 
-      Thanks for the suggestion.
+      Thanks.
 

### 2016-08-23 writeup/rocha_etal.tex
Small edits

@@
 small-scale turbulence de-stratifies the upper ocean, yielding mixed layers
-as deep as 300 m (Figure \ref{fig1}b). Summertime re-stratification yields mixed layers shallower
+deeper than 300 m (Figure \ref{fig1}b). Summertime re-stratification yields mixed layers shallower
 than 40 m (Figure \ref{fig1}c).

### 2016-08-23 writeup/RochatEtal_ResponseToReviewers.tex
Starts summary session

@@
 \maketitle
+
+\section*{Summary}
+
+We thank the reviewers for the speedy reviewer, constructive criticism, and
+great suggestions.
 
@@
             resolution of highest-resolution simulation. To be more explicit, we have now included
-            the effective resolution together with the nominal resolutions in the main text (Section 2).
+            the effective resolution together with the nominal resolutions in the main text (Section 2).\\
 
-            The ``unresolved energy'' does not come from anywhere, it is unresolved.
+            The ``unresolved energy'' does not come from anywhere; it is unresolved.
             To the extend that 1-km-scale are associated with a forward
-            energy transfer, then their effects on larger scales (10-100 km) is likely insignificant.
+            energy transfer, then their effects on the scales concerning our study
+            (10-100 km) are likely energetically insignificant.
             Also important to this discussion is that higher resolution simulations will resolve
             1-km-scale eddies and high-wavenumber IGWs that are generated through wave-wave
-            interactions and wave-vorticity interactions.
+            interactions and wave-vorticity interactions.\\
+
+            Buckingham et al. (2016) have a very good point that models and observations
+            to date may not resolve the mixed-layer deformation scale in summer; this is no
+            different for our simulations. But summer is also the season when lateral buoyancy
+            gradients (and thus available potential energy) are minimum within the mixed layer,
+            so the reviewer's conjecture that this is the main source of seasonally varying eddy
+            kinetic energy is not obvious.\\
+
+            For geostrophically balanced mixed-layer instabilities (which we call
+            shallow baroclinic instabilities), linear theory
+            predict maximum growth rates at about $2.6\times R_{ml}$, where
+            $R_{ml}$ is the mixed-layer deformation scale. But in equilibrated
+            QG turbulence driven by ML instabilities (Callies et al. 2016) the
+            deformation scale has only a catalytic role, where eddy baroclinic
+            energy is converted into (mixed-layer) eddy barotropic energy; as
+            in interior baroclinic geostrophic turbulence, most of the baroclinic
+            conversion occurs at larger scales (Callies et al. 2016).
+
+            Still, we recognize that this issue may be further investigated and
+            we included a paragraph to the discussion.
+
         \edp
@@
           But there seems to be no clear definition in the literature --- the
-          Stammer $\&$ Wunsch paper that the reviewer states that their
+          Stammer $\&$ Wunsch paper states that their
           definition of mesoscales is ``very rough'' and ``there is no generally
@@
           irritate others. But
-          besides unsettled definitons by our community and consistency with our recent
+          besides the unsettled definitions in our community and consistency with our recent
           work (Rocha et al. 2016), there are at least two reasons for sticking to

### 2016-08-23 writeup/RochatEtal_ResponseToReviewers.tex
Small fix

@@
           If the reviewer is using ``balanced'' as a short to ``geostrophically
+          and hydrostatically 
           balanced'', then the comment about the GM spectrum containing both balanced

### 2016-08-23 writeup/RochatEtal_ResponseToReviewers.tex
Adds calculation of MLD

@@
 
-We thank the reviewers for the speedy reviewer, constructive criticism, and
-great suggestions.
+In this revised version, we have addressed the reviewers
+comments. The most notable change is the inclusion of a wavenumber-frequency
+spectral analysis suggested by reviewer 2. This analysis helps us vindicate
+our claim that inertia-gravity waves, particularly super-inertial waves,
+account for most of the summertime SSH variance and surface KE at scales
+between 10 and 100 km. We now also place our study in the context of SWOT and discuss
+the implications of our findings for the estimation of geostrophic velocity from
+high-resolution altimeters. We thank the reviewers  for the speedy review, constructive
+criticism, and fine suggestions, which helped us improve our discussion and
+provide a better context for our study.
 

### 2016-08-23 writeup/RochatEtal_ResponseToReviewers.tex
Adds new figure 1

@@
         \bdp
-          We changed the order of the panels. All figures are provided in high-quality
-          resolution. We avoid plotting the mixed-layer depth because it is noisy,
-          hindering an important detail we would like the judicious reader to note,
-          particularly in figure 1c: the wiggles in the mixed-layer base
-          indicate internal waves.
+          We have changed the order of the panels and plotted the mixed layer depth
+          on the density plots.
         \edp

### 2016-08-23 writeup/rocha_etal.tex
Adds new figure 1

--- 2016-08-22 writeup/rocha_etal.tex
@@
 small-scale turbulence de-stratifies the upper ocean, yielding mixed layers
-as deep as 300 m (Figure \ref{fig1}b). Summertime re-stratification yields mixed layers shallower
+deeper than 200 m to the north (Figure \ref{fig1}b). Summertime re-stratification yields mixed layers shallower
 than 40 m (Figure \ref{fig1}c).
@@
           transects
-          of potential density at 165$^\circ$E (b and c) and  surface vorticity (d and e) . The snapshots were
-          taken at 00:00 UTC.}
+          of potential density at 165$^\circ$E (b and c) and  surface vorticity (d and e).
+          In (b) and (c) the white line represents the mixed layer depth. The snapshots were
+          taken at 12:00 UTC.}
 \vspace{-1.5cm}

### 2016-08-24 writeup/RochatEtal_ResponseToReviewers.tex
Small fixes

@@
 
-In this revised version, we have addressed the reviewers
-comments. The most notable change is the inclusion of a wavenumber-frequency
+In this revised manuscript, we have addressed the reviewers
+comments. The most fundamental change is the inclusion of a wavenumber-frequency
 spectral analysis suggested by reviewer 2. This analysis helps us vindicate
@@
 criticism, and fine suggestions, which helped us improve our discussion and
-provide a better context for our study.
+forced us to provide a better context for our study.
 
@@
             Also important to this discussion is that higher resolution simulations will resolve
-            1-km-scale eddies and high-wavenumber IGWs that are generated through wave-wave
+            1-km-scale eddies and high-wavenumber IGWs that are generated via wave-wave
             interactions and wave-vorticity interactions.\\
 
-            Buckingham et al. (2016) have a very good point that models and observations
+            Buckingham et al. (2016) made a very good point that models and observations
             to date may not resolve the mixed-layer deformation scale in summer; this is no
-            different for our simulations. But summer is also the season when lateral buoyancy
-            gradients (and thus available potential energy) are minimum within the mixed layer,
-            so the reviewer's conjecture that this is the main source of seasonally varying eddy
-            kinetic energy is not obvious.\\
+            different for our simulations. But the interpretation that this is the main
+            source of seasonally varying eddy kinetic energy is simplistic because
+            summer is also the season when lateral buoyancy
+            gradients (and thus available potential energy) are minimum within the mixed layer
+            (e.g., Callies et al. 2015).\\
 
-            For geostrophically balanced mixed-layer instabilities (which we call
+            For geostrophic mixed-layer instabilities (which we call
             shallow baroclinic instabilities), linear theory
@@
             deformation scale has only a catalytic role, where eddy baroclinic
-            energy is converted into (mixed-layer) eddy barotropic energy; as
-            in interior baroclinic geostrophic turbulence, most of the baroclinic
-            conversion occurs at larger scales (Callies et al. 2016).
+            energy is converted into (mixed-layer) eddy barotropic energy ---
+            most of the baroclinic
+            conversion occurs at scales much large than the mixed-layer deformation
+            scale (Callies et al. 2016).
 
@@
           with horizontal scales about 10-100 km\footnote{Bill Young suggested the
-          term submesoscale IGWs.}; IGWs have periods  The GM spectrum is not accurate
+          term submesoscale IGWs.}; IGWs have periods  The GM spectrum is inaccurate
           close to the surface (most of our results concern surface currents and SSH,
@@
           and unbalanced flows is incorrect.  GM used linear theory
-          of IGW to synthesize observations, mostly mooring data --- they
+          of IGW to synthesize observations, mostly mooring data, in the IGW band
+           --- they
           fitted an analytic model to (super-inertial) frequency spectra and
@@
       as a function of wavenumber, not the quantity in the y-axis.\footnote{This confusion is similar
-      to the pervasive axis mislabelling in plots of probability distribution function (PDF)
+      to the pervasive axis poor-labelling in plots of probability distribution function (PDF)
       --- the best label is \textit{probability density} not PDF.}

### 2016-08-26 writeup/RochatEtal_ResponseToReviewers.tex
Many changes to the text

@@
 \usepackage{natbib}
+\bibliographystyle{chicago}
+
+\usepackage{color}
 
@@
 
-\title{{\normalfont Response to reviewers of the manuscript \texttt{2016GL070261} }
+\title{{\normalfont Reply to reviewers of the manuscript \texttt{2016GL070261} }
        ``Seasonality of submesoscale
@@
 
-In this revised manuscript, we have addressed the reviewers
+In this revised manuscript, we have addressed the reviewers'
 comments. The most fundamental change is the inclusion of a wavenumber-frequency
-spectral analysis suggested by reviewer 2. This analysis helps us vindicate
+spectral analysis suggested by reviewer 3. The new results vindicate
 our claim that inertia-gravity waves, particularly super-inertial waves,
 account for most of the summertime SSH variance and surface KE at scales
-between 10 and 100 km. We now also place our study in the context of SWOT and discuss
-the implications of our findings for the estimation of geostrophic velocity from
-high-resolution altimeters. We thank the reviewers  for the speedy review, constructive
-criticism, and fine suggestions, which helped us improve our discussion and
-forced us to provide a better context for our study.
+between 10 and 100 km. We have also expanded the introductions and discussion
+to place our study in the context of Surface Water \& Ocean Topography (SWOT)
+satellite mission --- we discuss
+the implications of our findings for the estimation of surface velocity from
+high-resolution sea-surface height.
+
+We thank the reviewers  for the speedy review, constructive
+criticism, and fine suggestions, which has helped us improve our discussion and
+forced us to provide a better context for our study. We hope our manuscript is
+now suitable for publication and look forward to you reply.
 
@@
         spectra is indistinguishable within errorbars. In any event, we removed
-        that sentence and we believe the new Figure 4 will convey
+        that sentence from the text. We also believe the new Figure 4 will convey
         the phase cancellation idea more efficiently.
@@
 turbulence versus inertia-gravity waves.}
+
+  \bdp
+      The mesoscale eddy KE level compares well to gridded AVISO
+      geostrophic KE. \\\textcolor{purple}{Need to do the comparison with alon-track data}.
+  \edp
 
@@
           expression of IGWs, which is very sensitive to the near-surface stratification
-          (D'saro 1978). This is particular true for higher vertical-mode waves
-          (supplemental material).\\
+          (D'saro 1978; cf. second paragraph of discussion).\\
 
@@
           high wavenumbers by interaction with mesoscale-submesoscale turbulence and
-           PSI,  we added new references to the discussion (Ponte \& Klein 2015;
-          Mackinon \& Winters 2004; Alford et al 2016). \\
+           PSI,  we added new references to the discussion \citep{mackinnon_winters2005,
+           ponte_klein2015, alford_etal2016}. \\
 
@@
           horizontal resolution and barely resolves the submesoscales; its grid
-          spacing is 18 km, while its effective resolution larger than 60 km.
+          spacing is 18 km, while its effective resolution larger than 60 km.\\
+          \textcolor{purple}{Dimitris --- can I have access to the tideless 18-km LLC output?}
       \edp
@@
         \bdp
-            We are now citing the recent review on NIWs by Alford et al. (2016) that contain
-            a section dedicated to mesoscale-turbulence-NIW interactions, citing all the
+            We are now citing the recent review on NIWs by \cite{alford_etal2016} that contains
+            a section dedicated to mesoscale-turbulence-NIW interactions ---  \cite{alford_etal2016}
+            cite all the
             aforementioned papers. We believe the new discussion together with Figure 4
@@
             are mostly super-inertial (not near-inertial); the refracted/dispersed  NIWs
-            propagate into the interior and only weakly project onto SSH.
+            propagate into the interior and only weakly project onto SSH \citep{alford_etal2016}.
         \edp
@@
             Following reviewer's \#3 suggestion, we have added paragraphs about SWOT
-            to the introduction and conclusion. This adds context to the manuscript
-            and help justify why we focus only on 10-100 km scales. Our definition
+            to the introduction and conclusion. This discussion adds context to our study
+            and justifies the focus on the horizontal scales between 10-100 km. Our definition
             of submesoscales have pros and cons (see reply to reviewer's 3
@@
             We prefer to present bulk statistics first, and then discuss details of
-            different scales --- this makes even more sense now with the new wavenumber-frequency
+            different scales --- this makes more sense now with the new wavenumber-frequency
             analysis.
@@
              necessary for the spectral analysis. We decided to focus in
-             this subdomain to avoid the strong lateral inhomogeneities associated
+             this subdomain to avoid the strong lateral inhomogeneity and
+             anisotropies associated
              with the Kuroshio upstream and near its separation from Japan's coast.
-             We have consider larger and smaller domains and found insiginificant
+             We have considered larger and smaller domains and found insiginificant
              quantitative differences.\\
@@
              high-frequency modes. As discussed in the supporting information, the
-             comparison uses model output in the grid points closest
+             comparison uses model output at the grid points closest
              to the mooring.
@@
               likely small energy leak towards small scales. In any event, to avoid ambiguity, we
-              have rewritten that sentence and excluded repeated information.
+              have re-written that sentence and excluded repeated information.
           \edp
 
-  \item {\it X-10, section 4, 2nd paragraph, the last sentence: Please add some reference.}
-          The sentence has been rewritten.
+  \item {\it X-10, section 4, 2nd paragraph, the last sentence: Please add some reference.}\\
+          We have re-written that sentence.
 \end{enumerate}
@@
             We have included wavenumber-frequency
-            spectra of SSH variance (new figure 4) and KE (figure S7). Plotting bounds for the
-            dispersion relationship of free IGWs, as suggested by the reviewer, unambiguously
-            help us make that IGWs account for most of the super-inertial variability at
+            spectra of SSH variance (new figure 4) and KE (figure S7) with bounds for the
+            dispersion relationship of free IGWs, as suggested by the reviewer. The new
+            results unambiguously
+            indicate that IGWs account for most of the super-inertial variability at
             scales between 10-100 km. Thanks for the excellent suggestion!
@@
               We agree that the planning of SWOT is perhaps the most immediate
-              application of our results. We have added a paragraph dedicated to SWOT (new
+              application of our results. We have added an introductory paragraph dedicated to SWOT (new
               paragraph 2 of section 1). We have also added a discussion about the
@@
         \bdp
-            We are aware of the this issue --- the effective resolution based on the KE spectrum,
+            We are aware of the this issue --- the effective resolution based on the KE spectrum
             was already mentioned
@@
 
-            The ``unresolved energy'' does not come from anywhere; it is unresolved.
-            To the extend that 1-km-scale are associated with a forward
+            The ``unresolved energy'' does not come from anywhere --- flows smaller than
+            about 10 km are strongly damped.
+            To the extent that 1-km-scale flows are associated with a forward
             energy transfer, then their effects on the scales concerning our study
             (10-100 km) are likely energetically insignificant.
-            Also important to this discussion is that higher resolution simulations will resolve
-            1-km-scale eddies and high-wavenumber IGWs that are generated via wave-wave
+            Also important to this discussion is that higher resolution simulations will resolve {\it both}
+            1-km-scale eddies and 1-km-scale IGWs that are generated via wave-wave
             interactions and wave-vorticity interactions.\\
 
-            Buckingham et al. (2016) made a very good point that models and observations
+            \cite{buckingham_etal2016} made a very good point that models and observations
             to date may not resolve the mixed-layer deformation scale in summer; this is no
@@
             gradients (and thus available potential energy) are minimum within the mixed layer
-            (e.g., Callies et al. 2015).\\
+            \citep[e.g., ][]{callies_etal2015}.\\
 
@@
             $R_{ml}$ is the mixed-layer deformation scale. But in equilibrated
-            QG turbulence driven by ML instabilities (Callies et al. 2016) the
+            QG turbulence driven by ML instabilities \citep{callies_etal2016} the
             deformation scale has only a catalytic role, where eddy baroclinic
             energy is converted into (mixed-layer) eddy barotropic energy ---
-            most of the baroclinic
-            conversion occurs at scales much large than the mixed-layer deformation
-            scale (Callies et al. 2016).
+            most of the baroclinic (mean-to-eddy) energy
+            conversion occurs at scales much larger than the mixed-layer deformation
+            scale \citep{callies_etal2016}.\\
 
-            Still, we recognize that this issue may be further investigated and
-            we included a paragraph to the discussion.
+            \textcolor{purple}{Is it worth to add a paragraph to discuss this?}
 
@@
             We have integrated the wavenumber-frequency
-            spectrum of SSH (Figure 4) to estimate variance in a particular frequency
+            spectrum of SSH (Figure 4) to estimate variance in particular frequency
             bands.  We have included a table with the standard
             deviation of SSH for super-inertial and sub-inertial flows at scales
-            smaller than 100 km in different seasons. Thanks for the great suggestion!
+            smaller than 100 km in different seasons. Thanks for the suggestion!
         \edp
@@
           others would argue for $L_1 = 2\pi \times 30 \approx 190$ km
-          (Wunsch \& Ferrari 2004; Ferrari \& Wunsch 2009).
-          To add yet another controversy, Larichev \& Held 1995 argue
-          that the deformation length is not the panacea as linear theorists
+          \citep{wunsch_ferrari2004,ferrari_wunsch2009}.
+          To add yet another controversy, \cite{larichev_held1995} argue
+          that the deformation length is not the panacea that linear theorists
           suggest. In baroclinic geostrophic turbulence, the bulk of the
@@
           radius; the energy-containing scales of equilibrated baroclinic geostrophic
-          turbulence are also larger than the deformation radius (e.g., Larichev \& Held 1995).\\
+          turbulence are also larger than the deformation radius \cite{larichev_held1995}.\\
 
@@
           flow --- the eddies that contain most of the eddy kinetic energy in the ocean
-          (Ferrari \& Wunsch 2009). On observational
-          grounds, this scale varies roughly from 100-300 km (from the centroid of KE wavenumber
-          spectrum from current altimeters for example). We then \textit{define}
+          \citep{ferrari_wunsch2009}. From the the centroid of KE wavenumber
+          spectrum from current altimeters, this scale varies roughly from 100-300 km.
+          We then \textit{define} submesoscales as
           the energetically subdominant scales, the scales smaller than about 100 km.
@@
           of our higher resolution simulation ($\sim$10 km).
-          This definition is consistent with Callies \& Ferrari 2013 (they did not
+          This definition is consistent with \cite{callies_ferrari2013} (they did not
           define mesoscales/submesoscales based on scales that a sensor resolves
@@
           besides the unsettled definitions in our community and consistency with our recent
-          work (Rocha et al. 2016), there are at least two reasons for sticking to
+          work \citep{rocha_etal2016}, there are at least two reasons for sticking to
           our definition: (i) this is the definition used by the SWOT community (Fu \& Ferrari
               2008); incidentally current altimeters do not resolve
-              scales smaller then about 100 km; and (ii) our goal is to determine the flows
-              (geostrophically balanced turbulence, inertia-gravity waves, etc) that
+              scales smaller then about 100 km; and (ii) our goal is to determine which flows
+              (geostrophically balanced turbulence, inertia-gravity waves, etc)
               dominate the 10-100 km horizontal
@@
         \bdp
-          We now describe that there are implications for the accuracy of high-resolution
+          We now state that there are implications for the accuracy of high-resolution
           altimeters.
@@
           with horizontal scales about 10-100 km\footnote{Bill Young suggested the
-          term submesoscale IGWs.}; IGWs have periods  The GM spectrum is inaccurate
+          term submesoscale IGWs.}; IGWs have periods bettwen the inertial and buoyancy
+          periods.  The GM spectrum is inaccurate
           close to the surface (most of our results concern surface currents and SSH,
@@
         that exist in your model? Might you be attributing characteristics to these waves that
-        would be different in the real ocean?}
+        would be different in the real ocean?}\\
+
+        \textcolor{purple}{Dimitris --- do you want to pitch in a reply for this question?}
 
@@
       This might be worth a comment.}
+
+      \bdp
+          See major point $\#$ 3.
+      \edp
 
@@
 
+        \bdp
+            \textcolor{purple}{Ask Bill --- I haven't found any reference that mentions this
+            fact explicitly, but it follows from the linear boussinesq equations, which can
+            be combined to yield $\zeta_t= f_0 \delta$.}
+        \edp
+
 \item {\it (Section 4) I find “jPDF” odd. Try “joint-PDF” or “PDF” in general.}\\
@@
         predicted by ...'' Specifically mention convergence/downwelling.}
+
+        \bdp
+            OK.
+        \edp
 
@@
 
+        \bdp
+            At every snapshot, the horizontally-averaged divergence is nearly zero;
+            the strain rate horizontal average is not zero because by definition
+            $\alpha \ge 0$. The (normalized) RMS better characterizes the the bulk
+            gradient Rossby number of the flow.
+        \edp
+
 \item {\it (Figure 2) What is the sensitivity of the estimates to the stencil type? See Arbic et al.
         paper describing this sensitivity for satellite altimetry.}
+
+        \bdp
+            Our scheme likely slightly underestimates the vorticity. \textcolor{purple}{
+            I will perform a sensitivity analysis.}
+        \edp
 
@@
         readers may appreciate it but others may find it condescending.
-        We argue that most physical oceanographers interested in our paper will
+        We argue that most physical oceanographers  will likely
         likely know this material. Even readers from other areas, who did not take
@@
       term \textit{power spectral density}  appears to have been introduced
-      for frequency spectrum of various quantities in signal processing (Rob Pinkel, personal
+      for \textit{frequency} spectrum of various quantities in signal processing (Rob Pinkel, personal
       communication).}
@@
 
+\bibliography{rocha_etal.bib}
+
 \end{document}

### 2016-08-26 writeup/rocha_etal.tex
Many changes to the text

@@
 % Shorter version of title entered in capital letters:
-\titlerunninghead{Seasonality at submesoscales}
+\titlerunninghead{Seasonality at submesoscales (10-100 km)}
 
@@
  the vorticity is strongly positively skewed, whereas the divergence is moderately
- negatively skewed as predicted by frontogenesis \citep[e.g., ][]{mcwilliams2016}
+ negatively skewed (convergence/downwelling) as predicted by frontogenesis \citep[e.g., ][]{mcwilliams2016}
 . In late summer/early fall, the divergence is stronger, but PDFs are much less skewed,

### 2016-09-18 writeup/RochatEtal_ResponseToReviewers.tex
Minor edits by Sarah and Teri.

@@
 We thank the reviewers  for the speedy review, constructive
-criticism, and fine suggestions, which has helped us improve our discussion and
+criticism, and thoughtful suggestions that have helped us improve our discussion and
 forced us to provide a better context for our study. We hope our manuscript is
-now suitable for publication and look forward to you reply.
+now suitable for publication and look forward to your reply.
 
@@
         We agree that the emphasis on the similarity of hourly spectra can appear
-        out of contradictory if placed out of context. Those spectra (now Figure 4c
+        contradictory. Those spectra (now Figure 4c
         and supporting information) are similar because both submesoscale sub-inertial
@@
         phase --- there is a phase cancellation and the spectra based on hourly
-        spectra is indistinguishable within errorbars. In any event, we removed
-        that sentence from the text. We also believe the new Figure 4 will convey
-        the phase cancellation idea more efficiently.
+        data are indistinguishable within errorbars. In any event, we removed
+        that sentence from the text. The new Figure 4 should convey
+        the phase cancellation idea more effectively.
     \edp
@@
         \bdp
-          The generation mechanism generation of internal
-          waves are changes in the wind (NIWs) and barotropic tides sloshing over
+          The generation mechanism for internal
+          waves involves changes in the wind (NIWs) and barotropic tides sloshing over
           topographic features --- we are not claiming that there exists a different
@@
           spacing is 18 km, while its effective resolution larger than 60 km.\\
-          \textcolor{purple}{Dimitris --- can I have access to the tideless 18-km LLC output?}
       \edp
@@
         \bdp
-            Following reviewer's \#3 suggestion, we have added paragraphs about SWOT
+            Following reviewer \#3's suggestion, we have added paragraphs about SWOT
             to the introduction and conclusion. This discussion adds context to our study
             and justifies the focus on the horizontal scales between 10-100 km. Our definition
-            of submesoscales have pros and cons (see reply to reviewer's 3
+            of submesoscales have pros and cons (see reply to reviewer \#3's
             minor comment \#1) but is consistent with the definition used by
@@
             then there is no ambiguity in using 100 km as a cut-off scale for the filter.
-            We prefer to present bulk statistics first, and then discuss details of
+            We have chosen to present bulk statistics first, and then discuss details of
             different scales --- this makes more sense now with the new wavenumber-frequency
@@
             different for our simulations. But the interpretation that this is the main
-            source of seasonally varying eddy kinetic energy is simplistic because
+            source of seasonally varying eddy kinetic energy is unclear because
             summer is also the season when lateral buoyancy
@@
           decided to term mesoscales the flows with dynamics analogous to
-          the meteorological synoptic scale motions.\\
+          the meteorological synoptic scale motions \citep[Appendix C of ][]{wunsch2015}.\\
 
@@
 
-          We understand that our definition is imperfect --- we may amuse some readers but
+          We understand that our definition is imperfect --- we may satisfy some readers but
           irritate others. But
@@
               dominate the 10-100 km horizontal
-              scales and their seasonality. In any event,  readers that make to the first sentence of our
+              scales and their seasonality. In any event,  readers of the first sentence of our
               abstract will find our explicit definition of submesoscales.
@@
         \bdp
-          We explicit state that those are submesoscale IGWs. Using
+          We explicitly state that those are submesoscale IGWs. Using
           our definition of submesoscales, these are IGWs
           with horizontal scales about 10-100 km\footnote{Bill Young suggested the
-          term submesoscale IGWs.}; IGWs have periods bettwen the inertial and buoyancy
+          term submesoscale IGWs.}; IGWs have periods between the inertial and buoyancy
           periods.  The GM spectrum is inaccurate
@@
 
-          If the reviewer is using ``balanced'' as a short to ``geostrophically
+          If the reviewer is using ``balanced'' as a shorthand to ``geostrophically
           and hydrostatically
@@
           different vertical modes (see Walter Munk's chapter
-          in EPO 1981 and references therein).
+          in \textit{Evolution of Physical Oceanography} 1981 and references therein).
           One could argue
@@
       \bdp
-      Any reader familiar with basic geostrophy concepts is likely to understand the connection.
+      Readers familiar with basic geostrophy concepts are likely to understand the connection.
+      We make the connection between the Laplacian of SSH and relative vorticity
+      in the caption of figure 3.
       \edp
@@
         \bdp
-            OK.
+            We added ``convergence/downwelling'' to the text.
         \edp
@@
             At every snapshot, the horizontally-averaged divergence is nearly zero;
-            the strain rate horizontal average is not zero because by definition
+            the strain rate horizontal average is not zero, because by definition
             $\alpha \ge 0$. The (normalized) RMS better characterizes the the bulk
@@
         (skewness) for each month, etc.  As the reviewer mentions, the collapsed
-        PDFs for vorticity and divergence are shown in Figure S2, which is a plot
-        that compares April/October and hourly/daily-averaged in a single plot.\\
+        PDFs for vorticity and divergence are shown in Figure S2, which compares
+        April/October and hourly/daily-averaged in a single plot.\\
 
@@
       \bdp
-      We labelled the y-axis of the SSH variance spectrum as \textit{SSH variance density}
+      We label the y-axis of the SSH variance spectrum as \textit{SSH variance density}
       --- short for \textit{SSH variance spectral density} --- not \textit{SSH variance} as the reviewer
@@
       We further disagree that the y-axis should be labelled \textit{power spectral density of SSH}
-      or \textit{SSH spectra} [{\it sic}]. On physical grounds, \textit{power} adds
+      or \textit{SSH spectra}. On physical grounds, \textit{power} adds
       unnecessary ambiguity,
@@
       And ``SSH [wavenumber] spectrum'' is the plot of SSH variance spectral density
-      as a function of wavenumber, not the quantity in the y-axis.\footnote{This confusion is similar
-      to the pervasive axis poor-labelling in plots of probability distribution function (PDF)
-      --- the best label is \textit{probability density} not PDF.}
+      as a function of wavenumber, not the quantity in the y-axis.
       \edp

### 2016-09-18 writeup/rocha_etal.tex
Minor edits by Sarah and Teri.

@@
 challenge for SWOT is the presence of ageostrophic flows that straddle the submesoscales and project
-onto SSH \citep[e.g., ][]{richman_etal2012} --- the
-nuisance\footnote{``One man's noise is
-another man's signal'' (Walter Munk).} may upset the estimation
-of submesoscale surface geostrophic velocities.
+onto SSH \citep[e.g., ][]{richman_etal2012} thus making submesoscale geostrophic
+velocities difficult to extract from SSH.
 
@@
 effective resolution $\approx$8 km),   are
-forward Massachusetts Institute of Technology general circulation model \citep[MITgcm; ][]{marshall_etal1997}
-numerical solutions on a LLC grid  \citep{forget_etal2015} with
+forward solution of the Massachussetts Institute of Technology general circulation model \citep[MITgcm; ][]{marshall_etal1997}
+on a LLC grid  \citep{forget_etal2015} with
 90 vertical levels. The coarser-resolution LLC simulation was
@@
 a stronger enhancement of
-lateral divergence in late summer/early fall (RMS divergence $\sim0.22 f$ the 1/48$^\circ$
+lateral divergence in late summer/early fall (RMS divergence $\sim0.22 f$ in the 1/48$^\circ$
 simulation). Because the 1/48$^\circ$ simulation better resolves
@@
  the vorticity skewness reduces from 1.4 to 1.13 from hourly to daily-averaged.
- The April results are characteristic of winter, indicating that wintertime submesoscale
+ The April results indicate that wintertime submesoscale
  surface velocity is strongly dominated by submesoscale turbulence.
- The hourly velocity and sea-surface heigh fields have an important
+ The hourly velocity and sea-surface height fields have an important
 ageostrophic component as depicted by the joint-PDF of vorticity-Laplacian of SSH
@@
 the daily-averaged flow is essentially geostrophic as depicted by the 45$^\circ$-tilted
-ellipse in the vorticity-Laplacian of sea-surface height joint-PDF.
+ellipse in the vorticity-Laplacian of sea-surface height joint-PDF (Figure \ref{fig3}h).
 
@@
  the vorticity is strongly positively skewed, whereas the divergence is moderately
- negatively skewed (convergence/downwelling) as predicted by frontogenesis \citep[e.g., ][]{mcwilliams2016}
-. In late summer/early fall, the divergence is stronger, but PDFs are much less skewed,
+ negatively skewed (convergence/downwelling) as predicted by frontogenesis \citep[e.g., ][]{mcwilliams2016}.
+ In late summer/early fall, the divergence is stronger, but PDFs are much less skewed,
 consistent with linear IGWs.
@@
 To confirm that submesoscale IGWs account for most of the super-inertial SSH variance
-and surface KE, we calculate wavenumber-frequency spectrum of SSH variance. Focusing on
+and surface KE, we calculate the wavenumber-frequency spectrum of SSH variance. Focusing on
 the high-frequency
@@
 super-inertial SSH variance at submesoscales is larger in October. At scales smaller
-than 100 km (submesoscales), there the SSH variance is 66$\%$ larger in October than in April.
+than 100 km (submesoscales), the SSH variance is 66$\%$ larger in October than in April.
 IGWs, particularly modes two through four, account for most of the SSH variance
@@
 Similarly, the submesoscale super-inertial surface KE is 63$\%$ larger in October
-than the April (supporting information).
+than in April (supporting information).
 
@@
 Figure \ref{fig4}c shows the integral over frequency
-of wavenumber-frequency spectra in figures \ref{fig4}a-b. ,
+of wavenumber-frequency spectra in figures \ref{fig4}a-b.
 Both in April and October, the
@@
 sub-harmonic instability \citep[e.g., ][]{mackinnon_winters2005}.
-In the LLC simultions,
+In the LLC simulations,
 super-inertial IGWs account for most of the surface submesoscale KE and SSH variance;
@@
 
-Internal tide energy flux does not show a seasonal modulation \citep[e.g.,][]{alford2003},
+Internal tide energy flux does not show a seasonal modulation \citep[e.g.,][]{alford2003};
 hence one expects a strong seasonality in the near-surface
@@
 higher frequency IGWs if the near-surface stratification varies strongly.
-Windy near-inertial wave (NIW) generation is bursty and peaks in early winter, but
-NIW energy is similar energy level in October and April
+Windy near-inertial wave (NIW) generation is intermittent and peaks in early winter, but
+NIW energy has a similar energy level in October and April
 \citep{alford_etal2016}. Furthermore, in our simulations the higher frequency inertia-gravity-wave
@@
 Supercomputing (NAS) Division for their awesome support.
-This research was funded by NSF (OCE1357047) and NASA (NNX13AE44G,NNX13AE85G,NNX16AH67G).
+This research was funded by NSF (OCE1357047) and NASA (NNX13AE44G,NNX13AE85G,NNX16AH67G,NNX16AO5OH).
 The LLC output can be obtained from the ECCO project (\texttt{http://ecco2.org/llc$\_$hires}). The altimeter products were produced by Ssalto/Duacs

### 2016-09-18 writeup/rocha_etal_supporting-info.tex
Minor edits by Sarah and Teri.

@@
 
-Table \ref{tab:llc} shows details of the LLC spin-up hierarchy.
+Table 1 shows details of the LLC spin-up hierarchy.
 Both simulations analyzed in the present paper, LLC2160 and LLC4320,
@@
 we compare the model near-surface velocity field against two available
-mooring data: KEO (the data is available online at \texttt{http://www.pmel.noaa.gov/ocs/KEO})
-and KESS mooring $\#$7 (the data is available online at \texttt{http://uskess.whoi.edu}).
+mooring data records: KEO (the data are available online at \texttt{http://www.pmel.noaa.gov/ocs/KEO})
+and KESS mooring $\#$7 (the data are available online at \texttt{http://uskess.whoi.edu}).
 For proper comparison, we use model data closest to these two moorings (west of

### 2016-09-19 writeup/rocha_etal.tex
Adds paragraph to the discussion

@@
 account for most of the seasonality in super-inertial surface KE and SSH
-(Figure \ref{fig4}a-b),
+(Figure \ref{fig4}a-b).
+
+The seasonally varying MLD challenges observation and simulation of
+flows at ML-deformation-radius length scale in observational programs and models with
+fixed resolution \citep{buckingham_etal2016}. But the potential energy stored in the
+ML is lowest in summer owing to shallow MLD and weaker lateral buoyancy
+gradients \citep{callies_etal2015}. Moreover, in baroclinic turbulence, the deformation
+radius has a secondary, catalytic hole --- most of the baroclinic convertion occurs at
+larger scales \citep{larichev_held1995,callies_etal2016}. 
+Thus, the model failure to resolve the summertime
+ML deformation radius is unlikely to account for the vigorous seasonality in submesoscale
+turbulence.
 
@@
 seasonality in a future study.
-
-%The effects of smaller-scale/higher-frequency
-%``sub-submesoscale'' flows on the submesoscale surface velocity and SSH variability
-%are presently unknown.
-
-% That the  surface velocity and SSH submesoscale variability may be
-% dominated by ageostrophic flows in
-% summer/fall has consequences for the interpretation of data from
-% the future SWOT and COMPIRA altimeter missions,
-% which will deliver SSH measurements at submesoscales. To the extent that
-% high-frequency flows are dominated by spatially incoherent internal tides and other
-% internal waves, it may be very difficult (if not impossible) to
-% separate SSH submesoscale variability associated with geostrophic motions
-% from high-frequency, ageostrophic flows. Thus, previous claims that
-% one will be able to easily obtain submesoscale surface
-% geostrophic velocities and monitor such seasonal cycle on global scales
-% \citep{sasaki_etal2014,qiu_etal2014} are overstated.
-
-% Here we argue that those divergent motions are likely IGWs.
-% We can show a couple of spectra and argue that they roughly satisfy polarization
-% relations.
-%
-% We can also try and show some observations here. Any cruises with ADCP data.
-% While it may be hard to have enough data for the errorbars to be small, a
-% rough consistency may be better than nothing.
-%
-% Perhaps a mooring data should show strong seasonal modulation at high-frequencies?
-%
-% We must be able to present some observational evidence that the model is not
-% misleading at high-frequencies.
-
-% Here we argue that high-frequency (supra-daily) flows significantly project
-% on the sea-surface, and thus estimation of submesoscale (10-100 km) geostrophic
-% velocity from sea-surface height (SSH) is not warranted. Calculating spectral
-% KE fluxes from both velocity and geostrophic velocity (from SSH) would
-% contrast with the results in Sasaki et al.
-
-% \section{Projection on time scales}
-% To better understand the high-frequency/high-wavenumber flows  we calculate frequency
-% spectra of surface KE winter \ref{fig5}). To obtain statistically meaningful results,
-% we average spectral estimates along a section at 165$^\circ$E.  There is about 50
-% independent spectral realizations.
-%
-% The surface KE spectra of subinertial flows very roughly follow a $-2$ power-law.
-% The high-frequency flows are split near-inertial flows, tidal signals and their
-% higher harmonics. Hence these high-frequency flows are dominated by inertia-gravity
-% waves. (Note that the inertial peak nearly merges with the
-% diurnal tide peak.) There is statistically significant seasonality. The seasonal
-% cycle of the sub-inertial motions (at least for periods between 5-20 days)
-% is out-of-phase with the seasonal cycle of supra-inertial flows: the former flows
-% are more energetic in April and the latter in October.
-%
-% Most of these high-frequency IGWs project onto scales smaller
-% than 100 km. The surface KE spectra of 100-km-smoothed velocity field is significantly
-% suppressed in supra-inertial frequencies --- there are no statistically significant
-% changes at sub-inertial frequencies (see dashed lines in figure \ref{fig5}).
-
-% say something about relative energy in smoother vs.\\ unsmoothed fields.
-% also say something about comparison with observations.
+hing.
 

### 2016-09-19 writeup/RochatEtal_ResponseToReviewers.tex
Fixes typo.

@@
 In this revised manuscript, we have addressed the reviewers'
-comments. The most fundamental change is the inclusion of a wavenumber-frequency
+comments. Fundamentally, we have included a new wavenumber-frequency
 spectral analysis suggested by reviewer 3. The new results vindicate
@@
 to place our study in the context of Surface Water \& Ocean Topography (SWOT)
-satellite mission --- we discuss
-the implications of our findings for the estimation of surface velocity from
-high-resolution sea-surface height.
+satellite mission.
 
@@
             minor comment \#1) but is consistent with the definition used by
-            (at least part of) the SWOT community (Fu\& Ferrari 2008). Giving the
+            (at least part of) the SWOT community \citep{fu_ferrari2008}. Giving the
             definition, which we present in the first sentence of the introduction,
@@
 
-            \textcolor{purple}{Is it worth to add a paragraph to discuss this?}
+            We have added a paragraph discussing these issues (lines xx-yy).
 
@@
           fitted an analytic model to (super-inertial) frequency spectra and
-          different vertical modes (see Walter Munk's chapter
-          in \textit{Evolution of Physical Oceanography} 1981 and references therein).
+          different vertical modes \citep[see][and references therein]{munk1981}.
           One could argue

### 2016-09-19 writeup/rocha_etal.tex
Fixes typo.

@@
 \begin{abstract}
-    Two new high-resolution numerical simulations with embedded tides show a
-    strong modulation of near-surface dynamics at submesoscales
+    We present model-based evidence of
+    vigorous modulation of near-surface dynamics at submesoscales
     (10-100 km) in the Kuroshio Extension. Consistent with recent studies, deep late-winter mixed layers
     are prone to baroclinic instabilities, and submesoscale turbulence
-    prevails in late winter/early spring. While summertime
-    re-stratification weakens submesoscale turbulence, it also enhances submesoscale inertia-gravity
+    prevails in late winter/early spring. Summertime
+    re-stratification weakens submesoscale turbulence and enhances submesoscale inertia-gravity
      waves near the surface. In the Kuroshio Extension,
@@
 gradients \citep{callies_etal2015}. Moreover, in baroclinic turbulence, the deformation
-radius has a secondary, catalytic hole --- most of the baroclinic convertion occurs at
+radius has a secondary, catalytic role --- most of the baroclinic convertion occurs at
 larger scales \citep{larichev_held1995,callies_etal2016}.

### 2016-09-21 writeup/RochatEtal_ResponseToReviewers.tex
Minor changes
Small comments and corrections

@@
   \bdp
-      The mesoscale eddy KE level compares well to gridded AVISO
-      geostrophic KE. \\\textcolor{purple}{Need to do the comparison with alon-track data}.
+      The model smoothed fields have EKE levels similar to across-track (JASON II)
+      geostrophic EKE and gridded geostrophic EKE. We included this stantement with
+      the area-averaged EKE. Thanks for pointing this out.
   \edp
@@
         \bdp
-            Our scheme likely slightly underestimates the vorticity. \textcolor{purple}{
-            I will perform a sensitivity analysis.}
+            Wider stencils are more accurate because their order of accuracy is
+            higher. To asses the error of centered second-order scheme we periodized
+            our domain
         \edp

### 2016-09-21 writeup/rocha_etal.tex
Small comments and corrections

@@
 \begin{abstract}
-    We present model-based evidence of
+    We provide model-based evidence of
     vigorous modulation of near-surface dynamics at submesoscales
@@
     the accuracy of surface geostrophic currents diagnosed from SSH measured by
-    high-resolution
-    satellite altimeters.
+    high-resolution satellite altimeters.
 \end{abstract}
@@
 a comparison with Argo climatology shows that both simulations skillfully represent the Kuroshio
-Extension stratification and its seasonal variability  (supporting information).
+Extension stratification and its seasonal variability  (supporting information). Model smoothed
+fields have eddy kinetic energy (EKE) levels ($0.041$ m$^2$ s$^{-2}$) similar to Jason II across-track
+geostrophic EKE ($0.036$ m$^2$ s$^{-2}$) and gridded geostrophic EKE ($0.032$ m$^2$ s$^{-2}$).
 

### 2016-09-21 writeup/RochatEtal_ResponseToReviewers.tex
Minor changes

@@
 
-        \textcolor{purple}{Dimitris --- do you want to pitch in a reply for this question?}
+        The model resolves IGWs well, particularly those waves with periods larger
+        than 1h considered in this paper --- we are restricted to the output
+        frequency. Higher-frequency waves that project onto lateral scales smaller
+        than 10 km are potentially inaccurate but strongly damped in our simulations.
 
@@
       \bdp
-        Adding the elementary motivation above is yet another controversy: some
-        readers may appreciate it but others may find it condescending.
-        We argue that most physical oceanographers  will likely
-        likely know this material. Even readers from other areas, who did not take
-        basic fluid mechanics, will have the intuition that if the flow peaks
-        at mesoscales (energy-containing), then second-order statistics of the
-        velocity gradient will better highlight the submesoscales than the
-        KE (second-order statistic of the velocity).
+        We included a brief description/motivation for the use of second-order statistics
+        of those quantities. Thanks for pointing this out.
       \edp

### 2016-09-21 writeup/rocha_etal.tex
Minor changes

@@
 \end{equation}
-These diagnostics highlight the submesoscale structures in the flow
-\citep[e.g.,][]{shcherbina_etal2013}.
+Because vortices, fronts, and IGWs are ubiquitous at submesoscales, second-order
+statistics of vorticity, strain, and divergence are apt descriptors of these
+fine lateral scales \citep[e.g.,][]{shcherbina_etal2013}.
 

### 2016-09-21 writeup/rocha_etal.tex
Small clarification

@@
 a comparison with Argo climatology shows that both simulations skillfully represent the Kuroshio
-Extension stratification and its seasonal variability  (supporting information). Model smoothed
+Extension stratification and its seasonal variability  (supporting information). Moreover, model smoothed
 fields have eddy kinetic energy (EKE) levels ($0.041$ m$^2$ s$^{-2}$) similar to Jason II across-track

### 2016-09-22 writeup/rocha_etal.tex
Adds small comment about spectral differentiation

@@
 \begin{abstract}
-    We provide model-based evidence of
+    We provide evidence  of
     vigorous modulation of near-surface dynamics at submesoscales
-    (10-100 km) in the Kuroshio Extension. Consistent with recent studies, deep late-winter mixed layers
-    are prone to baroclinic instabilities, and submesoscale turbulence
+    (10-100 km) in the Kuroshio Extension. The analysis uses outputs of general
+    circulation models spun-up from a state estimate and forced by high-frequency
+    winds, surface fluxes, and tides.
+    Consistent with recent studies, deep late-winter mixed layers
+    are prone to shallow baroclinic instabilities and submesoscale turbulence
     prevails in late winter/early spring. Summertime
-    re-stratification weakens submesoscale turbulence and enhances submesoscale inertia-gravity
+    re-stratification, however, weakens submesoscale turbulence and enhances submesoscale inertia-gravity
      waves near the surface. In the Kuroshio Extension,
-    inertia-gravity waves strongly dominate the submesoscale surface kinetic energy and
+    inertia-gravity waves dominate the submesoscale surface kinetic energy and
     sea-surface height (SSH) variance in late summer/early fall, with implications for
@@
 % now SWOT
-The characterization of the partition between geostrophic and ageostrophic flows across
-the submesoscales and their seasonality has immediate applications for the planning
+The partition between geostrophic and ageostrophic flows across
+the submesoscales and their seasonality have immediate applications for the planning
 of the Surface and Water Ocean Topography (SWOT) satellite mission.
@@
 first global SSH measurements at submesoscales \citep{fu_ubelmann2014}. A
-challenge for SWOT is the presence of ageostrophic flows that straddle the submesoscales and project
+challenge --- and opportunity ---  for SWOT is the presence of ageostrophic flows
+that straddle the submesoscales and project
 onto SSH \citep[e.g., ][]{richman_etal2012} thus making submesoscale geostrophic
@@
 using a centered
-second-order finite difference scheme. We then diagnose
+second-order finite difference scheme --- we obtain qualitatively similar results
+using spectral differentiation of velocity fields periodized with reflections;
+statistics obtained via spectral differentiation are 10-15$\%$ larger.
+We then diagnose
 the vertical vorticity

### 2016-09-22 writeup/RochatEtal_ResponseToReviewers.tex
Significant changes to the abstract

@@
         \bdp
-            \textcolor{purple}{Ask Bill --- I haven't found any reference that mentions this
-            fact explicitly, but it follows from the linear boussinesq equations, which can
-            be combined to yield $\zeta_t= f_0 \delta$.}
+            We added a citation to \cite{garrett_munk1972}.
         \edp
@@
         \bdp
-            We added ``convergence/downwelling'' to the text.
+            We added ``convergence/downwelling'' to the text (line xx).
         \edp
@@
         \bdp
-            Wider stencils are more accurate because their order of accuracy is
-            higher. To asses the error of centered second-order scheme we periodized
-            our domain
+            To asses the error of centered second-order scheme we periodized
+            our domain with reflections. Statistics of lateral velocity gradients
+            obtained via spectral differentiation are 10-15$\%$ larger than the ones
+            obtained using second-order finite differences. But there are no qualitative
+            changes to Figure 2 and no changes in our conclusions. For simplicity, we
+            opt to use finite differences. We now report this estimate of accuracy of
+            our finite difference scheme (lines xx-xx).
         \edp

### 2016-09-22 writeup/rocha_etal.tex
Significant changes to the abstract

@@
 \begin{abstract}
-    We provide evidence  of
-    vigorous modulation of near-surface dynamics at submesoscales
-    (10-100 km) in the Kuroshio Extension. The analysis uses outputs of general
-    circulation models spun-up from a state estimate and forced by high-frequency
-    winds, surface fluxes, and tides.
-    Consistent with recent studies, deep late-winter mixed layers
-    are prone to shallow baroclinic instabilities and submesoscale turbulence
-    prevails in late winter/early spring. Summertime
-    re-stratification, however, weakens submesoscale turbulence and enhances submesoscale inertia-gravity
-     waves near the surface. In the Kuroshio Extension,
-    inertia-gravity waves dominate the submesoscale surface kinetic energy and
-    sea-surface height (SSH) variance in late summer/early fall, with implications for
-    the accuracy of surface geostrophic currents diagnosed from SSH measured by
-    high-resolution satellite altimeters.
+    Submesoscale turbulence, fueled by ample potential
+    energy stored in wintertime deep mixed layers, undergoes vigorous seasonality.
+    Here we provide model evidence that the seasonally-changing upper-ocean
+    stratification also modulates submesoscale --- here 10-100 km ---
+    inertia-gravity waves. Summertime re-stratification weakens submesoscale
+    turbulence but enhances inertia-gravity waves near the surface. Thus,
+    submesoscale turbulence and inertia-gravity waves undergo vigorous out-of-phase
+    seasonal cycles. The present results imply a strong seasonal modulation of
+    the skill and accuracy of geostrophic velocity diagnostics from submesoscale
+    sea-surface height delivered by the Surface and Water Ocean Topography (SWOT)
+    satellite mission
 \end{abstract}
@@
  In late summer/early fall, the divergence is stronger, but PDFs are much less skewed,
-consistent with linear IGWs.
+consistent with a superposition of random linear IGWs \citep[e.g.,][]{garrett_munk1972}.
 

### 2016-09-22 writeup/RochatEtal_ResponseToReviewers.tex
Minor changes

@@
 account for most of the summertime SSH variance and surface KE at scales
-between 10 and 100 km. We have also expanded the introductions and discussion
+between 10 and 100 km. We have also expanded the introduction and discussion
 to place our study in the context of Surface Water \& Ocean Topography (SWOT)
@@
           expression of IGWs, which is very sensitive to the near-surface stratification
-          (D'saro 1978; cf. second paragraph of discussion).\\
+          \citep[; cf. second paragraph of discussion]{dasaro1978}.\\
 
@@
 
-            \cite{buckingham_etal2016} made a very good point that models and observations
+            \cite{buckingham_etal2016} made a good point that models and observations
             to date may not resolve the mixed-layer deformation scale in summer; this is no
@@
 
-        Thanks for pointing that out.
+        We now reference to the recent paper byt Brannigan et al. (2015).
+        Ostrovskii (1995) can accessed via Buckingham et al. (2016).
 
@@
         \bdp
-            We added a citation to \cite{garrett_munk1972}.
+        Added.
         \edp

### 2016-09-22 writeup/rocha_etal.tex
Minor changes

@@
 \begin{abstract}
-    Submesoscale turbulence, fueled by ample potential
-    energy stored in wintertime deep mixed layers, undergoes vigorous seasonality.
+    The vigorous seasonal cycle of mixed layer modulates upper-ocean submesoscale turbulence.
     Here we provide model evidence that the seasonally-changing upper-ocean
-    stratification also modulates submesoscale --- here 10-100 km ---
+    stratification also modulates submesoscale  (here 10-100 km)
     inertia-gravity waves. Summertime re-stratification weakens submesoscale
@@
     submesoscale turbulence and inertia-gravity waves undergo vigorous out-of-phase
-    seasonal cycles. The present results imply a strong seasonal modulation of
-    the skill and accuracy of geostrophic velocity diagnostics from submesoscale
+    seasonal cycles. These results imply a strong seasonal modulation of
+    the skill and accuracy of geostrophic velocity diagnosed from submesoscale
     sea-surface height delivered by the Surface and Water Ocean Topography (SWOT)
@@
 of the Surface and Water Ocean Topography (SWOT) satellite mission.
-SWOT will provide
+SWOT will deliver
 sea-surface height (SSH) with about 15 km resolution, thereby providing the
-first global SSH measurements at submesoscales \citep{fu_ubelmann2014}. A
-challenge --- and opportunity ---  for SWOT is the presence of ageostrophic flows
-that straddle the submesoscales and project
-onto SSH \citep[e.g., ][]{richman_etal2012} thus making submesoscale geostrophic
-velocities difficult to extract from SSH.
-
-Using the output of $1/24^\circ$ and $1/48^\circ$ global
-numerical simulations with embedded tides, we show that IGWs undergo
-a strong seasonality near the surface in the Kuroshio Extension region.
+first global SSH measurements at submesoscales \citep{fu_ubelmann2014}.
+The projection of energetic submesoscale ageostrophic flows, such as
+inertia-gravity waves (IGWs), onto SSH represents
+a challenged --- and an opportunity --- for SWOT. The challenge: how to diagnose
+surface velocity from SSH. The opportunity: observations of submesoscale ageostrophic
+flows on global scale.
+
+Using the output of two comprehensive global numerical simulations, here we show
+that IGWs undergo
+a strong near-surface seasonal cycle in the Kuroshio Extension region.
 Interestingly, the seasonal cycle of IGWs is out of phase
 with the seasonal cycle of submesoscale turbulence. Consistent with previous studies,
-deep late-winter mixed layers are prone to
-shallow baroclinic instabilities that are roughly in geostrophic balance
-and flux energy upscale \citep{sasaki_etal2014,callies_etal2016},
-driving a mild seasonal modulation of the mesoscales \citep{sasaki_etal2014,qiu_etal2014}.
+submesoscale turbulence is strongest  in winter \citep{sasaki_etal2014,qiu_etal2014}.
 
@@
 whereas submesoscale IGWs prevail in late summer/early fall.
-Because submesoscale turbulence is weakest in late summer/early fall,
-IGWs account for most of the
-summertime submesoscale sea-surface height variability with implications for SWOT.
 
@@
 We use results of two latitude-longitude polar cap (LLC)
-realistic numerical simulations. The outputs
+comprehensive numerical simulations. The outputs
 analyzed here, LLC2160 (nominal resolution 1/24$^\circ$$\approx$4.7 km;
@@
 adjoint-method state estimate, constrained to millions
-of observations from 2009 through 2011. Both simulations were forced by
-tides hourly and by 6-hourly surface atmospheric fields. The LLC2160
+of observations from 2009 through 2011.
+
+The LLC2160
 output spans two years from March 2011 to April 2013; the LLC4320 was spun up from
 the LLC2160 simulation, spanning one year from September 2011 to October 2012.
+Both LLC2160 and LLC4320 simulations were forced by
+tides hourly and by 6-hourly surface atmospheric fields.
 The LLC4320 simulation is an extension of the 3-month long output used by
@@
 
-A key aspect of the LLC2160 and LLC4320 simulations is that they were forced by
-the 16 most-significant tidal components.
-Because barotropic tides interact with topography and generate internal
-tides that project onto mesoscales to submesoscales
-\citep[e.g., ][]{rocha_etal2016}, tidal forcing fundamentally distinguishes our analysis
+Tidal forcing by the 16 most-significant components represents a key feature LLC2160 and
+LLC4320 simulations. Barotropic tides flowing over topography convert significant tidal energy into the internal
+wave field. Those internal waves project onto mesoscales to submesoscales
+\citep[e.g., ][]{rocha_etal2016}. Hence tidal forcing fundamentally distinguishes our analysis
 from other modeling studies of seasonality \citep{sasaki_etal2014,qiu_etal2014}.
-
-To study seasonal variations in the upper-ocean dynamics, we focus on the northwest
-Pacific, in the vicinity of the Kuroshio
-Extension, where previous studies have shown strong mesoscale and submesoscale seasonality
-\citep{sasaki_etal2014,qiu_etal2014}.
-We analyze a sub-domain of the LLC4320 and LLC2160 simulations of about 2000 km$^2$
-spanning 155-175$^\circ$E; 25-40$^\circ$N (Figure \ref{fig1}a). The stratification
-in this mesoscale-rich subtropical region undergoes a vigorous seasonal cycle: wintertime-enhanced
-small-scale turbulence de-stratifies the upper ocean, yielding mixed layers
-deeper than 200 m to the north (Figure \ref{fig1}b). Summertime re-stratification yields mixed layers shallower
-than 40 m (Figure \ref{fig1}c).
-Fundamentally, the upper-ocean density structure is well-captured by both LLC simulations:
-a comparison with Argo climatology shows that both simulations skillfully represent the Kuroshio
-Extension stratification and its seasonal variability  (supporting information). Moreover, model smoothed
-fields have eddy kinetic energy (EKE) levels ($0.041$ m$^2$ s$^{-2}$) similar to Jason II across-track
-geostrophic EKE ($0.036$ m$^2$ s$^{-2}$) and gridded geostrophic EKE ($0.032$ m$^2$ s$^{-2}$).
 
@@
 
+To study seasonal variations in the upper-ocean dynamics, we focus on the northwest
+Pacific, in the vicinity of the Kuroshio
+Extension, where previous studies have shown strong mesoscale and submesoscale seasonality
+\citep{sasaki_etal2014,qiu_etal2014}.
+We analyze a sub-domain of the LLC4320 and LLC2160 simulations of about 2000 km$^2$
+spanning 155-175$^\circ$E; 25-40$^\circ$N (Figure \ref{fig1}a). The stratification
+in this mesoscale-rich subtropical region undergoes a vigorous seasonal cycle: wintertime-enhanced
+small-scale turbulence de-stratifies the upper ocean, yielding mixed layers
+deeper than 200 m to the north (Figure \ref{fig1}b). Summertime re-stratification yields mixed layers shallower
+than 40 m (Figure \ref{fig1}c).
+
+Fundamentally, the upper-ocean density structure is well-captured by both LLC simulations:
+a comparison with Argo climatology shows that both simulations skillfully represent the Kuroshio
+Extension stratification and its seasonal variability  (supporting information). Moreover, model smoothed
+fields have eddy kinetic energy (EKE) levels ($0.041$ m$^2$ s$^{-2}$) similar to Jason II across-track
+geostrophic EKE ($0.036$ m$^2$ s$^{-2}$) and gridded geostrophic EKE ($0.032$ m$^2$ s$^{-2}$).
+
 \section{Bulk statistics of the surface lateral velocity gradient tensor}
@@
 using spectral differentiation of velocity fields periodized with reflections;
-statistics obtained via spectral differentiation are 10-15$\%$ larger.
+statistics of fields obtained via spectral differentiation are 10-15$\%$ larger.
 We then diagnose
@@
 Because vortices, fronts, and IGWs are ubiquitous at submesoscales, second-order
-statistics of vorticity, strain, and divergence are apt descriptors of these
+statistics of vorticity, strain, and divergence highlight these
 fine lateral scales \citep[e.g.,][]{shcherbina_etal2013}.
 
-Figures \ref{fig1}b-c show snapshots of vertical vorticity $\zeta$ in early spring
-(April 15) and fall (October 15) in the LLC4320 (1/48$^\circ$) simulation.
-The model solutions depict seasonality in vorticity: large values of
+The model solutions depict seasonality in vorticity (Figures \ref{fig1}b-c): large values of
 fine-grained vertical vorticity are observed in early spring with maximum
@@
 
-The bulk of vertical vorticity and strain rate are associated with subinertial flows
+Subinertial flows account for the bulk of vertical vorticity and strain rate
 ($T_{32.5} = 2\pi/f_{32.5}\approx 22.3$h, where $f_{32.5}$ is the inertial frequency at
@@
 cycle remains strong (see
-dashed lines in figures \ref{fig2}a-b). Indeed, most of this seasonal cycle is associated
-with submesoscale flows: smoothing the velocity fields
+dashed lines in figures \ref{fig2}a-b). Indeed, submesoscale flows account for
+most of this seasonal cycle: smoothing the velocity fields
 with a Hanning filter with cut off scale of 100 km dramatically reduces the RMS
@@
 The seasonal cycle of the horizontal divergence, however, showcases the complexity
-of the upper-ocean annual variability. If submesoscale eddies and fronts dominated
+of the upper-ocean annual variability. If submesoscale turbulence dominated
 the near-surface variability all year, then the seasonal cycle of horizontal divergence,
@@
   \begin{center}
-    \includegraphics[width=.75\textwidth]{figs/fig3.pdf}
+    \includegraphics[width=.8\textwidth]{figs/fig3.pdf}
  \caption{\small Seasonal variation of joint probability distributions of
@@
  surface velocity is strongly dominated by submesoscale turbulence.
- The hourly velocity and sea-surface height fields have an important
+ The daily-averaged fields
+ are largely in  geostrophic balance,
+but the hourly velocity and sea-surface height fields have an important
 ageostrophic component as depicted by the joint-PDF of vorticity-Laplacian of SSH
-winter (Figure \ref{fig3}e). The daily-averaged fields
-are largely
-in  geostrophic balance (Figure \ref{fig3}f).
+winter (Figure \ref{fig3}d-e).
 
@@
 
-Figures \ref{fig4}a-b show the wavenumber-frequency of SSH variance for the LLC4320
-simulation. Mesoscales flows and the lunar semi-diurnal tide contain most of the
-SSH variance. But there is significant SSH variance at submesoscales, both sub-inertial
+\begin{figure*}[ht]
+  \begin{center}
+    \includegraphics[width=.9\textwidth]{figs/fig4.pdf}
+ \caption{\small LLC4320 wavenumber-frequency spectrum of SSH variance in (a) April and (b) October.
+          (c) Wavenumber spectrum of SSH variance --- the integral of (a) and (b) over frequency.
+          In (a) and (b), the light gray shaded region depicts the dispersion relations for inertia-gravity
+          waves from mode 1 through mode 4 across the latitudinal domain; horizontal dashed lines represent the semi-diurnal
+          lunar tidal frequency ($M_2$), and the inertial frequency at mid-latitude
+          ($f_{32.5}$) and its first harmonic (2$f_{32.5}$). }
+  \label{fig4}
+ \end{center}
+\end{figure*}
+
+Mesoscales flows and the lunar semi-diurnal tide contain most of the
+SSH variance (Figures \ref{fig4}a-b). But there is significant SSH variance at
+submesoscales, both sub-inertial
 and super-inertial. Sub-inertial SSH variance at submesoscales is larger in April, whereas
@@
  sub-inertial frequencies is dramatic in October when super-inertial flows account for
- about 80\% of the SSH variance and KE at submesoscales. We obtain similar results by computing
+ about 80\% of the SSH variance and KE at submesoscales --- we obtain similar results by computing
 SSH variance and KE wavenumber spectra directly from hourly and daily-averaged
 SSH velocity fields.
-
-\begin{figure*}[ht]
-  \begin{center}
-    \includegraphics[width=.8\textwidth]{figs/fig4.pdf}
- \caption{\small LLC4320 wavenumber-frequency spectrum of SSH variance in (a) April and (b) October.
-          (c) Wavenumber spectrum of SSH variance --- the integral of (a) and (b) over frequency.
-          In (a) and (b), the light gray shaded region depicts the dispersion relations for inertia-gravity
-          waves from mode 1 through mode 4 across the latitudinal domain; horizontal dashed lines represent the semi-diurnal
-          lunar tidal frequency ($M_2$), and the inertial frequency at mid-latitude
-          ($f_{32.5}$) and its first harmonic (2$f_{32.5}$). }
-  \label{fig4}
- \end{center}
-\end{figure*}
 
@@
 \section{Discussion and conclusion}
-Our work adds to recent modeling \citep[e.g., ][]{sasaki_etal2014}
+Our work adds to recent model \citep[e.g., ][]{sasaki_etal2014}
 and observational \citep[e.g., ][]{callies_etal2015,buckingham_etal2016} evidence of vigorous seasonality in
-submesoscale turbulence. In particular, our main finding is that in global simulations with embedded tides the
-near-surface submesoscale IGWs in the Kuroshio Extension undergo
+submesoscale turbulence. In particular ---
+in global simulations with embedded tides --- near-surface submesoscale IGWs undergo
 a strong seasonal cycle that is out of phase with the seasonal cycle of
 submesoscale turbulence. IGW energy and SSH variance are significantly spread across
-the submesoscales, likely through mesoscale refraction and dispersion \citep[e.g.,][and references therein]{ponte_klein2015,alford_etal2016}
+the submesoscales, likely through mesoscale advection and refraction, and
+dispersion \citep[e.g.,][and references therein]{ponte_klein2015,alford_etal2016}
  and parametric
@@
 higher frequency IGWs if the near-surface stratification varies strongly.
-Windy near-inertial wave (NIW) generation is intermittent and peaks in early winter, but
+The intermittent windy near-inertial wave (NIW) generation peaks in early winter, but
 NIW energy has a similar energy level in October and April
@@
 
-The seasonally varying MLD challenges observation and simulation of
-flows at ML-deformation-radius length scale in observational programs and models with
-fixed resolution \citep{buckingham_etal2016}. But the potential energy stored in the
-ML is lowest in summer owing to shallow MLD and weaker lateral buoyancy
+The seasonally varying mixed-layer depth challenges observation and simulation of
+flows at mixed-layer deformation-radius length scale in observational programs and models with
+fixed resolution \citep{buckingham_etal2016}. But the mixed layer available potential energy
+is lowest in summer owing to shallow mixed-layer depths and weak lateral buoyancy
 gradients \citep{callies_etal2015}. Moreover, in baroclinic turbulence, the deformation
-radius has a secondary, catalytic role --- most of the baroclinic convertion occurs at
+radius has only a secondary, catalytic role --- most of the baroclinic convertion occurs at
 larger scales \citep{larichev_held1995,callies_etal2016}.
@@
 
-In the context of SWOT, our results suggest that if one is interested
+In the context of SWOT, these results imply that if one is interested
 only in the geostrophic flow, so that surface velocity can be diagnosed
@@
  William R. Young and the reviewers --- Christian Buckingham and two anonymous referees ---
-  provided exceptionally helpful and constructive comments. Greg Wagner
+  provided helpful and constructive comments. Greg Wagner
  suggested that the near-surface shape of the baroclinic modes changes seasonally.

### 2016-09-23 writeup/rocha_etal.tex
Fix typo

@@
 \begin{abstract}
-    The vigorous seasonal cycle of mixed layer modulates upper-ocean submesoscale turbulence.
+    The vigorous seasonal cycle of the mixed layer modulates upper-ocean submesoscale turbulence.
     Here we provide model evidence that the seasonally-changing upper-ocean
-    stratification also modulates submesoscale  (here 10-100 km)
+    stratification in the Kuroshio Extension also modulates submesoscale  (here 10-100 km)
     inertia-gravity waves. Summertime re-stratification weakens submesoscale

### 2016-09-25 writeup/RochatEtal_ResponseToReviewers.tex
Fixes small typos

@@
     The convergence of meridians (the $\beta$-effect) accounts for the
-    small divergence depicted by the black line in Figure 2c.
+    small divergence depicted by the black line in Figure 2c. We now added
+    a brief explanation to the caption of figure 2.
   \edp
@@
             cite all the
-            aforementioned papers. We believe the new discussion together with Figure 4
-            and Figure S7 will help clear any ambiguity. Those figures unambiguously
+            aforementioned papers. The new discussion together with Figure 4
+            and Figure S7 should clear any ambiguity. Those figures unambiguously
             show that most of the IGWs with horizontal scales between 10-100 km
@@
           our definition: (i) this is the definition used by the SWOT community (Fu \& Ferrari
-              2008); incidentally current altimeters do not resolve
+              2008); incidentally current Ku-band altimeters do not resolve
               scales smaller then about 100 km; and (ii) our goal is to determine which flows

### 2016-09-25 writeup/rocha_etal.tex
Fixes small typos

@@
     The vigorous seasonal cycle of the mixed layer modulates upper-ocean submesoscale turbulence.
-    Here we provide model evidence that the seasonally-changing upper-ocean
+    Here we provide model-based evidence that the seasonally-changing upper-ocean
     stratification in the Kuroshio Extension also modulates submesoscale  (here 10-100 km)
@@
     the skill and accuracy of geostrophic velocity diagnosed from submesoscale
-    sea-surface height delivered by the Surface and Water Ocean Topography (SWOT)
+    sea-surface height delivered by the Surface Water and Ocean Topography (SWOT)
     satellite mission
@@
 the submesoscales and their seasonality have immediate applications for the planning
-of the Surface and Water Ocean Topography (SWOT) satellite mission.
+of the Surface Water and Ocean Topography (SWOT) satellite mission.
 SWOT will deliver
@@
 inertia-gravity waves (IGWs), onto SSH represents
-a challenged --- and an opportunity --- for SWOT. The challenge: how to diagnose
+a challenge --- and an opportunity --- for SWOT. The challenge: how to diagnose
 surface velocity from SSH. The opportunity: observations of submesoscale ageostrophic
-flows on global scale.
+flows on a global scale.
 
@@
 effective resolution $\approx$8 km),   are
-forward solution of the Massachussetts Institute of Technology general circulation model \citep[MITgcm; ][]{marshall_etal1997}
+forward solutions of the Massachussetts Institute of Technology general circulation model \citep[MITgcm; ][]{marshall_etal1997}
 on a LLC grid  \citep{forget_etal2015} with
@@
 
-Tidal forcing by the 16 most-significant components represents a key feature LLC2160 and
+Tidal forcing by the 16 most-significant components represents a key feature of the LLC2160 and
 LLC4320 simulations. Barotropic tides flowing over topography convert significant tidal energy into the internal
@@
   \caption{\small Time series of the root-mean-square (RMS) of surface (a) vorticity,
-  (b) rate of strain, and (c) horizontal divergence in the LLC outputs and gridded AVISO data.}
+  (b) rate of strain, and (c) horizontal divergence in the LLC outputs and gridded AVISO data.
+  The convergence of meridians account for the small RMS divergence in AVISO line in (c).}
   \label{fig2}
@@
 ageostrophic component as depicted by the joint-PDF of vorticity-Laplacian of SSH
-winter (Figure \ref{fig3}d-e).
+winter (Figure \ref{fig3}e).
 
@@
           lunar tidal frequency ($M_2$), and the inertial frequency at mid-latitude
-          ($f_{32.5}$) and its first harmonic (2$f_{32.5}$). }
+          ($f_{32.5}$) and its first harmonic (2$f_{32.5}$). For reference, the gray curvish line represents
+          the SWOT baseline requirement $2\times 10^{-4} + 1.25\times 10^{-7} k^{-2}$ m$^2$/cpkm \citep{rodriguez2012}.
+          }
   \label{fig4}
@@
         estimated from the wavenumber-frequency spectrum of SSH variance in figure \ref{fig4}.
+        The SWOT baseline requirement at 15 km is approximately $\sqrt{2}$ cm.
         }
@@
 seasonal cycle of submesoscale turbulence and inertia-gravity waves. The sub-inertial flow
-has steeper submesoscale SSH variance wavenumber spectrum (dashed lines in figure \ref{fig4}c).
+has a steeper submesoscale SSH variance wavenumber spectrum (dashed lines in figure \ref{fig4}c).
  The difference between the SSH variance wavenumber spectrum for all frequencies and
@@
 NIW energy has a similar energy level in October and April
-\citep{alford_etal2016}. Furthermore, in our simulations the higher frequency inertia-gravity-wave
+\citep{alford_etal2016}. Furthermore, in our simulations the higher frequency inertia-gravity waves
 account for most of the seasonality in super-inertial surface KE and SSH
@@
 gradients \citep{callies_etal2015}. Moreover, in baroclinic turbulence, the deformation
-radius has only a secondary, catalytic role --- most of the baroclinic convertion occurs at
+radius has only a secondary, catalytic role --- most of the baroclinic conversion occurs at
 larger scales \citep{larichev_held1995,callies_etal2016}.
@@
 from SSH, then the noise-to-signal ratio will have a strong seasonality owing
-to the presence of IGWs --- the coherent fraction of internal tides may be removed
-efficiently.
+to the presence of IGWs (the coherent fraction of internal tides may be removed
+efficiently).
 An important caveat is that the Kuroshio Extension may be typical of a mesoscale-rich
@@
 seasonality in a future study.
-hing.
 
@@
  William R. Young and the reviewers --- Christian Buckingham and two anonymous referees ---
-  provided helpful and constructive comments. Greg Wagner
+  provided helpful and stimulating comments. Greg Wagner
  suggested that the near-surface shape of the baroclinic modes changes seasonally.

### 2016-09-25 writeup/RochatEtal_ResponseToReviewers.tex
Updates resply with line numbers

@@
       The model smoothed fields have EKE levels similar to across-track (JASON II)
-      geostrophic EKE and gridded geostrophic EKE. We included this stantement with
-      the area-averaged EKE. Thanks for pointing this out.
+      geostrophic EKE and gridded geostrophic EKE. We included this statement with
+      the area-averaged EKE (lines 78-80). Thanks for pointing this out.
   \edp
@@
           expression of IGWs, which is very sensitive to the near-surface stratification
-          \citep[; cf. second paragraph of discussion]{dasaro1978}.\\
+          \citep[; lines 191-196]{dasaro1978}.\\
 
@@
             cite all the
-            aforementioned papers. The new discussion together with Figure 4
+            aforementioned papers (lines 185-190). The new discussion together with Figure 4
             and Figure S7 should clear any ambiguity. Those figures unambiguously
@@
              with the Kuroshio upstream and near its separation from Japan's coast.
-             We have considered larger and smaller domains and found insiginificant
+             We have considered larger and smaller domains and found insignificant
              quantitative differences.\\
@@
         \bdp
-          We included the definition --- thanks for pointing that out.
+          We included the definition (line 41). thanks for pointing that out.
         \edp
@@
               We agree that the planning of SWOT is perhaps the most immediate
-              application of our results. We have added an introductory paragraph dedicated to SWOT (new
-              paragraph 2 of section 1). We have also added a discussion about the
+              application of our results. We have added an introductory paragraph dedicated to SWOT (
+              lines 27-35). We have also added a discussion about the
               implication for SWOT in section 6. Thanks for making us provide a better
@@
 
-            We have added a paragraph discussing these issues (lines xx-yy).
+            We have added a paragraph discussing these issues (lines 204-210).
 
@@
             spectrum of SSH (Figure 4) to estimate variance in particular frequency
-            bands.  We have included a table with the standard
+            bands.  We have included a table (table 1) with the standard
             deviation of SSH for super-inertial and sub-inertial flows at scales
@@
           close to the surface (most of our results concern surface currents and SSH,
-          and it does not account for tidal and inertial peaks. We believe that the
+          and it does not account for tidal and inertial peaks. The
           new plot of the wavenumber-frequency spectrum of the SSH
           variance (new Figure 4) --- following the reviewer's suggestion ---
-          will more clearly convey the idea that there is no horizontal scale
+          should more clearly convey the idea that there is no horizontal scale
           separation between submesoscale IGWs and submesoscale turbulence.\\
@@
             opt to use finite differences. We now report this estimate of accuracy of
-            our finite difference scheme (lines xx-xx).
+            our finite difference scheme (see text after equation 1).
         \edp
@@
         We included a brief description/motivation for the use of second-order statistics
-        of those quantities. Thanks for pointing this out.
+        of those quantities (lines 81-83). Thanks for pointing this out.
       \edp

### 2016-09-25 writeup/rocha_etal.tex
Updates resply with line numbers

@@
     seasonal cycles. These results imply a strong seasonal modulation of
-    the skill and accuracy of geostrophic velocity diagnosed from submesoscale
+    the accuracy of geostrophic velocity diagnosed from submesoscale
     sea-surface height delivered by the Surface Water and Ocean Topography (SWOT)
@@
 first global SSH measurements at submesoscales \citep{fu_ubelmann2014}.
-The projection of energetic submesoscale ageostrophic flows, such as
+Thus, the projection of energetic submesoscale ageostrophic flows, such as
 inertia-gravity waves (IGWs), onto SSH represents

### 2016-09-25 writeup/rocha_etal_draft-like.tex
Adds draft version
Updates resply with line numbers

--- 2016-07-15 writeup/rocha_etal.tex
@@
 
-%\documentclass[draft,grl]{agutex2015}
-\documentclass[grl]{agutex2015}
+\documentclass[draft,grl]{agutex2015}
+%\documentclass[grl]{agutex2015}
 
@@
 
-% \usepackage{lineno}
-% \linenumbers*[1]
-%  To add line numbers to lines with equations:
+\usepackage{lineno}
+\linenumbers*[1]
+%To add line numbers to lines with equations:
 %  \begin{linenomath*}
@@
 
-\keypoints{\item  Upper-ocean submesoscale (10-100 km) turbulence and inertia-gravity waves
+\keypoints{\item  Upper-ocean submesoscale (here 10-100 km) turbulence and inertia-gravity
+                  waves
                   undergo strong seasonal cycles that are out of phase.
@@
 \begin{abstract}
-    Two new high-resolution numerical simulations with embedded tides show a
-    strong modulation of near-surface dynamics at submesoscales
-    (roughly 10-100 km) in the Kuroshio Extension. Consistent with recent studies, deep late-winter mixed layers
-    are prone to baroclinic instabilities, and submesoscale turbulence
-    prevails in late winter/early spring. While summertime
-    re-stratification weakens submesoscale turbulence, it also enhances submesoscale inertia-gravity
-     waves near the surface. In the Kuroshio Extension,
-    inertia-gravity waves strongly dominate the submesoscale surface kinetic energy and
-    sea-surface height variance in late summer/early fall.
+    The vigorous seasonal cycle of the mixed layer modulates upper-ocean submesoscale turbulence.
+    Here we provide model-based evidence that the seasonally-changing upper-ocean
+    stratification in the Kuroshio Extension also modulates submesoscale  (here 10-100 km)
+    inertia-gravity waves. Summertime re-stratification weakens submesoscale
+    turbulence but enhances inertia-gravity waves near the surface. Thus,
+    submesoscale turbulence and inertia-gravity waves undergo vigorous out-of-phase
+    seasonal cycles. These results imply a strong seasonal modulation of
+    the accuracy of geostrophic velocity diagnosed from submesoscale
+    sea-surface height delivered by the Surface Water and Ocean Topography (SWOT)
+    satellite mission
 \end{abstract}
@@
 
-Using the output of $1/24^\circ$ and $1/48^\circ$ global
-numerical simulations with embedded tides, we show that inertia-gravity waves undergo
-a strong seasonality near the surface in the Kuroshio Extension region.
-Interestingly, the seasonal cycle of inertia-gravity waves is out of phase
-with the seasonal cycle of submesoscale turbulence. On one hand, consistent with previous studies,
-deep late-winter mixed layers are prone to
-shallow baroclinic instabilities that are roughly in geostrophic balance
-and flux energy upscale \citep{sasaki_etal2014,callies_etal2016},
-driving a mild seasonal modulation of the mesoscales \citep{sasaki_etal2014,qiu_etal2014}.
-
-Inertia-gravity waves, on the other hand, peak in late summer/early fall,
+% now SWOT
+The partition between geostrophic and ageostrophic flows across
+the submesoscales and their seasonality have immediate applications for the planning
+of the Surface Water and Ocean Topography (SWOT) satellite mission.
+SWOT will deliver
+sea-surface height (SSH) with about 15 km resolution, thereby providing the
+first global SSH measurements at submesoscales \citep{fu_ubelmann2014}.
+Thus, the projection of energetic submesoscale ageostrophic flows, such as
+inertia-gravity waves (IGWs), onto SSH represents
+a challenge --- and an opportunity --- for SWOT. The challenge: how to diagnose
+surface velocity from SSH. The opportunity: observations of submesoscale ageostrophic
+flows on a global scale.
+
+Using the output of two comprehensive global numerical simulations, here we show
+that IGWs undergo
+a strong near-surface seasonal cycle in the Kuroshio Extension region.
+Interestingly, the seasonal cycle of IGWs is out of phase
+with the seasonal cycle of submesoscale turbulence. Consistent with previous studies,
+submesoscale turbulence is strongest  in winter \citep{sasaki_etal2014,qiu_etal2014}.
+
+IGWs with horizontal scales between 10-100 km (hereafter submesoscale IGWs),
+however, peak in late summer/early fall,
 when the upper ocean is strongly stratified. Thus there exists a
 strong seasonal modulation of upper-ocean submesoscale dynamics:
-submesoscale turbulence dominates the upper-ocean dynamics in late winter/early,
-whereas submesoscale inertia-gravity waves prevail in late summer/early fall.
-Because submesoscale turbulence is weakest in late summer/early fall,
-the present results indicate that inertia-gravity waves account for most of the
-summertime submesoscale sea-surface height variability.
+submesoscale turbulence dominates the upper-ocean dynamics in late winter/early fall,
+whereas submesoscale IGWs prevail in late summer/early fall.
 
@@
 We use results of two latitude-longitude polar cap (LLC)
-realistic numerical simulations. The outputs
-analyzed here, LLC2160 (1/24$^\circ$) and LLC4320 (1/48$^\circ$),   are
-forward Massachusetts Institute of Technology general circulation model \citep[MITgcm; ][]{marshall_etal1997}
-numerical solutions on a LLC grid  \citep{forget_etal2015} with
-90-vertical-levels. The coarser-resolution LLC simulation was
+comprehensive numerical simulations. The outputs
+analyzed here, LLC2160 (nominal resolution 1/24$^\circ$$\approx$4.7 km;
+effective resolution $\approx$ 20 km) and LLC4320
+(nominal resolution 1/48$^\circ$$\approx$2.3 km;
+effective resolution $\approx$8 km),   are
+forward solutions of the Massachussetts Institute of Technology general circulation model \citep[MITgcm; ][]{marshall_etal1997}
+on a LLC grid  \citep{forget_etal2015} with
+90 vertical levels. The coarser-resolution LLC simulation was
 spun up from an Estimating the Circulation and Climate of the Ocean, Phase II \citep[ECCO2; ][]{menemenlis_etal2008}
 adjoint-method state estimate, constrained to millions
-of observations from 2009 through 2011. Both simulations were forced by
-tides hourly and by 6-hourly surface atmospheric fields. The LLC2160
+of observations from 2009 through 2011.
+
+The LLC2160
 output spans two years from March 2011 to April 2013; the LLC4320 was spun up from
 the LLC2160 simulation, spanning one year from September 2011 to October 2012.
+Both LLC2160 and LLC4320 simulations were forced by
+tides hourly and by 6-hourly surface atmospheric fields.
 The LLC4320 simulation is an extension of the 3-month long output used by
-\citet{rocha_etal2016}.  Details of the LLC simulations are provided in the supplemental material.
-
-A key aspect of the LLC2160 and LLC4320 simulations is that they were forced by
-the 16 most-significant tidal components.
-Because barotropic tides interact with topography and generate internal
-tides that project onto mesoscales to submesoscales
-\citep[e.g., ][]{rocha_etal2016}, tidal forcing fundamentally distinguishes our analysis
+\citet{rocha_etal2016} --- for details, see supporting information.
+
+Tidal forcing by the 16 most-significant components represents a key feature of the LLC2160 and
+LLC4320 simulations. Barotropic tides flowing over topography convert significant tidal energy into the internal
+wave field. Those internal waves project onto mesoscales to submesoscales
+\citep[e.g., ][]{rocha_etal2016}. Hence tidal forcing fundamentally distinguishes our analysis
 from other modeling studies of seasonality \citep{sasaki_etal2014,qiu_etal2014}.
@@
 small-scale turbulence de-stratifies the upper ocean, yielding mixed layers
-as deep as 300 m (Figure \ref{fig1}d). In late spring/early summer enhanced solar radiation and mixed-layer
-instabilities re-stratify the upper ocean, yielding mixed layers as shallow as 40 m (Figure \ref{fig1}e).
+deeper than 200 m to the north (Figure \ref{fig1}b). Summertime re-stratification yields mixed layers shallower
+than 40 m (Figure \ref{fig1}c).
+
 Fundamentally, the upper-ocean density structure is well-captured by both LLC simulations:
 a comparison with Argo climatology shows that both simulations skillfully represent the Kuroshio
-Extension stratification and its seasonal variability  (supporting information).
-
-\begin{figure*}[ht]
-\begin{center}
-\hspace{-1.25cm}\includegraphics[width=.75\textwidth]{figs/fig1_1.pdf}\\
-\vspace{-.125cm}
-\includegraphics[width=.75\textwidth]{figs/fig1_2.pdf}
- \caption{(a) The study region with the subregion where the LLC outputs are
-          analyzed. Colors represent the topography and white lines are contours of absolute
-          dynamic topography every 0.1 m from AVISO. LLC 4320 (1/48$^\circ$) snapshots of surface vorticity (b and c) and transects
-          of potential density at 165$^\circ$E (d and e). The snapshots were
-          taken at 00:00 UTC.}
-\vspace{-1.5cm}
- \label{fig1}
- \end{center}
- \end{figure*}
-
-\section{Statistics of the surface lateral velocity gradient tensor}
+Extension stratification and its seasonal variability  (supporting information). Moreover, model smoothed
+fields have eddy kinetic energy (EKE) levels ($0.041$ m$^2$ s$^{-2}$) similar to Jason II across-track
+geostrophic EKE ($0.036$ m$^2$ s$^{-2}$) and gridded geostrophic EKE ($0.032$ m$^2$ s$^{-2}$).
+
+\section{Bulk statistics of the surface lateral velocity gradient tensor}
 To study the seasonality in the surface velocity, we calculate the lateral velocity gradient tensor
@@
 using a centered
-second-order finite difference scheme. We then diagnose
+second-order finite difference scheme --- we obtain qualitatively similar results
+using spectral differentiation of velocity fields periodized with reflections;
+statistics of fields obtained via spectral differentiation are 10-15$\%$ larger.
+We then diagnose
 the vertical vorticity
@@
 \end{equation}
-These diagnostics highlight the submesoscale structures in the flow
-\citep[e.g.,][]{capet_etal2008a,shcherbina_etal2013}.
-
-Figures \ref{fig1}b-c show snapshots of vertical vorticity $\zeta$ in early spring
-(April 15) and fall (October 15) in the LLC4320 (1/48$^\circ$) simulation.
-The model solutions depict seasonality in vorticity: large values of
+Because vortices, fronts, and IGWs are ubiquitous at submesoscales, second-order
+statistics of vorticity, strain, and divergence highlight these
+fine lateral scales \citep[e.g.,][]{shcherbina_etal2013}.
+
+The model solutions depict seasonality in vorticity (Figures \ref{fig1}b-c): large values of
 fine-grained vertical vorticity are observed in early spring with maximum
@@
 
- \begin{figure*}[ht]
-   \begin{center}
-     \includegraphics[width=.75\textwidth]{figs/fig2.pdf}
-  \caption{Time series of the root-mean-square (RMS) of (a) vorticity,
-  (b) rate of strain, and (c) horizontal divergence in the LLC outputs and gridded AVISO data.}
-  \label{fig2}
-  \end{center}
-\end{figure*}
-
-The bulk of vertical vorticity and strain rate are associated with subinertial flows
-($T_f = 2\pi/f_0\approx 23.5$h, where $f_0$ is the inertial frequency at
-the mean latitude): daily-averaging the velocity fields suppresses super-inertial
+Subinertial flows account for the bulk of vertical vorticity and strain rate
+($T_{32.5} = 2\pi/f_{32.5}\approx 22.3$h, where $f_{32.5}$ is the inertial frequency at
+the  mid-latitude $32.5^\circ$N): daily-averaging the velocity fields suppresses super-inertial
 motions and reduces the RMS
@@
 relatively fast, and there is no clear
-temporal and spatial scale separation between those motions and inertia-gravity waves
-\citep{mcwilliams2016}:
+temporal and spatial scale separation between those motions and IGWs
+\citep[][]{mcwilliams2016}:
 daily-averaging the velocity fields efficiently suppresses the summertime horizontally
@@
 
+\section{Joint probability density functions}
+
+The results of figure \ref{fig2}a-c show that submesoscale surface
+ variability stems from different dynamics in summer than in winter.
+ To characterize these differences, we calculate
+joint probability distributions (joint-PDF) of
+vorticity-strain and vorticity-Laplacian of sea-surface height (Figure \ref{fig3}).
+
+The April vorticity-strain joint-PDF has a shape characteristic of submesoscale turbulence
+ (Figures \ref{fig3}a-b). The alignment of vorticity and strain $\alpha \sim \pm\zeta$
+ with strong positive skewness are fingerprints of submesoscale fronts
+ \citep{shcherbina_etal2013,mcwilliams2016}. The shape of the vorticity-strain
+ joint-PDF is similar for hourly and daily-averaged fields, although
+ the vorticity skewness reduces from 1.4 to 1.13 from hourly to daily-averaged.
+ The April results indicate that wintertime submesoscale
+ surface velocity is strongly dominated by submesoscale turbulence.
+ The daily-averaged fields
+ are largely in  geostrophic balance,
+but the hourly velocity and sea-surface height fields have an important
+ageostrophic component as depicted by the joint-PDF of vorticity-Laplacian of SSH
+winter (Figure \ref{fig3}e).
+
+The October vorticity-strain joint-PDF shows much weaker skewness (the vorticity skewness
+is 0.68 and 0.67 for hourly and daily-averaged velocities). The shape of the
+vorticity-strain joint-PDF  appears to be a combination of two half-ellipses centered
+about $\zeta=0$, one with a 45$^\circ$ slope (characteristic of submesocale fronts
+that persist in summer) and one with a very steep slope.
+That the submesoscale dynamics in October are mainly ageostrophic is clearly depicted
+in the shape of joint-PDF of vorticity-Laplacian of sea-surface height for hourly fields
+ (Figure \ref{fig3}g), which is an ellipse aligned in the vertical axis.
+Daily-averaging the model suppresses the ageostrophic, super-inertial flows, and, therefore,
+the daily-averaged flow is essentially geostrophic as depicted by the 45$^\circ$-tilted
+ellipse in the vorticity-Laplacian of sea-surface height joint-PDF (Figure \ref{fig3}h).
+
+ Time series of PDFs of vorticity and divergence (supporting information) show
+ a strong oscillation between these two regimes. In late winter/early spring
+ the vorticity is strongly positively skewed, whereas the divergence is moderately
+ negatively skewed (convergence/downwelling) as predicted by frontogenesis \citep[e.g., ][]{mcwilliams2016}.
+ In late summer/early fall, the divergence is stronger, but PDFs are much less skewed,
+consistent with a superposition of random linear IGWs \citep[e.g.,][]{garrett_munk1972}.
+
+\section{Wavenumber-frequency spectrum}
+To confirm that submesoscale IGWs account for most of the super-inertial SSH variance
+and surface KE, we calculate the wavenumber-frequency spectrum of SSH variance. Focusing on
+the high-frequency
+content, we compute the wavenumber-frequency spectrum  every 10 days and average
+the results to obtain a spectral estimate. Before calculating the discrete Fourier
+transform, we
+remove linear trends and multiply the data by a three-dimensional Hanning window.
+We azimuthally-average the spectrum in wavenumber.
+
+Mesoscales flows and the lunar semi-diurnal tide contain most of the
+SSH variance (Figures \ref{fig4}a-b). But there is significant SSH variance at
+submesoscales, both sub-inertial
+and super-inertial. Sub-inertial SSH variance at submesoscales is larger in April, whereas
+super-inertial SSH variance at submesoscales is larger in October. At scales smaller
+than 100 km (submesoscales), the SSH variance is 66$\%$ larger in October than in April.
+IGWs, particularly modes two through four, account for most of the SSH variance
+increase in summer --- higher vertical
+modes are more sensitive to changes in near-surface stratification (supporting information).
+In October, super-inertial IGWs account for a SSH standard deviation of about 4.9 cm,
+comparable to the April sub-inertial SSH standard deviation of $\sim$5.2 cm (Table \ref{tab1}).
+The
+Similarly, the submesoscale super-inertial surface KE is 63$\%$ larger in October
+than in April (supporting information).
+
+Figure \ref{fig4}c shows the integral over frequency
+of wavenumber-frequency spectra in figures \ref{fig4}a-b.
+Both in April and October, the
+SSH wavenumber spectrum approximately follows a -2 power-law at submesoscales
+(solid lines in figure \ref{fig4}c)  --- there is
+small difference between the spectra owing to the phase cancellation between the
+seasonal cycle of submesoscale turbulence and inertia-gravity waves. The sub-inertial flow
+has a steeper submesoscale SSH variance wavenumber spectrum (dashed lines in figure \ref{fig4}c).
+ The difference between the SSH variance wavenumber spectrum for all frequencies and
+ sub-inertial frequencies is dramatic in October when super-inertial flows account for
+ about 80\% of the SSH variance and KE at submesoscales --- we obtain similar results by computing
+SSH variance and KE wavenumber spectra directly from hourly and daily-averaged
+SSH velocity fields.
+
+%\section{Projection onto horizontal scales}
+%
+%To quantify the projection of these flows onto different horizontal
+%scales, we calculate wavenumber spectra of kinetic energy and sea-surface height
+% variance.  Before calculating the spectra, time-mean and spatial linear
+%trends were removed, and the resulting fields were multiplied by a two-dimensional Hanning window.
+%The two-dimensional spectra were averaged azimuthaly \citep[e.g., ][]{rocha_etal2016} .
+%We discuss only spectra for the 1/48$^\circ$ simulation, which extends the 1/24$^\circ$ simulation
+%towards smaller scales (supporting information).
+%
+%
+
+%
+%
+%\begin{figure*}[ht]
+%  \begin{center}
+%    \includegraphics[width=.65\textwidth]{figs/fig5.pdf}
+% \caption{\small Surface (horizontal) kinetic energy (a) and sea-surface height variance wavenumber spectra (b)
+% in the 1/48$^\circ$ simulation. Solid lines
+% are spectra based on hourly snapshots, and dashed lines are spectra based on daily-averaged
+% fields.}
+% \label{fig5}
+% \end{center}
+%\end{figure*}
+%
+%
+%Figure \ref{fig5}a depicts the horizontal wavenumber spectra of surface KE in
+%April and October. At scales larger than 20 km, the April and October spectra based on hourly
+%velocity snapshots (solid lines) are nearly indistinguishable from each other within 95$\%$ confidence level.
+%Consistent with the results of
+%\citet{rocha_etal2016}, who analyzed a 3-month output of the LLC4320 simulation in Drake Passage,
+%there is significant high-frequency variabilty at submesoscales. Daily-averaging
+%the velocity field suppresses spatial variability at scales smaller than about 250
+%km, both in April and October (compare solid lines against dashed lines in Figure
+%\ref{fig5}a). But this suppression is dramatic in October, when the IGWs peak. At scales
+%smaller than 100 km, 39$\%$ of the surface KE in April is accounted for by super-inertial
+%flows as opposed to 79$\%$ in October. The seasonality of subinertial submesoscale flows
+%is strong, consistent
+%with the results of \citet{sasaki_etal2014}, which are
+%based on daily-averaged velocity fields of a different model without tidal forcing
+%(P. Klein, personal communication).
+%
+%The IGWs project on the sea-surface.
+%There is a dramatic difference between the spectra
+%based on hourly and daily-averaged sea-surface height in October (Figure \ref{fig5}b): at scales smaller than 100 km, the spectra
+%of hourly sea-surface height roughly follows a $-2$ power-law, whereas the spectra of daily-averaged
+%sea-surface height roughly follows a $-5$ power-law. At these scales, 33$\%$ of the sea-surface height variance in April is accounted for by super-inertial
+%flows as opposed to 83$\%$ in October. Curiously, the out-of-phase seasonal cycle of submesoscale
+%turbulence and near-surface submesoscale IGWs conspire to yield weak seasonality
+%in the spectra of KE and sea-surface height variance based on hourly fields.
+%
+%
+\section{Discussion and conclusion}
+Our work adds to recent model \citep[e.g., ][]{sasaki_etal2014}
+and observational \citep[e.g., ][]{callies_etal2015,buckingham_etal2016} evidence of vigorous seasonality in
+submesoscale turbulence. In particular ---
+in global simulations with embedded tides --- near-surface submesoscale IGWs undergo
+a strong seasonal cycle that is out of phase with the seasonal cycle of
+submesoscale turbulence. IGW energy and SSH variance are significantly spread across
+the submesoscales, likely through mesoscale advection and refraction, and
+dispersion \citep[e.g.,][and references therein]{ponte_klein2015,alford_etal2016}
+ and parametric
+sub-harmonic instability \citep[e.g., ][]{mackinnon_winters2005}.
+In the LLC simulations,
+super-inertial IGWs account for most of the surface submesoscale KE and SSH variance;
+near-inertial IGWs project onto larger scales, possibly because submesoscale
+NIWs quickly propagate into the interior and weakly project onto SSH.
+
+ %We further conjecture that the summertime dominance of IGWs
+ %\citep{callies_etal2015} is a consequence both of suppression of submesoscale turbulence and
+%enhancement of IGWs due to re-stratification of
+%the upper ocean.
+
+\cite{dasaro1978} showed that the velocity of linear internal waves
+in the mixed layer strongly depends on the density jump at the mixed layer
+base, with largest velocities when the jump is strongest.
+In summer, the shallow mixed layer overlays a strong seasonal pycnocline,
+and thus the internal waves projection onto the mixed layer may be stronger
+according to \cite{dasaro1978}'s arguments. An
+alternative explanation is that the shape of the baroclinic modes
+changes seasonally (supporting information).
+
+Internal tide energy flux does not show a seasonal modulation \citep[e.g.,][]{alford2003};
+hence one expects a strong seasonality in the near-surface
+expression of internal tides and
+higher frequency IGWs if the near-surface stratification varies strongly.
+The intermittent windy near-inertial wave (NIW) generation peaks in early winter, but
+NIW energy has a similar energy level in October and April
+\citep{alford_etal2016}. Furthermore, in our simulations the higher frequency inertia-gravity waves
+account for most of the seasonality in super-inertial surface KE and SSH
+(Figure \ref{fig4}a-b).
+
+The seasonally varying mixed-layer depth challenges observation and simulation of
+flows at mixed-layer deformation-radius length scale in observational programs and models with
+fixed resolution \citep{buckingham_etal2016}. But the mixed layer available potential energy
+is lowest in summer owing to shallow mixed-layer depths and weak lateral buoyancy
+gradients \citep{callies_etal2015}. Moreover, in baroclinic turbulence, the deformation
+radius has only a secondary, catalytic role --- most of the baroclinic conversion occurs at
+larger scales \citep{larichev_held1995,callies_etal2016}.
+Thus, the model failure to resolve the summertime
+ML deformation radius is unlikely to account for the vigorous seasonality in submesoscale
+turbulence.
+
+In the context of SWOT, these results imply that if one is interested
+only in the geostrophic flow, so that surface velocity can be diagnosed
+from SSH, then the noise-to-signal ratio will have a strong seasonality owing
+to the presence of IGWs (the coherent fraction of internal tides may be removed
+efficiently).
+An important caveat is that the Kuroshio Extension may be typical of a mesoscale-rich
+subtropical regions, but it is unlikely to be
+representative of other regions such as low-eddy-kinetic-energy eastern boundary
+currents.  We plan to report on the geographic variability of submesoscale
+seasonality in a future study.
+
+%
+%  ACKNOWLEDGMENTS
+
+\begin{acknowledgments}
+ William R. Young and the reviewers --- Christian Buckingham and two anonymous referees ---
+  provided helpful and stimulating comments. Greg Wagner
+ suggested that the near-surface shape of the baroclinic modes changes seasonally.
+ We thank the MITgcm community and our colleagues at the NASA Advanced
+Supercomputing (NAS) Division for their awesome support.
+This research was funded by NSF (OCE1357047) and NASA (NNX13AE44G,NNX13AE85G,NNX16AH67G,NNX16AO5OH).
+The LLC output can be obtained from the ECCO project (\texttt{http://ecco2.org/llc$\_$hires}). The altimeter products were produced by Ssalto/Duacs
+and distributed by AVISO, with support from CNES (\texttt{http://www.aviso.altimetry.fr/duacs/}).
+Codes and output files are available online at the project repository
+ (\texttt{https://github.com/crocha700/UpperOceanSeasonality}).
+\end{acknowledgments}
+
+%\bibliographystyle{agufull08}
+%\bibliography{rocha_etal}
+
  \begin{figure*}[ht]
  \begin{center}
-    \includegraphics[width=.95\textwidth]{figs/fig3.pdf}
- \caption{Seasonal variation of joint probability distributions:  vorticity vs.\ strain rate (a through d),
+ \hspace{-1.25cm}\includegraphics[width=.9\textwidth]{figs/fig1_1.pdf}\\
+ \vspace{-.125cm}
+ \includegraphics[width=.9\textwidth]{figs/fig1_2.pdf}
+  \caption{\small (a) The study region with the subregion where the LLC outputs are
+           analyzed. Colors represent the topography and white lines are contours of absolute
+           dynamic topography every 0.1 m from AVISO. LLC 4320 (1/48$^\circ$) snapshots of
+           transects
+           of potential density at 165$^\circ$E (b and c) and  surface vorticity (d and e).
+           In (b) and (c) the white line represents the mixed layer depth. The snapshots were
+           taken at 12:00 UTC.}
+ \vspace{-1.5cm}
+  \label{fig1}
+  \end{center}
+  \end{figure*}
+
+  \begin{figure*}[ht]
+    \begin{center}
+      \includegraphics[width=.65\textwidth]{figs/fig2.pdf}
+   \caption{\small Time series of the root-mean-square (RMS) of surface (a) vorticity,
+   (b) rate of strain, and (c) horizontal divergence in the LLC outputs and gridded AVISO data.
+   The convergence of meridians account for the small RMS divergence in AVISO line in (c).}
+   \label{fig2}
+   \end{center}
+ \end{figure*}
+
+ \begin{figure*}[ht]
+   \begin{center}
+     \includegraphics[width=.8\textwidth]{figs/fig3.pdf}
+  \caption{\small Seasonal variation of joint probability distributions of
+           surface lateral velocity gradient tensor:  vorticity vs.\
+           strain rate (a through d),
            and vorticity vs.\ Laplacian of sea-surface height (e through h) in April (a, b,
@@
 
-\section{Seasonality of submesoscale dynamics}
-The results of figure \ref{fig2}a-c show that submesoscale surface
- variability stems from different dynamics in summer than in winter.
- To characterize these differences, we calculate
-joint probability distributions (jPDF) of
-vorticity-strain and vorticity-Laplacian of sea-surface height (Figure \ref{fig3}).
-
-The April vorticity-strain jPDF has a shape characteristic of submesoscale turbulence
- (Figures \ref{fig3}a-b). The alignment of vorticity and strain $\alpha \sim \pm\zeta$
- with strong positive skewness are fingerprints of submesoscale fronts
- \citep{shcherbina_etal2013,mcwilliams2016}. The shape of the vorticity-strain
- jPDF is similar for hourly and daily-averaged fields, although
- the vorticity skewness reduces from 1.4 to 1.13 from hourly to daily-averaged.
- The April results are characteristic of winter, indicating that wintertime submesoscale
- surface velocity is strongly dominated by submesoscale turbulence.
- The hourly velocity and sea-surface heigh fields have an important
-ageostrophic component as depicted by the jPDF of vorticity-Laplacian of SSH
-winter (Figure \ref{fig3}e). In other words, even in April, when submesoscale turbulence
-prevails, only the daily-averaged fields
-are largely
-in  geostrophic balance (the jPDF of daily-averaged vorticity vs.\\ Laplacian of sea-surface height
-is an ellipse with large eccentricity and main axis tilted by 45$^\circ$; Figure \ref{fig3}f).
-
-The October vorticity-strain jPDF shows much weaker skewness (the vorticity skewness
-is 0.68 and 0.67 for hourly and daily-averaged velocities). The shape of the
-vorticity-strain jPDF  appears to be a combination of two half-ellipses centered
-about $\zeta=0$, one with a 45$^\circ$ slope (characteristic of submesocale fronts
-that persist in summer) and one with a very steep slope.
-That the submesoscale dynamics in October are mainly ageostrophic is clearly depicted
-in the shape of jPDF of vorticity-Laplacian of sea-surface height for hourly fields
- (Figure \ref{fig3}g), which is an ellipse aligned in the vertical axis.
-Daily-averaging the model suppresses the ageostrophic, super-inertial flows, and, therefore,
-the daily-averaged flow is essentially geostrophic as depicted by the 45$^\circ$-tilted
-ellipse in the vorticity-Laplacian of sea-surface height jPDF.
-
- Time series of PDFs of vorticity and divergence (supporting information) show
- a strong oscillation between these two regimes. In late winter/early spring
- the vorticity is strongly positively skewed, whereas the divergence is moderately
- negatively skewed as predicted by frontogenesis \citep[e.g., ][]{capet_etal2008a,mcwilliams2016}
-. In late summer/early fall, the divergence is stronger, but PDFs are much less skewed,
-consistent with linear inertia-gravity waves.
-
-\section{Projection onto horizontal scales}
-
-To better quantify the projection of these flows onto different horizontal
-scales, we calculate wavenumber spectra of kinetic energy and sea-surface height
- variance.  Before calculating the spectra, time-mean and spatial linear
-trends were removed, and the resulting fields were multiplied by a two-dimensional Hanning ``spectral window''.
-The two-dimensional spectra were averaged azimuthaly \citep[e.g., ][]{rocha_etal2016} .
-We discuss only spectra for the 1/48$^\circ$ simulation, which extends the 1/24$^\circ$ simulation
-towards smaller scales (supporting information).
-
  \begin{figure*}[ht]
    \begin{center}
-    \includegraphics[width=1.\textwidth]{figs/fig4.pdf}
- \caption{Surface (horizontal) kinetic energy (a) and sea-surface height variance wavenumber spectra (b)
- in the 1/48$^\circ$ simulation. Solid lines
- are spectra based on hourly snapshots, and dashed lines are spectra based on daily-averaged
- fields.}
+     \includegraphics[width=.9\textwidth]{figs/fig4.pdf}
+  \caption{\small LLC4320 wavenumber-frequency spectrum of SSH variance in (a) April and (b) October.
+           (c) Wavenumber spectrum of SSH variance --- the integral of (a) and (b) over frequency.
+           In (a) and (b), the light gray shaded region depicts the dispersion relations for inertia-gravity
+           waves from mode 1 through mode 4 across the latitudinal domain; horizontal dashed lines represent the semi-diurnal
+           lunar tidal frequency ($M_2$), and the inertial frequency at mid-latitude
+           ($f_{32.5}$) and its first harmonic (2$f_{32.5}$). For reference, the gray curvish line represents
+           the SWOT baseline requirement $2\times 10^{-4} + 1.25\times 10^{-7} k^{-2}$ m$^2$/cpkm \citep{rodriguez2012}.
+           }
    \label{fig4}
@@
 
-Figure \ref{fig4}a depicts the horizontal wavenumber spectra of surface KE in
-April and October. At scales larger than 20 km, the April and October spectra based on hourly
-velocity snapshots (solid lines) are nearly indistinguishable from each other within 95$\%$ confidence level.
-Consistent with the results of
-\citet{rocha_etal2016}, who analyzed a 3-month output of the LLC4320 simulation in Drake Passage,
-there is significant high-frequency variabilty at submesoscales. Daily-averaging
-the velocity field suppresses spatial variability at scales smaller than about 250
-km, both in April and October (compare solid lines against dashed lines in Figure
-\ref{fig4}a). But this suppression is dramatic in October, when the inertia-gravity waves peak. At scales
-smaller than 100 km, 39$\%$ of the surface KE in April is accounted for by super-inertial
-flows as opposed to 79$\%$ in October. The seasonality of subinertial submesoscale flows
-is strong, consistent
-with the results of \citet{sasaki_etal2014}, which are
-based on daily-averaged velocity fields of a different model without tidal forcing
-(P. Klein, personal communication).
-
-The inertia-gravity waves project on the sea-surface.
-There is a dramatic difference between the spectra
-based on hourly and daily-averaged sea-surface height in October (Figure \ref{fig4}b): at scales smaller than 100 km, the spectra
-of hourly sea-surface height roughly follows a $-2$ power-law, whereas the spectra of daily-averaged
-sea-surface height roughly follows a $-5$ power-law. At these scales, 33$\%$ of the sea-surface height variance in April is accounted for by super-inertial
-flows as opposed to 83$\%$ in October! Curiously, the out-of-phase seasonal cycle of submesoscale
-turbulence and near-surface submesoscale inertia-gravity waves conspire to yield weak seasonality
-in the spectra of KE and sea-surface height variance based on hourly fields.
-
-\section{Summary and Conclusion}
-Our work adds to recent modeling \citep{sasaki_etal2014}
-and observational \citep{callies_etal2015,buckingham_etal2016} evidence of vigorous seasonality in
-submesoscale turbulence. In particular, our main finding is that in global simulations with embedded tides the
-near-surface submesoscale inertia-gravity waves in the Kuroshio Extension undergo
-a strong seasonal cycle that is out of phase with the seasonal cycle of
-submesoscale turbulence. We conjecture that the summertime dominance of inertia-gravity waves
- \citep{callies_etal2015} is a consequence both of suppression of submesoscale turbulence and
-enhancement of inertia-gravity waves due to re-stratification of
-the upper ocean.
-
-\cite{dasaro1978} showed that the velocity of linear internal waves
-in the mixed layer strongly depends on the density jump at the mixed layer
-base, with largest mixed layer velocities when the jump is strongest.
-In summer, the shallow mixed layer overlays a strong seasonal pycnocline,
-and thus the internal waves projection onto the mixed layer may be stronger
-according to \cite{dasaro1978}'s arguments. An
-alternative explanation is that the shape of the baroclinic modes
-changes seasonally. In particular the baroclinic modes are
-significantly more surface intensified in late summer/early fall
- (supporting information).
-If the internal wave source has weak seasonal dependence, then a strong
-seasonal cycle in the near-surface expression of internal waves is expected.
-Internal tide generation does not show a seasonal modulation \citep[e.g.,][]{alford2003},
-and therefore one expects a strong seasonality in the near-surface
-expression of internal tides and
-other submesoscale inertia-gravity waves generated through internal tide interactions.
-Near-inertial wave generation peaks in winter \citep[e.g.,][]{alford2003}, but
-those waves project onto large horizontal scales \citep[e.g, ][]{qi_etal1995}.
-
-An important caveat is that this study focuses on a single patch of ocean in the
-vicinity of the Kuroshio Extension. This may be typical of a mesoscale-rich
-subtropical region, but it is unlikely to be
-representative of other regions such as low-eddy-kinetic-energy eastern boundary currents
-and the middle of the subtropical gyre.  We plan to report on the geographic variability of submesoscale
- seasonality in a future study.
-
-%The effects of smaller-scale/higher-frequency
-%``sub-submesoscale'' flows on the submesoscale surface velocity and SSH variability
-%are presently unknown.
-
-% That the  surface velocity and SSH submesoscale variability may be
-% dominated by ageostrophic flows in
-% summer/fall has consequences for the interpretation of data from
-% the future SWOT and COMPIRA altimeter missions,
-% which will deliver SSH measurements at submesoscales. To the extent that
-% high-frequency flows are dominated by spatially incoherent internal tides and other
-% internal waves, it may be very difficult (if not impossible) to
-% separate SSH submesoscale variability associated with geostrophic motions
-% from high-frequency, ageostrophic flows. Thus, previous claims that
-% one will be able to easily obtain submesoscale surface
-% geostrophic velocities and monitor such seasonal cycle on global scales
-% \citep{sasaki_etal2014,qiu_etal2014} are overstated.
-
-% Here we argue that those divergent motions are likely inertia-gravity waves.
-% We can show a couple of spectra and argue that they roughly satisfy polarization
-% relations.
-%
-% We can also try and show some observations here. Any cruises with ADCP data.
-% While it may be hard to have enough data for the errorbars to be small, a
-% rough consistency may be better than nothing.
-%
-% Perhaps a mooring data should show strong seasonal modulation at high-frequencies?
-%
-% We must be able to present some observational evidence that the model is not
-% misleading at high-frequencies.
-
-% Here we argue that high-frequency (supra-daily) flows significantly project
-% on the sea-surface, and thus estimation of submesoscale (10-100 km) geostrophic
-% velocity from sea-surface height (SSH) is not warranted. Calculating spectral
-% KE fluxes from both velocity and geostrophic velocity (from SSH) would
-% contrast with the results in Sasaki et al.
-
-% \section{Projection on time scales}
-% To better understand the high-frequency/high-wavenumber flows  we calculate frequency
-% spectra of surface KE winter \ref{fig4}). To obtain statistically meaningful results,
-% we average spectral estimates along a section at 165$^\circ$E.  There is about 50
-% independent spectral realizations.
-%
-% The surface KE spectra of subinertial flows very roughly follow a $-2$ power-law.
-% The high-frequency flows are split near-inertial flows, tidal signals and their
-% higher harmonics. Hence these high-frequency flows are dominated by inertia-gravity
-% waves. (Note that the inertial peak nearly merges with the
-% diurnal tide peak.) There is statistically significant seasonality. The seasonal
-% cycle of the sub-inertial motions (at least for periods between 5-20 days)
-% is out-of-phase with the seasonal cycle of supra-inertial flows: the former flows
-% are more energetic in April and the latter in October.
-%
-% Most of these high-frequency inertia-gravity waves project onto scales smaller
-% than 100 km. The surface KE spectra of 100-km-smoothed velocity field is significantly
-% suppressed in supra-inertial frequencies --- there are no statistically significant
-% changes at sub-inertial frequencies (see dashed lines in figure \ref{fig4}).
-
-% say something about relative energy in smoother vs.\\ unsmoothed fields.
-% also say something about comparison with observations.
-
-%\begin{figure}[ht]
-%  \begin{center}
-%    \includegraphics[width=.45\textwidth]{figs/fig4.pdf}
-% \caption{Surface (horizontal) KE frequency spectra. Solid lines
-% are spectra based on the 1/24$^\circ$ fields, dashed line are spectra
-% based on 100-km-smoothed fields.}
-% \label{fig4}
-% \end{center}
-%\end{figure}
-
-%
-%  ACKNOWLEDGMENTS
-
-\begin{acknowledgments}
- William R. Young provided helpful feedback on the first draft. Greg Wagner first
- suggested that the near-surface shape of the baroclinic modes may change seasonally.
- We thank the MITgcm community and our colleagues at the NASA Advanced
-Supercomputing (NAS) Division for their awesome support.
-This research was funded by NSF (OCE1357047) and NASA (NNX13AE44G,NNX13AE85G,NNX16AH67G).
-The LLC output can be obtained from the ECCO project (\texttt{http://ecco2.org/llc$\_$hires}). The altimeter products were produced by Ssalto/Duacs
-and distributed by AVISO, with support from CNES (\texttt{http://www.aviso.altimetry.fr/duacs/}).
-Codes and output files are available online at the project repository
- (\texttt{https://github.com/crocha700/UpperOceanSeasonality}).
-\end{acknowledgments}
-
-\bibliographystyle{agufull08}
-%\bibliography{rocha_etal}
+ % table
+ \begin{table}
+ \label{tab1}
+ \caption{The square root of the mean-square SSH at submesoscales (10-100 km)
+         estimated from the wavenumber-frequency spectrum of SSH variance in figure \ref{fig4}.
+         The SWOT baseline requirement at 15 km is approximately $\sqrt{2}$ cm.
+         }
+   \begin{center}
+     \begin{tabular}{  l | c | c }
+     \hline
+     & April & October\\
+     \hline
+   sub-inertial ($<0.8f$)   & 5.2 cm & 2.0 cm  \\
+   super-inertial ($>1.2f$) & 2.9 cm  & 4.9 cm\\
+   \hline
+     \end{tabular}
+   \end{center}
+ \end{table}
 
@@
 % Text
-\begin{thebibliography}{18}
+% Or,
+%
+% If you use BiBTeX for your references, please use the agufull08.bst file (available at % ftp://ftp.agu.org/journals/latex/journals/Manuscript-Preparation/) to produce your .bbl
+% file and copy the contents into your paper here.
+%
+% Follow these steps:
+% 1. Run LaTeX on your LaTeX file.
+%
+% 2. Make sure the bibliography style appears as \bibliographystyle{agufull08}. Run BiBTeX on your LaTeX
+% file.
+%
+% 3. Open the new .bbl file containing the reference list and
+%   copy all the contents into your LaTeX file here.
+%
+% 4. Comment out the old \bibliographystyle and \bibliography commands.
+%
+% 5. Run LaTeX on your new file before submitting.
+%
+% AGU does not want a .bib or a .bbl file. Please copy in the contents of your .bbl file here.
+
+\begin{thebibliography}{2}
+
 \providecommand{\natexlab}[1]{#1}
@@
   \textit{423}(6936), 159--162.
+
+\bibitem[{\textit{Alford et~al.}(2016)\textit{Alford, MacKinnon, Simmons, and
+  Nash}}]{alford_etal2016}
+Alford, M.~H., J.~A. MacKinnon, H.~L. Simmons, and J.~D. Nash (2016),
+  Near-inertial internal gravity waves in the ocean, \textit{Annual review of
+  marine science}, \textit{8}, 95--123.
+
+\bibitem[{\textit{Brannigan et~al.}(2015)\textit{Brannigan, Marshall,
+  Naveira-Garabato, and Nurser}}]{brannigan_etal2015}
+Brannigan, L., D.~P. Marshall, A.~Naveira-Garabato, and A.~G. Nurser (2015),
+  The seasonal cycle of submesoscale flows, \textit{Ocean Modelling},
+  \textit{92}, 69--84.
 
@@
 
-\bibitem[{\textit{Capet et~al.}(2008)\textit{Capet, McWilliams, Molemaker, and
-  Shchepetkin}}]{capet_etal2008a}
-Capet, X., J.~C. McWilliams, M.~J. Molemaker, and A.~Shchepetkin (2008),
-  {Mesoscale to submesoscale transition in the California Current System. Part
-  I: Flow structure, eddy flux, and observational tests}, \textit{Journal of
-  Physical Oceanography}, \textit{38}(1), 29--43.
-
 \bibitem[{\textit{D'Asaro}(1978)}]{dasaro1978}
@@
 
+\bibitem[{\textit{Fu and Ubelmann}(2014)}]{fu_ubelmann2014}
+Fu, L.-L., and C.~Ubelmann (2014), On the transition from profile altimeter to
+  swath altimeter for observing global ocean surface topography,
+  \textit{Journal of Atmospheric and Oceanic Technology}, \textit{31}(2),
+  560--568.
+
+\bibitem[{\textit{Garrett and Munk}(1972)}]{garrett_munk1972}
+Garrett, C., and W.~Munk (1972), Space-time scales of internal waves,
+  \textit{Geophysical \& Astrophysical Fluid Dynamics}, \textit{3}(1),
+  225--264.
+
+\bibitem[{\textit{Larichev and Held}(1995)}]{larichev_held1995}
+Larichev, V.~D., and I.~M. Held (1995), Eddy amplitudes and fluxes in a
+  homogeneous model of fully developed baroclinic instability, \textit{Journal
+  of physical oceanography}, \textit{25}(10), 2285--2297.
+
+\bibitem[{\textit{MacKinnon and Winters}(2005)}]{mackinnon_winters2005}
+MacKinnon, J., and K.~Winters (2005), Subtropical catastrophe: Significant loss
+  of low-mode tidal energy at 28.9, \textit{Geophysical Research Letters},
+  \textit{32}(15).
+
 \bibitem[{\textit{Marshall et~al.}(1997)\textit{Marshall, Hill, Perelman, and
@@
 Menemenlis, D., J.-M. Campin, P.~Heimbach, C.~Hill, T.~Lee, A.~Nguyen,
-    M.~Schodlok, and H.~Zhang (2008), ECCO2: High resolution global ocean and sea
-    ice data synthesis, \textit{Mercator Ocean Quarterly Newsletter},
+  M.~Schodlok, and H.~Zhang (2008), {ECCO2: High resolution global ocean and
+  sea ice data synthesis}, \textit{Mercator Ocean Quarterly Newsletter},
   \textit{31}, 13--21.
 
-\bibitem[{\textit{Qi et~al.}(1995)\textit{Qi, De~Szoeke, Paulson, and
-  Eriksen}}]{qi_etal1995}
-Qi, H., R.~A. De~Szoeke, C.~A. Paulson, and C.~C. Eriksen (1995), {The
-  structure of near-inertial waves during ocean storms}, \textit{Journal of
-  Physical Oceanography}, \textit{25}, 2853--2871.
+\bibitem[{\textit{Ponte and Klein}(2015)}]{ponte_klein2015}
+Ponte, A.~L., and P.~Klein (2015), Incoherent signature of internal tides on
+  sea level in idealized numerical simulations, \textit{Geophysical Research
+  Letters}, \textit{42}(5), 1520--1526.
 
@@
 
+\bibitem[{\textit{Rodriguez}(2014)}]{rodriguez2012}
+Rodriguez, E. (2014), {Surface Water and Ocean Topography Mission (SWOT):
+  Science Requirements Document}.
+
 \bibitem[{\textit{Sasaki et~al.}(2014)\textit{Sasaki, Klein, Qiu, and
@@
 
-% Or,
-%
-% If you use BiBTeX for your references, please use the agufull08.bst file (available at % ftp://ftp.agu.org/journals/latex/journals/Manuscript-Preparation/) to produce your .bbl
-% file and copy the contents into your paper here.
-%
-% Follow these steps:
-% 1. Run LaTeX on your LaTeX file.
-%
-% 2. Make sure the bibliography style appears as \bibliographystyle{agufull08}. Run BiBTeX on your LaTeX
-% file.
-%
-% 3. Open the new .bbl file containing the reference list and
-%   copy all the contents into your LaTeX file here.
-%
-% 4. Comment out the old \bibliographystyle and \bibliography commands.
-%
-% 5. Run LaTeX on your new file before submitting.
-%
-% AGU does not want a .bib or a .bbl file. Please copy in the contents of your .bbl file here.
-
-%\begin{thebibliography}{}
-
-%\providecommand{\natexlab}[1]{#1}
-%\expandafter\ifx\csname urlstyle\endcsname\relax
-%  \providecommand{\doi}[1]{doi:\discretionary{}{}{}#1}\else
-%  \providecommand{\doi}{doi:\discretionary{}{}{}\begingroup
-%  \urlstyle{rm}\Url}\fi
-%
-%\bibitem[{\textit{Atkinson and Sloan}(1991)}]{AtkinsonSloan}
-%Atkinson, K., and I.~Sloan (1991), The numerical solution of first-kind
-%  logarithmic-kernel integral equations on smooth open arcs, \textit{Math.
-%  Comp.}, \textit{56}(193), 119--139.
-%
-%\bibitem[{\textit{Colton and Kress}(1983)}]{ColtonKress1}
-%Colton, D., and R.~Kress (1983), \textit{Integral Equation Methods in
-%  Scattering Theory}, John Wiley, New York.
-%
-%\bibitem[{\textit{Hsiao et~al.}(1991)\textit{Hsiao, Stephan, and
-%  Wendland}}]{StephanHsiao}
-%Hsiao, G.~C., E.~P. Stephan, and W.~L. Wendland (1991), On the {D}irichlet
-%  problem in elasticity for a domain exterior to an arc, \textit{J. Comput.
-%  Appl. Math.}, \textit{34}(1), 1--19.
-%
-%\bibitem[{\textit{Lu and Ando}(2012)}]{LuAndo}
-%Lu, P., and M.~Ando (2012), Difference of scattering geometrical optics
-%  components and line integrals of currents in modified edge representation,
-%  \textit{Radio Sci.}, \textit{47},  RS3007, \doi{10.1029/2011RS004899}.
-
-%\end{thebibliography}
-
 %Reference citation examples:
@@
  %\end{figure}
+
 %

### 2016-09-25 writeup/rocha_etal.tex
Adds missing period.

@@
     sea-surface height delivered by the Surface Water and Ocean Topography (SWOT)
-    satellite mission
+    satellite mission.
 \end{abstract}

### 2016-09-26 writeup/rocha_etal_draft-like.tex
Adds missing refs to main text

@@
 Fundamentally, the upper-ocean density structure is well-captured by both LLC simulations:
-a comparison with Argo climatology shows that both simulations skillfully represent the Kuroshio
+a comparison with Argo climatology \citep[updated from ][]{roemmich_gilson2009} shows that both simulations skillfully represent the Kuroshio
 Extension stratification and its seasonal variability  (supporting information). Moreover, model smoothed
@@
 alternative explanation is that the shape of the baroclinic modes
-changes seasonally (supporting information).
+changes seasonally  \citep[see World Ocean Atlas 2013,
+][modal shapes in the
+supporting information]{levitus_etal2013}.
 
@@
   Science Requirements Document}.
+
+  \bibitem[{\textit{Roemmich and Gilson}(2009)}]{roemmich_gilson2009}
+    Roemmich, D., and J.~Gilson (2009), {The 2004--2008 mean and annual cycle of
+      temperature, salinity, and steric height in the global ocean from the Argo
+      Program}, \textit{Progress in Oceanography}, \textit{82}(2), 81--100.
 

### 2016-09-26 writeup/rocha_etal_supporting-info.tex
Adds missing refs to main text

@@
    \caption{Potential density: a comparison between Roemmich-Gilson Argo
-           climatology [updates from \cite{roemmich_gilson2009}] (a, b) and the 1/24$^\circ$ (c, d)
+           climatology [updated from \cite{roemmich_gilson2009}] (a, b) and the 1/24$^\circ$ (c, d)
            and  1/48$^\circ$ (e, f) LLC simulations. The Argo climatology is smoother

### 2016-10-11 writeup/rocha_etal_track-changes.tex
Adds track changes.

--- 2016-09-25 writeup/rocha_etal.tex
@@
 Fundamentally, the upper-ocean density structure is well-captured by both LLC simulations:
-a comparison with Argo climatology shows that both simulations skillfully represent the Kuroshio
+a comparison with Argo climatology \citep[updated from ][]{roemmich_gilson2009}  shows that both simulations skillfully represent the Kuroshio
 Extension stratification and its seasonal variability  (supporting information). Moreover, model smoothed
@@
 alternative explanation is that the shape of the baroclinic modes
-changes seasonally (supporting information).
+changes seasonally \citep[see World Ocean Atlas 2013,
+][modal shapes in the
+supporting information]{levitus_etal2013}.
 

### 2016-10-11 writeup/RochatEtal_ResponseToReviewers_BIS.tex
Response to minor comments.

--- 2016-09-22 writeup/RochatEtal_ResponseToReviewers.tex
@@
 
-In this revised manuscript, we have addressed the reviewers'
-comments. Fundamentally, we have included a new wavenumber-frequency
-spectral analysis suggested by reviewer 3. The new results vindicate
-our claim that inertia-gravity waves, particularly super-inertial waves,
-account for most of the summertime SSH variance and surface KE at scales
-between 10 and 100 km. We have also expanded the introduction and discussion
-to place our study in the context of Surface Water \& Ocean Topography (SWOT)
-satellite mission.
-
-We thank the reviewers  for the speedy review, constructive
-criticism, and thoughtful suggestions that have helped us improve our discussion and
-forced us to provide a better context for our study. We hope our manuscript is
-now suitable for publication and look forward to your reply.
-
 \section{Reviewer \#1}
 
-{\it This is a timely investigation into the submesoscale sea surface height (SSH) and
-velocity variations in the Kuroshio Extension region using global high-resolution
-OGCM simulations. The main conclusion of the study that the surface ocean submesoscale
-turbulence and inertia-gravity waves modulate out-of-phase seasonally, is novel and
-important. Overall, the manuscript is well presented and clearly written and I
-recommend its publication in GRL after the following comments are addressed.}\\
+{\it The authors have adequately addressed all the questions raised in my
+previous review. I recommend the manuscript be accepted for publication in GRL.}
 
-We thank reviewer \#1 for the speedy review. Below we address the reviewer’s
-comments and suggestions.
+\bdp
+  \item We thank the reviewer for constructive review.
+\edp
+
+\section{Reviewer \#2}
+
+{\it I have read the revised manuscript and the authors have adequately addressed
+     my concerns. Here I would like to point out things I noticed.
+     Please ignore those if you don't like.}
+
+\bdp
+   \item We thank the reviewer for constructive review.
+\edp
 
@@
 
-\item {\it In the bottom paragraph on P.10, the authors emphasized that "At scales larger
-than 20km, the April and October spectra based on hourly velocity snapshots are
-nearly indistinguishable from each other". This emphasis appears to be contradictory
-to the statements in other places of the manuscript that the seasonality of subinertial
-submesoscale flows is strong (e.g., those relating to Figure 1b-c and Figure 2).}\\
+\item {\it Line 166. "super-inertial IGWs" I think IGW represents any oscillation whose
+                frequency is between f and N. If so, super-inertial is not needed.}
 
-    \bdp
-        We agree that the emphasis on the similarity of hourly spectra can appear
-        contradictory. Those spectra (now Figure 4c
-        and supporting information) are similar because both submesoscale sub-inertial
-        and super-inertial flows undergo a strong seasonal cycle that is out of
-        phase --- there is a phase cancellation and the spectra based on hourly
-        data are indistinguishable within errorbars. In any event, we removed
-        that sentence from the text. The new Figure 4 should convey
-        the phase cancellation idea more effectively.
-    \edp
+\item {\it Line 174. "inertia gravity waves" should be IGWs.}
 
-\item {\it Does the mesoscale eddy variability in the Kuroshio Extension region of the model
-have a similar energy level as those in the along-track altimeter data? This point
-is worth mentioning as it impacts on the relative importance between the instability-induced
-turbulence versus inertia-gravity waves.}
+\item {\it Line 187. I am not sure if PSI is important near surface and wonder if it
+                is resolved in this model because of the vertical resolution.
+                Authors could discuss about PSI in the model with Jenn Mackinnon there.}
 
-  \bdp
-      The model smoothed fields have EKE levels similar to across-track (JASON II)
-      geostrophic EKE and gridded geostrophic EKE. We included this stantement with
-      the area-averaged EKE. Thanks for pointing this out.
-  \edp
+\item {\it Line 196. Remove "?"}
 
-\item {\it P.4, line 2 from bottom: Missing a word after "early".}
+\item {\it In supplement, page x-3 first line. Remove "r"}
 
-  Fixed.
-
-\item {\it P.6, line 5-7: Should October 15 be part of "late spring/early summer"?}
-
-  We rewrote that sentence to avoid ambiguity.
-
-\item {\it P.7, line 12: Change "KE spectra" to "KE spectrum"?}
-
-Fixed.
-
-\item {\it P.8, bottom: The black line in Figure 2c shows a non-zero divergence value.
-      Shouldn't the gridded SSH data be non-divergent by definition?}
-
-  \bdp
-    The convergence of meridians (the $\beta$-effect) accounts for the
-    small divergence depicted by the black line in Figure 2c.
-  \edp
-
-\item {\it P.11, line 4-6: Figure 4a reveals that daily averaging not only reduces the
-      kinetic energy level for scales smaller than 250km, but also those longer
-      than 500km. What part of the large-scale flow signals are being removed?}
-
-  \bdp
-  The large-scale super-inertial flows are barotropic tides. The new figure 4
-  supports this interpretation --- there is significant semi-diurnal lunar
-  tidal SSH variance (and KE) at scales larger than 500 km.
-  \edp
+\item {\it To separate IGWs and other submesoscale motion, Ertel vorticity could be
+      used (e.g. Kunze 1993, JPO, 2567-2588). This paper is about deep ocean though.}
 
@@
 
-%
-% Reviewer 2
-%
-
-\section{Reviewer \#2}
-{\it In this paper, the authors describe seasonality of submesoscale (10-100 km)
-activities in Kuroshio Extension region by using outputs from their new high
-resolution numerical simulations. Authors found that submesoscale turbulence
-dominates in late winter and early spring because of mixed layer instabilities, and
-inertial-internal waves in late summer and early fall because of internal tides and
-stratification near the surface. This new simulation, which includes tides and a high
-frequency wind forcing, and authors' attempts to analyze surface submesoscale
-turbulence and near-inertial waves are potentially important to oceanographic
-community. However, some discussions seem weak, possibly because it is very
-difficult to separate sub- and super-inertial motions in the submeso dynamics as
-authors mentioned.}\\
-
-We thank reviewer \#2 for the speedy review. Below we address the reviewer’s
-comments and suggestions.
-
-\subsection{Major points}
-
-\begin{enumerate}
-
-  \item {\it First of all, generation mechanism of submesoscale inertial-gravity waves in summer
-        and early fall is unclear to me. In the paper, authors attribute the mechanism to
-        internal tides (X-12, section 6, last sentence) and emphasize using the tidal forcing in
-        their model (X-5, section 2, 2nd paragraph). However, authors do not explain or show
-        any interaction mechanism of internal tides (please at least add some reference). With
-        the fixed frequencies of tidal constituents and existence of the critical latitude for the
-        diurnal tides, I am not sure what mechanism can represent the surface near-inertial
-        wave generation within the whole analyzed region by internal tides (with the
-        horizontal scales of 10-100km). If authors compare Figures 2 and 4 with those
-        estimated from the non-tidal forcing run in the Supplemental material, authors'
-        conclusion could be strengthened.}\\
-
-        \bdp
-          The generation mechanism for internal
-          waves involves changes in the wind (NIWs) and barotropic tides sloshing over
-          topographic features --- we are not claiming that there exists a different
-          generation mechanism
-          in summer compared to winter; neither are we claiming that surface NIWs
-          are generated by internal tides. What is different is the near-surface
-          expression of IGWs, which is very sensitive to the near-surface stratification
-          \citep[; cf. second paragraph of discussion]{dasaro1978}.\\
-
-          To reference potential mechanisms of how energy can quickly be transferred to
-          high wavenumbers by interaction with mesoscale-submesoscale turbulence and
-           PSI,  we added new references to the discussion \citep{mackinnon_winters2005,
-           ponte_klein2015, alford_etal2016}. \\
-
-          The simulation without tides (ECCO2 Adjoint) has much coarser
-          horizontal resolution and barely resolves the submesoscales; its grid
-          spacing is 18 km, while its effective resolution larger than 60 km.\\
-      \edp
-
-  \item {\it I also wonder why authors ignore wind-induced near-inertial waves (X-13, the first
-        line). It is well studied that wind-induced near-inertial motion in mixed layer in back
-        ground flow fields disperse quickly and propagate into the ocean interior (e.g., Young
-        and Ben-Jelloul 1997, J. Mar. Res., 55, 735-66). The distribution of those waves
-        depends on the distribution of vorticity (or Laplacian of vorticity) of background
-        flows (Klein and Llewellyn Smith 2001, J. Mar. Res., 59, 697-723; Klein et al. 2004, Q.
-        J. R. Meteorol. Soc., 130, 1153-1166). Therefore, those wind-induced near-inertial
-        waves could have the horizontal scales of 10-100km. Note that those studies are based
-        on geostrophically balanced or quasigeostrophic turbulent field. However, authors
-        may need to study interactions between submeso dynamics regime (Rossby number
-        and Richardson number are close to 1) and near-inertial waves in winter and spring
-        cases where I guess that non-linear terms may play a role. I don't think that authors
-        need to explain all internal wave dynamics in this GRL paper. I would rather suggest
-        to remove all discussions related to submesoscale inertial-gravity waves and focus on
-        describing differences between October and April.}\\
-
-        \bdp
-            We are now citing the recent review on NIWs by \cite{alford_etal2016} that contains
-            a section dedicated to mesoscale-turbulence-NIW interactions ---  \cite{alford_etal2016}
-            cite all the
-            aforementioned papers. We believe the new discussion together with Figure 4
-            and Figure S7 will help clear any ambiguity. Those figures unambiguously
-            show that most of the IGWs with horizontal scales between 10-100 km
-            are mostly super-inertial (not near-inertial); the refracted/dispersed  NIWs
-            propagate into the interior and only weakly project onto SSH \citep{alford_etal2016}.
-        \edp
-
-  \item {\it Finally, I don't understand why authors define 10-100 km as a submesoscale and used
-        this scale for the spatial filtering (X-7, section 3). Authors should discuss Figure 4 first
-        and then give a reason why they define 10-100 km as a submesoscale and use this
-        scale for the filtering. In X-11, section 5, author point out that a part of the dynamics
-        around 250 km scale should be explained by near-inertial waves. Then, I wonder why
-        authors decide to discuss internal waves in 10-100 km scale.}
-
-        \bdp
-            Following reviewer \#3's suggestion, we have added paragraphs about SWOT
-            to the introduction and conclusion. This discussion adds context to our study
-            and justifies the focus on the horizontal scales between 10-100 km. Our definition
-            of submesoscales have pros and cons (see reply to reviewer \#3's
-            minor comment \#1) but is consistent with the definition used by
-            (at least part of) the SWOT community \citep{fu_ferrari2008}. Giving the
-            definition, which we present in the first sentence of the introduction,
-            then there is no ambiguity in using 100 km as a cut-off scale for the filter.
-            We have chosen to present bulk statistics first, and then discuss details of
-            different scales --- this makes more sense now with the new wavenumber-frequency
-            analysis.
-        \edp
-
-\end{enumerate}
-
-\subsection{Minor points}
-
-\begin{enumerate}
-
-  \item {\it X-4, section 1, 2nd paragraph and X-13, section 6 last: Authors claimed that they
-        analyze the Kuroshio Extension region. However, their study region does not
-        include between $~$140E and $~$155E where the Kuroshio Extension is strongest in
-        Figure 1(a). Please explain why you exclude this region in the paper. Since those
-        longitudes includes the Izu-Ogasawara ridge (internal tide generation) and
-        formation region of Subtropical mode water (strong mixed layer front), I would
-        expect that including these latitudes would strengthen authors conclusions. I
-        would also like to point out that positions of KEO and KESS moorings in the
-        supplemental material are outside of the study region.}\\
-
-        \bdp
-             The upper bound on
-             the size of the domain is chosen for lateral homogeneity purposes,
-             necessary for the spectral analysis. We decided to focus in
-             this subdomain to avoid the strong lateral inhomogeneity and
-             anisotropies associated
-             with the Kuroshio upstream and near its separation from Japan's coast.
-             We have considered larger and smaller domains and found insiginificant
-             quantitative differences.\\
-
-             The supporting information explicitly states that both KEO and KESS
-             mooring were outside our chosen domain. We analyzed data from those
-             mooring simply to assess the skill of numerical model in representing
-             high-frequency modes. As discussed in the supporting information, the
-             comparison uses model output at the grid points closest
-             to the mooring.
-        \edp
-
-  \item {\it X-4, section 1, 3rd paragraph: “submesoscale inertial-gravity waves”. I feel that this
-        nomenclature is too sudden. It might be better to write “inertial-gravity waves
-        with a horizontal scales of 10-100 km (hereafter submesoscale inertial-gravity
-        waves)”?}\\
-
-        \bdp
-          We included the definition --- thanks for pointing that out.
-        \edp
-
-  \item {\it X-9, section 4, 2nd paragraph: I wonder what we could learn from the sentence “In
-          other words, even in April, when submesoscale turbulence prevails, only the
-          daily-averaged fields are largely in geostrophic balance”. This suggests that the
-          balanced and upward cascade motion (originated from the submesoscale
-          instability) scenario in Sasaki et al., (2014, nature comm. 5) could be wrong
-          because they use the daily averaged data (X-11, section 5, last of 1st paragraph)?
-          Or it just means that the balanced regime is actually the submeso-dynamics
-          regime?}\\
-
-          \bdp
-              This simply means that there is significant ageostrophic components,
-              even in April. Of course, daily-averaging the fields filters most of
-              ageostrophic. This does not imply that Sasaki et al's mechanism is
-              incorrect; it does suggests, however, a more complicated picture, with
-              likely small energy leak towards small scales. In any event, to avoid ambiguity, we
-              have re-written that sentence and excluded repeated information.
-          \edp
-
-  \item {\it X-10, section 4, 2nd paragraph, the last sentence: Please add some reference.}\\
-          We have re-written that sentence.
-\end{enumerate}
-
-\section{Reviewer \# 3}
-        {\it This manuscript provides model-based evidence of a seasonal cycle in submesoscale motions
-         in the western North Pacific (Kuroshio Extension). The manuscript summarizes the results of
-         a study in which two years of moderate-resolution (1/24th degree or 4.6 km) and one year of
-         high-resolution (1/48th degree or 2.3 km) model output were examined for evidence of
-         submesoscale activity. Measures of submesoscale activity include vorticity, divergence and
-         strain rate and the authors demonstrate that these quantities are enhanced in winter. Kinetic
-         energy and sea surface height (SSH) spectra also show seasonal trends. These results are
-         consistent with an existing suite of studies reporting seasonality to submesoscale flows.
-
-         What is relevant for the present study and distinguishes it from previous studies is that the
-         authors report an increase in internal wave activity during summer and which (they suggest)
-         projects onto the SSH variance. Because of the relevance of this study to challenges facing
-         the upcoming Surface Water Ocean Topography (SWOT) mission—which aims at measuring
-         submesoscale flows from space from nearly-balanced flows—this study will be of interest to
-         readers of GRL. The major error that should be addressed prior to publication is a more direct
-         connection of divergence and increased SSH variance to inertia-gravity waves. At present,
-         this connection is entirely speculative since any ageostrophic flows can give rise to
-        divergence and SSH variance. Edits should be made to the text and figures, as well.}\\
-
-        We thank  Christian Buckingham for the speedy and thorough review. Below we address
-                  Buckingham's comments and suggestions.
-
-\subsection{Major comments}
-
-\begin{enumerate}
-
-  \item {\it The authors must clearly demonstrate that increased inertia-gravity waves (IGWs) are
-        responsible for the seasonal signal in horizontal divergence and SSH spectra at high
-        wavenumbers. The authors argue that this is so but it is not demonstrated with sufficient
-        clarity. It is true that a portion of this energy/variance will be attributed to internal waves
-        modified by rotation but other flows give rise to divergence (and thus, SSH variance). Thus,
-        the seasonal changes reported by the authors might be the result of seasonal changes in these
-        other flows. Given that this is the main message of the paper, the authors must make this
-        modification.}\\
-
-        \bdp
-            We have included wavenumber-frequency
-            spectra of SSH variance (new figure 4) and KE (figure S7) with bounds for the
-            dispersion relationship of free IGWs, as suggested by the reviewer. The new
-            results unambiguously
-            indicate that IGWs account for most of the super-inertial variability at
-            scales between 10-100 km. Thanks for the excellent suggestion!
-        \edp
-
-  \item {\it There is very little introduction in this manuscript. The authors must address this
-          shortcoming prior to publication. The obvious method of addressing this issue is to discuss
-          the seasonality of IGWs in the context of SWOT and why this might matter. The introduction
-          should, therefore, place your study in the greater context and answer the question,
-          ``Why does this study matter?''}\\
-
-          \bdp
-              We agree that the planning of SWOT is perhaps the most immediate
-              application of our results. We have added an introductory paragraph dedicated to SWOT (new
-              paragraph 2 of section 1). We have also added a discussion about the
-              implication for SWOT in section 6. Thanks for making us provide a better
-              context for our study.
-          \edp
-
-  \item {\it There needs to be a discussion about the effective resolution of your model. The fact that
-        the model has a resolution of dx = 2.3 km and 4.6 km does not mean it can resolve oceanic
-        phenomena at this scale. Its effective horizontal resolution is closer to 8 times the grid
-        resolution (Soufflet et al. 2013). This corresponds to 18 km and 35 km for LLC4320 and
-        LLC2160 output, respectively. [A more exact value could be obtained by estimating zonal
-        and meridional spectra (of u, v or SSH) and identifying the high wavenumbers / low
-        wavelengths at which the spectral energy falls precipitously.] This means the model is not
-        resolving a number of submesoscale coherent vortices (SCVs) that may occur as a result of
-        ML baroclinic instability; these have diameters close to 5 km. So the follow-up questions
-        from a dynamical point-of-view become (1) where does the unresolved energy come from or
-        go to? and (2) how does this affect your results? These questions need to be addressed or at
-        least discussed if the main topic of your paper is submesoscale dynamics.}\\
-
-        \bdp
-            We are aware of the this issue --- the effective resolution based on the KE spectrum
-            was already mentioned
-            in the supporting information to the first draft.  In
-            fact, the smallest scale discussed in our paper (10 km) was set by the effective
-            resolution of highest-resolution simulation. To be more explicit, we have now included
-            the effective resolution together with the nominal resolutions in the main text (Section 2).\\
-
-            The ``unresolved energy'' does not come from anywhere --- flows smaller than
-            about 10 km are strongly damped.
-            To the extent that 1-km-scale flows are associated with a forward
-            energy transfer, then their effects on the scales concerning our study
-            (10-100 km) are likely energetically insignificant.
-            Also important to this discussion is that higher resolution simulations will resolve {\it both}
-            1-km-scale eddies and 1-km-scale IGWs that are generated via wave-wave
-            interactions and wave-vorticity interactions.\\
-
-            \cite{buckingham_etal2016} made a good point that models and observations
-            to date may not resolve the mixed-layer deformation scale in summer; this is no
-            different for our simulations. But the interpretation that this is the main
-            source of seasonally varying eddy kinetic energy is unclear because
-            summer is also the season when lateral buoyancy
-            gradients (and thus available potential energy) are minimum within the mixed layer
-            \citep[e.g., ][]{callies_etal2015}.\\
-
-            For geostrophic mixed-layer instabilities (which we call
-            shallow baroclinic instabilities), linear theory
-            predict maximum growth rates at about $2.6\times R_{ml}$, where
-            $R_{ml}$ is the mixed-layer deformation scale. But in equilibrated
-            QG turbulence driven by ML instabilities \citep{callies_etal2016} the
-            deformation scale has only a catalytic role, where eddy baroclinic
-            energy is converted into (mixed-layer) eddy barotropic energy ---
-            most of the baroclinic (mean-to-eddy) energy
-            conversion occurs at scales much larger than the mixed-layer deformation
-            scale \citep{callies_etal2016}.\\
-
-            We have added a paragraph discussing these issues (lines xx-yy).
-
-        \edp
-
- \item {\it If the authors could estimate the magnitude of the sea surface height signature within a
-        particular band as a function of season that might be helpful. For example, the authors could
-        integrate SSH spectra (Figure 4b) from hourly fields between 10 km and 33 km (i.e.,
-        wavenumber = 3 x 10-2 km-1 which appears to be the cross-over point for the two curves) to
-        obtain SSH variances within a particular wavenumber band. Additionally, under the
-        assumption that SSH variance in winter is dominated by balanced motions while SSH
-        variance in summer is the sum of balanced + unbalanced motions, the difference between
-        these two quantities yields a bound1 for the SSH variance due to unbalanced motions.}\\
-
-        \bdp
-            We have integrated the wavenumber-frequency
-            spectrum of SSH (Figure 4) to estimate variance in particular frequency
-            bands.  We have included a table with the standard
-            deviation of SSH for super-inertial and sub-inertial flows at scales
-            smaller than 100 km in different seasons. Thanks for the suggestion!
-        \edp
-
-\end{enumerate}
-
-\subsection{Minor comments}
-
-\begin{enumerate}
-
-  \item {\it (Section 1, Paragraph 1) The authors use the term ``submesoscale'' and ``mesoscale'' as
-        descriptive of motions having (1-100-km) and (100-300 km) horizontal scales,
-        respectively. I find these definitions misleading.
-
-        One way to define mesoscale and submesoscale is to define what scales of motion a
-        particular sensor will resolve. While this would not be optimal, it is objective. This is
-        probably the motivation for the definition used by Callies and Ferrari (2013), since the
-        swath altimeter (as initially proposed) would resolve scales greater than 1 km but
-        smaller than the present-day, profiling satellite altimeters – i.e., eddy-like features
-        with diameters greater than 80-100 km (see Chelton et al. 2011, Progress).
-        Unfortunately, this definition has the adverse effect of changing for each sensor.
-
-        A better definition of the submesoscale might be founded on dynamical arguments.
-        Wunsch and Stammer (1998) introduce the geostrophic equations and note that the
-        lower limit for which this approximation is strictly valid is ~30 km. This conveniently
-        defines the lower limit of the mesoscale regime. I would therefore define the
-        submesoscale regime as anything smaller than this but one for which the hydrostatic
-        approximation remains valid (e.g., 1 km).
-
-      The motivation for wanting to use the term ``submesoscale'' in the title is clear –
-      submesoscale motions are a hot topic within the community and it catches individuals’
-      eyes. This is fine. But what must be made clear within the manuscript is what scales of
-      motion and what force balances (i.e., dynamical regimes) are being described.}\\
-
-      \bdp
-          We respect the reviewer's opinion about the definition of submesoscales.
-          But there seems to be no clear definition in the literature --- the
-          Stammer $\&$ Wunsch paper states that their
-          definition of mesoscales is ``very rough'' and ``there is no generally
-          agreed upon definition''. There is also irritation that oceanographers
-          decided to term mesoscales the flows with dynamics analogous to
-          the meteorological synoptic scale motions \citep[Appendix C of ][]{wunsch2015}.\\
-
-          Some investigators define mesoscales as flows with scales near the
-          1st baroclinic deformation radius. But there is significant confusion
-          owing to factors of  $2 \pi$ --- e.g., some investigators would use $R_1 = 30$ km,
-          others would argue for $L_1 = 2\pi \times 30 \approx 190$ km
-          \citep{wunsch_ferrari2004,ferrari_wunsch2009}.
-          To add yet another controversy, \cite{larichev_held1995} argue
-          that the deformation length is not the panacea that linear theorists
-          suggest. In baroclinic geostrophic turbulence, the bulk of the
-          energy production occurs on scales much larger than the deformation
-          radius; the energy-containing scales of equilibrated baroclinic geostrophic
-          turbulence are also larger than the deformation radius \cite{larichev_held1995}.\\
-
-          According
-          to our definition the mesoscales are the energy-containing scales of the
-          flow --- the eddies that contain most of the eddy kinetic energy in the ocean
-          \citep{ferrari_wunsch2009}. From the the centroid of KE wavenumber
-          spectrum from current altimeters, this scale varies roughly from 100-300 km.
-          We then \textit{define} submesoscales as
-          the energetically subdominant scales, the scales smaller than about 100 km.
-          Our lower wavelength definition is limited by the effective resolution
-          of our higher resolution simulation ($\sim$10 km).
-          This definition is consistent with \cite{callies_ferrari2013} (they did not
-          define mesoscales/submesoscales based on scales that a sensor resolves
-          as the reviewer suggests).\\
-
-          We understand that our definition is imperfect --- we may satisfy some readers but
-          irritate others. But
-          besides the unsettled definitions in our community and consistency with our recent
-          work \citep{rocha_etal2016}, there are at least two reasons for sticking to
-          our definition: (i) this is the definition used by the SWOT community (Fu \& Ferrari
-              2008); incidentally current altimeters do not resolve
-              scales smaller then about 100 km; and (ii) our goal is to determine which flows
-              (geostrophically balanced turbulence, inertia-gravity waves, etc)
-              dominate the 10-100 km horizontal
-              scales and their seasonality. In any event,  readers of the first sentence of our
-              abstract will find our explicit definition of submesoscales.
-      \edp
-
-  \item {\it (Abstract) The abstract could be rewritten to emphasize its connection to SWOT.}\\
-
-        \bdp
-          We now state that there are implications for the accuracy of high-resolution
-          altimeters.
-        \edp
-
-  \item {\it (Section 1, last paragraph) Adding “with implications for SWOT” at the end of the
-          introduction would also help this article.}
-
-          \bdp
-            Added.
-          \edp
-
-  \item {\it (Section 1, Paragraph 1) The authors mention a suite of papers documenting
-        seasonality at the submesoscale but fail missed a few: Ostrovskii (1995) and
-        Brannigan et al. (2015). These are observation- and model-based studies, resp.}\\
-
-        We now reference to the recent paper byt Brannigan et al. (2015).
-        Ostrovskii (1995) can accessed via Buckingham et al. (2016).
-
-  \item {\it (Section 1, Paragraph 2) Model resolutions: it would be helpful for readers to place in
-        parentheses nominal horizontal resolutions corresponding to these spatial resolutions.}
-
-        \bdp
-          We have added both nominal and effective resolutions.
-        \edp
-
-  \item {\it (Section 1, Paragraph 3) What types of horizontal and temporal scales for IGWs are
-        we talking about here? How would this fall onto the Garrett-Munk (GM) internal
-        wave spectrum? (Recall: the GM spectrum contains both balanced and unbalanced
-        flows, not just internal waves, despite that the name suggests this.)}\\
-
-        \bdp
-          We explicitly state that those are submesoscale IGWs. Using
-          our definition of submesoscales, these are IGWs
-          with horizontal scales about 10-100 km\footnote{Bill Young suggested the
-          term submesoscale IGWs.}; IGWs have periods between the inertial and buoyancy
-          periods.  The GM spectrum is inaccurate
-          close to the surface (most of our results concern surface currents and SSH,
-          and it does not account for tidal and inertial peaks. We believe that the
-          new plot of the wavenumber-frequency spectrum of the SSH
-          variance (new Figure 4) --- following the reviewer's suggestion ---
-          will more clearly convey the idea that there is no horizontal scale
-          separation between submesoscale IGWs and submesoscale turbulence.\\
-
-          If the reviewer is using ``balanced'' as a shorthand to ``geostrophically
-          and hydrostatically
-          balanced'', then the comment about the GM spectrum containing both balanced
-          and unbalanced flows is incorrect.  GM used linear theory
-          of IGW to synthesize observations, mostly mooring data, in the IGW band
-           --- they
-          fitted an analytic model to (super-inertial) frequency spectra and
-          different vertical modes \citep[see][and references therein]{munk1981}.
-          One could argue
-          that other flows project onto similar temporal scales, but those motions
-          are super-inertial (e.g., stratified turbulence), not (sub-inertial)
-          balanced flows.
-        \edp
-
- \item {\it Discretization of the underlying equations of motion. How do you expect the
-        combination of vertical and horizontal resolution modifies internal wave properties
-        that exist in your model? Might you be attributing characteristics to these waves that
-        would be different in the real ocean?}\\
-
-        The model resolves IGWs well, particularly those waves with periods larger
-        than 1h considered in this paper --- we are restricted to the output
-        frequency. Higher-frequency waves that project onto lateral scales smaller
-        than 10 km are potentially inaccurate but strongly damped in our simulations.
-
-\item {\it There should be some discussion of the seasonality of the mixed layer deformation
-      radius. Buckingham et al. (2016) point out that all modeling and observational studies
-      to date necessarily introduce seasonally varying energy by virtue of the deformation
-      radius changing with season and falling below the grid resolution in certain seasons.
-      This might be worth a comment.}
-
-      \bdp
-          See major point $\#$ 3.
-      \edp
-
-\item {\it (Section 2, Paragraph 1) ``The LLC4320 simulation is an extension of the 3-month
-      long output used by Rocha et al. (2016).'' I think this is good to mention, here, so
-      please keep. In contrast, in (Section 5, Paragraph 2) the authors note a consistency
-      between the results in the Kuroshio Extension and those in the Drake Passage. This is
-      not a consistency as it is a different location, different time. The author has already
-      made the reader aware of the previous study so there is no reason to cite again.}\\
-
-        Removed.
-
-\item {\it (Section 5, Paragraph 1) eliminate ``spectral'' after Hanning window. Also, no need for
-        quotation marks.}\\
-
-        Removed.
-
-\item {\it (Section 5, Paragraph 1) ``the projection of these flows onto different horizontal scales
-            ...'' Have you thought about computing scale-dependent vorticity, strain rate and
-            divergence? This is really what you would be doing if you did a breakdown
-            of the aforementioned flows into spectral space.}
-
-            \bdp
-                Following the reviewer's suggestion, we have replaced the old section
-                5 with a discussion of the wavenumber-frequency spectrum of SSH variance.
-                To better illustrate the projection of different flows across
-                different horizontal scales, figure 5c shows the integral of the
-                wavenumber-frequency spectrum over frequencies --- the result is
-                very similar to calculating the wavenumber spectrum directly from
-                hourly and daily-averaged fields (old figure 4b).
-            \edp
-
-\item {\it (Section 5, Paragraph 3) Remove the exclamation mark.}\\
-
-    Removed.
-
-\item {\it (Section 4, first paragraph) I understand the motivation for examining the Laplacian of
-      sea surface height but the average reader will not understand the connection.}
-
-      \bdp
-      Readers familiar with basic geostrophy concepts are likely to understand the connection.
-      We make the connection between the Laplacian of SSH and relative vorticity
-      in the caption of figure 3.
-      \edp
-
-\item {\it (Section 4, second paragraph) Last sentence. This is good. I understand this but to
-        make this legible for the general audience it should be explained why we expect a near
-        one-to-one line for the joint-PDF of vorticity and the Laplacian of SSH.}
-
-      As above.
-
-\item {\it (Section 4, last paragraph) ``consistent with linear inertia-gravity waves''.
-        It would be
-        helpful to have a reference after this statement.}
-
-        \bdp
-        Added.
-        \edp
-
-\item {\it (Section 4) I find “jPDF” odd. Try “joint-PDF” or “PDF” in general.}\\
-
-We changed jPDF with joint-PDF; it does look much better.
-
-\item {\it (Section 4, last paragraph) ``...whereas divergence is moderately, negatively skewed as
-        predicted by ...'' Specifically mention convergence/downwelling.}
-
-        \bdp
-            We added ``convergence/downwelling'' to the text (line xx).
-        \edp
-
-\item {\it (Figure 1) ** Place panels (d) and (e) before panels (b) and (c) since this is the order
-        in which you refer to them in the manuscript. ** Difficult to see contours in (d) and
-        (e). ** Consider overlaying mixed layer depth on these transects since it is mentioned
-        in the text.}
-
-        \bdp
-          We have changed the order of the panels and plotted the mixed layer depth
-          on the density plots.
-        \edp
-
-\item {\it (Figure 2) Is the mean of the strain rate zero? Is the mean of the divergence zero? The
-        variance should be used rather than the root-mean-squared (RMS).}
-
-        \bdp
-            At every snapshot, the horizontally-averaged divergence is nearly zero;
-            the strain rate horizontal average is not zero, because by definition
-            $\alpha \ge 0$. The (normalized) RMS better characterizes the the bulk
-            gradient Rossby number of the flow.
-        \edp
-
-\item {\it (Figure 2) What is the sensitivity of the estimates to the stencil type? See Arbic et al.
-        paper describing this sensitivity for satellite altimetry.}
-
-        \bdp
-            To asses the error of centered second-order scheme we periodized
-            our domain with reflections. Statistics of lateral velocity gradients
-            obtained via spectral differentiation are 10-15$\%$ larger than the ones
-            obtained using second-order finite differences. But there are no qualitative
-            changes to Figure 2 and no changes in our conclusions. For simplicity, we
-            opt to use finite differences. We now report this estimate of accuracy of
-            our finite difference scheme (lines xx-xx).
-        \edp
-
-\item {\it (Section 3 and Figures 2 \& 3) At what depth are you computing these quantities?}\\
-
-  All quantities are computed at the surface as indicated in the title of the
-  section. We now explicitly indicate that in the caption of figures 2 \& 3.
-
-\item {\it (Section 2, last paragraph) It is not just solar radiation that enters the surface buoyancy
-      flux calculation.}
-
-      We rewrote that sentence.
-
-\item {\it (Section 2, last paragraph) ``as shallow as 40 m'' – I bet that the mixed layer depths get
-      even shallower than this. At midlatitudes, 20-30 m is not uncommon.}\\
-
-      We rewrote that sentence.
-
-\item {\it 24. (Section 3, Paragraph 1) ``These diagnostic highlight the submesoscale structures in
-      the flow.'' It might help the authors to motivate their calculations using the following:}\\
-
-      {\it ``Any horizontal velocity field, uh, can be expressed in terms of vorticity, divergence,
-      strain rate and mean flow. This can be seen by expressing uh as a Taylor series
-      expansion and then decomposing the velocity gradient into symmetric and anti-
-      symmetric parts [Landau and Lifshitz, 1987]. Because vortices, vertical motion and
-      fronts are ubiquitous at the submesoscale, this is a natural decomposition and
-      variances of these quantities are apt descriptors of submesoscale turbulence.''}\\
-
-      \bdp
-        We included a brief description/motivation for the use of second-order statistics
-        of those quantities. Thanks for pointing this out.
-      \edp
-
-\item {\it (Figure 3) Would be nice to include a third row that contains the collapsed (i.e., one-
-      dimensional) PDFs of normalized vorticity to illustrate the skewness of hourly and
-      daily-averaged flow. I believe you already have this information in the supplementary
-      material but would be nice to illustrate to the reader. Also, in your normalization of
-      vorticity, are you accounting for the fact that the Coriolis parameter changes with
-      latitude? Your domain spans considerable changes in f and the relevant dynamical
-      parameter is the gradient Rossby number.}
-
-      \bdp
-        We have experimented with a new figure 3, including a third row to show the
-        vorticity PDFs. But the extra plots add little extra, if any, information
-        and take a lot of space. For what it is worth, we explicitly give the relevant statistic
-        (skewness) for each month, etc.  As the reviewer mentions, the collapsed
-        PDFs for vorticity and divergence are shown in Figure S2, which compares
-        April/October and hourly/daily-averaged in a single plot.\\
-
-        Yes, as we mention in the text, we have normalized by those
-        quantities by the \textit{local} inertial frequency, not the inertial
-        frequency at mid-latitude.
-      \edp
-
-\item {\it (Figure 4b) Strictly speaking, the label on the y-axis should not be SSH variance. SSH
-      variance would be the integral of this curve. The y-axis should be labelled as the
-      power spectral density of sea surface height (SSH), or SSH Spectra.}\\
-
-      \bdp
-      We label the y-axis of the SSH variance spectrum as \textit{SSH variance density}
-      --- short for \textit{SSH variance spectral density} --- not \textit{SSH variance} as the reviewer
-      suggests.  We chose to use the short term to avoid cluttering the label
-      since there is no ambiguity because we include the
-      units of SSH variance per unit wavenumber.
-      We further disagree that the y-axis should be labelled \textit{power spectral density of SSH}
-      or \textit{SSH spectra}. On physical grounds, \textit{power} adds
-      unnecessary ambiguity,
-      because SSH variance density has nothing to do with \textit{power}.\footnote{The
-      term \textit{power spectral density}  appears to have been introduced
-      for \textit{frequency} spectrum of various quantities in signal processing (Rob Pinkel, personal
-      communication).}
-      And ``SSH [wavenumber] spectrum'' is the plot of SSH variance spectral density
-      as a function of wavenumber, not the quantity in the y-axis.
-      \edp
-
-\item {\it The authors may find the work of Brüggeman and Eden (2015) helpful in addressing
-      some of the aforementioned questions.}\\
-
-      Thanks.
-
-\end{enumerate}
+\section{Reviewer \#3}
 

### 2016-10-11 writeup/rocha_etal.tex
Starts review

@@
 \begin{acknowledgments}
- William R. Young and the reviewers --- Christian Buckingham and two anonymous referees ---
+ William R. Young and three reviewers
   provided helpful and stimulating comments. Greg Wagner

### 2016-10-11 writeup/rocha_etal.tex
Adds reviewers reviwer2

@@
 small difference between the spectra owing to the phase cancellation between the
-seasonal cycle of submesoscale turbulence and inertia-gravity waves. The sub-inertial flow
+seasonal cycle of submesoscale turbulence and IGWs. The sub-inertial flow
 has a steeper submesoscale SSH variance wavenumber spectrum (dashed lines in figure \ref{fig4}c).
@@
 changes seasonally \citep[see World Ocean Atlas 2013,
-][modal shapes in the
-supporting information]{levitus_etal2013}.
+][modal shapes in the supporting information]{levitus_etal2013}.
 

### 2016-10-11 writeup/rocha_etal_supporting-info.tex
Adds reviewers reviwer2

@@
 using a linearized equation. While the nominal horizontal resolution  of the
-LLC4320 and LLC2160 simulations are 1/48$^\circ$ and 1/24$^\circ$, r
+LLC4320 and LLC2160 simulations are 1/48$^\circ$ and 1/24$^\circ$,
 horizontal wavenumber spectra suggests an effective resolution of about 8 km and 20 km,

### 2016-10-11 writeup/RochatEtal_ResponseToReviewers_BIS.tex
Starts review
Adds reviewers reviwer2
Start minor revision

@@
 
+\usepackage{changes}
+
 \usepackage{fourier, heuristica}
@@
 \usepackage{graphicx}
-\usepackage[x11names]{xcolor}
-\usepackage{colortbl}
-\usepackage{caption}
-\DeclareCaptionFont{blue}{\color{LightSteelBlue3}}
-
-\newcommand{\foo}{\color{LightSteelBlue3}\makebox[0pt]{\textbullet}\hskip-0.5pt\vrule width 1pt\hspace{\labelsep}}
 
@@
 
+\bdp
+   \item Changed.
+\edp
+
 \item {\it Line 187. I am not sure if PSI is important near surface and wonder if it
@@
 
+\bdp
+  \item A reference  was amiss --- the ``?'' was raised during the\, \LaTeX\, compilation. Fixed.
+\edp
+
 \item {\it In supplement, page x-3 first line. Remove "r"}
+
+\bdp
+  \item Removed.
+\edp
 
@@
       used (e.g. Kunze 1993, JPO, 2567-2588). This paper is about deep ocean though.}
+
+      \bdp
+        \item Thanks. Better to use APV.
+      \edp
 
@@
 
+{\it The manuscript provides model-based evidence of a seasonal cycle in submesoscale motions
+in the western North Pacific (Kuroshio Extension). The manuscript summarizes the results of
+a study in which two years of moderate-resolution (1/24th degree or 4.6 km) and one year of
+high-resolution (1/48th degree or 2.3 km) model output were examined for evidence of
+submesoscale activity. Measures of submesoscale activity include vorticity, divergence and
+strain rate and the authors demonstrate that these quantities are enhanced in winter. Kinetic
+energy and sea surface height (SSH) spectra also show seasonal trends. These results are
+consistent with an existing suite of studies reporting seasonality to submesoscale flows. What
+is relevant for the present study and what distinguishes it from previous studies is that the
+authors report an increase in internal wave activity during summer and which projects onto
+the SSH variance. Because of the relevance of this study to challenges facing the upcoming
+Surface Water Ocean Topography (SWOT) mission—which aims at measuring submesoscale
+flows from space from nearly-balanced flows—this study will be of interest to readers of
+GRL.
+
+This is the second submission of the manuscript. The errors in the previous manuscript have
+been addressed. The authors provide a direct connection between inertia-gravity waves and
+SSH variance (Figure 4 of their manuscript) and illustrate that these undergo seasonality.
+They have also framed the manuscript in terms of SWOT, which helps the reader understand
+the importance of the results. There is no explicit connection between divergence and inertia gravity
+waves (IGWs), making the connection between the two segments of the manuscript
+tentative. (Part 1: analysis of the gradient tensor. Part 2: spectral analysis of SSH.) That is,
+other ageostrophic flows can contribute to this divergence. Nevertheless, the seasonality in the
+internal wave field is evident, complementing the observed seasonality in vorticity,
+divergence and strain rate. In summary, despite the potential for varying internal wave
+activity to be specifically tied to the Kuroshio Extension region, the implication of seasonal
+variation in internal wave activity is new, and I recommend publication pending a few
+revisions listed below.}
+
+\bdp
+   \item We thank the reviewer for the thorough and constructive review.
+\edp
+
+\begin{enumerate}
+ \item {\it I appreciate the changes made by the authors. I am happy to know that the suggestions in
+the previous review helped. Incidentally, I also appreciated the response by the authors
+regarding the resolution of the model and scales energized by baroclinic instability, citing
+work by Callies et al. (2016) and Larichev and Held (1995). Lastly, I am grateful for the
+authors’ acknowledgement toward the end of the manuscript. However, I would rather the
+authors remove me from the acknowledgements since I am involved in the formal review
+process. There is no problem with this and I appreciate the authors’ efforts to thank me.}
+
+\bdp
+   \item We have removed the referee's name from the acknowledgements.
+\edp
+
+\bdp
+
+\item {\it 2. I have been looking at the $\omega-k$ spectra in Figure 4. There is a lot of interesting
+information in these data. In particular, I have found three characteristics of interest. See
+attached figure for my definition of corresponding regions.}
+
+    \subitem (a) {\it Region 1: If one interprets submesoscale but larger-period waves (e.g., $\omega$ < f32.5 \& $k$ >
+10-2 km-1
+) as associated with gradient wind- or geostrophically-balanced flows and/or
+front-related flows (note: I am not certain of this interpretation1
+), there is considerable
+change in energy from April to October, reflective of frontogenetic activity being
+enhanced in late winter/early spring. This is consistent with the authors’ depiction of
+surface fields in Figure 1d and 1e.}
+
+  \subitem (b) {\it Region 2: Another interesting feature is the lack of energy in the inertia-gravity wave
+(IGW) band (i.e., between mode-1 at f40 and mode-4 at f25) in April but presence of
+this energy in October. This is a nice result.}
+
+\subitem (c) {\it Region 3: Finally, the last feature I note is the change in energy at $\omega$ = 2f32.5 and at
+horizontal wavenumbers $k$ > 8 x10-3 km-1 (i.e., horizontal wavelengths < 125 km).
+Although these are IGWs themselves and may result from increased stratification as
+mentioned by the authors, there still does seem to be elevated energy at 2f during
+October. As one does not expect internal waves at the M2 frequency to change with
+season, this seasonal change in energy is most likely explained by wave-wave
+interactions. I confess I don’t know much about this topic but there is the suggestion
+of such waves generated as a result of near-inertial wave interaction with quasigeostrophic
+flows (Wagner and Young 2016). Perhaps the authors could discuss this
+with GW or BY since they are mentioned in the acknowledgements. Niwa and Hibiya
+(1999) also document such motions generated by an atmospheric disturbance and
+which Olbers (1983) suggests can propagate very long distances. (Is it possible that
+typhoon-related depressions that cross over the Kuroshio during October but not in
+April excite inertial motions of frequency 2f and that then propagate into the analysis
+region? The authors need not answer this in their paper ... Food for thought.) In
+summary, I think (1) it would be worth mentioning that the energy at $\omega$ = 2f is
+seasonally varying, as the authors are documenting, (2) that this projects onto SSH
+variance at small horizontal scales, and (3) will be a source of energy seen by the
+SWOT altimeter. This may be a result specific to the Kuroshio region but is
+nevertheless interesting.}
+
+\edp
+
+\bdp
+\item {\it I recommend the authors use a different colormap when plotting spectra in Figure so that the
+reader can see changes in internal wave energy more easily. Or perhaps contour levels?}
+\edp
+
+\bdp
+\item {\it As a side note, Reviewer 2 made a comment in the previous draft about near-inertial waves
+(NIWs). S/he hypothesized that they may be an important source of SSH variance. Given the
+above conversation, this Reviewer may be at least partly correct. It seems, though, as you say
+that the waves near f penetrate deep into the bottom, since there is virtually no SSH variance
+here}
+\edp
+
+\item{\it In the previous submission, I suggested estimating the magnitude of the sea surface height
+signature within a particular band as a function of season. You have done this and it is nice.
+Table 1 is also helpful for placing these quantities into context.
+
+2 ne question I have though is that, given that the bands described above have rather
+complicated shape in $\omega-k$ space, might you instead integrate the PSD over these contoured
+regions to determine the SSH variance found within each band and corresponding to each
+type of motion? It would be interesting, for example, to know what fraction of the SSH signal
+could be directly attributed to (1) IGWs motions within the dispersion relation curves and for
+$k$ > 10-2 km-1 (region 2) (2) motions near $\omega$ = 2f32.5 \& $k$ > 10-2 km-1 (region 3) and (3) the
+other portion of the spectra (region 1), which may reflect balanced and nearly-balanced
+submesoscale motions. While the authors may not anticipate understanding energy within
+these bands completely, future scientists may cite these motions as contributing to an
+appreciable amount of the submesoscale SSH signal. If the authors feel this is not necessary
+to include within the manuscript, this could be placed within the supporting material since it
+would be interesting to the SWOT community}
+
+\item{\it  There are a few wording mistakes that the authors should address prior to resubmission. I
+don’t list them all but here are a few:}
+
+  \subitem (a) {\it Figure 4, caption: “For reference, the gray curvish line represents” ==> Reword as,
+  “For reference, we provide the baseline requirement as specified by the SWOT
+  science team”, and provide a formal reference to where the reader can find this.}
+
+  \subitem (b) {\it   Line 168: “The Similarly” ==> “Similarly...''}
+
+  \subitem (c) {\it Lines 31-35: The revisit coverage of the SWOT spacecraft is on the order of 11 days
+  (https://swot.jpl.nasa.gov/files/swot/SRD$\_$021215.pdf). I don’t really see how the
+  inability to decouple geostrophic and ageostrophic motions presents an opportunity.
+  But that being said, I would eliminate hyphens from line 33. It seems very informal.
+  The colons in the sentence following this are fine.}
+
+\end{enumerate}
+
 \bibliography{rocha_etal.bib}

### 2016-10-11 writeup/RochatEtal_ResponseToReviewers_BIS.tex
Minor fix in caption

@@
 
-\bdp
-
 \item {\it 2. I have been looking at the $\omega-k$ spectra in Figure 4. There is a lot of interesting
@@
 
-\edp
-
-\bdp
 \item {\it I recommend the authors use a different colormap when plotting spectra in Figure so that the
 reader can see changes in internal wave energy more easily. Or perhaps contour levels?}
-\edp
 
-\bdp
 \item {\it As a side note, Reviewer 2 made a comment in the previous draft about near-inertial waves
@@
 here}
-\edp
 
@@
 
-2 ne question I have though is that, given that the bands described above have rather
+One question I have though is that, given that the bands described above have rather
 complicated shape in $\omega-k$ space, might you instead integrate the PSD over these contoured

### 2016-10-13 writeup/RochatEtal_ResponseToReviewers_BIS.tex
Small fixed

@@
 
-    \subitem (a) {\it Region 1: If one interprets submesoscale but larger-period waves (e.g., $\omega$ < f32.5 \& $k$ >
-10-2 km-1
+    \subitem (a) {\it Region 1: If one interprets submesoscale but larger-period waves (e.g., $\omega < f_{32.5}$ \& $k >
+10-2$ km$^{-1}$
 ) as associated with gradient wind- or geostrophically-balanced flows and/or
@@
   \subitem (b) {\it Region 2: Another interesting feature is the lack of energy in the inertia-gravity wave
-(IGW) band (i.e., between mode-1 at f40 and mode-4 at f25) in April but presence of
+(IGW) band (i.e., between mode-1 at $f_{40}$ and mode-4 at $f_{25}$) in April but presence of
 this energy in October. This is a nice result.}
@@
 \subitem (c) {\it Region 3: Finally, the last feature I note is the change in energy at $\omega$ = 2f32.5 and at
-horizontal wavenumbers $k$ > 8 x10-3 km-1 (i.e., horizontal wavelengths < 125 km).
+horizontal wavenumbers $k > 8 \times 10-3$ km$^{-1}$ (i.e., horizontal wavelengths $< 125$ km).
 Although these are IGWs themselves and may result from increased stratification as
-mentioned by the authors, there still does seem to be elevated energy at 2f during
+mentioned by the authors, there still does seem to be elevated energy at $2f$ during
 October. As one does not expect internal waves at the M2 frequency to change with
@@
 typhoon-related depressions that cross over the Kuroshio during October but not in
-April excite inertial motions of frequency 2f and that then propagate into the analysis
+April excite inertial motions of frequency $2f$ and that then propagate into the analysis
 region? The authors need not answer this in their paper ... Food for thought.) In
-summary, I think (1) it would be worth mentioning that the energy at $\omega$ = 2f is
+summary, I think (1) it would be worth mentioning that the energy at $\omega$ = $2f$ is
 seasonally varying, as the authors are documenting, (2) that this projects onto SSH
@@
 could be directly attributed to (1) IGWs motions within the dispersion relation curves and for
-$k$ > 10-2 km-1 (region 2) (2) motions near $\omega$ = 2f32.5 \& $k$ > 10-2 km-1 (region 3) and (3) the
+$k > 10-2$ km$^{-1}$ (region 2) (2) motions near $\omega$ = 2$f_{32.5}$ \& $k$ > 10-2 km$^{-1}$ (region 3) and (3) the
 other portion of the spectra (region 1), which may reflect balanced and nearly-balanced
@@
 
+  \bdp
+      \item \noindent Done.
+  \edp
+
   \subitem (b) {\it   Line 168: “The Similarly” ==> “Similarly...''}
+
+  \bdp
+      \item \noindent Fixed.
+  \edp
 
@@
 
+  \bdp
+    \item {\noindent We removed the dashes. There are many opportunities associated with observations
+          of ageostrophic submesoscale SSH from SWOT. At very least, the projection of both geostrophic
+          and ageostrophic flows onto SSH --- and the interest in separating those flows --- will boost research
+          on interactions between those types of motion. We have changed that sentence to make this point clear. }
+  \edp
+
 \end{enumerate}

### 2016-10-13 writeup/rocha_etal.tex
Minor fix in caption
Small fixed

@@
 \begin{abstract}
-    The vigorous seasonal cycle of the mixed layer modulates upper-ocean submesoscale turbulence.
+    Recent studies show that the vigorous seasonal cycle of the mixed layer modulates
+    upper-ocean submesoscale turbulence.
     Here we provide model-based evidence that the seasonally-changing upper-ocean
@@
 inertia-gravity waves (IGWs), onto SSH represents
-a challenge --- and an opportunity --- for SWOT. The challenge: how to diagnose
-surface velocity from SSH. The opportunity: observations of submesoscale ageostrophic
-flows on a global scale.
+a challenge and an opportunity for the SWOT community. The challenge: how to diagnose submesoscale
+surface velocity from SSH. The opportunity: stimulation of research on interactions
+between geostrophic and ageostrohic flows and their projection onto SSH.
 
@@
           lunar tidal frequency ($M_2$), and the inertial frequency at mid-latitude
-          ($f_{32.5}$) and its first harmonic (2$f_{32.5}$). For reference, the gray curvish line represents
-          the SWOT baseline requirement $2\times 10^{-4} + 1.25\times 10^{-7} k^{-2}$ m$^2$/cpkm \citep{rodriguez2012}.
+          ($f_{32.5}$) and its first harmonic (2$f_{32.5}$). For reference, the gray line shows
+          the baseline requirement as specified by the SWOT science team \citep{rodriguez2012}.
           }
@@
 comparable to the April sub-inertial SSH standard deviation of $\sim$5.2 cm (Table \ref{tab1}).
-The
 Similarly, the submesoscale super-inertial surface KE is 63$\%$ larger in October

### 2016-10-13 writeup/RochatEtal_ResponseToReviewers_BIS.tex
Addresses minor points

@@
 
+In this revised version we have addressed the remaining comments and suggestions
+by reviewers 2 and 3.
+These minor changes to the text are highlighted in the file \texttt{rocha$\_$etal$\_$diff.pdf}.
+We thank the three referees for their questions, comments, and valuable suggestions,
+and look forward to your reply.
+
 \section{Reviewer \#1}
@@
 
+\bdp
+  \item  \noindent{The reviewer is correct that free IGWs have frequencies between the inertial
+         and buoyancy frequencies.  In line 166 (and elsewhere) we use the
+         qualifier super-inertial to emphasize the part of the IGW spectrum that have
+         frequencies larger than $f$ (as opposed to near-inertial
+         IGWs that have frequencies close to $f$).}
+\edp
+
 \item {\it Line 174. "inertia gravity waves" should be IGWs.}
@@
                 Authors could discuss about PSI in the model with Jenn Mackinnon there.}
+
+\bdp
+  \item The reviewer is correct. The model may only partially resolve this mechanism ---
+        the model numerical dissipation may strongly damp the high-vertical shear
+        waves created via PSI. We have added a sentence reporting this.
+\edp
 
@@
       \bdp
-        \item Thanks. Better to use APV.
+        \item {\noindent Thanks for the suggestion. We note, however, that linear waves do
+        project onto Ertel's potential vorticity. Linear waves have no \textit{available potential vorticity}
+        \citep[APV, ][]{wagner_young2015}. But APV is a new concept and its use
+        for separating IGWs and submesoscale turbulence needs to be tested in idealized
+        scenarios, which is beyond the scope of our study.}
       \edp
@@
 
-    \subitem (a) {\it Region 1: If one interprets submesoscale but larger-period waves (e.g., $\omega < f_{32.5}$ \& $k >
+    \subitem {\it Region 1: If one interprets submesoscale but larger-period waves (e.g., $\omega < f_{32.5}$ \& $k >
 10-2$ km$^{-1}$
@@
 
-  \subitem (b) {\it Region 2: Another interesting feature is the lack of energy in the inertia-gravity wave
+  \subitem {\it Region 2: Another interesting feature is the lack of energy in the inertia-gravity wave
 (IGW) band (i.e., between mode-1 at $f_{40}$ and mode-4 at $f_{25}$) in April but presence of
@@
 
-\subitem (c) {\it Region 3: Finally, the last feature I note is the change in energy at $\omega$ = 2f32.5 and at
+\subitem {\it Region 3: Finally, the last feature I note is the change in energy at $\omega$ = 2f32.5 and at
 horizontal wavenumbers $k > 8 \times 10-3$ km$^{-1}$ (i.e., horizontal wavelengths $< 125$ km).
@@
 
+\bdp
+    \item \noindent Our interpretation of the seasonally varying upper-ocean at
+                    scales between 10-100 km focuses two regions of the spectrum:
+                    sub-inertial flows (region 1) and super-inertial motions that
+                    are dominated by IGWs (region 2). These two regions have
+                    out-of-phase seasonal cycle.\\
+
+          \noindent We have discussed these spectra with Young and Wagner.
+          The $2f$ wave generation is indeed a possibility (it's the first harmonic
+          of the energetic near-inertial waves). But we note that $2 f_{32.5}$ is
+          close to energetic $M_2$ tide, and the strong seasonally varying signal
+          at high-wavenumber appears to be associated with $M_2$ tide. The $2 f_{32.5}$
+          line was added to those plots because \textit{parametric subharmonic instability}
+          may take place, but PSI may not be fully resolved by this model. To avoid
+          misleading the reader, we have removed the $f_{32.5}$ reference line from the spectra.
+\edp
+
 \item {\it I recommend the authors use a different colormap when plotting spectra in Figure so that the
 reader can see changes in internal wave energy more easily. Or perhaps contour levels?}
+
+\bdp
+  \item \noindent We changed the colormap --- the differences are insignificant, however.
+        But, perhaps subjectively, this new colormap seems better. More importantly,
+        both the previous and the new colormaps are \textit{perpetually uniform}
+        --- they accurately depict the data.
+\edp
 
@@
 here}
+
+\bdp
+  \item \noindent Exactly --- that's our interpretation.
+\edp
 
@@
 would be interesting to the SWOT community}
+
+\bdp
+    Because the submesoscale super-inertial is dominated by IGWs, there's insignificant changes
+    between integrating within the shaded region and the whole submesoscale super-inertial region.
+    We agree, however, that integrating over different bands may interest the community. Hence,
+     we have extended table 1 to account for near$-f$ band and near-$M_2$ bands. This extended
+     table shows that there is insignificant submesoscale near-$f$ waves and that waves with
+     frequency near $M_2$ --- which contain $2 f_{32.2}$ waves --- undergo strong seasonality (
+     see reply to point 1). Thanks for the suggestion!
+
+\edp
+
+\item {\it There is no explicit connection between divergence and inertia-gravity waves (IGWs). That
+is, other ageostrophic flows can contribute to divergence. I don’t think this is a deal-breaker
+for the study. It would help with the read-ability of the two portions of the paper – i.e.,
+reporting the divergence and then computing spectra to show that IGWs matter. The authors
+could address this portion of their manuscript by bandpass-filtering the horizontal velocity u-
+fields using frequency-wavenumber bands and quantifying what fraction of the divergence is
+explicitly associated with these waves. If the authors no longer have velocity fields, there may
+be a way to do this from the polarization equations for linear internal waves but I’m not sure.}
+
+\bdp
+\item {\noindent This point is addressed in figure 2, which shows that super-inertial submesoscale
+      flows account for most of the divergence. The wavenumber-frequency spectra of
+      SSH variance (figure 4) and KE (supplementary figure S7) confirms that IGWs account
+      for most of this super-inertial signal. We chose to show the SSH variance spectra in
+      the main text (as opposed to the KE spectra) because of its applications for SWOT.}
+\edp
 
@@
 
-\bibliography{rocha_etal.bib}
+\bibliography{refs.bib}
 

### 2016-10-13 writeup/rocha_etal.tex
Addresses minor points

@@
           waves from mode 1 through mode 4 across the latitudinal domain; horizontal dashed lines represent the semi-diurnal
-          lunar tidal frequency ($M_2$), and the inertial frequency at mid-latitude
-          ($f_{32.5}$) and its first harmonic (2$f_{32.5}$). For reference, the gray line shows
+          lunar tidal frequency ($M_2$) and the inertial frequency at mid-latitude
+          ($f_{32.5}$). For reference, the gray line shows
           the baseline requirement as specified by the SWOT science team \citep{rodriguez2012}.
@@
   super-inertial ($>1.2f$) & 2.9 cm  & 4.9 cm\\
+  near-$f$ ($0.8-1.2 f$) & 0.3 cm  & 0.4 cm\\
+  near-$M_2$ ($0.8-1.2 M_2$) & 1.5 cm  & 3.1 cm\\
   \hline
@@
  and parametric
-sub-harmonic instability \citep[e.g., ][]{mackinnon_winters2005}.
+sub-harmonic instability \citep[PSI, e.g., ][]{mackinnon_winters2005}, though
+the LLC simulations unlikely fully resolve PSI.
 In the LLC simulations,

### 2016-10-13 writeup/RochatEtal_ResponseToReviewers_BIS.tex
Finish first take on minor reviews

@@
 In this revised version we have addressed the remaining comments and suggestions
-by reviewers 2 and 3.
-These minor changes to the text are highlighted in the file \texttt{rocha$\_$etal$\_$diff.pdf}.
-We thank the three referees for their questions, comments, and valuable suggestions,
-and look forward to your reply.
+by reviewers 2 and 3. We highlighted these minor changes in the file \texttt{rocha$\_$etal$\_$diff.pdf}.\\
+
+\noindent We thank the three referees for their questions, comments, and valuable suggestions.
 
@@
 \bdp
-  \item  \noindent{The reviewer is correct that free IGWs have frequencies between the inertial
-         and buoyancy frequencies.  In line 166 (and elsewhere) we use the
-         qualifier super-inertial to emphasize the part of the IGW spectrum that have
-         frequencies larger than $f$ (as opposed to near-inertial
-         IGWs that have frequencies close to $f$).}
+  \item  \noindent{The reviewer is correct that free IGWs have frequencies between
+          the inertial and buoyancy frequencies. In line 166 (and elsewhere) we use
+          the qualifier super-inertial to emphasize the part of the IGW spectrum
+          with frequencies larger than f (as opposed to near-inertial IGWs with
+          frequencies close to $f$ ).}
 \edp
@@
                 is resolved in this model because of the vertical resolution.
-                Authors could discuss about PSI in the model with Jenn Mackinnon there.}
+                Authors could discuss about PSI in the model with Jen Mackinnon there.}
 
 \bdp
-  \item The reviewer is correct. The model may only partially resolve this mechanism ---
-        the model numerical dissipation may strongly damp the high-vertical shear
-        waves created via PSI. We have added a sentence reporting this.
+  \item \noindent The reviewer is correct. The model likely only partially resolve
+        this mechanism — numerical dissipation may strongly damp the high vertical
+        shear waves created via PSI. We have added a sentence reporting this.
 \edp
@@
 \bdp
-   \item We have removed the referee's name from the acknowledgements.
+   \item We have removed the referee's name from the acknowledgments.
 \edp
@@
     \item \noindent Our interpretation of the seasonally varying upper-ocean at
-                    scales between 10-100 km focuses two regions of the spectrum:
-                    sub-inertial flows (region 1) and super-inertial motions that
-                    are dominated by IGWs (region 2). These two regions have
-                    out-of-phase seasonal cycle.\\
+          scales between 10-100 km focuses two regions of the spectrum: sub-inertial
+          flows (region 1) and super- inertial motions (region 2). IGWs dominate
+          both SSH variance and KE in region 2. These two regions have an out-of-phase seasonal cycle.\\
 
-          \noindent We have discussed these spectra with Young and Wagner.
-          The $2f$ wave generation is indeed a possibility (it's the first harmonic
-          of the energetic near-inertial waves). But we note that $2 f_{32.5}$ is
-          close to energetic $M_2$ tide, and the strong seasonally varying signal
-          at high-wavenumber appears to be associated with $M_2$ tide. The $2 f_{32.5}$
-          line was added to those plots because \textit{parametric subharmonic instability}
-          may take place, but PSI may not be fully resolved by this model. To avoid
-          misleading the reader, we have removed the $f_{32.5}$ reference line from the spectra.
+          \noindent We have discussed these spectra with Young and Wagner. The $2f$
+           wave generation is indeed a possibility (it’s the first harmonic of the
+          energetic near-inertial waves). But we note that $2f_{32.5}$ is close to
+          the $M_2$ tide, and the seasonally varying signal at high-wavenumber appears
+          to be associated with the $M_2$ tide. The $2f_{32.5}$ line was added to those plots
+          because parametric subharmonic instability may take place, but this model
+          may not fully resolve PSI. To avoid misleading readers, we have removed
+          the $f_{32.5}$ reference line from the spectra.
 \edp
@@
         But, perhaps subjectively, this new colormap seems better. More importantly,
-        both the previous and the new colormaps are \textit{perpetually uniform}
+        both the previous and the new colormaps are \href{https://en.wikipedia.org/wiki/Color_difference#Tolerance}{\tt perpetually uniform}
         --- they accurately depict the data.
@@
 \bdp
-    Because the submesoscale super-inertial is dominated by IGWs, there's insignificant changes
-    between integrating within the shaded region and the whole submesoscale super-inertial region.
-    We agree, however, that integrating over different bands may interest the community. Hence,
-     we have extended table 1 to account for near$-f$ band and near-$M_2$ bands. This extended
-     table shows that there is insignificant submesoscale near-$f$ waves and that waves with
-     frequency near $M_2$ --- which contain $2 f_{32.2}$ waves --- undergo strong seasonality (
-     see reply to point 1). Thanks for the suggestion!
-
+\item \noindent Because IGWs dominate the submesoscale super-inertial, there are insignificant
+changes between integrating within the shaded region and the whole submesoscale
+super-inertial region. We agree, however, that integrating over different bands
+will interest the community. Hence, we have extended table 1 to account for near$-f$
+band and near$-M_2$ bands. This extended table shows that there is insignificant
+submesoscale near$-f$ waves and that waves with frequencies near $M_2$ — which contain
+$2f_{32.2}$ waves — undergo strong seasonality (see reply to point 1). We also added
+a similar supporting table (Table S2) based on the KE spectra. Thanks for the suggestion!
 \edp
@@
 \bdp
-\item {\noindent This point is addressed in figure 2, which shows that super-inertial submesoscale
-      flows account for most of the divergence. The wavenumber-frequency spectra of
-      SSH variance (figure 4) and KE (supplementary figure S7) confirms that IGWs account
-      for most of this super-inertial signal. We chose to show the SSH variance spectra in
-      the main text (as opposed to the KE spectra) because of its applications for SWOT.}
+\item {\noindent
+We address this point in figure 2, which shows that super-inertial submesoscale
+flows account for most of the divergence. The wavenumber-frequency spectra of SSH
+variance (figure 4) and KE (supplementary figure S7) confirms that IGWs account
+for most of this super-inertial signal. We chose to show the SSH variance spectra
+in the main text (as opposed to the KE spectra) because of its applications for SWOT.
+}
 \edp

### 2016-10-13 writeup/rocha_etal_supporting-info.tex
Finish first take on minor reviews

@@
 \item Figure S7 Wavenumber-frequency spectrum of KE for LLC4320 simulation.
+\item Table S2 The mean-square velocity at submesoscales (10-100 km)
+        estimated from the wavenumber-frequency spectrum of SSH variance in figure S7.
 %if Tables are larger than 1 page, upload as separate excel file
@@
 
+ % table
+ \begin{table}
+ \label{tab1}
+ \caption{The mean-square velocity at submesoscales (10-100 km)
+         estimated from the wavenumber-frequency spectrum of SSH variance in figure S7.
+         }
+   \begin{center}
+     \begin{tabular}{  l | c | c }
+     \hline
+     & April & October\\
+     \hline
+   sub-inertial ($<0.8f$)   & 0.32 m s$^{-1}$ & 0.20 m s$^{-1}$  \\
+   super-inertial ($>1.2f$) & 0.17 m s$^{-1}$  & 0.46 m s$^{-1}$\\
+   near-$f$ ($0.8-1.2 f$) & 0.25 m s$^{-1}$  & 0.12 m s$^{-1}$\\
+   near-$M_2$ ($0.8-1.2 M_2$) & 0.05 m s$^{-1}$  & 0.10 m s$^{-1}$\\
+   \hline
+     \end{tabular}
+   \end{center}
+ \end{table}
+
 % ---------------

### 2016-10-17 writeup/rocha_etal.tex
Updates draft-like verson

@@
 
-The model solutions depict seasonality in vorticity (Figures \ref{fig1}b-c): large values of
+The model solutions depict seasonality in vorticity (Figures \ref{fig1}d-e): large values of
 fine-grained vertical vorticity are observed in early spring with maximum
@@
  and parametric
-sub-harmonic instability \citep[PSI, e.g., ][]{mackinnon_winters2005}, though
-the LLC simulations unlikely fully resolve PSI.
-In the LLC simulations,
+sub-harmonic instability \citep[PSI, e.g., ][]{mackinnon_winters2005}, though the
+LLC simulations are unlikely to fully resolve PSI. In the LLC simulations,
 super-inertial IGWs account for most of the surface submesoscale KE and SSH variance;

### 2016-10-17 writeup/rocha_etal_draft-like.tex
Updates draft-like verson

@@
 %   when using Draft:
-\setkeys{Gin}{draft=false}
+%\setkeys{Gin}{draft=false}
 %
@@
 comparable to the April sub-inertial SSH standard deviation of $\sim$5.2 cm (Table \ref{tab1}).
-The
 Similarly, the submesoscale super-inertial surface KE is 63$\%$ larger in October
 than in April (supporting information).
+
+% table
+\begin{table}
+\label{tab1}
+\caption{The square root of the mean-square SSH at submesoscales (10-100 km)
+        estimated from the wavenumber-frequency spectrum of SSH variance in figure \ref{fig4}.
+        The SWOT baseline requirement at 15 km is approximately $\sqrt{2}$ cm.
+        }
+  \begin{center}
+    \begin{tabular}{  l | c | c }
+    \hline
+    & April & October\\
+    \hline
+  sub-inertial ($<0.8f$)   & 5.2 cm & 2.0 cm  \\
+  super-inertial ($>1.2f$) & 2.9 cm  & 4.9 cm\\
+  near-$f$ ($0.8-1.2 f$) & 0.3 cm  & 0.4 cm\\
+  near-$M_2$ ($0.8-1.2 M_2$) & 1.5 cm  & 3.1 cm\\
+  \hline
+    \end{tabular}
+  \end{center}
+\end{table}
 
@@
 
-%\section{Projection onto horizontal scales}
-%
-%To quantify the projection of these flows onto different horizontal
-%scales, we calculate wavenumber spectra of kinetic energy and sea-surface height
-% variance.  Before calculating the spectra, time-mean and spatial linear
-%trends were removed, and the resulting fields were multiplied by a two-dimensional Hanning window.
-%The two-dimensional spectra were averaged azimuthaly \citep[e.g., ][]{rocha_etal2016} .
-%We discuss only spectra for the 1/48$^\circ$ simulation, which extends the 1/24$^\circ$ simulation
-%towards smaller scales (supporting information).
-%
-%
-
-%
-%
-%\begin{figure*}[ht]
-%  \begin{center}
-%    \includegraphics[width=.65\textwidth]{figs/fig5.pdf}
-% \caption{\small Surface (horizontal) kinetic energy (a) and sea-surface height variance wavenumber spectra (b)
-% in the 1/48$^\circ$ simulation. Solid lines
-% are spectra based on hourly snapshots, and dashed lines are spectra based on daily-averaged
-% fields.}
-% \label{fig5}
-% \end{center}
-%\end{figure*}
-%
-%
-%Figure \ref{fig5}a depicts the horizontal wavenumber spectra of surface KE in
-%April and October. At scales larger than 20 km, the April and October spectra based on hourly
-%velocity snapshots (solid lines) are nearly indistinguishable from each other within 95$\%$ confidence level.
-%Consistent with the results of
-%\citet{rocha_etal2016}, who analyzed a 3-month output of the LLC4320 simulation in Drake Passage,
-%there is significant high-frequency variabilty at submesoscales. Daily-averaging
-%the velocity field suppresses spatial variability at scales smaller than about 250
-%km, both in April and October (compare solid lines against dashed lines in Figure
-%\ref{fig5}a). But this suppression is dramatic in October, when the IGWs peak. At scales
-%smaller than 100 km, 39$\%$ of the surface KE in April is accounted for by super-inertial
-%flows as opposed to 79$\%$ in October. The seasonality of subinertial submesoscale flows
-%is strong, consistent
-%with the results of \citet{sasaki_etal2014}, which are
-%based on daily-averaged velocity fields of a different model without tidal forcing
-%(P. Klein, personal communication).
-%
-%The IGWs project on the sea-surface.
-%There is a dramatic difference between the spectra
-%based on hourly and daily-averaged sea-surface height in October (Figure \ref{fig5}b): at scales smaller than 100 km, the spectra
-%of hourly sea-surface height roughly follows a $-2$ power-law, whereas the spectra of daily-averaged
-%sea-surface height roughly follows a $-5$ power-law. At these scales, 33$\%$ of the sea-surface height variance in April is accounted for by super-inertial
-%flows as opposed to 83$\%$ in October. Curiously, the out-of-phase seasonal cycle of submesoscale
-%turbulence and near-surface submesoscale IGWs conspire to yield weak seasonality
-%in the spectra of KE and sea-surface height variance based on hourly fields.
-%
-%
 \section{Discussion and conclusion}
@@
  and parametric
-sub-harmonic instability \citep[e.g., ][]{mackinnon_winters2005}.
-In the LLC simulations,
+sub-harmonic instability \citep[PSI, e.g., ][]{mackinnon_winters2005}, though the
+LLC simulations are unlikely to fully resolve PSI. In the LLC simulations,
 super-inertial IGWs account for most of the surface submesoscale KE and SSH variance;
@@
 
-%\bibliographystyle{agufull08}
-%\bibliography{rocha_etal}
+%  this should go in a supplementary material
+\clearpage
 
@@
 \begin{center}
- \hspace{-1.25cm}\includegraphics[width=.9\textwidth]{figs/fig1_1.pdf}\\
+\hspace{-1.25cm}\includegraphics[width=.7\textwidth]{figs/fig1_1.pdf}\\
 \vspace{-.125cm}
- \includegraphics[width=.9\textwidth]{figs/fig1_2.pdf}
+\includegraphics[width=.7\textwidth]{figs/fig1_2.pdf}
  \caption{\small (a) The study region with the subregion where the LLC outputs are
@@
           waves from mode 1 through mode 4 across the latitudinal domain; horizontal dashed lines represent the semi-diurnal
-           lunar tidal frequency ($M_2$), and the inertial frequency at mid-latitude
-           ($f_{32.5}$) and its first harmonic (2$f_{32.5}$). For reference, the gray curvish line represents
-           the SWOT baseline requirement $2\times 10^{-4} + 1.25\times 10^{-7} k^{-2}$ m$^2$/cpkm \citep{rodriguez2012}.
+          lunar tidal frequency ($M_2$) and the inertial frequency at mid-latitude
+          ($f_{32.5}$). For reference, the gray line shows
+          the baseline requirement as specified by the SWOT science team \citep{rodriguez2012}.
           }
@@
 \end{figure*}
-
- % table
- \begin{table}
- \label{tab1}
- \caption{The square root of the mean-square SSH at submesoscales (10-100 km)
-         estimated from the wavenumber-frequency spectrum of SSH variance in figure \ref{fig4}.
-         The SWOT baseline requirement at 15 km is approximately $\sqrt{2}$ cm.
-         }
-   \begin{center}
-     \begin{tabular}{  l | c | c }
-     \hline
-     & April & October\\
-     \hline
-   sub-inertial ($<0.8f$)   & 5.2 cm & 2.0 cm  \\
-   super-inertial ($>1.2f$) & 2.9 cm  & 4.9 cm\\
-   \hline
-     \end{tabular}
-   \end{center}
- \end{table}
-
-%  this should go in a supplementary material
-%\clearpage
-
-%\section{Supplemental material 1: Spectral errors}
 
@@
 
-\begin{thebibliography}{2}
-
+\begin{thebibliography}{26}
 \providecommand{\natexlab}[1]{#1}
@@
 
+\bibitem[{\textit{Levitus et~al.}(2013)\textit{Levitus, Antonov, Baranova,
+  Boyer, Coleman, Garcia, Grodsky, Johnson, Locarnini, Mishonov
+  et~al.}}]{levitus_etal2013}
+Levitus, S., J.~Antonov, O.~K. Baranova, T.~Boyer, C.~Coleman, H.~Garcia,
+  A.~Grodsky, D.~Johnson, R.~Locarnini, A.~V. Mishonov, et~al. (2013), The
+  world ocean database, \textit{Data Science Journal}, \textit{12}(0),
+  WDS229--WDS234.
+
 \bibitem[{\textit{MacKinnon and Winters}(2005)}]{mackinnon_winters2005}
@@
 
+%\begin{thebibliography}{}
+
+%\providecommand{\natexlab}[1]{#1}
+%\expandafter\ifx\csname urlstyle\endcsname\relax
+%  \providecommand{\doi}[1]{doi:\discretionary{}{}{}#1}\else
+%  \providecommand{\doi}{doi:\discretionary{}{}{}\begingroup
+%  \urlstyle{rm}\Url}\fi
+%
+%\bibitem[{\textit{Atkinson and Sloan}(1991)}]{AtkinsonSloan}
+%Atkinson, K., and I.~Sloan (1991), The numerical solution of first-kind
+%  logarithmic-kernel integral equations on smooth open arcs, \textit{Math.
+%  Comp.}, \textit{56}(193), 119--139.
+%
+%\bibitem[{\textit{Colton and Kress}(1983)}]{ColtonKress1}
+%Colton, D., and R.~Kress (1983), \textit{Integral Equation Methods in
+%  Scattering Theory}, John Wiley, New York.
+%
+%\bibitem[{\textit{Hsiao et~al.}(1991)\textit{Hsiao, Stephan, and
+%  Wendland}}]{StephanHsiao}
+%Hsiao, G.~C., E.~P. Stephan, and W.~L. Wendland (1991), On the {D}irichlet
+%  problem in elasticity for a domain exterior to an arc, \textit{J. Comput.
+%  Appl. Math.}, \textit{34}(1), 1--19.
+%
+%\bibitem[{\textit{Lu and Ando}(2012)}]{LuAndo}
+%Lu, P., and M.~Ando (2012), Difference of scattering geometrical optics
+%  components and line integrals of currents in modified edge representation,
+%  \textit{Radio Sci.}, \textit{47},  RS3007, \doi{10.1029/2011RS004899}.
+
+%\end{thebibliography}
+
 %Reference citation examples:
@@
  %\end{figure}
-
 %
